### 2017-09-29 manuscripts/AmNat_response_to_rev.Rmd
WIP

--- 
title: "Response to reviewers"
author:
  - "Milad Memarzadeh"
  - "Carl Boettiger"
output: 
  pdf_document:
    keep_tex: true
    fig_caption: true
bibliography: citations.bib
---

ASSOCIATE EDITOR'S RECOMMENDATION

Dear Professor Bronstein; Prof Dwyer,

> My recommendation is that manuscript 57722 be declined without prejudice.   The authors use a state of the art computer algorithm known as ``POMDP'' to determine, on the fly, how fisheries should be managed, in terms of the total amount of stock that can be harvested, and the profit accruing from that harvest.   This is an important problem in ecology, and the authors approach appears to offer significant insight.  The reviewers nevertheless point out some important ways in which the text could be improved, and reviewer 1 in particular argues that the authors's approach may be significantly flawed.  These problems would therefore need to be addressed for a revision to be successful.

We appreciate the careful and constructive comments of the editor and reviewers, which have allowed us to re-write the manuscript to both better communicate both the method and broader implications of the results shown here.  We also now address directly the significant differences between the extensive literature on the optimal management of fisheries to

We appreciate the time and effort of reviewer 1 in exploring these results in further simulation, but were unable to understand precisely what the reviewer had implemented. From the description provided, it appears the reviewer's results rather underscore our point about the importance of observation uncertainty, an uncertainty which does not appear to be fully present in the reviewer's simulation, as we discuss below.  In short, if the stock size is unknown (due to measurement uncertainty), how does one enforce the safe-guard that total allowable catch quota or harvest not exceed 90% the stock?  As our work shows, it precisely the presence of such assumptions

We have expanded our results to include a comparisons to policies that much more closely resemble the implementation in actual fisheries

and believe our revisions should now address each of the concerns raised.  In particular,

> One of Reviewer 2's main comments is that the article focusses too narrowly on fisheries and stock assessment, and the Reviewer argues that the authors could instead frame their work more broadly in terms of applied ecology in general, perhaps by citing problems of measurement uncertainty in other applications besides fisheries.  These points are very well taken, especially given that The American Naturalist covers biological problems in a very broad sense.   For the work to be of general interest to the readership of The American Naturalist, the authors would therefore have to explain more clearly the relevance of their work to applied ecology in general.

we now emphasize that making decisions in face of imperfect information is by no means an issue isolated solely to fisheries, and that our focus on fisheries example is merely for the same reason introductory texts on ecological management have so often turned to fisheries: no other example of ecological managment has a rich or foundational a history in our understanding of ecosystem management and decision-theory.  We now discuss the broader concept of uncertain decision making in the introduction and highlight examples where these issues most readily appear in so many other ecological applications.

> Reviewer 1 in contrast makes a  specific and detailed comment about a very particular issue.   In discussing this comment, an important point is that I am sufficiently far from being a fisheries biologist that I am not entirely understand sure that I understand the objection, emphasizing the importance of Reviewer 2's comment that The American Naturalist is a general interest journal, not a fisheries journal.  This caveat aside, my understanding of Reviewer 1's comment is that the Reviewer ran a set of simulations which were essentially the same as in the manuscript, except that the Reviewer explicitly excluded the possibility of 90 percent harvest of the stock population.  In the reviewer's simulations, the optimal escapement level is lower in the presence of measurement error, whereas in fig. 5 of the manuscript,  the optimal escapement level is instead higher in the presence of measurement error.

We are afraid that Reviewer 1's concern is mistaken

> Reviewer 1's description of their simulations is brief enough that I would not expect the authors to build their entire revision around an effort to respond to this comment.   If the authors have indeed allowed for 90 percent harvest, however, it is important that they test whether their results are different if they place an upper limit on the fraction of the stock that they allow to be harvested.

When measurements are *uncertain*, it is impossible *not* to allow for a 90% harvest of the stock.  Indeed, this cuts to  the very heart of why measurement uncertainty in *decisions* is so important!  If X is a random number from a log-normal distribution of parameters mu and sigma, how can one choose a number Y to be certain that Y < 90% X?   Y = 0 is the only (non-negative) solution; you cannot be sure that *any* harvest level is guatenteed to be safe.

> I also agree with Reviewer 1's comment that the methods are not that easy to follow.  That said, to my understanding, the escapements in fig. 2 are not constant, and so Reviewer 1's comment that a constant escapement cannot lead to the collapse of a stock seems to me to be irrelevant, but again I may be wrong, and my uncertainty emphasizes the lack of clarity in the writing, especially with respect to the Methods.

The editor is absolutely correct, POMDP solutions do not result in a constant escapement, as we stated.  Unfortunately, Reviewer 1 does not provide any description of the methods they used.

We have revised the

> Page 8-12 in particular are written in a very formal style that I associate with journals that focus on mathematics rather than on biology, with the difficulty problem that much of this text has small grammatical errors that exacerbate the clarity problem.  Meanwhile, both Reviewers argued that the authors have not accurately characterized the literature in this area, and that problem would also have to be solved in a revised version.

We have revised this section significantly to address both grammar and presentation.

The formal definitions and methodology of POMDPs has already been well-documented in the appopriate literature, and there is no reason for us to rehash the all the precise technical details here, which only serves to obscure the most salient concept.  The point here is that there is an immensely well-developed area of theory and tools to address this problem of decision-making under imperfect information.  Though much of that research has come from other fields and is not currently familiar to the readers, we believe we now present enough context to understand the problem.

> In spite of these negative comments, it seems like the overall approach has a great deal of potential, as Reviewer 2 argues, and if Reviewer 1's concerns prove to be minor, then I believe that a revised version would be a substantial contribution.  That said, the complexity of the technical issues is quite high, and so it may be simpler to submit the manuscript to a journal that specializes in mathematical biology or in fisheries management.

Greg Dwyer

xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

Reviewer #1:

For complete review, see the attached PDF, which begins as follows:

> I attempted to evaluate the papers overall conclusions by a very simple simulation and these results
seem to challenge the papers conclusions.
I simulated a population with the model the authors describe with both observation and process error
lognormally distributed and using the authors parameters. For a range of escapement goals I calculated
the net present value of using that escapement goal, and compared the consequences when there was
no observation error to the case where the lognormal cv was 0.5 – that is very high observation error.
The one difference I made was I did not let the harvest exceed 90% of the total population size ... this
seems reasonable as I know of few fisheries where 90% of the population could be harvested in a single
year.

As the editor observes, this short discription does not permit us to properly understand what the reviewer has done to obtain
different results, but as the editor already noted, the comments make it clear that it is not analgous to our approach,
since our approach does not assume nor does it find a constant escapement policy to be optimal.  
Moreover, the remark about restricting harvest to 90% of population size is very instructive, as it cuts to the heart of the problem of measurement uncertainty. 
Clearly such a restriction is only possible if the manager has access to perfect observations of the stock size,
an thus any simulation that enforces this restriction is implicitly ignoring the reality of measurement error.

There are only two ways a manager can be certain that harvest does not exceed 90% of stock: (1) don't harvest, or (2) measure the stock size exactly, (or at least, exactly measure some absolute lower bound of possible stock sizes).

As the literature has observed consistently for some time, (e.g. see an eloquent discussion in Clark & Kirkwood 1987) 
if we only consider policies in which accidental population extinction by overharvesting is impossible, 
(e.g. because we assume a priori harvest cannot exceed 90% of the stock), 
there is every reason to suspect this will result in artifically less cautious policies.  If we begin with 
the assumption that there's just no possible way overharvesting could destroy a stock, then the risks of
overharvesting are drastically reduced: net present value need only reflect the loss of revenue while a stock
rebuilds, and not the much greater loss that would result from extinction.  Because even the stochastic theory
such as Reed 1979 made effectively the same assumption (e.g. assumes the stock is measured perfectly before
harvest, so that it would never be rational to attempt a harvest)

> I can’t really understand some of the results in the paper ... for instance I don’t see how a fixed
escapement policy could lead to the collapses shown in Figure 2 – using fixed escapement policies you
can’t have you population continue to decline.

As we discuss above, this is once again fundamental feature of measurement uncertainty -- and again one
that was observed long before our paper and is well established by the literature we discuss at some length.
If a fixed escapement policy is applied to a measurement that is uncertain, the true stock will be somewhat
larger or somewhat smaller than the observed value, and thus the true escapement can be somewhat smaller
or somewhat larger than the 'target' escapement.  Without perfect measurements, it is clearly impossible
to achieve the target escapement exactly.  Overestimates of the stock size can lead to stock collapse

Without perfect measurements, the true escapement will vary under a

The novelty of our result is how to deal with this issue.

> Finally it is certainly not true that fixed escapement policies are commonly used. To my knowledge
almost all marine fish populations are managed with policies more resembling fixed harvest rates with
caveats that target harvest rate declines when stocks drop below a target stock size. Certainly some
salmon stocks are managed with an escapement goal, but they are not in any sense common or the
default in marine fisheries management.

The reviewer is entirely correct about this -- while a rigorous constant escapement policy shuts down
all harvest below B_MSY (the constant escapement level), most fisheries actually continue to harvest
even when the stock drops below the target stock size, though potentially at some decreasing rate,
and do not shut down harvest until the population is significantly below this level, commonly B_MSY / 2, if at all.

As mentioned above, we have now revised all of our analyses to consider these realistic harvest policies,
and have added a substantial discussion addressing the differences between such constant harvest (or constant effort /
fishing mortality) policies and the constant-escapement policies we have hitherto focused upon.  
Because these policies permit harvesting below B_MSY, it is not surprising that we find them even more
suseptible

Moreover, many fisheries do not

xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

Reviewer #2:

> In this clearly written paper the authors show that fish population dynamic models that ignore measurement error can result in significant overharvesting, decreased economic returns, and the potential for complete collapse of the stock. The key advance is the use of an algorithm for partially observed Markov decision processes (POMDP) to provide a complete solution rather than relying upon inferences from a deterministic model or a model of optimal control that does not update belief states during each time step.

> The treatment of this issue is remarkably comprehensive with respect to structural uncertainty of the model formulation. In fact, the authors in some way bury what I see as the lead in the Appendix. They include analysis of a population subject to an Allee effect, and produce the same general conclusion as they do via analysis of models with more standard dynamics: the POMDP approach is less likely to cause overharvest and population collapse than alternative approaches, even if the location of the unstable steady state population size (tipping point) is unknown. This finding is very important given the increasingly voluminous literature on thresholds and tipping points, suggesting that proper accounting for measurement uncertainty may be a practical alternative to investing more in reducing such uncertainty in the budget-constrained world of natural resource management.

> As is clear from my summary, this paper is framed for a narrower audience focused on fisheries and stock assessment. To draw in a broader readership, I suggest that the authors motivate the problem they address and interpret their results in the context of the more general ecological, wildlife, and forestry literatures. Specific examples of how measurement uncertainty has been found problematic in fisheries and wildlife assessments would be helpful.

We thank the reviewer for this careful and constructive review, and agree entirely with the recommendation that the fundamental issues here (managing with imperfect information) as well as some other themes (tipping points, management with constrained resources) are far more general. We have now tried to state these implications more clearly.

> The authors mischaracterize the extent to which measurement uncertainty is left untreated in stock assessments. Dichmont et al. 2016 (Fisheries Research) provide a nice review of how some of these measurement uncertainty issues are treated in practice in US stock assessments, and an entry point to the literature. If the authors intend to focus on theoretical models and general solutions for population dynamics, their assertion about failure to incorporate measurement uncertainty completely may be valid but this should be clarified.

We had no intention of suggesting that measurement uncertainty was under-represented in stock assessments; indeed it is precisely the *prevalance* with which stock assessment models account for uncertainty that motivates

> Unlike the classic pretty good yield paper mentioned in the paper (line 440), the authors do not adequately address how conservative harvest is likely to be met with opposition by harvesters and how this issue can be addressed directly by their analysis.

This is an excellent point which we now discuss properly. Our approach highlights that strategies such as Pretty Good Yield, taking 80% of the maximum sustainable yield, are not sufficient solutions to the challenge of measurement uncertainty, since they are still based on fundamentally deterministic premises.

As our analysis shows, unless it is possible to be certain beyond any shadow of a doubt that one has

How do we determine the maximum sustainable yield in the first place?  This number comes from a deterministic stock-recruitment model.  In stochastic models, and even in many deterministic models with age and stage structure, there is no such thing as a maximum sustainable yield, since the optimal yield must fluctuate with the population dynamics.  But at least in those cases one can reasonably assert an average

Stock assessments give us

Measurement errors accumulate.

Pretty Good Yield, taking 80% of the maximum sustainable yield, reflects the same

is defined as a percentage of maximum sustainable yield, which is an essentially deterministic concept.  It is impossible to be sure that

> Similarly, rather than dismiss management strategy evaluation as a less preferable alternative to the optimal control approach taken here, it would be nice to see the authors include a formal contrast. Might the offline POMDP approach be framed this way and results inferred without additional analyses? This would be nice because in the real world, optimal control with annual updating of harvest regulations (the POMDP approach) may not practical. The offline POMDP seems more realistic, perhaps with an online phase implemented such that belief states could be updated in 5-10 year intervals.

 As the reviewer suggests here and as was recognized by Fischer et al (2009) among others, one advantage of the optimal control approach is that it is generally more transparent and unbiased algorithmic way to reach a decision, which to some extent sidesteps the inherently more subjective existing methods, where this issue is acutely felt.  For instance, our value function is a cold-blooded profit equation which places no value in leaving fish in the sea beyond their future return as harvest.  We also ignore more realistic supply-and-demand effects as a conservative assumption, since the decreasing price of higher supply would dampen the incentive for larger harvests (see comparisons in appendix).

> Minor comments:
The analysis related to Fig 5 made me wonder how POMDP compares to alternative frameworks when a stock is increasing or decreasing in abundance. The authors mention this in their interpretation, but I think it could be highlighted more as it seems that the POMDP approach (at least for plaice) allows a more careful tracking of catch to stock size during collapse and rebuilding, a result that would be viewed favorably by fishermen and conservationists alike.

Excellent observation! We have added the following line to emphasize this more clearly:

> Some reorganization may be warranted. Several portions of the Results section describe Methods, which would be nice to signal earlier in the MS (e.g., lines 266-281, lines 298-300 and lines 321-324 could be included in Methods to clarify that the POMDP overestimates measurement uncertainty, etclines 348-351, etc).

> In several places the authors use loose language by referring to marine ecosystem dynamics when in fact they are treating population dynamics.

Thank you, fixed. (TBD)

> The authors discuss volatility of different solutions and would be served well to quantify it. E.g., lines 365-369

Thank you, fixed (TBD)

### 2017-09-29 manuscripts/manuscript.Rmd
WIP

---
title: "The importance of measurement uncertainty in ecological management"
author:
  - name: "Milad Memarzadeh"
    affiliation: a
  - name: "Carl Boettiger"
    affiliation: a, 1
address:
  - code: a
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"

abstract: |
  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.  Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to the intuition of the precautionary principle. Through the use of cutting-edge point-based POMDP methods, we demonstrate that the standard management theory which ignores measurement uncertainty may lead to over-exploitation of fisheries and increased probability of economic and ecological collapse.  In contrast, POMDP solutions which account for measurement error are overall more conservative, avoiding over-exploitation while generating higher economic value.

keywords:
  - Measurement uncertainty
  - POMDP
  - Decision theory
  - Fisheries
  - Conservation

lineno: true

bibliography: citations.bib
output: rticles::elsevier_article
---

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on forecasting, i.e. the uncertainty of the future [@Petchy2015], management decisions must also contend with uncertainty of the present: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in *decisions* -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.  In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.

Approaches for decision making under uncertainty in ecological systems can be divided into two camps: approaches based in "optimal control" and usually favored by natural resource economists, and approaches based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].  While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknolwedged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.  Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigour of optimization to resolve a long-standing paradox and suggest a more robust approach to management.

The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters], ecosystem-based management [@Levin2008] to resilience thinking [@Holling; @May] and back to the birth of resource economics [@Beverton; @Scheffer], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.

## Methods

We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.  To facilitate tractability and intepretation, we will consider a simple and well-studied model of population growth (stock recruitment) in marine fisheries, the Graham-Schaefer model:

Though this and very similar models are widely used in large scale analyses across diverse stocks (e.g. [@Costello2016; @Britten2017]), and common in other conservation contexts where richer models are unavailable, neither the existing approaches nor the one we introduce here are limited to such one-dimensional stochastic models with known parameters. In practice, this model may include age or stage structure and will reflect uncertainty not only in the stochastic recruitment but in the parameters themselves, usually as the result of a heirarchical model estimation process.   All that is required of the model is that it defines the probability of transitioning to a new state given the current state (here, simply the stock size) and the action taken (here, just the harvest).  Uncertainty either in the parameters, in the model, or inherent in the stochastic process can all be reflected in such probabilities.

Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

$$y_t = Z_m x_t$$
for some measurement error $Z_m$.  Given the estimated recruitment model along with this estimate of the stock size, the manager must set a harvest quota $h_t$ for that year.  The methods we consider here all differ primarily in their handling of these uncertainties. If this system were completely deterministic so that there was no uncertainty, all would produce identical long term policies and thus achieve identical outcomes.

In the case of no uncertainty, the optimal harvest strategy is to reduce the population to the size at which it obtains the maximum growth rate, known as Biomass at Maximum Sustainable Yield, $B_{MSY}$ (this is adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @ClarkTextbook). For the Graham-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$, with a corresponding annual harvest of $H_{MSY} = r K / 4$.

This equilibrium analysis says nothig about the appropriate harvest at stock sizes above or below $B_{MSY}$, which highlights some essential differences between the optimal control policy and more typical fisheries management.  The optimal policy is impatient: it seeks to achieve the optimal yield as quickly as possible.  Below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, a harvest greater than $H_{MSY}$ can be optimal for that year.  This strategy is known as "constant escapement," since a constant sock size $B_MSY$ escapes harvest each year.  In practice, it is more common to pursue a policy of constant fishing effort (fishing mortality), $F_MSY = H_MSY / B_MSY$.  Continuing to harvest at $F_{MSY}$ when a stock is above or below $B_{MSY}$ is economically inefficient and causes depleted stocks to rebuild more slowly, but, in the absense of uncertainty, ensures a return in the long run to stock sizes of $B_MSY$ and harvests of $H_{MSY}$.  These differences will be important in understanding how these approaches perform once uncertainty is introduced.

## Pardoxes of uncertainty

The literature of optimal harvests under uncertainty is one of paradoxes, which can only be resolved by a more comprehensive treatment of measurement uncertainty.

### Reed's Paradox: $S = D$

The first of these we shall refer to as Reed's Paradox, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.  This suprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.  This also provides the manager with a remarkably convenient mechanism for determining the optimal harvest policy: rather than rely on computationally intensive Stochastic Dynamic Programming to determine the optimal policy, a simple derivative will suffice, which is all that is needed to solve the maximum of the deterministic problem.

### Clark's Paradox: $M < D$

@Clark1987 was among the first attempts to resolve Reed's Paradox. @Clark1987 (quite correctly, as we will see), identified the crux of Reed's Paradox as the absence of measurement uncertainty:

> An important tacit assumption in Reed’s analysis, as in the other works referred
to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.

Unfortunately, @Clark1987 was unable to solve the resulting problem exactly, but had to adopt an almost equally troublesome assumption:

> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
of the problem increases markedly if this assumption is
relaxed.)

The resulting solutions only deepened the paradox, rather than recovering a more cautious policy in response to the uncertainty, @Clark1987 finds policies that are often *less* cautious:

> [Our] results appear to contradict the conventional wisdom of renewable resource management, under which high uncertainty would call for increased caution in the setting of quotas.

Using different but still implicitly approximate results nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observe with some concern:

> It may seem counter-intuitive that a measurement error causes lower expected escapements below the deterministic fishery closure threshold.

While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@.....].

## Uncertainty in practice

Fortunately, management practice has so far ignored this paradoxical suggestion that uncertainty should be met with greater rather than reduced fishing quotas.  Instead, most well-managaged fisheries operate on rules that are more the result of common sense than mathematical optimization.  While details differ, the basic approach can be thought of as starting with the deterministic notion of $B_{MSY}$ and corresponding optimal fishing level $F_{MSY}$ as a way of determining an Overfishing Limit, (OFL), which is then adjusted to some lower level, Allowable Biological Catch (ABC) to reflect uncertainty in the stock assessment.  ABC is then adjusted lower once again, to determine a Total Allowable Catch, (TAC), reflecting uncertainty in reported catch or other uncertainty.

------

A fundamental concept to the management of any natural population is the notion of Maximum Sustainable Yield, MSY. The notion of MSY is rooted in the determinstic models dating back to foundational work in resource economics and ecological management half a century ago [@Berverton; @Scheffer].  Though the concept has been the focus of significant criticism for some time [e.g. @Botsford_nature_piece] that ctitique has been largely on the "Maximum" part, while the existence of a "Sustainable Yield."

## Reed's Paradox: $S = D$

In 1979, a fundamental paper by William Reed [@Reed1979] proved that, under a very general assumptions, the harvest policy for a stochastically fluctuating population was identical to that of the equivalent, deterministic population.  More precisely, Reed shows that the optimal harvest of the stochastically flutuation population should always aim to reduce the population back to a constant size $S$ (termed by Reed as "escapement," while in most fisheries literature uses the term $B_{\text{MSY}$, the biomass at Maximum Sustainable Yield}).  Moreover, this value, $S$, is no more conservative than had we been able to predict future populations with perfect accuracy, as in the determinstic case of constant escapement, $D$.  A notable exception to this result is when the population is not "self-sustaining," that is,
the stochasticity is severe enough to pose an extinction risk to the unhavested population.  In such cases Reed proves there is still a single constant target population size $S$, but that $S > D$, though often by very small amount, reflecting the small probability of a purely chance extinction.

An important corrolary of this result is that suggested a way to side-step the laborious dynamic programming required to find an optimal policy for stochastic populations: the deterministic solution is found trivially from maximum of the stock-recruitment curve. Thus much subsequent attention has focused on the estimating stock-recruitment, including methods that address considerable uncertainty in the data used for this estimate [@], while largely ignoring the subsequent step of how to determine an optimal policy in light of that uncertainty.

The lack of any intuitively precautionary response to the introduction of uncertainty did not go unnoticed: in 1986 Clark and Kirkwood first pointed to the absence of measurement error in the formulation used by Reed and others as fundamental to understanding this paradox.

> An important tacit assumption in Reed’s analysis, as in the other works referred
to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.

Under such an assumption, uncertainty introduced by stochastic growth is effectively purged by each subsequent observation rather than accumulating, and stock extinction can only occur if the manager explicitly decides to harvest the entire stock.

## Clark's Paradox: M < D

Clark and Kirkwood set out to address this problem by considering the case in which stochastic population growth occurs after population size is measured but before harvest, such that manager must act without perfect knowledge of the current system state.  Unfortunately, this problem proved easier to recognize than to solve, as Clark & Kirkwood recognized at the outset:

> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
of the problem increases markedly if this assumption is
relaxed.)

Rather than resolve Reed's Paradox, this assumption compounds the problem, leading to optimal harvests that are generally even higher (i.e. less cautious) than the analgous deterministic policy.  The same counter-intuitive result can also be found in subsequent work [@Sethi2005] which makes a different but equally limiting simplifying assumption (supplement).  (Both @Clark1986 and @Sethi2005 remark on the counter-intuitive nature of their results)

<!--

Clark: 
> The above results appear to contradict the conventional wisdom of renewable resource management, under which high uncertainty would call for increased
caution in the setting of quotas.

Sethi: 
> It may seem counter-intuitive that a measurement error causes lower expected escapements below the deterministic fishery closure threshold.

-->

## Precautionary approaches

Perhaps recognizing

## A solution

Here, we explore how uncertainty over the present impacts decisions

We shall focus on examples from fisheries, which have long been both incubator and proving ground for most of our more general theories of ecological management, from the and

Data are readily avialable and a compartively strong connection between scientific process of stock assessment and the implementation of policy is already established.  It is also an area which, as we will see, though many models already account for uncertainty, decision strategy still reflects fundamentally more deterministic ideas.

Conclusions:

- failing to account for uncertainty properly can lead to erroneous conclusions
- Prior beliefs (reflecting prior observations) matter.

<!-- FIXME: In practice, MSY and Reed are very different -->

<!-- Big picture: fisheries is a foundation and testing ground for conservation
and management. Yet at it's heart lies a controdiction.  -->

<!-- MSY and Constant Escapement -->

Yet fisheries are not managed at their maximum ecological or economic potential.  
Taken as a whole, most ocean fisheries yield a net negative economic return
[@FAO2009], while some 68% of global fisheries are not in good biological condition [@Costello2016].
Nevertheless, fisheries management in both theory and practice are

In a deterministic world at equilibrium (without economic discounting), 
the policies of constant escapement and maximum
sustainable yield look identical: the stock is maintained at a fixed stock
size or biomass, called $B_{MSY}$ in the MSY/fisheries jargon, which is 
equivalent to the escapement $S^*$ in the optimal control jargon and equal
to half the carrying capacity, or $K/2$, in the familiar Graham-Schaefer 
(logistic growth followed by harvest) model. Likewise, the harvest is identical,
occuring after recruitment and bringing the population back to this optimal
level; an annual harvest of $H_{MSY} = rK/4$ in Graham-Schafer model for either policy.
This similarity disguises significant differences in how the policies behave
when the biomass is either above or below $B_{MSY}$.

<!-- Ways in which constant escapement deviates from MSY -->
The economically optimal amount to harvest when the population is below 
$B_{MSY}$ is zero -- no harvest.  This permits the population to return
as rapidly as possible to the stock size at which it generates the maximum
yield, and thus the maximum economic return.  This result holds even when
recruitment is stochastic and thus future stock sizes are uncertain, 
as rigorously proven by @Reed1979.  Perhaps surprisingly, this is not what
fisheries management thinks of as a maximum sustainable yield (MSY) policy,
and does not reflect how fisheries are managed today. Rather, an MSY policy
is intepreted as fishing at the *effort*, $F_{MSY} = H_{MSY} / B_{MSY}$ that would generate the maximum
sustainable yield if the population was at the optimal biomass, $B_{MSY}$,
*regardless* of the current stock size. It is important to realize this policy
is framed in terms of *effort* $F$: for example, the length of the year open to fishing,
or equivalently, the per-capita mortality rate of fish, and not in terms of harvest, $H$.
A population below $B_{MSY}$ that remains harvested at a constant $H_{MSY}$ will eventually be
driven to extinction, but harvested at a constant $H = F_{MSY} \cdot B$ for $B < B_{MSY}$,
the stock will eventually recover to $B_{MSY}$.  Consequently, this simple policy should
converge to the optimal policy in the long run, assuming there is no uncertainty in
recruitment, measurement, or implementation.

The economically optimal policy of constant-escapement also differs from MSY
when stocks are above $B_{MSY}$.  As before, the MSY policy continues to assume
a fixed mortality rate, $F_{MSY}$, which should eventually reduce the stock to
the level of $B_{MSY}$.  Also as before, the optimal policy is not so patient --
to derive maximum yield it seeks to return the population to the $B_{MSY}$ not
*eventually*, but as quickly as possible, and thus will recommend a harvest larger
than $H_{MSY}$ to accomplish this. These differences become more salient in the case
of a population under stochastic growth, since the resulting variation leads
to a population that frequently above or below $B_{MSY}$.  The greater responsiveness
of the constant-escapement policy to the current stock size thus necessarily 
results in both the faster recovery of overexploited stocks and higher economic
returns.

The variation in the harvest required to achieve this constant escapement policy
can seem impractical from a top-down management perspective.  However, it is 
important to realize that the application and implications of this policy are
not purely academic. Economists have for some time observed these inherent 
inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
Approaches such as individual transferable quotas (ITQs) would incentivise
fisheries behavior to converge towards the more dynamic harvests of a constant
escapement policy, as doing so would increase the value of an individual's share.
It is precisely this reasoning which explains the broad patterns of recovery
observed in the landmark paper @Costello2016, which compares future of global
fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
approach should achieve the economically optimal policy of constant-escapement,
which alone explains why RBFM should see both faster recovery and higher 
economic yields than MSY, as we have just described. The substantial detail their
analysis introduces with respect to the dynamics of over 4000 individual stocks
as well as additional economic knock-on effects from managing based on quotas
rather than fishing effort permit them to offer quantitatively precise 
estimates regarding the time required and economic value generated by such
recovery, but are not essential to understand the broad pattern that suggests
that nearly all stocks should recover under either an RBFM or MSY-style 
management, but that the recovery would be fastest and the economic benefits
greatest under the former.  That conclusion was already implicit in how
these models are defined.  Unfortunately,

-------

Measuring and understanding collapse and over-exploitation of fisheries
world-wide remains an essential issue for the ecological sciences
[@Costello2016].  Effective management practices in marine ecosystems also
inform conservation and resource management in less well-studied
systems.

<!-- Set up broad relevance of understanding this issue -->

Central to both fisheries research and
policy, including international agreements such as the United Nations
Law of the Sea [@UNCLOS1982], is the principle of  maximal sustainable yield (MSY).
MSY aims to maintain a population at a fixed population size which should result in
the greatest long-term harvest (often referred to as Biomass at MSY,
$B_{MSY}$). While the theory dates back to classic work
of 1950s such as Beverton [@Beverton1957] and Schaefer [@Schaefer1957]
in deterministic models, it is Reed 1979 paper [@Reed1979] which proves
that the intuitive result of the deterministic models remains the optimal
policy for maximizing expected value of a fishery subject to stochastic
growth (i.e. environmental noise). Reed's result has allowed management
to largely sidestep the complexity of optimal control of stochastic
systems.

This has long begged the question: if stochastic growth can be
safely ignored in determining the optimal harvest, is the same true of
measurement error?  Thanks to recent computational advances for solving
the decision theory problem in the case of imperfect observations
(measurement error), we show that -- contrary to both the intuition
from Reed's result and previous, approximate solutions -- ignoring
measurement error can result in significant over-harvesting, decreased
economic returns, and the potential for complete collapse of the stock.

Some historical perspective is necessary to put this result in the
appropriate context with respect to both management practice as well
as previous research in optimal control of fisheries and other natural
systems under uncertainty.  When population growth
varies stochastically, any fixed annual harvest such as MSY is no longer optimal:
instead, management must respond by dynamically harvesting more in years
of high growth and less in years of low growth.  Reed's proof [@Reed1979]
nevertheless demonstrated that the optimal harvest can still be captured
by a simple rule: harvesting just enough to reduce the population back
down to $B_{MSY}$.  This result is known as a "constant-escapement" policy 
because the population escaping harvest is always the same, 
even while the harvest itself varies dynamically with each stochastic shock. 
Moreover, Reed shows that under some relatively generic assumptions, this escapement
is identical to that predicted by the deterministic theory  ($B_{MSY}$).
Because the optimal harvest policy in the deterministic theory follows directly
from the growth rate (just find the stock size for which recruitment function
is maximized, no need for dynamic programming), attention shifted to inferring
that state equation from data.  Our focus is not on inferring this state
equation, but on the subsequent step of: having done so, how does one determine
the optimal harvest policy while also accounting for measurement uncertainty? 
Previous attempts to address this issue over the past three decades
have relied on various approximations which have usually led to the paradox
that harvests should increase under additional measurement uncertainty, contrary 
to our intuition that we should respond to greater uncertainty with greater caution.
Using more recent optimal control algorithms first developed in the artificial
intelligence literature, we provide a solution which avoids such assumptions
and resolves this paradox in favor of common sense intuition.

Despite it's generality, Reed's proof makes a subtle assumption which
reduces the the impact uncertainty and underlies his conclusion
that the deterministic intuition of an average MSY is still optimal in 
the stochastic case.
In Reed's construction, the manager measures the stock
size without error during each time step before setting the harvest policy.
An important consequence of this construction is that this model can
thus never lead to accidental extinction of the stock.
Researchers have long recognized the weakness of this assumption of
perfect knowledge of the stock size and have sought to understand how it
may impact the resulting management policy [@Ludwig1981].  Because there is no risk of
accidental extinction in this scenario, it is possible that the optimal
policy may be more aggressive than is warranted.  Unfortunately, due
to the complexity of the analysis required to account for measurement
uncertainty, this question has never been satisfactorily answered.
Moreover, previous attempts at an approximate solution have consistently
found the surprising conclusion that the Reed solution is usually less
aggressive (dictating higher escapement or lower catch quotas) than
attempts to account for measurement uncertainty.  We briefly review the
main results in the ecological and resource economics literature that
have attempted to address this through various approximations before
presenting our solution, which draws on substantial algorithmic advances
that have been made in other fields.

The important way in which perfect observations diminish the role of
uncertainty in the model of Reed [@Reed1979] was first recognized in the classic
work of Clark & Kirkwood [@Clark1986].  They acknowledge the importance of measurement
uncertainty to the optimal solution, while observing that if
this assumption was relaxed, the "difficulty of the problem increases
markedly." Instead, @Clark1986 consider a perfectly observed model in
which harvest takes place after stochastic recruitment but before the
stock is measured, making accidental extinction a possibility
should realized growth fall sufficiently short of expectation.
In this case, @Clark1986 demonstrate that the optimal solution deviates
from the constant escapement rule, with the resulting optimal
policy is often significantly less cautious than the constant escapement policy.
In certain cases, @Clark1986 demonstrate that harvesting to extinction
is actually optimal. Other recent work [@Holden2015; @Johnson2016] has taken 
a similar approach to @Clark1986 in which measurement becomes uncertain due to
stochastic stock dynamics occurring between measurement and harvest, which
can also suggest less conservative (higher harvest) escapement rates.

Rougharden [@Roughgarden1996] take a complementary approach, introducing measurement
uncertainty but searching only for the optimal constant escapement,
despite presenting no evidence (and indeed contrary to @Clark1986)
that the optimal solution under this circumstance should be a
constant escapement policy. By applying the rule-of-thumb assumption
that the optimal policy can be expressed simply as a constant escapement
(or equivalently, target-stock), the set of possible policies is reduced
dramatically and the calculation is relatively trivial.  As expected,
this restriction results in a policy that is always more conservative (lower harvest) than
the Reed solution which ignores measurement error, but both economically
inefficient and, as we shall see later, still less conservative than
the truly optimal policy for large stocks.   Other studies 
have likewise attempted to include uncertainties in model parameters and measurements
by making this same assumption that a constant escapement solution is optimal and 
thus avoiding solving the full sequential decision optimization problem [@Ludwig1981; @Moxnes2003; @Engen1997].

Sethi et al. [@Sethi2005] seek to address
the limitations of @Roughgarden1996 by searching for the true optimal
policy without the strong assumption of constant escapement.
Like @Clark1986 and others, @Sethi2005 finds support for the counter-intuitive result
of an optimal policy that is significantly less cautious when accounting
for measurement uncertainty than the corresponding constant escapement
solution which ignores it. 
Unfortunately, the approach taken by @Sethi2005 once again avoids solving
the full optimal decision-making problem by means of a simplification
which we might term the amnesic assumption: during each subsequent
decision the manager forgets any previous knowledge of the system.
This allows them to use the standard Stochastic Dynamic Programming
(SDP) solution method used by @Reed1979 and @Clark1986. This method
assumes that problem in question is a Markov Decision Process (MDP):
one in which given the state $x$ and action $a$, the probability of a
transition to state $x'$ is conditionally independent of all previous
states and actions (known as the Markov property).  This holds only if the state
is known without error -- once we introduce any amount of measurement
uncertainty about the state $x$, the process with respect to the observed
state is no longer Markovian but depends on all previous observations.

This problem can be solved by considering
the 'belief' state rather than the observed state, which restores
the Markov property at the cost of much greater problem complexity:
one must consider a (continuously-valued) probability
that the system is in each of the $|X|$ possible states of such a 
Partially Observed Markov Decision Process (POMDP). This is the
key feature of the POMDP approach: all previous observations are
collectively reflected in this belief distribution, and decisions
are based not only on the most recent observation (as in MDP), but
are also influenced by this prior belief.

The mechanics of 
updating these beliefs, while technical, have been well
illustrated elsewhere even in conservation applications
[e.g. @Lane1989; @Chades2008; @Haight2010; @Chades2011;
@Williams2011; @Springborn2013; @Fackler2014; @Fackler22014; @Baggio2016].
However, these previous studies have relied on relatively direct algorithms 
which are not amenable to the large number of states and actions required by the optimal
fisheries harvest problem.  By leveraging newer point-based algorithms developed in the artificial 
intelligence community [@Kurniawati2008; @Shani2013; see SI; R code], we are able to solve
the POMDP models for fisheries management directly. 
These solutions illustrate how incorporating the information
of all previous observations into a prior belief allows the
POMDP approach to handle measurement error which would 
cause a stock collapse under MDP or deterministic methods.
We also evaluate the performance of proposed method
on a real-world time series of population estimates of
marine ecosystems from RAM Legacy Stock
Assessment Database [@Ricard2011].

## Model formulation

Our formulation of the fisheries management problem closely follows that of 
earlier work [e.g. @Sethi2005], which we summarize here. Alternate formulations are provided in the SI,
illustrating that our conclusions are not constrained to these particular choices.

The fisheries manager seeks to maximize the expected net present value of the fishery
over an infinite time horizon under two sources of uncertainty: population dynamics (i.e. growth of the stock),
and stock measurements, given an model estimate of those dynamics:

\begin{equation}\label{optim}
\max_{\{h_t\} \geq 0} \mathbf{E} \left[ \sum_0^{\infty} \gamma^t U(x_t, h_t) \right]
\end{equation}

The state of the system, $x_t$ corresponds to the (true) fish stock at time $t$. 
The harvest is $h_t = \min \left(x_t,a_t\right)$, where $a_t$ is the attempted harvest at time step $t$.
For simplicity, we assume the reward (utility) is directly proportional to the
harvest, $U(x_t, h_t) = \min(h_t, x_t)$, with future rewards discounted by multiplicative discount factor $\gamma = 0.95$;
alternate harvest functions and alternate discount rate are considered in the SI.

The population dynamics is formulated as:

\begin{equation}
x_t = \epsilon_{g} f(s_{t-1})
\end{equation}

where $s_{t-1} = x_{t-1} - h_{t-1}$ is the escapement at time $t-1$, and
$\epsilon_{g}$ corresponds to stochastic component of recruitment (growth noise).

While it is theoretically possible to learn about this model and its parameters during the
decision process itself (so-called adaptive management, e.g. @Johnson2015 for MDP case; 
we discuss the potential for POMDP adaptive management in *Future Directions*), it is
typical to consider that state-space dynamics $f$ have been previously estimated from data.  
The goal of a decision process is to determine the optimal sequence of actions given this
model. From the population dynamics state
equation we derive a set of transition matrices $T(x_t, a_t, x_{t+1})$ describing the
probability of a transition from $x_t$ to $x_{t+1}$ under action $a_t$ (SI Methods).
While such transition matrices can be defined even when $f$ represents only a posterior
distribution of model parameter values [e.g. @Boettiger2015] or an age-structured 
model [e.g. @Holden2015], for simplicity of presentation we will focus on the case
in which $f$ is a fixed model with known parameters, given by the Ricker equation [@Ricker1954]:

\begin{equation}\label{ricker}
f(s_t) = s_t \exp\left(r \left(1-\frac{s_t}{K}\right)\right),
\end{equation}

where $K$ denotes the carrying capacity of the system, and $r$ the growth rate.
The optimal management of this model in the deterministic and growth-noise-only (MDP)
case is already well understood and thus will provide a useful foil against
which to examine the role of measurement uncertainty.  
To demonstrate that our results are not unique to this choice of \eqref{ricker}, two alternate models,
the Beverton-Holt model [@Beverton1957], and the Allen model [@Allen2005], are evaluated in the SI.

The POMDP problem is distinguished by the introduction of measurement error, in which 
the observed stock $z_t$ is measured with uncertainty, $z_t = \epsilon_{m} x_t$. 
In general this is defined by observation matrices $O(z_t, x_t, a_t)$ giving the 
probability of observing $z_t$ given a system in state $x_t$ and having taken action $a_t$,
though in our formulation this probability is independent of the action.

We will assume for simplicity that the stochastic growth noise $\epsilon_g$ and measurement error $\epsilon_m$ 
are distributed log-normally with mean $1$ and scale parameter $\sigma_{g}$ 
and $\sigma_m$ respectively. Larger $\sigma$ corresponds to higher variance,
scaled to be comparable to the uniform noise distribution shown in the SI,
(see *Materials and Methods*).  We evaluate the effect of
different noise levels in both growth and measurements, $\sigma_{g},
\sigma_{m} \in \{0, 0.1, 0.5\}$.

As the numerical solutions of the MDP and POMDP problems require discrete states,
we consider a system of 50 states on a uniform grid of 0 to 1 with $K = 0.66$ and
$r = 1$; (alternative growth rates shown in the SI).   
The transition matrices $T$, together with the observation matrix $O$ and the utility
function, completely specify a POMDP problem which can be solved with a numerical
algorithm (*Materials and Methods*).

# Results {#results .unnumbered}

To illustrate the impact of measurement error on management outcomes,
we compare two scenarios: with measurement error ($\sigma_m = 0.5$,
see *Materials and Methods*) and without measurement error (Fig \ref{sim}).
In each scenario, we compute the optimal policy using three different
solution methods: POMDP, MDP, and a policy we refer to as Offline POMDP.
In POMDP, the optimal policy at each time point depends not only on the
most recent observation, but on the manager's prior belief of the state,
which depends on all prior observations. As such, the optimal POMDP
policy for a particular observation cannot be computed ahead of time
but must be calculated "online", as observations come in.  In contrast,
in MDP, the optimal policy depends only on the most recent observation.
The "Offline" policy uses the POMDP solution given a fixed, uniform
prior belief, rather than updating the prior belief over states after
each observation.  For each scenario (with and without measurement error), 
we generate 100 simulations of  the fishery being managed under each 
strategy (MDP, Offline, POMDP) for 100 time steps.  We compute the average and standard
deviation of the stock size over these 100 replicates, Fig. \ref{sim},
as well as the mean economic return of each case (Table \ref{Table1})
and the fraction of simulated stocks that have collapsed by the end
of the simulation (Table \ref{Table2}). While we perform these
simulations over a range of models, parameter values and noise distributions, for
simplicity of presentation we focus on simulations
from a Ricker model with log-normally distributed noise (see *Materials and Methods*);
comparable results under the alternative choices are presented in the SI.

\begin{figure}[h]
\centering
\caption{Forward simulations of managing a marine ecosystem under POMDP and MDP models in the presence ( $\sigma_m = 0.5$, top panel) or absence ( $\sigma_m = 0$, bottom panel) of measurement error.  Lines show the mean stock size over 100 replicate simulations, error bands show +/- 1 standard deviation.  Many of the replicate simulations under MDP quickly collapse the stock by over-harvesting when measurement error is present.  In contrast, the POMDP management is robust to the level of measurement error, avoiding stock collapse whether or not measurement error is actually present.  The growth model is Ricker, the noise is log-normally distributed, the carrying capacity of the system is $K = 0.66$, the maximum per capita growth rate is $r = 1$, growth noise is $\sigma_g = 0.5$, and the discount factor is $\gamma = 0.95$.}\label{sim}
\end{figure}

Fig. \ref{sim}, top panel, illustrates that ignoring or underestimating
the measurement error result in dramatic collapse of the stock and
significant loss in revenue: the mean MDP state declines as subsequent
replicate simulations collapse (94/100 collapse within 100 time steps)
when measurement error leads to over-harvesting.  The POMDP method
retains a steady, sustainable harvest that keeps the population below
carrying capacity but results in minimal population collapse (4/100). The
offline solution results in some stock collapse (34/100) as a result of
not updating the manager's belief state over subsequent observations.
Unsurprisingly, average economic returns are also much lower for MDP
when measurement error is present, returning an average of 66% of the net present value returned by
POMDP (Table \ref{Table1}), and would be lower still with less aggressive discounting of future
rewards (Here $\gamma = 0.95$ makes later collapses less relevant to the total).
Fig. S1 shows stock trajectories from several individual replicates, which 
show relatively little decline before collapsing suddenly when managed under MDP
when measurement error is present.  Averaging over an ensemble of such trajectories
gives the appearance of a smooth decline.

On the other hand, overestimating measurement error might result in a
trivial loss in revenue but does not result in collapse of the stock,
(Fig. \ref{sim}, lower panel). In this scenario, since no
measurement error is present, the MDP solution is indeed optimal, but the
POMDP policy which assumes a measurement error of $\sigma_m = 0.5$ and
thus overestimates the measurement uncertainty performs nearly as well
(average of 94% of the MDP net present value and likewise no collapses of the stock, Table \ref{Table2}).
In the SI we report the same patterns for Ricker model
with uniform noise (Fig. S2), Beverton-Holt model with log-normal noise
(Fig. S3), and Allen model with log-normal noise (Fig. S4).

<!-- Perhaps in-text quotations of these numbers is sufficient?-->

\begin{table}
\centering
\caption{Ecological and economic performance when
measurement error is present, (\(\sigma_m = 0.5\))}
\begin{tabular}{lccc}
& MDP & Offline & POMDP \\
\midrule
Fraction collapsed & 94\% & 34\% & 4\% \\
Economic return & 66\% & 99\% & 100\% \\
\bottomrule
\end{tabular}
\label{Table1}
\end{table}

\begin{table}
\centering
\caption{Ecological and economic performance when
measurement are perfect, (\(\sigma_m = 0\))}
\begin{tabular}{lccc}
& MDP & Offline & POMDP \\
\midrule
Fraction collapsed & 0\% & 0\% & 0\% \\
Economic return & 100\% & 96\% & 94\% \\
\bottomrule
\end{tabular}
\label{Table2}
\end{table}

To facilitate comparison with previous research and management practice,
we plot the corresponding optimal action in terms of *expected escapement*: the
stock the manager expects will be left in the sea after harvest $z_t - h_t$.
Recall that introducing measurement error means that this value can differ
from realized escapement, $x_t - h_t$. Fig. \ref{policy} shows the optimal expected 
escapement as a function of observed stock sizes over a range of growth 
noise and observation error intensities, assuming a uniform prior belief over
states $x_t \in (0,1)$ for $\sigma_m > 0$. These policies are determined numerically
using the SARSOP algorithm for POMDP [ @Kurniawati2008], as discussed in *Materials and Methods* and
illustrated in the supplementary code provided.

In the MDP case ($\sigma_m = 0$),
we see the classical result of constant escapement [@Reed1979], where the stock is harvested
back down to the same escapement regardless of observed stock or growth noise.
When growth noise is large ($\sigma_g = 0.5$) relative to measurement
error ($\sigma_m = 0.1$), the policy converges towards the the constant
optimal escapement solution of @Reed1979, as expected. These patterns
hold independent of the population growth model (e.g. Beverton-Holt and
Allen) and the shape of the noise distribution (Fig. S5--S6).
The most obvious difference between the MDP and POMDP policies arises in the region of
linear or nearly linear increases in expected escapement occurring at very large observed stock sizes,
which arises from any policy whose harvest remains *constant* over that range. Not increasing
harvest in response to higher observed stock sizes always result in a higher expected escapement, $z_t - h_t$.  
This corresponds to a policy which believes such high observations are almost surely the 
result of measurement error, and suggests that overestimated stock sizes are responsible for MDP
over-harvesting and the subsequent collapse of the stock. At smaller stock sizes, the policies
shown in Fig. \ref{policy} appear more similar over any measurement error, but in fact this is only a consequence of the
uniform prior belief. While Fig. \ref{policy} completely specifies the management policy under MDP,
this is not true for the POMDP solutions ($\sigma_m > 0$), which also depend on all
prior observations through the influence of the belief state.

\begin{figure}[h]
\centering
\caption{Optimal escapement policy as a function of measured stock for Ricker model, with different measurement errors and growth noise. Noises are log-normally distributed, and the prior belief state is uniform among all stocks. The carrying capacity of the system is $K = 0.66$, the growth rate is $r = 1$ and the discount factor is $\gamma = 0.95$.}\label{policy}
\end{figure}

To illustrate this key difference, we plot (Fig. \ref{bel})
the POMDP policy under two alternative prior belief states: (1) Low
Stock: where the prior belief is centered at $x = 0.2$ , and (2) High Stock:
where the prior belief is centered at the location of carrying capacity $x =
0.66$. This figure is generated in an identical manner to Fig. \ref{policy} 
except for the different choice of prior belief.  Visualizations of these 
prior distributions are shown in Fig. S12. When the prior belief skews towards low stocks, the
probability that an observation of a high stock is an overestimate due to
measurement error is greater, and consequently the policy becomes more conservative
(higher escapement). This ability to incorporate the knowledge from prior
observations accounts for the difference
in performance between the POMDP solution and the Offline solution in
Fig. \ref{sim} when measurement error is present. 
The evolution of the belief state in response to subsequent observations
over a given simulation is shown for the different models in Figs S14 - S16.
Corresponding plots for alternate models are shown in Fig. S9.
The patterns shown here are further robust to the reward function (including
linear cost to harvest and a diminishing returns model, Fig. S8) and the
discount rate Fig. S13.

\begin{figure}[h]
\centering
\caption{Optimal escapement policy as a function of measured stock and different prior belief states: (1) Uniform: uniform prior belief over all stocks, (2) Low Stock: the prior belief is centered at $x = 0.2$, and (3) High Stock: the prior belief is centered at the location of carrying capacity $x =
0.66$. In this figure, growth model is Ricker, measurement and growth noises are $\sigma_m = \sigma_g = 0.5$, and are log-normally distributed, the growth rate is $r = 1$, and the discount factor is $\gamma = 0.95$.}\label{bel}
\end{figure}

## Application to real-world fisheries

Using historical time series of Argentine Hake on the Patagonian Shelf and
European Plaice on the Celtic-Biscay Shelf from the the RAM Legacy Stock
Assessment database [@Ricard2011], we compare the catch levels recommended
by POMDP and MDP methods against historical values (Fig. \ref{RAM2}). We have selected
these two examples for having both adequately long records and show very
different trends in both population biomass and catch. We non-dimensionalize
the data and then use Maximum
Likelihood Estimation (MLE) through NIMBLE [@nimble-software:2016] to
fit a Ricker model for both these examples (details of MLE fits can be
found in the SI). Based on this model estimate, we then determine MDP
and POMDP policies using the appropriate algorithms.  In the case of MDP,
the optimal harvest is determined directly from that year's estimated
stock abundance.  In the case of POMDP, the optimal harvest depends both
on this stock estimate and the prior belief, which is sequentially updated
based on each previous observation (stock estimate) and action (catch).

Both examples illustrate the contrast between the ignoring (MDP) and accounting for measurement error (POMDP).
The POMDP solution tends to recommend a lower harvest than the MDP solution
in both cases, but not strictly so.  In Plaice example, we see the the POMDP solution is less 
volatile, increasing harvest more modestly in response to observing stock 
sizes while fishing pressure remains constant or increases (e.g. 1980-1990), 
while also decreasing less if stocks decline under constant or decreasing fishing
pressure (e.g. 1990-2000). Under the MDP model, only sharp changes in catch
can drive these sharp shifts in stock levels, while the POMDP model permits that these
can arise in part by measurement error and thus require less volatility in the harvest. 
In the Hake time series, the continued decline of the stock leads both MDP
and POMDP solutions to recommend smaller harvests than historically observed.
Again the POMDP solution is less volatile than the MDP, starting with a more
modest harvest at the beginning of the time-series (where a uniform prior over
states is assumed), and at first decreasing more gradually but then retaining
the moratorium on fishing when a small rebound (circa 2005) leads the MDP to a
renewed harvest.

\begin{figure}[h]
\centering
\caption{This figure compares the actual stock biomass and catch of European Plaice on the Celtic-Biscay Shelf and Argentine Hake on the Patagonian Shelf with actions suggested by MDP and POMDP frameworks. In the case of POMDP we have used measurement error of $\sigma_m = 0.5$.}\label{RAM2}
\end{figure}

# Discussion {#discussion .unnumbered}

We have shown that fully accounting for measurement error in calculating
the optimal fishing harvest leads to management that is usually more conservative
(higher expected escapement) as the measurement error increases. This
contrasts with the classic solution of constant escapement, in which
larger estimates of the stock size justify proportionally larger harvests.
The importance of our intuitively sensible conclusion is underscored
by the contrast to both much of current fisheries management practice
and previous research into measurement uncertainty.  Most fishing policy
still reflects the results @Reed1979 driven in
the absence of measurement error, and even much previous work on the role
of measurement uncertainty  has either assumed constant escapement-type
solutions [@Roughgarden1996; @Moxnes2003; @Engen1997] or relied on approximations which
have argued that accounting for measurement uncertainty should result
in an *increased* harvest (lower target stock size) [@Ludwig1981; @Clark1986; @Sethi2005].

Using simulations (Fig. \ref{sim}, S2-S4), we have shown that failing to account for measurement error
can lead to dramatic ecological and economic consequences. MDP solutions, which
ignore measurement error, collapse 94% of simulated time series within 100 time
steps. This does not mean that measurement error in the simulations is so large
as to make informed management impossible -- in fact the optimal POMDP solution,
which accounts for measurement error, experiences a collapse in only 4% of
simulations shown here. While the precise degree of measurement uncertainty may
itself be unknown, these results also show that it is better to overestimate
than underestimate this uncertainty. Even when no measurement error is present,
a POMDP solution which assumes a large measurement error ($\sigma_m = 0.5$) performs
nearly as well as the optimal MDP solution in all of our simulations. Fig. \ref{policy}
and \ref{bel} provide the intuition behind why MDP approach results in such a high probability
of collapse: when measurement error results in a chance over-estimation of the 
stock, the resulting harvest under MDP can be much larger than under POMDP. A single
large or a few consecutive smaller over-estimations of the stock size can thus 
quickly drive the stock to zero.  In contrast, the POMDP solution will harvest
much less under the same observation error, particularly when it compares that
observation against the accumulated prior belief that the current state should
be lower, Fig \ref{bel}.

We have shown that this dramatic reduction in the probability of a 
collapse is not trivially the result of a more cautious strategy; e.g. one which always
harvests less than MDP.  Rather, it is the result of a making full use
of all of the data through the role of the belief state.
The POMDP solution, unlike the MDP solution, depends on
the entire history of previous observations, which inform the current
belief state.  As shown in Fig \ref{bel}, this can correspond
to higher harvests if prior observations give sufficient evidence that 
the stock is higher than suggested by the most recent measurement. As a result,
the POMDP policy can outperform the MDP policy economically as well as 
ecologically.

We have also demonstrated these patterns by comparing harvests recommended
by MDP methods against those recommended by POMDP methods using two separate
historical time series.  These examples provide a further illustration
of how accounting for measurement uncertainty results in very different policy
outcomes than ignoring it.  Note that this analysis does not seek to explain
the differences between either the MDP and POMDP solutions from the historical
catches actually observed.  While it is very possible that the observed harvests
were governed by some sort of constant-escapement policy, actual catch quotas
or other policy information is not available in this data. Fluctuations in historical
harvest may reflect many other factors than such a policy, such as different stock
models, external constraints such as the cost of adjusting policy [@Boettiger2016], 
or other economic, political or environmental factors.  While we are free to
make comparisons between MDP and POMDP solutions on the historical data, understanding
why historical harvests differed from either of these would need an understanding of 
these contextual issues that are beyond our data and our scope here.
Given the stock-recruitment model for the fishery, the MDP optimum harvest depends only
on the most recent stock estimate, while both the full time series of historical harvests
and historical stock estimates influence the POMDP optimum, by altering the belief
state at the time.  We have
relied on estimating reasonable candidate models and parameters from the time series
a priori (specifically MLE parameters to the Ricker model). Ideally model
and parameters would also be inferred over the course of the sequential decisions,
which would integrate over such uncertainty in models or parameters.
While MDP methods to address these issues of model uncertainty
and parameter uncertainty have been developed elsewhere [e.g. @Johnson2015] this remains
a subject for future research in POMDP applications (see *Future directions*).

We have shown that our result is generally robust to our assumptions:
this pattern holds for differing models (Ricker, Beverton-Holt, and Allen),
differing magnitudes and also different shapes of noise distributions
(log-normal, as considered by @Clark1986 or uniform, as considered by
@Sethi2005), the choice of the reward/profit equation, or the details of
the discretization (see SI).  This robustness is not surprising, being consistent
with sensitivity observed in previous work.  Indeed, in the absence of
measurement error most of these scenarios were already covered by Reed's
proof, (with the exception of the Allen model whose change in concavity
violates the theorem and likewise shows deviations here).
Thus it is perhaps all the more surprising that
constant escapement is not robust to the assumption of measurement error.

The optimization approach considered here can be distinguished from more heuristic
approaches to address the uncertainty, such as pretty good yield (PGY)[@Hilborn2010; @Rindorf2016] or management strategy evaluation (MSE).  Methods based on MSE have been applied in fishery management for providing the decision maker with a rationale to base their decisions given their specific optimization objectives [@Smith1994; @Mapstone2008; @Bunnefeld2011]. The main difference between MSE and methods based on dynamic programming is that MSE does not solve the optimization problem to provide the optimal strategy, rather it evaluates consequences of a set of strategies defined based on management objectives and provides the outcome of those simulations to the decision maker. In that sense, MSE can be compared to the policy evaluation step of optimization in dynamic programming using policy iteration methods [@Sutton1998], while the policy improvement step is being ignored due to computational complexity. With advancements in efficient solving of dynamic programming optimization under several sources of uncertainty [@Shani2013], the traditional MSE approach can be interpreted as an internal step in the optimization process that tries to identify the best management strategy, searching through all viable and possible strategies. MSE can be also compared to Monte Carlo search algorithms within the space of possible policies that the manager can explore and evaluating the policies based on the outcome [@Smith1999]. For example, @Smith1999 evaluates the effect of several sources of uncertainty such as structural and parameter uncertainty in the models, errors in data and observation systems, estimation uncertainty, and management implementation uncertainty in partnership with Australian Fisheries Management Authority.

Our results have important implications for management of fisheries.
The most evident of these is that given the unavoidable realities of measurement
error in stock estimates, the wide use of management based on constant-escapement
solutions and derivative concepts such as maximum sustainable yield (MSY), maximum
sustainable biomass ($B_{MSY}$),
which are central to scientific fisheries management and policy
included the United Nations Law of the Sea [@UNCLOS1982], are insufficient to
guarantee either maximum sustainable economic yield or persistence of fish stocks.
Instead, target harvest or target escapement must reflect current and previous 
stock estimates and the degree of measurement uncertainty present. While precise 
estimates of the uncertainty in measurements, $\sigma_m$ may be difficult to obtain,
we have shown that it better both ecologically (in the probability of stock collapse)
and economically (in the profits from fishing) to overestimate than underestimate this
quantity.

This important qualitative impact that measurement error has on the
optimal solution of an important and well-studied ecological management
problem has more general implications.  The role of measurement error
is most often acknowledged in the breach -- a mathematically
convenient assumption only.  After all, inherent stochasticity has often been found
to make only small differences from deterministic solutions, as @Reed1979
proved in this case. Our result demonstrates how frail such an assumption
regarding measurement uncertainty may be, particularly when determining
optimal policy.  No doubt the mathematical and computational difficulty of
accounting for measurement uncertainty has forestalled greater attention
to this issue.  The growing wave of available data makes such concerns
about measurement error all the more pressing, as imperfect but readily
available observations make such data-driven optimization possible in more
applications.

These results have both tactical and strategic implications for ecological 
modeling and fisheries management [sensu @May1973]. At a strategic level,
this highlights the danger in ignoring the potentially asymmetric impact of
measurement error, which can lead to significant over-harvesting and potential
collapse of the stock.  At a tactical level, the precise amount by which
harvest should be adjusted to account for a given measurement error to maximize
the expected value of a stock will require a POMDP solution for the specific 
fishery in question, which is now computationally feasible using the approach
we have presented here. We hope that the recent advances in efficient and accurate
POMDP solvers, together with greater attention to these challenges,
will encourage a wider consideration of the role of measurement error
in determining appropriately precautionary management policy.

In conclusion, we have demonstrated the importance of measurement error
in the long-standing question of optimal harvests in fisheries.  We have 
shown that overestimating the measurement error can still result in policies
that are nearly optimal both economically and ecologically, while ignoring or
underestimating the measurement noise can cause dramatic collapses of fish
stocks even when no tipping point is present.  This suggests that the failure
to account for measurement error in this way in even the most advanced 
fisheries management could contribute to long-term trends in declining
stocks [e.g. @Costello2016]. We believe the approach taken
here both highlights the importance of accounting for measurement error and
provides an example of how this can be accomplished in the POMDP framework
using modern algorithms on complex problems. We have also implemented the
proposed method on a real-world problem of managing a marine ecosystem and show
that the policy suggested by POMDP is significantly different with respect to the
actual catch landings.

## Future directions

Several limitations that have been studied in MDP models, such as parameter 
and model uncertainty and adaptive management [e.g. @Johnson2015] remain
intriguing future challenges for POMDP problems. Future work could extend the examples
shown here to include more realistic models, such as those with age structure, as
has only recently been done for the MDP case [@Holden2015].
There remains a need for further improvement of the proposed method to include such model uncertainty. 
Previous attempts that adapt POMDP tools to include model uncertainty [i.e. @Chades2012] have had to
assume full observability of state space, ignoring measurement uncertainty. 
A second limiting assumption common to MDPs and POMDPs is that of stationary dynamics: 
that the population dynamics equation itself is not changing over
time. In reality, forces such as climate change and other forms of environmental variations violate this
assumption. @Fackler22014 attempts to address this problem by 
adaptation of Mixed Observability MDP [@Ong2010] framework, although the proposed algorithm
for finding optimal solution to such non-stationary systems is intractable, except for 
extremely simplified examples used in that study. Our future work involve relaxing
these assumptions by proposing computationally feasible approximate algorithms within the 
POMDP framework to deal both with model uncertainty and non-stationarity of system's dynamics.
Lastly, the introduction of POMDP approaches raises a central question for managers: when is it
worthwhile to actively reduce measurement error to derive a better policy, and when is it
best simply to account for current measurement error?  This issue can be addressed through
a Value of Information framework, as @Johnson2015 illustrates in the case of MDPs.

# Materials and Methods {#methods .unnumbered}

A Partially Observed Markov Decision Process (POMDP) [@Smallwood1973;
@Sondik1978] is a generalization of a Markov Decision Process (MDP),
which is a fundamental model for sequential decision making.  In the
ecological literature, MDPs are often referred to by their solution
method: Stochastic Dynamic Programming [SDP see @Marescot2013].
The MDP approach has been widely used in behavioral ecology
[@Mangel1988] as well as conservation ecology and resource economics
[e.g. @Mangel1985]. Introductions to MDP and solving them through dynamic
programming are provided in @Mangel1985, @Bertsekas1996 and @Sutton1998.
Specifically, POMDPs overcome one of the main limitations of MDP which
is the assumption about the full observability of the system's state.
In POMDP, the exact state of the system cannot be
observed directly, but can be inferred by indirect measurements.
POMDPs can be represented as a probabilistic graphical models
[@Barber2012] as in Fig. S11.  For further details regard POMDP framework,
its parameters, and details of solving the optimization problem,
refer to the SI.

Our approach utilizes the SARSOP algorithm of @Kurniawati2008. Computational 
requirements are still significantly more demanding than the equivalent MDP solution.
Precise computational time and memory requirements depend not only on system size but on
the precision requested of the approximate, point-based solution. Reasonable
solutions for the problems considered here can be reached within an hour of computation,
though were run for five hours each to confirm convergence. 
As a supplement to this paper we provide example scripts for replicating
the analyses shown here, which can be found at https://github.com/boettiger-lab/sarsop.

### 2017-09-29 manuscripts/manuscript.Rmd
WIP

@@
 
-@Clark1987 was among the first attempts to resolve Reed's Paradox. @Clark1987 (quite correctly, as we will see), identified the crux of Reed's Paradox as the absence of measurement uncertainty:
+@Clark1986 was among the first attempts to resolve Reed's Paradox. @Clark1986 (quite correctly, as we will see), identified the crux of Reed's Paradox as the absence of measurement uncertainty:
 
@@
 
-Unfortunately, @Clark1987 was unable to solve the resulting problem exactly, but had to adopt an almost equally troublesome assumption:
+Unfortunately, @Clark1986 was unable to solve the resulting problem exactly, but had to adopt an almost equally troublesome assumption:
 
@@
 ABC is then adjusted lower once again, to determine a Total Allowable Catch, (TAC), reflecting uncertainty in reported catch or other uncertainty.
+
+## POMDPs: An optimal treatment of measurement uncertainty
+
+As discussed above, absent the short-cut provided by Reed's theorems, solving for the optimal management policy in a stochastically fluctuating population involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP).
+Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).
+As in the case of MDP, such problems can be solved by existing and well-studied algorithms.
+While not unknown to the conservation literature [@Williams2009], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
+@Chades, @Fackler, ....].
+POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurtwani] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+<!--  -->
+
+We will not summarize all of the technical details here which are already well-developed in the existing POMDP literature, but rather we focus on providing the essential intuition behind decisions under a POMDP as compared to a perfectly observed system.
+For readers interested in exploring or applying these methods in more depth, we have developed a preliminary R package which can be used to replicate any of the results presented here [@Rsarsop].
+
+## Results
+
+Figure 1 shows the results of simulation
+
+-------------
 

### 2017-09-30 manuscripts/manuscript.Rmd
WIP

@@
 
-## Methods
+## Model
 
@@
 
-While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@.....].
+While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
+Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.
+First we will introduce the more heuristic alternative to addressing uncertainty, which more accurately reflects both intuition and actual management.
 
@@
 ABC is then adjusted lower once again, to determine a Total Allowable Catch, (TAC), reflecting uncertainty in reported catch or other uncertainty.
+To capture this process, we will consider a policy in which the harvest quota is set at a TAC of 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
+We will refer to both MSY policies and this more cautious adjustment of the MSY policy as "constant mortality" policies, since they are defined by a constant fishing mortality, $F$, rather than by a constant escapement, $S$.
+
+Note that compared to Reed's constant escapement, this policy still permits some level of harvest even when stocks are measured below $B_{MSY}$.
 
@@
 Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).
-As in the case of MDP, such problems can be solved by existing and well-studied algorithms.
+The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process.
+The POMDP problem for fisheries question considered here can be summarized as follows:
+
+- Transition function (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.
+In our context, this relationship is given by the Graham-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $\mathcal{N_T}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.
+
+- Observation function: $O(x_t,y_t,a_t)$ the probability of observing state $y$ given a system in state $x$.
+In principle, the action chosen can influence the precision of the observation.
+In our case, we simply assume normally distributed errors around the true state, $y \sim \mathcal{N_T}(x, \sigma_m)$, truncated at zero to exclude negative population sizes.
+
+- Utility function: $U(x_t,a_t)$, the utility recieved at time $t$ for taking action $a_t$, given that the system is in state $x_t$.
+For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.
+This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation.
+More realistic utility functions may include diminishing returns with increasing harvest (supply and demand effects), and the cost of fishing, both of which act to suppress large harvests.
+By focusing on a simple utility we can be sure that our comparison to MSY is driven by the treatment of uncertainty rather than merely differing economic assumptions.
+
+The optimization problem is to select the action $a_t$ that will maximize the net present utility over all time.
+Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.
+
+Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.
+As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms.
 While not unknown to the conservation literature [@Williams2009], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
 @Chades, @Fackler, ....].
-POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurtwani] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
-<!--  -->
-
-We will not summarize all of the technical details here which are already well-developed in the existing POMDP literature, but rather we focus on providing the essential intuition behind decisions under a POMDP as compared to a perfectly observed system.
-For readers interested in exploring or applying these methods in more depth, we have developed a preliminary R package which can be used to replicate any of the results presented here [@Rsarsop].
+POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
+For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop, which can be used to replicate any of the results presented here and explore other variations.
 
@@
 
-Figure 1 shows the results of simulation
+- Replicate simulations
+
+- non-pomdp policies
+
+- pomdp policies
+
+- historical data.
+
+**A POMDP policy reflects all prior observations**.
+The essential difference in a POMDP policy compared to either that of constant escapement $S$ or constant mortality $F$ is that nothing is just a constant -- instead, the harvest quota recommended by POMDP will always depend not only on the most recent observation of the stock size, but on all previous stock sizes.
+This dependency can be expressed in terms of a *prior belief* -- if all prior observations have led one to expect that the stock should be relatively small, a single observation of a very large stock is met by some skepticism under a POMDP, and the resulting harvest recommended by POMDP can differ substantially from that recommended by constant escapement or constant mortaility, which by definition consider only the most recent observation.
+If the observation is largely consistent with prior expectations, a POMDP policy will be more similar to that of the constant escapement policy, though with a natural level of precaution built in to allow for the possibility of error in measurement coupled with a negative shock to growth.
 
@@
 
-------
+## Discussion
+
+### A practical implementation?
+
+The year-to-year variation in quotas necessary under either a constant escapement policy
+or a policy in which quotas are set by POMDP can seem impractical from a top-down
+management perspective, particularly when compared to the simplicity of fixing a mortality of fishing $F_{MSY}$.
+However, the economist's perspective reveals that this may not be as unrealistic as it may first appear.
+
+Fisheries today are not managed at either their maximum ecological or economic potential.  
+Taken as a whole, most ocean fisheries yield a net negative economic return
+[@FAO2009], with fishing persisting only with the support of government subsidies [@Arnason2012], 
+meanwhile some 68% of global fisheries are not in good biological condition [@Costello2016].
+While the causes for this are both complex and numerous, we have seen that both the economic
+inefficiency and ecological health are hurt by current policies, including
+both MSY and it's more precautionary variants, as these all permit some level of harvest
+below $B_{MSY}$.  Economists have for some time observed these inherent 
+inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
+Approaches such as individual transferable quotas (ITQs) would incentivise
+fisheries behavior to converge towards the more dynamic harvests of a constant
+escapement policy, as doing so would increase the value of an individual's share.
+@Arnason2012 argues that properly implemented, this would lead fisheries to desire
+an even more dynamic adjustment of the TAC to maximize the value of their shares,
+rather than to oppose the  @Costello2016 refers to this as Rights Based Fisheries 
+Mangement, (RBFM), and it is precisely the willingness to shut down harvests for 
+over-exploited fisheries that explains the faster recovery they predict under RBFM
+relative to MSY. Though the @Costello2016 RBFM projections are based on a simple 
+constant escapement policy which ignores uncertainty, the logic is quite general:
+given property rights to establish an efficient market, it is not unrealistic to
+expect fisheries adopting such dynamic polices. Indeed, the greater adoption of 
+RBFM should spur the adoption of POMDP approaches as rights-holders seek strategies
+that maximize the share values.  This is not to dismiss the immense practical
+challenges in such an implementation, but to observe that such dynamic, better-performing
+policies can be seen to embrace rather than ignore economic realities.  Notably,
+the optimization approach assumed in POMDP can be easily adapted to reflect other
+constraints such as the cost of idle fleets if a fishery is closed, or include
+the value of ecosystem services from fish not harvested [@Halpern2013].  <!--Alternate utility functions do not increase
+the complexity of the solution.  Note that fixed cost of a fleet doesn't alter the answer at all:
+if you have the strategy that optimizes your profits when you have no expenses, you don't do a 
+different strategy that gets less profit unless it will also lower costs.  So whether or not
+you have to pay a fixed fee each year for boats, even if those boats are idle, is no reason
+not to leave those boats idle.  only if leaving them idle is *more expensive* than operating them 
+does this make sense, which seems unlikely.  -->
+
+<!-- True whether or not it is implemented in effort-based control or as a TAC-->
+
+<!--
+Moreover, regulating fishing effort rather than fishing catch, i.e. opening
+a fishery for only certain days of the year rather than limiting dockside catch, is both
+economically less efficient: for instance, it increasing costs due to the incentive to use more or larger fishing boats
+to maximize catch in that time, and it decreases profits since much of the year's catch
+is landed in a short window where it drives down prices, instead of being sold fresh throughout
+the year. Because regulating effort is both less direct and more difficult to enforce, the uncertainty
+in the expected harvest is also larger than in regulating catch.  
+-->
+
+It is precisely this reasoning which explains the broad patterns of recovery
+observed in the landmark paper @Costello2016, which compares future of global
+fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
+and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
+approach should achieve the economically optimal policy of constant-escapement,
+which alone explains why RBFM should see both faster recovery and higher 
+economic yields than MSY, as we have just described. The substantial detail their
+analysis introduces with respect to the dynamics of over 4000 individual stocks
+as well as additional economic knock-on effects from managing based on quotas
+rather than fishing effort permit them to offer quantitatively precise 
+estimates regarding the time required and economic value generated by such
+recovery, but are not essential to understand the broad pattern that suggests
+that nearly all stocks should recover under either an RBFM or MSY-style 
+management, but that the recovery would be fastest and the economic benefits
+greatest under the former.  That conclusion was already implicit in how
+these models are defined.
+
+-------------
 

### 2017-10-02 appendix/appendix.Rmd
WIP

---
title: "POMDP comparisons across sigma_m, sigma_g"
author: "Carl Boettiger"
date: "`r Sys.Date()`"
output: 
  github_document: default
  pdf_document: default
---

```{r include=FALSE}
## aesthetics
knitr::opts_chunk$set(comment="", message=FALSE)
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r message=FALSE}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(sarsop)       # the main POMDP package
library(tidyverse)    # for munging and plotting
library(parallel)
```

## Basic deterministic model

```{r}
r <- 0.75
K <- 1

## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
f <- function(x, h){ 
    x + x * r * (1 - x / K) - pmin(x,h)
}
```

Utility (reward) function:

```{r}
reward_fn <- function(x,h) pmin(x,h)
discount <- 1
```

## Calculating MSY

This uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.

```{r}
## A generic routine to find stock size (x) which maximizes 
## growth rate (f(x,0) - x, where x_t+1 = f(x_t))
S_star <- optimize(function(x) -f(x,0) + x / discount, 
                   c(0, 2*K))$minimum
```

```{r}
B_MSY <- S_star
MSY <- f(B_MSY,0) - B_MSY

F_MSY <- MSY / B_MSY  
F_PGY = 0.8 * F_MSY
```

### Optional checks:

```{r}
## Make sure these match the known analytic solutions for Graham-Schaefer model:
testthat::expect_equal(B_MSY, K / 2)
testthat::expect_equal(MSY, r * K / 4)
testthat::expect_equal(F_MSY, r / 2)
```

As a basic reference point, simulate these three policies in a purely deterministic world.  Unlike later simulations, here we consider all states an actions exactly (that is, within floating point precision).  Later, states and actions are limited to a discrete set, so solutions can depend on resolution and extent of that discretization.

We first define functions that correspond to each policy:

```{r}
msy_policy <- function(x) F_MSY * x
pgy_policy <- function(x) F_PGY * x

escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
  if(f(x,0) > B_MSY)
    f(x,0) - B_MSY
  else
    0
}
```

```{r}
# inits
x0 <- K/6
Tmax = 20

# a simulation function
det_sim <- function(policy, f, x0, Tmax){
    action <- state <- obs <- as.numeric(rep(NA,Tmax))
    state[1] <- x0
    for(t in 1:(Tmax-1)){
      action[t] <- policy(state[t])
      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
      state[t+1] <- f(state[t], action[t]) 
    }
    data.frame(time = 1:Tmax, state, action, obs)
  }

# simulate each policy, returning all results in a single unified data frame 
sims <- 
  list(msy = msy_policy, 
       det = escapement_policy, 
       pgy = pgy_policy) %>% 
  map_df(det_sim, 
         f, x0, Tmax, 
         .id = "method") 
```

```{r}
# plot simulation results 
sims %>%
  ggplot(aes(time, state, col=method)) + 
  geom_line() + 
  geom_hline(aes(yintercept=B_MSY), lwd=1) 
```

-----------

##  Introduce a discrete grid

We compute the above policies on this grid for later comparison.

```{r}
states <- seq(0,2, length=100)
actions <- states

## Harvest after recruitment
det_policy <- sapply(states, function(x) pmax(f(x,0) - S_star,0))
det_action <- sapply(det_policy, function(x) which.min(abs(actions - x)))
msy_action <- sapply(states, function(x) which.min(abs(actions - x * F_MSY)))
pgy_action <- sapply(states, function(x) which.min(abs(actions - x * F_PGY)))

```

# POMDP Model

We compute POMDP matrices for a range of `sigma_g` and `sigma_m` values:

```{r}
observations <- states # same discretization, for convenience

meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE)

models <- 
  parallel::mclapply(1:dim(meta)[1], 
           function(i){
  fisheries_matrices(
  states = states,
  actions = actions,
  observed_states = observations,
  reward_fn = reward_fn,
  f = f,
  sigma_g = meta[i,"sigma_g"][[1]],
  sigma_m = meta[i,"sigma_m"][[1]],
  noise = "normal")
})

```

## Simulating the static policies under uncertainty

```{r}
# inits
x0 <- which.min(abs(states - K/6))
Tmax <- 100
reps <- 100
options(mc.cores = parallel::detectCores())

static_sims <-
  map_dfr(models, 
          function(m){
            do_policy_sim <- function(policy){
              sim <- sim_pomdp(m$transition, m$observation, m$reward, discount, 
                               x0 = x0, Tmax = Tmax, policy = policy, reps = reps)
              sim$df %>% 
                mutate(state = states[state], action = actions[action])
            }

            list(msy = msy_action, 
                 det = det_action, 
                 pgy = pgy_action) %>% 
              map_df(do_policy_sim, .id = "method")
          },
          .id = "scenario") 
```

```{r}
# add parameter metadata to the output data.frame to track which simulation had which parameters
meta_simple <-  
  meta %>% 
  select(sigma_m, sigma_g) %>%
  mutate(scenario  = as.character(1:dim(meta)[1]))
static_sims <- right_join(meta_simple, static_sims)
```

```{r}
# plot
static_sims %>%
  select(time, state, rep, method, sigma_m, sigma_g) %>%
  group_by(time, method, sigma_m, sigma_g) %>%
  summarise(mean = mean(state), sd = sd(state)) %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line() + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
  facet_grid(sigma_m ~ sigma_g)
```

## POMDP solution

```{r eval=TRUE}
## log metadata associated with each run
log_data <- data.frame(id = id, 
                       model = "gs", 
                       r = r, 
                       K = K, 
                       sigma_g = sigma_g, 
                       sigma_m = sigma_m, 
                       noise = "normal")

dir.create(log_dir)

system.time({
alpha <- sarsop(m$transition, m$observation, m$reward, discount, 
                log_data = log_data, log_dir = log_dir,
                precision = .000002, timeout = 1000)
})
```

```{r}
meta <- meta_from_log(data.frame(id = id), log_dir)
alpha <- alphas_from_log(meta, log_dir)[[1]] ## bc fn returns a list with all matching alphas, we need [[1]]
```

## Combined plot

## Policy plots

## Historical data estimation

## Historical data POMDP & other solutions

## Historical combined plot

### 2017-10-02 appendix/solve_pomdp.R

@@ -1,0 +1,59 @@
+# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
+library(sarsop)       # the main POMDP package
+library(tidyverse)    # for munging and plotting
+library(parallel)
+options(mc.cores = 3) # Reserve ~ 10 GB per core
+log_dir <- "pomdp_intro"
+
+r <- 0.75
+K <- 1
+## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
+f <- function(x, h){ 
+  x + x * r * (1 - x / K) - pmin(x,h)
+}
+reward_fn <- function(x,h) pmin(x,h)
+discount <- 1
+
+## Discretize space
+states <- seq(0,2, length=100)
+actions <- states
+observations <- states
+
+## All parameter values combinations for which we want solutions
+meta <- expand.grid(sigma_g = c(0.01, 0.05, 0.1), 
+                    sigma_m = c(0, 0.1, 0.2),
+                    stringsAsFactors = FALSE)
+
+## Create the models
+models <- 
+  parallel::mclapply(1:dim(meta)[1], 
+                     function(i){
+                       fisheries_matrices(
+                         states = states,
+                         actions = actions,
+                         observed_states = observations,
+                         reward_fn = reward_fn,
+                         f = f,
+                         sigma_g = meta[i,"sigma_g"][[1]],
+                         sigma_m = meta[i,"sigma_m"][[1]],
+                         noise = "normal")
+                     })
+
+## log metadata associated with each run
+log_data <- data.frame(id = id, 
+                       model = "gs", 
+                       r = r, 
+                       K = K, 
+                       sigma_g = sigma_g, 
+                       sigma_m = sigma_m, 
+                       noise = "normal")
+dir.create(log_dir)
+
+## POMDP solution (slow, >10,000 seconds per loop memory intensive)
+system.time({
+  alpha <- sarsop(m$transition, m$observation, m$reward, discount, 
+                  log_data = log_data, log_dir = log_dir,
+                  precision = .000002, timeout = 10000)
+})
+
+

### 2017-10-02 appendix/solve_pomdp.R
fix script

@@ -40,20 +40,32 @@ models <-
                      })
 
 ## log metadata associated with each run
-log_data <- data.frame(id = id, 
-                       model = "gs", 
-                       r = r, 
-                       K = K, 
-                       sigma_g = sigma_g, 
-                       sigma_m = sigma_m, 
-                       noise = "normal")
+
 dir.create(log_dir)
 
 ## POMDP solution (slow, >10,000 seconds per loop memory intensive)
-system.time({
-  alpha <- sarsop(m$transition, m$observation, m$reward, discount, 
-                  log_data = log_data, log_dir = log_dir,
-                  precision = .000002, timeout = 10000)
-})
+system.time(
+  alphas <- 
+    mclapply(1:length(models), 
+    function(i){
+      log_data <- data.frame(id = "pomdp_intro", 
+                             model = "gs", 
+                             r = r, 
+                             K = K, 
+                             sigma_g = meta[i,"sigma_g"][[1]], 
+                             sigma_m = meta[i,"sigma_m"][[1]], 
+                             noise = "normal")
+      
+      sarsop(models[[i]]$transition,
+             models[[i]]$observation,
+             models[[i]]$reward,
+             discount = discount,
+             precision = 0.00000002,
+             timeout = 10000,
+             log_dir = log_dir,
+             log_data = log_data)
+    },
+    mc.cores = mc.cores)
+)
 
 

### 2017-10-02 appendix/solve_pomdp.R
fix typo

@@ -46,7 +46,7 @@ dir.create(log_dir)
 ## POMDP solution (slow, >10,000 seconds per loop memory intensive)
 system.time(
   alphas <- 
-    mclapply(1:length(models), 
+    parallel::mclapply(1:length(models), 
     function(i){
       log_data <- data.frame(id = "pomdp_intro", 
                              model = "gs", 
@@ -64,8 +64,7 @@ system.time(
              timeout = 10000,
              log_dir = log_dir,
              log_data = log_data)
-    },
-    mc.cores = mc.cores)
+    })
 )
 
 

### 2017-10-02 appendix/appendixB.Rmd
updated appendices

---
output: github_document
---

Computes model parameter estimates on each `stockid` in RAM (after normalizing data) using nimble. Then, estimate common stock measurement (Bmsy, MSY, PGY).

- Currently parameter estimates are just MLE. Should do Bayesian version (slower of course, and tricky to summarize; maybe just as sds around estimates.)

```{r aethetics, include=FALSE}
library(printr)
knitr::opts_chunk$set(message=FALSE, comment="")
```

```{r libraries}
library(tidyverse)
library(nimble)
```

```{r}
download.file("https://depts.washington.edu/ramlegac/wordpress/databaseVersions/RLSADB_v3.0_(assessment_data_only)_excel.zip",
              "ramlegacy.zip")
```

```{r}
path <- unzip("ramlegacy.zip")
sheets <- readxl::excel_sheets(path)
ram <- lapply(sheets, readxl::read_excel, path = path)
names(ram) <- sheets
unlink("ramlegacy.zip")
unlink("RLSADB v3.0 (assessment data only).xlsx")
ramlegacy <- 
  ram$timeseries_values_views %>%
  select(assessid, stockid, stocklong, year, SSB, TC) %>%
  left_join(ram$stock) %>%
  left_join(ram$area) %>%
  select(assessid, stockid, scientificname, 
         commonname, areaname, country, year, 
         SSB, TC) %>%
  left_join(ram$timeseries_units_views %>%
              rename(TC_units = TC, SSB_units = SSB)) %>%
  select(scientificname, commonname, 
         stockid, areaname, country, year, 
         SSB, TC, SSB_units, TC_units)
```

Let's filter out missing data, non-matching units, and obvious reporting errors (catch exceeding total spawning biomass), then we rescale each series into the 0,1 by appropriate choice of units:

```{r}
df2 <- ramlegacy %>% 
  filter(!is.na(SSB), !is.na(TC)) %>%
  filter(SSB_units == "MT", TC_units=="MT") %>% 
  filter(SSB > TC) %>%
  select(-SSB_units, -TC_units) %>% 
  group_by(stockid) %>%
  mutate(scaled_catch = TC / max(SSB),
         scaled_biomass = SSB / max(SSB)) 
```

```{r subset}
stock_ids <- c("PLAICNS", "ARGHAKENARG")
examples <- df2 %>% filter(stockid %in% stock_ids)
```

```{r}
## Model does not estimate sigma_m; data is insufficient to do so.
## Note that RAM stock estimates are not always direct measurements, can be artifically smooth, underestimating sigma_g

gs_code  <- nimble::nimbleCode({

  r ~ dunif(0, 2)
  K ~ dunif(0, 2)
  sigma ~ dunif(0, 1)

  x[1] <- x0
  for(t in 1:(N-1)){
    mu[t] <- x[t] + x[t] * r * (1 - x[t] / K) - min(a[t],x[t])
    x[t+1] ~ dnorm(mu[t], sd = sigma)
  }

})

fit_models <- function(fish, code){
  # fish <- examples %>% filter(stockid == stock_ids[1])

  ## Rescale data
  N <- dim(fish)[1]
  scaled_data <- data.frame(t = 1:N, y = fish$scaled_biomass, a = fish$scaled_catch)
  data = data.frame(x = scaled_data$y)

  ## Compile  model
  constants <- list(N = N, a = scaled_data$a)
  inits <- list(r = 0.5, K = 0.5, sigma = 0.02, x0 = scaled_data$y[1])
  model <- nimbleModel(code, constants, data, inits)
  C_model <- compileNimble(model)

  mcmcspec <- configureMCMC(model, thin = 1e2)
  mcmc <- buildMCMC(mcmcspec)
  Cmcmc <- compileNimble(mcmc, project = model)
  Cmcmc$run(1e6)

  samples <- as.data.frame(as.matrix(Cmcmc$mvSamples))
  burnin <- 1:(0.05 * dim(samples)[1]) # drop first 5%
  samples <- samples[-burnin,1:(length(inits) - 1)] # drop raised vars, burnin

  #gather(samples) %>% ggplot() + geom_density(aes(value)) + facet_wrap(~key, scale='free')

  ## Return fit
  data.frame(stockid = fish$stockid[1],
             species = fish$commonname[1],
             r = mean(samples$r),
             K = mean(samples$K),
             sigma_g = mean(samples$sigma),
             r_sd = sd(samples$r),
             K_sd = sd(samples$K),
             sigma_g_sd = sd(samples$sigma),
             stringsAsFactors = FALSE)

}
```

```{r}
fits <- examples %>% do(fit_models(., code=gs_code))
fits 
```

Do the estimates make sense?

```{r}
examples %>% ungroup() %>%
  select(year, scaled_biomass, scaled_catch, commonname) %>%
  gather(stock, biomass, -year, -commonname) %>%
  ggplot(aes(year, biomass, col=stock)) + 
  geom_line() + facet_wrap(~commonname, scales = "free")
```

```{r}
pars <- fits %>% ungroup() %>% select(species, r, K, sigma_g) 
```

```{r}
# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
library(sarsop)       # the main POMDP package
library(tidyverse)    # for munging and plotting
library(parallel)
options(mc.cores = 6) # Reserve ~ 10 GB per core
log_dir <- "pomdp_intro"

## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
f <- function(x, h){ 
  x + x * r * (1 - x / K) - pmin(x,h)
}
reward_fn <- function(x,h) pmin(x,h)
discount <- 1

## Discretize space
states <- seq(0,4, length=150)
actions <- states
observations <- states

## All parameter values combinations for which we want solutions
meta <- expand.grid(species = pars$species, 
                    sigma_m = c(0, 0.1, 0.2),
                    stringsAsFactors = FALSE) %>%
  left_join(pars)

## Create the models
models <- 
  parallel::mclapply(1:dim(meta)[1], 
                     function(i){
                       fisheries_matrices(
                         states = states,
                         actions = actions,
                         observed_states = observations,
                         reward_fn = reward_fn,
                         f = f,
                         sigma_g = meta[i,"sigma_g"][[1]],
                         sigma_m = meta[i,"sigma_m"][[1]],
                         noise = "normal")
                     })

## log metadata associated with each run

dir.create(log_dir)

## POMDP solution (slow, >20,000 seconds per loop memory intensive)
system.time(
  alphas <- 
    parallel::mclapply(1:length(models), 
    function(i){
      log_data <- data.frame(id = "pomdp_historical", 
                             model = "gs", 
                             r = meta[i, "r"][[1]], 
                             K = meta[i, "K"][[1]], 
                             sigma_g = meta[i,"sigma_g"][[1]], 
                             sigma_m = meta[i,"sigma_m"][[1]], 
                             noise = "normal")

      sarsop(models[[i]]$transition,
             models[[i]]$observation,
             models[[i]]$reward,
             discount = discount,
             precision = 0.00000002,
             timeout = 20000,
             log_dir = log_dir,
             log_data = log_data)
    })
)

```

### 2017-10-02 appendix/appendixB.Rmd
define f

@@
 log_dir <- "pomdp_intro"
+```
 
+```{r}
 ## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
-f <- function(x, h){ 
+gs <- function(r,K){
+  function(x, h){ 
     x + x * r * (1 - x / K) - pmin(x,h)
+  }
 }
@@
 discount <- 1
+```
 
 ## Discretize space
+
+Note that the large values of $K$ require we carry the numerical grid out further.
+
+```{r}
 states <- seq(0,4, length=150)
@@
 observations <- states
+```
 
 ## All parameter values combinations for which we want solutions
+
+```{r}
 meta <- expand.grid(species = pars$species, 
@@
   left_join(pars)
+meta
+```
 
 ## Create the models
+
+```{r}
 models <- 
@@
                          reward_fn = reward_fn,
-                         f = f,
+                         f = gs(meta[i, "r"][[1]], meta[i, "K"][[1]]),
                          sigma_g = meta[i,"sigma_g"][[1]],
@@
 ```
+
+## Comparison to the static models
+
+```{r eval=FALSE}
+## A generic routine to find stock size (x) which maximizes 
+## growth rate (f(x,0) - x, where x_t+1 = f(x_t))
+
+f <- gs(pars[i,"r"][[1]], pars[i, "K"][[1]])
+
+S_star <- optimize(function(x) -f(x,0) + x / discount, 
+                   c(0, 2*K))$minimum
+
+B_MSY <- S_star
+MSY <- f(B_MSY,0) - B_MSY
+
+F_MSY <- MSY / B_MSY  
+F_PGY = 0.8 * F_MSY
+
+policies <- list(
+  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
+  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
+  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
+```

### 2017-10-04 appendix/solve_pomdp.R
more reasonable precision

@@ -2,7 +2,7 @@
 library(sarsop)       # the main POMDP package
 library(tidyverse)    # for munging and plotting
 library(parallel)
-options(mc.cores = 3) # Reserve ~ 10 GB per core
+options(mc.cores = 2) # Reserve ~ 10 GB per core
 log_dir <- "pomdp_intro"
 
 r <- 0.75
@@ -60,7 +60,7 @@ system.time(
              models[[i]]$observation,
              models[[i]]$reward,
              discount = discount,
-             precision = 0.00000002,
+             precision = 2e-6,
              timeout = 10000,
              log_dir = log_dir,
              log_data = log_data)

### 2017-10-04 appendix/appendix.Rmd

@@
 
-```{r include=FALSE}
-## aesthetics
+```{r aethetics, include=FALSE}
 knitr::opts_chunk$set(comment="", message=FALSE)
 ggplot2::theme_set(ggplot2::theme_bw())
+```
+
+```{r helper_routines, include = FALSE}
+det_simulations <- function(x0 = 1/6, Tmax = 20, 
+                              B_MSY = K / 2, 
+                              F_MSY = r / 2 , 
+                              F_PGY = 0.8 * F_MSY){
+
+  msy_policy <- function(x) F_MSY * x
+  pgy_policy <- function(x) F_PGY * x
+  escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
+    if(f(x,0) > B_MSY)
+      f(x,0) - B_MSY
+    else
+      0
+  }
+
+  # a simulation function
+  det_sim <- function(policy, f, x0, Tmax){
+    action <- state <- obs <- as.numeric(rep(NA,Tmax))
+    state[1] <- x0
+    for(t in 1:(Tmax-1)){
+      action[t] <- policy(state[t])
+      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
+      state[t+1] <- f(state[t], action[t]) 
+    }
+    data.frame(time = 1:Tmax, state, action, obs)
+  }
+
+  # simulate each policy, returning all results in a single unified data frame 
+  sims <- 
+    list(msy = msy_policy, 
+         det = escapement_policy, 
+         pgy = pgy_policy) %>% 
+    map_df(det_sim, 
+           f, x0, Tmax, 
+           .id = "method")
+
+  sims
+}
 ```
@@
 
-We first define functions that correspond to each policy:
-
 ```{r}
-msy_policy <- function(x) F_MSY * x
-pgy_policy <- function(x) F_PGY * x
-
-escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
-  if(f(x,0) > B_MSY)
-    f(x,0) - B_MSY
-  else
-    0
-}
-```
-
-```{r}
-# inits
-x0 <- K/6
-Tmax = 20
-
-# a simulation function
-det_sim <- function(policy, f, x0, Tmax){
-    action <- state <- obs <- as.numeric(rep(NA,Tmax))
-    state[1] <- x0
-    for(t in 1:(Tmax-1)){
-      action[t] <- policy(state[t])
-      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
-      state[t+1] <- f(state[t], action[t]) 
-    }
-    data.frame(time = 1:Tmax, state, action, obs)
-  }
-
-# simulate each policy, returning all results in a single unified data frame 
-sims <- 
-  list(msy = msy_policy, 
-       det = escapement_policy, 
-       pgy = pgy_policy) %>% 
-  map_df(det_sim, 
-         f, x0, Tmax, 
-         .id = "method") 
-```
-
-```{r}
-# plot simulation results 
-sims %>%
+det_simulations(x0 = 1/6, Tmax = 20, B_MSY, F_MSY, F_PGY) %>%
   ggplot(aes(time, state, col=method)) + 
@@
 
+```{r}
+## Discretize space
+states <- seq(0,2, length=100)
+actions <- states
+observations <- states
+```
+
 We compute the above policies on this grid for later comparison.
@@
 ```{r}
-states <- seq(0,2, length=100)
-actions <- states
-
-## Harvest after recruitment
-det_policy <- sapply(states, function(x) pmax(f(x,0) - S_star,0))
-det_action <- sapply(det_policy, function(x) which.min(abs(actions - x)))
-msy_action <- sapply(states, function(x) which.min(abs(actions - x * F_MSY)))
-pgy_action <- sapply(states, function(x) which.min(abs(actions - x * F_PGY)))
-
+policies <- list(
+  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
+  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
+  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
 ```
@@
 ```{r}
-observations <- states # same discretization, for convenience
-
 meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
                     sigma_m = c(0, 0.1, 0.2),
-                    stringsAsFactors = FALSE)
+                    stringsAsFactors = FALSE) %>%
+        mutate(scenario  = as.character(1:length(sigma_m)))
+```
 
+```{r}
 models <- 
@@
 
+## POMDP solution
+
+The POMDP solution is represented by a collection of alpha-vectors and values, returned in a `*.policyx` file.
+Each scenario (parameter combination of `sigma_g`, `sigma_m`, and so forth) results in a separate solution file.
+
+Because this solution is computationally somewhat intensive, we provide
+
+```{r eval=FALSE}
+log_dir <- "appendix_alphas" # Store the computed solution files here
+dir.create(log_dir)
+id = "pomdp_intro"    # an optional id for these runs
+
+## POMDP solution (slow, >10,000 seconds per scenario, & memory intensive)
+system.time(
+  alphas <- 
+    parallel::mclapply(1:length(models), 
+    function(i){
+      log_data <- data.frame(id = id, 
+                             model = "gs", 
+                             r = r, 
+                             K = K, 
+                             sigma_g = meta[i,"sigma_g"][[1]], 
+                             sigma_m = meta[i,"sigma_m"][[1]], 
+                             noise = "normal")
+
+      sarsop(models[[i]]$transition,
+             models[[i]]$observation,
+             models[[i]]$reward,
+             discount = discount,
+             precision = 0.00000002,
+             timeout = 10000,
+             log_dir = log_dir,
+             log_data = log_data)
+    })
+)
+
+```
+
+We can read the stored solution from the log:
+
+```{r}
+#alphas <- alphas_from_log(meta, log_dir)
+```
+
 ## Simulating the static policies under uncertainty
 
-```{r}
-# inits
-x0 <- which.min(abs(states - K/6))
+```{r static_sims}
 Tmax <- 100
+x0 <- which.min(abs(K/6 - states))
 reps <- 100
-options(mc.cores = parallel::detectCores())
 
 static_sims <- 
-  map_dfr(models, 
-          function(m){
-            do_policy_sim <- function(policy){
-              sim <- sim_pomdp(m$transition, m$observation, m$reward, discount, 
-                               x0 = x0, Tmax = Tmax, policy = policy, reps = reps)
-              sim$df %>% 
-                mutate(state = states[state], action = actions[action])
-            }
+ map_dfr(models, function(m){
+            do_sim <- function(policy) sim_pomdp(
+                        m$transition, m$observation, m$reward, discount, 
+                        x0 = x0, Tmax = Tmax, policy = policy, reps = reps)$df
+            map_dfr(policies, do_sim, .id = "method")
+          }, .id = "scenario")
 
-            list(msy = msy_action, 
-                 det = det_action, 
-                 pgy = pgy_action) %>% 
-              map_df(do_policy_sim, .id = "method")
-          },
-          .id = "scenario") 
 ```
 
-```{r}
-# add parameter metadata to the output data.frame to track which simulation had which parameters
-meta_simple <-  
-  meta %>% 
-  select(sigma_m, sigma_g) %>%
-  mutate(scenario  = as.character(1:dim(meta)[1]))
-static_sims <- right_join(meta_simple, static_sims)
+## Simulating the POMDP policies under uncertiainty
+
+```{r pomdp_sims}
+# slower!
+pomdp_sims <- 
+  map2_dfr(models, alphas, function(.x, .y){
+             sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
+                       unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
+                       reps=reps)$df %>% 
+              mutate(method = "pomdp") # include a column labeling method
+           },
+           .id = "scenario") %>% left_join(meta)
 ```
 
+Combine the resulting data frames
+
 ```{r}
-# plot
-static_sims %>%
+sims <- bind_rows(static_sims, pomdp_sims) %>%
+    left_join(meta) %>%    ## combine output with scenarios
+    mutate(state = states[state], action = actions[action]) # index -> value
+
+write_csv(sims, "sims.csv")
+```
+
+##  Figure S1
+
+We the results varying over different noise intensities, sigma_g, and sigma_m.  Figure 1 of the main text considers the case of sigma_g = 0.05, sigma_m = 0.1
+
+```{r}
+sims %>%
   select(time, state, rep, method, sigma_m, sigma_g) %>%
@@
 
-## POMDP solution
-
-```{r eval=TRUE}
-## log metadata associated with each run
-log_data <- data.frame(id = id, 
-                       model = "gs", 
-                       r = r, 
-                       K = K, 
-                       sigma_g = sigma_g, 
-                       sigma_m = sigma_m, 
-                       noise = "normal")
-
-dir.create(log_dir)
-
-system.time({
-alpha <- sarsop(m$transition, m$observation, m$reward, discount, 
-                log_data = log_data, log_dir = log_dir,
-                precision = .000002, timeout = 1000)
-})
-```
-
-```{r}
-meta <- meta_from_log(data.frame(id = id), log_dir)
-alpha <- alphas_from_log(meta, log_dir)[[1]] ## bc fn returns a list with all matching alphas, we need [[1]]
-```
-
-## Combined plot
-
 ## Policy plots

### 2017-10-04 appendix/appendixB.Rmd
no printr

@@
 ```{r libraries}
+# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
 library(tidyverse)
+library(sarsop)
 library(nimble)
+library(parallel)
 ```
@@
 stock_ids <- c("PLAICNS", "ARGHAKENARG")
-examples <- df2 %>% filter(stockid %in% stock_ids)
+examples <- df2 %>% 
+  filter(stockid %in% stock_ids) %>% 
+  ungroup() %>% 
+  group_by(commonname)
 ```
@@
   data.frame(stockid = fish$stockid[1],
-             species = fish$commonname[1],
+             commonname = fish$commonname[1],
              r = mean(samples$r),
@@
 ```{r}
-pars <- fits %>% ungroup() %>% select(species, r, K, sigma_g) 
+pars <- fits %>% ungroup() %>% select(commonname, r, K, sigma_g) 
 ```
@@
 ```{r}
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
-library(sarsop)       # the main POMDP package
-library(tidyverse)    # for munging and plotting
-library(parallel)
+
 options(mc.cores = 6) # Reserve ~ 10 GB per core
@@
 ```{r}
-meta <- expand.grid(species = pars$species, 
+meta <- expand.grid(commonname = pars$commonname, 
                     sigma_m = c(0, 0.1, 0.2),
                     stringsAsFactors = FALSE) %>%
-  left_join(pars)
+  left_join(pars) %>%
+  mutate(scenario  = as.character(1:length(sigma_m)))
+
 meta
@@
                      })
+```
 
-## log metadata associated with each run
-
+```{r}
 dir.create(log_dir)
@@
 
-```{r eval=FALSE}
-## A generic routine to find stock size (x) which maximizes 
-## growth rate (f(x,0) - x, where x_t+1 = f(x_t))
+```{r}
+pars <- examples %>% 
+  group_by(commonname) %>% 
+  summarise(N = max(SSB)) %>% 
+  right_join(pars)
+```
 
-f <- gs(pars[i,"r"][[1]], pars[i, "K"][[1]])
+<!--- ICK, find a more streamlined way to do this stuff species by species -->
 
-S_star <- optimize(function(x) -f(x,0) + x / discount, 
-                   c(0, 2*K))$minimum
+Convert example data into discrete index space.
 
+```{r}
+ex <- examples %>%
+  mutate(biomass = index(scaled_biomass, states),          
+         catch = index(scaled_catch, actions))  %>% 
+  select(year, biomass, catch, commonname)
+```
+
+Add corresponding static policy levels on:
+
+```{r}
+statics <- function(P){
+  f <- gs(P$r, P$K)
+  S_star <- optimize(function(x) -f(x,0) + x / discount, c(0, 2* P$K))$minimum
   B_MSY <- S_star
@@
 
-F_MSY <- MSY / B_MSY  
-F_PGY = 0.8 * F_MSY
+  tibble(S_star, F_MSY = MSY / B_MSY, F_PGY = 0.8 * F_MSY, 
+         commonname = P$commonname, N = P$N)
+  }
 
-policies <- list(
-  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
-  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
-  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
+  pars %>% 
+  rowwise() %>% 
+  do(statics(.)) %>% 
+  right_join(ex) %>% 
+  ungroup()  -> ex2
+
+  ex2
 ```
+
+```{r} 
+det_f <- function(S_star, i) 
+  map_int(pmax(f(states,0) - S_star[[1]],0), ~ which.min(abs(actions - .x)))[i]
+msy_f <- function(F_MSY, i) 
+  map_int(states, ~ which.min(abs(actions - .x * F_MSY[[1]])))[i]
+pgy_f <- function(F_PGY, i) 
+  map_int(states, ~ which.min(abs(actions - .x * F_PGY[[1]])))[i]
+
+rescale <- function(x, N) states[x]*N
+
+historical <- ex2 %>% 
+  group_by(commonname) %>% 
+  mutate(det =  det_f(S_star, biomass),
+         msy =  msy_f(F_MSY, biomass),
+         pgy =  pgy_f(F_PGY, biomass)) %>%
+  mutate_if(is.integer, rescale, N=N) %>%
+  select(year, biomass, catch, det, msy, pgy, commonname)
+```
+
+```{r}
+historical %>%
+select(year, commonname, pgy, msy, det) %>%
+gather(model, stock, -year, -commonname) %>%
+  ggplot(aes(year, stock, col=model)) + 
+  geom_line(data = historical, aes(year, biomass), col="grey", lwd=1) +
+  geom_line(data = historical, aes(year, catch), col="black", lwd=1) +
+  geom_line(lwd=1) + facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
+```
+
+## POMDP historical
+
+```{r pomdp_sims}
+## FIXME obs, action is in index space!
+pomdp_sims <- 
+  map2_dfr(models, alphas, function(.x, .y){
+             hindcast_pomdp(.x$transition, .x$observation, .x$reward, discount, 
+                            obs = historical$scaled_biomass, actions = historical$scaled_catch,
+                            state_prior = unif_prior,
+                            alpha = .y)$df %>% 
+              mutate(method = "pomdp") # include a column labeling method
+           },
+           .id = "scenario") %>% left_join(meta)
+```

### 2017-10-04 appendix/appendixB.Rmd
fewer cores

--- 2017-10-02 appendix/appendixB.Rmd
@@
 ```{r aethetics, include=FALSE}
-library(printr)
+#library(printr)
 knitr::opts_chunk$set(message=FALSE, comment="")
@@
 library(parallel)
-options(mc.cores = 6) # Reserve ~ 10 GB per core
+options(mc.cores = 3) # Reserve ~ 10 GB per core
 log_dir <- "pomdp_intro"

### 2017-10-05 appendix/appendix.Rmd
solution with discount 0.99 resolves initialization issues

@@
 ```{r}
-#alphas <- alphas_from_log(meta, log_dir)
+log_dir <- "../pomdp_intro"
+meta <- meta_from_log("pomdp_intro", log_dir)
+alphas <- alphas_from_log(meta, log_dir)
 ```
@@
 # slower!
+unif_prior <- rep(1, length(states)) / length(states)
 pomdp_sims <- 

### 2017-10-06 appendix/appendixB.Rmd
cleaner code

--- 2017-10-04 appendix/appendixB.Rmd
@@
 
-<!--- ICK, find a more streamlined way to do this stuff species by species -->
-
-Convert example data into discrete index space.
-
-```{r}
-ex <- examples %>%
-  mutate(biomass = index(scaled_biomass, states),          
-         catch = index(scaled_catch, actions))  %>% 
-  select(year, biomass, catch, commonname)
-```
-
 Add corresponding static policy levels on:
@@
 
+policy_pars <- 
   pars %>% 
   rowwise() %>% 
-  do(statics(.)) %>% 
-  right_join(ex) %>% 
-  ungroup()  -> ex2
+  do(statics(.))
+```  
+  Convert example data into discrete index space.
 
-  ex2
+```{r}
+ex <- examples %>%
+  mutate(biomass = index(scaled_biomass, states),          
+         catch = index(scaled_catch, actions))  %>% 
+  left_join(policy_pars) %>% 
+  ungroup()
+
 ```
@@
 
-historical <- ex2 %>% 
+historical <- ex %>% 
   group_by(commonname) %>% 
@@
 ```
+
+Plot
 

### 2017-10-06 appendix/appendixB.Rmd

--- 2017-10-04 appendix/appendixB.Rmd
@@
 ```{r}
-
-options(mc.cores = 6) # Reserve ~ 10 GB per core
-log_dir <- "pomdp_intro"
+# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
+library(sarsop)       # the main POMDP package
+library(tidyverse)    # for munging and plotting
+library(parallel)
+options(mc.cores = 3) # Reserve ~ 10 GB per core
+log_dir <- "pomdp_historical"
 ```
@@
 reward_fn <- function(x,h) pmin(x,h)
-discount <- 1
+discount <- .99
 ```
@@
              discount = discount,
-             precision = 0.00000002,
+             precision = 2e-6,
              timeout = 20000,

### 2017-10-06 appendix/appendix.Rmd
cannot fix id, overwrites outputs

@@
 
-```{r eval=FALSE}
+```{r eval=TRUE}
 log_dir <- "appendix_alphas" # Store the computed solution files here
 dir.create(log_dir)
-id = "pomdp_intro"    # an optional id for these runs
 
@@
     function(i){
-      log_data <- data.frame(id = id, 
-                             model = "gs", 
+      log_data <- data.frame(model = "gs", 
                              r = r, 
@@
 ```{r}
-log_dir <- "../pomdp_intro"
-meta <- meta_from_log("pomdp_intro", log_dir)
+meta <- meta_from_log(data.frame(model="gs"), log_dir)
 alphas <- alphas_from_log(meta, log_dir)

### 2017-10-07 appendix/appendix.Rmd
fix meta read

--- 2017-10-05 appendix/appendix.Rmd
@@
 log_dir <- "../pomdp_intro"
-meta <- meta_from_log("pomdp_intro", log_dir)
+meta <- meta_from_log("pomdp_intro", log_dir) %>% mutate(scenario=as.character(1:length(models)))
 alphas <- alphas_from_log(meta, log_dir)
@@
 ```{r}
-sims %>%
+pomdp_sims %>%
   select(time, state, rep, method, sigma_m, sigma_g) %>%

### 2017-10-07 appendix/appendix.Rmd

--- 2017-10-06 appendix/appendix.Rmd
@@
 ```{r}
-meta <- meta_from_log(data.frame(model="gs"), log_dir)
+meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% mutate(scenario=as.character(1:length(models)))
 alphas <- alphas_from_log(meta, log_dir)

### 2017-10-07 appendix/appendixB.Rmd

@@
 library(parallel)
-options(mc.cores = 3) # Reserve ~ 10 GB per core
+options(mc.cores = 4) # Reserve ~ 10 GB per core
 log_dir <- "pomdp_historical"
@@
     function(i){
-      log_data <- data.frame(id = "pomdp_historical", 
-                             model = "gs", 
+      log_data <- data.frame(model = "gs", 
                              r = meta[i, "r"][[1]], 

### 2017-10-08 appendix/appendixB.Rmd
revised appendixB sims

@@
 #library(printr)
-knitr::opts_chunk$set(message=FALSE, comment="")
+knitr::opts_chunk$set(message=FALSE, comment="", cache=TRUE)
 ```
@@
 ```{r}
+set.seed(123)
 fits <- examples %>% do(fit_models(., code=gs_code))
@@
 
-```{r}
+```{r eval=FALSE}
 dir.create(log_dir)
@@
                              sigma_m = meta[i,"sigma_m"][[1]], 
-                             noise = "normal")
+                             noise = "normal",
+                             commonname = meta[i, "commonname"][[1]])
 
@@
 
+```{r}
+meta <- meta_from_log(data.frame(model="gs"), log_dir) %>%
+  mutate(scenario  = as.character(1:length(sigma_m)))
+alphas <- alphas_from_log(meta, log_dir)
+#models <- models_from_log(meta)
+```
+
 ## Comparison to the static models
@@
   summarise(N = max(SSB)) %>% 
-  right_join(pars)
+  right_join(meta)
 ```
@@
 ```{r}
+index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
+
 ex <- examples %>%
@@
   ungroup() 
-
+ex
 ```
@@
 ```{r} 
-det_f <- function(S_star, i) 
-  map_int(pmax(f(states,0) - S_star[[1]],0), ~ which.min(abs(actions - .x)))[i]
-msy_f <- function(F_MSY, i) 
-  map_int(states, ~ which.min(abs(actions - .x * F_MSY[[1]])))[i]
-pgy_f <- function(F_PGY, i) 
-  map_int(states, ~ which.min(abs(actions - .x * F_PGY[[1]])))[i]
+det_f <- function(S_star, i) index(pmax(f(states,0) - S_star[[1]],0), actions)[i]
+msy_f <- function(F_MSY, i) index(states * F_MSY[[1]], actions)[i]
+pgy_f <- function(F_PGY, i) index(states * F_PGY[[1]], actions)[i]
 
@@
          pgy =  pgy_f(F_PGY, biomass)) %>%
-  mutate_if(is.integer, rescale, N=N) %>%
-  select(year, biomass, catch, det, msy, pgy, commonname)
+  select(year, biomass, catch, det, msy, pgy, commonname, N) %>%
+  gather(model, stock, -year, -commonname, -N) %>%
+  mutate(stock = states[stock] * N) %>% 
+  select(-N)
 ```
@@
 historical %>%
-select(year, commonname, pgy, msy, det) %>%
-gather(model, stock, -year, -commonname) %>%
   ggplot(aes(year, stock, col=model)) +
-  geom_line(data = historical, aes(year, biomass), col="grey", lwd=1) +
-  geom_line(data = historical, aes(year, catch), col="black", lwd=1) +
-  geom_line(lwd=1) + facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
+  geom_line(lwd=1) +
+  scale_color_manual(values = c("black", "grey", "#D9661F", "#00B0DA", "#CFDD45")) +
+  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
 ```
@@
 
-```{r pomdp_sims}
-## FIXME obs, action is in index space!
+```{r}
   pomdp_sims <- 
-  map2_dfr(models, alphas, function(.x, .y){
+  pmap_dfr(list(models, alphas, 1:dim(meta)[[1]]), 
+           function(.x, .y, .z){
+
+             who <- ex$commonname == meta[.z,"commonname"]
+             df <- ex[who,]
+
              hindcast_pomdp(.x$transition, .x$observation, .x$reward, discount, 
-                            obs = historical$scaled_biomass, actions = historical$scaled_catch,
-                            state_prior = unif_prior,
+                            obs = index(df$scaled_biomass, states), 
+                            action = index(df$scaled_catch,states),
                             alpha = .y)$df %>% 
@@
 ```
+```{r}
+rescale <- function(x) states[x]
+
+pomdp_sims %>% 
+  select(time, obs, action, optimal, commonname, sigma_m) %>%
+  mutate_at(c("action", "optimal"), rescale) %>%
+  ggplot(aes(time, optimal, col=as.character(sigma_m))) + 
+  geom_line() + 
+  geom_line(aes(time, action), col="black") +
+  facet_wrap(~commonname, scales="free") 
+```

### 2017-10-09 appendix/appendixB.Rmd
tweaking appendixB

@@
   summarise(N = max(SSB)) %>% 
-  right_join(meta)
+  right_join(
+    meta %>% 
+      select(commonname, r, K) %>% 
+      distinct())
 ```
@@
 index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
-
+## repeats each series for each static model
 ex <- examples %>%
@@
 
-Plot
+## Plots
 
@@
 
-             who <- ex$commonname == meta[.z,"commonname"]
+             ## FIXME need to get obs for unique commmonname, wrong length?
+             who <- (ex$commonname == meta[.z,"commonname"]) 
              df <- ex[who,]
@@
                             alpha = .y)$df %>% 
-              mutate(method = "pomdp") # include a column labeling method
+              mutate(method = "pomdp") %>% # include a column labeling method
+              mutate(year = ex[who, "year"][[1]]) 
            },
-           .id = "scenario") %>% left_join(meta)
+           .id = "scenario") 
 ```
+
 ```{r}
-rescale <- function(x) states[x]
+pomdp_sims <- 
+  meta %>% 
+  select(scenario, commonname,sigma_m) %>% 
+  left_join(pars) %>%
+  right_join(pomdp_sims)
 
-pomdp_sims %>% 
-  select(time, obs, action, optimal, commonname, sigma_m) %>%
-  mutate_at(c("action", "optimal"), rescale) %>%
-  ggplot(aes(time, optimal, col=as.character(sigma_m))) + 
-  geom_line() + 
-  geom_line(aes(time, action), col="black") +
-  facet_wrap(~commonname, scales="free") 
 ```
+
+```{r}
+sims <- pomdp_sims %>% 
+  mutate(optimal = states[optimal] * N) %>% # original scale
+  select(year, optimal, commonname, sigma_m) %>%
+  rename(stock = optimal) %>% 
+  ## treat each sigma_m value as separate 'model'
+  mutate(sigma_m = as.factor(sigma_m)) %>%
+  mutate(model = recode(sigma_m, 
+                        "0" = "reed",
+                        "0.1" = "pomdp_0.1", 
+                        "0.2" = "pomdp_0.2")) %>%
+  select(-sigma_m) %>%
+  bind_rows(historical)
+```
+
+```{r}
+sims %>%
+  filter(model %in% c("biomass", "catch", "pomdp_0.2", "reed", "pgy")) %>%
+  ggplot(aes(year, stock, col=model)) +
+  geom_line(lwd=1) +
+  scale_color_manual(values = c("black", "grey", "#D9661F", "#FDB515", "#6C3302", "#00B0DA", "#CFDD45", "#3B7EA1")) +
+  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
+```

### 2017-10-10 appendix/appendix.Rmd
tweaks for savio
analytic checks only if discount = 1

@@
 
+```{r}
+options(mc.cores=9)
+```
+
 ## Basic deterministic model
@@
 reward_fn <- function(x,h) pmin(x,h)
-discount <- 1
+discount <- 0.99
 ```
@@
 
-### Optional checks:
-
-```{r}
-## Make sure these match the known analytic solutions for Graham-Schaefer model:
+```{r optional_checks}
+## Make sure these match the known analytic solutions for Graham-Schaefer model (only true of discount = 1)
+if(discount == 1){
   testthat::expect_equal(B_MSY, K / 2)
@@
   testthat::expect_equal(F_MSY, r / 2)
+}
 ```

### 2017-10-11 appendix/appendix.Rmd
fix order of meta

@@
         mutate(scenario  = as.character(1:length(sigma_m)))
+log_dir <- "appendix_alphas" # Store the computed solution files here
+meta
 ```
@@
 
-```{r eval=TRUE}
-log_dir <- "appendix_alphas" # Store the computed solution files here
+```{r eval=FALSE}
 dir.create(log_dir)
@@
                              sigma_m = meta[i,"sigma_m"][[1]], 
-                             noise = "normal")
+                             noise = "normal",
+                             scenario = meta[i, "scenario"][[1]])
 
@@
 ```{r}
-meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% mutate(scenario=as.character(1:length(models)))
+meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% left_join(meta) %>% arrange(scenario)
 alphas <- alphas_from_log(meta, log_dir)
@@
                        unif_prior, x0 = x0, Tmax = Tmax, alpha = .y,
-                       reps=reps)$df %>% 
+                       reps = reps)$df %>% 
               mutate(method = "pomdp") # include a column labeling method
            },
-           .id = "scenario") %>% left_join(meta)
+           .id = "scenario")
 ```
@@
 ```{r}
-pomdp_sims %>%
+sims %>%
   select(time, state, rep, method, sigma_m, sigma_g) %>%

### 2017-10-11 appendix/appendixB.Rmd
tidy up appendixB
ensure alignment on scenarios

@@
 
+# Section 1: Model estimates from data
+
 Computes model parameter estimates on each `stockid` in RAM (after normalizing data) using nimble. Then, estimate common stock measurement (Bmsy, MSY, PGY).
-
-- Currently parameter estimates are just MLE. Should do Bayesian version (slower of course, and tricky to summarize; maybe just as sds around estimates.)
 
@@
 
-Do the estimates make sense?
+<!--
+Do the estimates make sense?  Consider simulation from models, under this historical harvests, compared to historical trajectories
 
@@
 
+-->
+
 ```{r}
 pars <- fits %>% ungroup() %>% select(commonname, r, K, sigma_g) 
+pars
 ```
 
+-------
+
+# Decision Policies
+
 ```{r}
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
-library(sarsop)       # the main POMDP package
-library(tidyverse)    # for munging and plotting
-library(parallel)
 options(mc.cores = 4) # Reserve ~ 10 GB per core
@@
                              noise = "normal",
-                             commonname = meta[i, "commonname"][[1]])
+                             commonname = meta[i, "commonname"][[1]],
+                             scenario = meta[i, "scenario"][[1]])
 
@@
 meta <- meta_from_log(data.frame(model="gs"), log_dir)  %>% 
-  mutate(scenario  = as.character(1:length(sigma_m)))
+  left_join(
+    select(meta, sigma_m, commonname, scenario), 
+    by = c("sigma_m", "commonname")) %>% 
+  arrange(scenario)
 alphas <- alphas_from_log(meta, log_dir)

### 2017-10-11 appendix/appendixB.Rmd
tweaking fig

--- 2017-10-09 appendix/appendixB.Rmd
@@
 
-             ## FIXME need to get obs for unique commmonname, wrong length?
+             ## avoid NSE
              who <- (ex$commonname == meta[.z,"commonname"]) 
@@
 sims %>%
-  filter(model %in% c("biomass", "catch", "pomdp_0.2", "reed", "pgy")) %>%
+  filter(model %in% c("biomass", "catch", "pomdp_0.1", "reed", "pgy")) %>%
   ggplot(aes(year, stock, col=model)) +
   geom_line(lwd=1) +
-  scale_color_manual(values = c("black", "grey", "#D9661F", "#FDB515", "#6C3302", "#00B0DA", "#CFDD45", "#3B7EA1")) +
+  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
   facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  

### 2017-10-21 appendix/appendix.Rmd
WIP policies

@@
 ```{r aethetics, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE)
+knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE)
 ggplot2::theme_set(ggplot2::theme_bw())
@@
 ```{r helper_routines, include = FALSE}
+## map cts values to nearest grid indices 
+index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
+
 det_simulations <- function(x0 = 1/6, Tmax = 20, 
@@
   pgy_policy <- function(x) F_PGY * x
-  escapement_policy <- function(x){  ## ASSUMES harvest takes place *after* recruitment
-    if(f(x,0) > B_MSY)
-      f(x,0) - B_MSY
-    else
-      0
-  }
+  det_policy <- function(x) pmax(f(x,0) - B_MSY,0) ## ASSUMES harvest takes place *after* recruitment
+  reed_policy  <- function(x) pmax(x - B_MSY,0) ## Optimal when harvest takes place *before* recruitment, assumes we are seeing a pre-harvest size.
+  ## consequently, if we use such a policy when we are seeing a the post-harvest size, we give too much buffer.
 
@@
     for(t in 1:(Tmax-1)){
-      action[t] <- policy(state[t])
-      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
+      obs[t] <- state[t]
+
+      action[t] <- policy(obs[t])
       state[t+1] <- f(state[t], action[t]) 
@@
     list(msy = msy_policy, 
-         det = escapement_policy, 
-         pgy = pgy_policy) %>% 
+         det = det_policy, 
+         pgy = pgy_policy
+         #opt = reed_policy
+         ) %>% 
     map_df(det_sim, 
@@
 
+If we observe->recruit->harvest, $X_{t+1} = f(X_t) - h_t$, then we see the stock right after a harvest and before recuitment, when it is at $X_t ~ $B_{MSY}$.  
+If we instead did observe->harvest->recruit, we would see the stock at its pre-harvest size, $X_t ~ B_{MSY} + $H_{MSY}$.
+
 ```{r}
@@
 ```{r}
-policies <- list(
-  det = map_int(pmax(f(states,0) - S_star,0), ~ which.min(abs(actions - .x))),
-  msy = map_int(states, ~ which.min(abs(actions - .x * F_MSY))),
-  pgy = map_int(states, ~ which.min(abs(actions - .x * F_PGY))) )
+
+policies <- data.frame(
+  det = index(pmax(f(states,0) - S_star,0), actions), # assumes obs->recruit->harvest, f(x_t) - h_t
+ # det = index(pmax(states - S_star,0),     actions), # assumes obs->harvest->recruit, f(x_t - h_t)
+  msy = index(states * F_MSY,               actions),
+  pgy = index(states * F_PGY,               actions))
 ```
@@
 
+Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (deterministic) recruitment would land us right at it $B_{MSY}$.
+This corresponds to constant escapement.
+The discrete grid makes these appear slightly stepped.
+
+Note that the deterministic solution crosses the MSY solution at an observed value of $B_{MSY}$ (i.e.
+$K/2 = 0.5$).
+PGY harvests are always smaller than MSY harvests, but unlike the deterministic optimal solution, PGY and MSY solutions never go to zero.
+
+```{r}
+policy_table <- tibble(state = 1:length(states)) %>%
+  bind_cols(policies) %>%
+  gather(policy, harvest, -state) %>%
+  mutate(harvest = actions[harvest], state = states[state])
+
+policy_table %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+
+```
+
+Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles.
+
+```{r}
+policy_table  %>%
+  mutate(escapement = f(state,harvest))  %>% 
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line() + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+
+```
+
+### Adding the POMDP policy solution
+
 ## Historical data estimation

### 2017-11-30 manuscripts/manuscript.Rmd
WIP
add drone
fix typo

@@
 Below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, a harvest greater than $H_{MSY}$ can be optimal for that year.
-This strategy is known as "constant escapement," since a constant sock size $B_MSY$ escapes harvest each year.
-In practice, it is more common to pursue a policy of constant fishing effort (fishing mortality), $F_MSY = H_MSY / B_MSY$.
-Continuing to harvest at $F_{MSY}$ when a stock is above or below $B_{MSY}$ is economically inefficient and causes depleted stocks to rebuild more slowly, but, in the absense of uncertainty, ensures a return in the long run to stock sizes of $B_MSY$ and harvests of $H_{MSY}$.
+This strategy is known as "constant escapement," since a constant sock size $B_{MSY}$ escapes harvest each year.
+In practice, it is more common to pursue a policy of constant fishing effort (fishing mortality), $F_{MSY} = H_{MSY} / B_{MSY}$.
+Continuing to harvest at $F_{MSY}$ when a stock is above or below $B_{MSY}$ is economically inefficient and causes depleted stocks to rebuild more slowly, but, in the absense of uncertainty, ensures a return in the long run to stock sizes of $B_{MSY}$ and harvests of $H_{MSY}$.
 These differences will be important in understanding how these approaches perform once uncertainty is introduced.
@@
 
+Moderate / modest sigma_g, sigma_m
+
 - non-pomdp policies
@@
 
--------------
-
 ## Discussion
+
+Is it surprising that POMDP model should do better?  Recall that Reed found stochasticity to make no difference, the deterministic model performs identically.
 
@@
 However, the economist's perspective reveals that this may not be as unrealistic as it may first appear.
-
 Fisheries today are not managed at either their maximum ecological or economic potential.  
@@
 constraints such as the cost of idle fleets if a fishery is closed, or include
-the value of ecosystem services from fish not harvested [@Halpern2013].  <!--Alternate utility functions do not increase
+the value of ecosystem services from fish not harvested [@Halpern2013].
+
+<!--Alternate utility functions do not increase
 the complexity of the solution.  Note that fixed cost of a fleet doesn't alter the answer at all:
@@
 -->
+
+-----------------
+<!-- MISC draft text -->
 
@@
 In 1979, a fundamental paper by William Reed [@Reed1979] proved that, under a very general assumptions, the harvest policy for a stochastically fluctuating population was identical to that of the equivalent, deterministic population.
-More precisely, Reed shows that the optimal harvest of the stochastically flutuation population should always aim to reduce the population back to a constant size $S$ (termed by Reed as "escapement," while in most fisheries literature uses the term $B_{\text{MSY}$, the biomass at Maximum Sustainable Yield}).
+More precisely, Reed shows that the optimal harvest of the stochastically flutuation population should always aim to reduce the population back to a constant size $S$ (termed by Reed as "escapement," while in most fisheries literature uses the term $B_{MSY}$, the biomass at Maximum Sustainable Yield}).
 Moreover, this value, $S$, is no more conservative than had we been able to predict future populations with perfect accuracy, as in the determinstic case of constant escapement, $D$.

### 2017-12-13 appendix/appendixB.Rmd
set cores

@@
 ```{r}
-options(mc.cores = 4) # Reserve ~ 10 GB per core
+options(mc.cores = parallel::detectCores()) # Reserve ~ 10 GB per core
 log_dir <- "pomdp_historical"
@@
   left_join(
+    ## use the estimated sigma_m, r, K pars recorded in the logged meta.csv.  Technically
+    ## these should be the same as the above, but without random seed re-running the nimble
+    ## code but not re-running these can create mis-matches.  
     select(meta, sigma_m, commonname, scenario),   
@@
          catch = index(scaled_catch, actions))  %>% 
-  left_join(policy_pars) %>% 
+  left_join(policy_pars) %>% left_join(pars) %>%
   ungroup() 
@@
 ```{r} 
-det_f <- function(S_star, i) index(pmax(f(states,0) - S_star[[1]],0), actions)[i]
+det_f <- function(S_star, r, K, i) index(pmax(gs(r[[1]],K[[1]])(states,0) - S_star[[1]],0), actions)[i]
 msy_f <- function(F_MSY, i) index(states * F_MSY[[1]], actions)[i]
@@
   group_by(commonname) %>% 
-  mutate(det =  det_f(S_star, biomass),
+  mutate(det =  det_f(S_star, r, K, biomass),
          msy =  msy_f(F_MSY, biomass),

### 2017-12-13 appendix/appendix.Rmd
set cores dynamically

@@
 ```{r}
-options(mc.cores=9)
+options(mc.cores=parallel::detectCores())
 ```
@@
 
-Because this solution is computationally somewhat intensive, we provide
+Because this solution is computationally somewhat intensive, be sure to have ~ 4 GB RAM per core if running the 9 models in parallel.
+Alternately, readers can skip the evaluation of this code chunk and read the cached solution from the `policyx` file using the `*_from_log` functions that follow:
 

### 2017-12-13 manuscripts/manuscript.Rmd
separate out old / scratch drafts from current working draft

@@
 Imperfect information is ubiquitous in ecological management and conservation decision making.
-While the pressing concerns of global change have put the spotlight on forecasting, i.e.
-the uncertainty of the future [@Petchy2015], management decisions must also contend with uncertainty of the present: How many fish are in the sea today?
+While the pressing concerns of global change have put the spotlight on *forecasting*, i.e.
+the uncertainty of the future [e.g.
+@Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today?
 What regions harbor the most fragile biodiversity?
@@
 A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.
-In sharp contrast to that shift, the treatment of uncertainty in *decisions* -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.
+In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.
 In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.
+This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.
 
@@
 
-The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters], ecosystem-based management [@Levin2008] to resilience thinking [@Holling; @May] and back to the birth of resource economics [@Beverton; @Scheffer], thanks to its global relevance, long history, and readily available data.
+The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.
 Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.
@@
 While not unknown to the conservation literature [@Williams2009], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
-@Chades, @Fackler, ....].
+@Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
 POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
@@
 - historical data.
+
+```{r}
+```
 
@@
 -->
-
------------------
-<!-- MISC draft text -->
-
-It is precisely this reasoning which explains the broad patterns of recovery
-observed in the landmark paper @Costello2016, which compares future of global
-fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
-and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
-approach should achieve the economically optimal policy of constant-escapement,
-which alone explains why RBFM should see both faster recovery and higher 
-economic yields than MSY, as we have just described. The substantial detail their
-analysis introduces with respect to the dynamics of over 4000 individual stocks
-as well as additional economic knock-on effects from managing based on quotas
-rather than fishing effort permit them to offer quantitatively precise 
-estimates regarding the time required and economic value generated by such
-recovery, but are not essential to understand the broad pattern that suggests
-that nearly all stocks should recover under either an RBFM or MSY-style 
-management, but that the recovery would be fastest and the economic benefits
-greatest under the former.  That conclusion was already implicit in how
-these models are defined.
-
--------------
-
-A fundamental concept to the management of any natural population is the notion of Maximum Sustainable Yield, MSY.
-The notion of MSY is rooted in the determinstic models dating back to foundational work in resource economics and ecological management half a century ago [@Berverton; @Scheffer].
-Though the concept has been the focus of significant criticism for some time [e.g.
-@Botsford_nature_piece] that ctitique has been largely on the "Maximum" part, while the existence of a "Sustainable Yield."
-
-## Reed's Paradox: $S = D$
-
-In 1979, a fundamental paper by William Reed [@Reed1979] proved that, under a very general assumptions, the harvest policy for a stochastically fluctuating population was identical to that of the equivalent, deterministic population.
-More precisely, Reed shows that the optimal harvest of the stochastically flutuation population should always aim to reduce the population back to a constant size $S$ (termed by Reed as "escapement," while in most fisheries literature uses the term $B_{MSY}$, the biomass at Maximum Sustainable Yield}).
-Moreover, this value, $S$, is no more conservative than had we been able to predict future populations with perfect accuracy, as in the determinstic case of constant escapement, $D$.
-A notable exception to this result is when the population is not "self-sustaining," that is,
-the stochasticity is severe enough to pose an extinction risk to the unhavested population.
-In such cases Reed proves there is still a single constant target population size $S$, but that $S > D$, though often by very small amount, reflecting the small probability of a purely chance extinction.
-
-An important corrolary of this result is that suggested a way to side-step the laborious dynamic programming required to find an optimal policy for stochastic populations: the deterministic solution is found trivially from maximum of the stock-recruitment curve.
-Thus much subsequent attention has focused on the estimating stock-recruitment, including methods that address considerable uncertainty in the data used for this estimate [@], while largely ignoring the subsequent step of how to determine an optimal policy in light of that uncertainty.
-
-The lack of any intuitively precautionary response to the introduction of uncertainty did not go unnoticed: in 1986 Clark and Kirkwood first pointed to the absence of measurement error in the formulation used by Reed and others as fundamental to understanding this paradox.
-
-> An important tacit assumption in Reed’s analysis, as in the other works referred
-to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.
-
-Under such an assumption, uncertainty introduced by stochastic growth is effectively purged by each subsequent observation rather than accumulating, and stock extinction can only occur if the manager explicitly decides to harvest the entire stock.
-
-## Clark's Paradox: M < D
-
-Clark and Kirkwood set out to address this problem by considering the case in which stochastic population growth occurs after population size is measured but before harvest, such that manager must act without perfect knowledge of the current system state.
-Unfortunately, this problem proved easier to recognize than to solve, as Clark & Kirkwood recognized at the outset:
-
-> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
-of the problem increases markedly if this assumption is
-relaxed.)
-
-Rather than resolve Reed's Paradox, this assumption compounds the problem, leading to optimal harvests that are generally even higher (i.e.
-less cautious) than the analgous deterministic policy.
-The same counter-intuitive result can also be found in subsequent work [@Sethi2005] which makes a different but equally limiting simplifying assumption (supplement).
-(Both @Clark1986 and @Sethi2005 remark on the counter-intuitive nature of their results)
-
-<!--
-
-Clark: 
-> The above results appear to contradict the conventional wisdom of renewable resource management, under which high uncertainty would call for increased
-caution in the setting of quotas.
-
-Sethi: 
-> It may seem counter-intuitive that a measurement error causes lower expected escapements below the deterministic fishery closure threshold.
-
--->
-
-## Precautionary approaches
-
-Perhaps recognizing
-
-## A solution
-
-Here, we explore how uncertainty over the present impacts decisions
-
-We shall focus on examples from fisheries, which have long been both incubator and proving ground for most of our more general theories of ecological management, from the and
-
-Data are readily avialable and a compartively strong connection between scientific process of stock assessment and the implementation of policy is already established.
-It is also an area which, as we will see, though many models already account for uncertainty, decision strategy still reflects fundamentally more deterministic ideas.
-
-Conclusions:
-
-- failing to account for uncertainty properly can lead to erroneous conclusions
-- Prior beliefs (reflecting prior observations) matter.
-
-<!-- FIXME: In practice, MSY and Reed are very different -->
-
-<!-- Big picture: fisheries is a foundation and testing ground for conservation
-and management. Yet at it's heart lies a controdiction.  -->
-
-<!-- MSY and Constant Escapement -->
-
-Yet fisheries are not managed at their maximum ecological or economic potential.  
-Taken as a whole, most ocean fisheries yield a net negative economic return
-[@FAO2009], while some 68% of global fisheries are not in good biological condition [@Costello2016].
-Nevertheless, fisheries management in both theory and practice are
-
-In a deterministic world at equilibrium (without economic discounting), 
-the policies of constant escapement and maximum
-sustainable yield look identical: the stock is maintained at a fixed stock
-size or biomass, called $B_{MSY}$ in the MSY/fisheries jargon, which is 
-equivalent to the escapement $S^*$ in the optimal control jargon and equal
-to half the carrying capacity, or $K/2$, in the familiar Graham-Schaefer 
-(logistic growth followed by harvest) model. Likewise, the harvest is identical,
-occuring after recruitment and bringing the population back to this optimal
-level; an annual harvest of $H_{MSY} = rK/4$ in Graham-Schafer model for either policy.
-This similarity disguises significant differences in how the policies behave
-when the biomass is either above or below $B_{MSY}$.
-
-<!-- Ways in which constant escapement deviates from MSY -->
-The economically optimal amount to harvest when the population is below 
-$B_{MSY}$ is zero -- no harvest.  This permits the population to return
-as rapidly as possible to the stock size at which it generates the maximum
-yield, and thus the maximum economic return.  This result holds even when
-recruitment is stochastic and thus future stock sizes are uncertain, 
-as rigorously proven by @Reed1979.  Perhaps surprisingly, this is not what
-fisheries management thinks of as a maximum sustainable yield (MSY) policy,
-and does not reflect how fisheries are managed today. Rather, an MSY policy
-is intepreted as fishing at the *effort*, $F_{MSY} = H_{MSY} / B_{MSY}$ that would generate the maximum
-sustainable yield if the population was at the optimal biomass, $B_{MSY}$,
-*regardless* of the current stock size. It is important to realize this policy
-is framed in terms of *effort* $F$: for example, the length of the year open to fishing,
-or equivalently, the per-capita mortality rate of fish, and not in terms of harvest, $H$.
-A population below $B_{MSY}$ that remains harvested at a constant $H_{MSY}$ will eventually be
-driven to extinction, but harvested at a constant $H = F_{MSY} \cdot B$ for $B < B_{MSY}$,
-the stock will eventually recover to $B_{MSY}$.  Consequently, this simple policy should
-converge to the optimal policy in the long run, assuming there is no uncertainty in
-recruitment, measurement, or implementation.
-
-The economically optimal policy of constant-escapement also differs from MSY
-when stocks are above $B_{MSY}$.  As before, the MSY policy continues to assume
-a fixed mortality rate, $F_{MSY}$, which should eventually reduce the stock to
-the level of $B_{MSY}$.  Also as before, the optimal policy is not so patient --
-to derive maximum yield it seeks to return the population to the $B_{MSY}$ not
-*eventually*, but as quickly as possible, and thus will recommend a harvest larger
-than $H_{MSY}$ to accomplish this. These differences become more salient in the case
-of a population under stochastic growth, since the resulting variation leads
-to a population that frequently above or below $B_{MSY}$.  The greater responsiveness
-of the constant-escapement policy to the current stock size thus necessarily 
-results in both the faster recovery of overexploited stocks and higher economic
-returns.
-
-The variation in the harvest required to achieve this constant escapement policy
-can seem impractical from a top-down management perspective.  However, it is 
-important to realize that the application and implications of this policy are
-not purely academic. Economists have for some time observed these inherent 
-inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
-Approaches such as individual transferable quotas (ITQs) would incentivise
-fisheries behavior to converge towards the more dynamic harvests of a constant
-escapement policy, as doing so would increase the value of an individual's share.
-It is precisely this reasoning which explains the broad patterns of recovery
-observed in the landmark paper @Costello2016, which compares future of global
-fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
-and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
-approach should achieve the economically optimal policy of constant-escapement,
-which alone explains why RBFM should see both faster recovery and higher 
-economic yields than MSY, as we have just described. The substantial detail their
-analysis introduces with respect to the dynamics of over 4000 individual stocks
-as well as additional economic knock-on effects from managing based on quotas
-rather than fishing effort permit them to offer quantitatively precise 
-estimates regarding the time required and economic value generated by such
-recovery, but are not essential to understand the broad pattern that suggests
-that nearly all stocks should recover under either an RBFM or MSY-style 
-management, but that the recovery would be fastest and the economic benefits
-greatest under the former.  That conclusion was already implicit in how
-these models are defined.  Unfortunately,
-
--------
-
-Measuring and understanding collapse and over-exploitation of fisheries
-world-wide remains an essential issue for the ecological sciences
-[@Costello2016].  Effective management practices in marine ecosystems also
-inform conservation and resource management in less well-studied
-systems.
-
-<!-- Set up broad relevance of understanding this issue -->
-
-Central to both fisheries research and
-policy, including international agreements such as the United Nations
-Law of the Sea [@UNCLOS1982], is the principle of  maximal sustainable yield (MSY).
-MSY aims to maintain a population at a fixed population size which should result in
-the greatest long-term harvest (often referred to as Biomass at MSY,
-$B_{MSY}$). While the theory dates back to classic work
-of 1950s such as Beverton [@Beverton1957] and Schaefer [@Schaefer1957]
-in deterministic models, it is Reed 1979 paper [@Reed1979] which proves
-that the intuitive result of the deterministic models remains the optimal
-policy for maximizing expected value of a fishery subject to stochastic
-growth (i.e. environmental noise). Reed's result has allowed management
-to largely sidestep the complexity of optimal control of stochastic
-systems.
-
-This has long begged the question: if stochastic growth can be
-safely ignored in determining the optimal harvest, is the same true of
-measurement error?  Thanks to recent computational advances for solving
-the decision theory problem in the case of imperfect observations
-(measurement error), we show that -- contrary to both the intuition
-from Reed's result and previous, approximate solutions -- ignoring
-measurement error can result in significant over-harvesting, decreased
-economic returns, and the potential for complete collapse of the stock.
-
-Some historical perspective is necessary to put this result in the
-appropriate context with respect to both management practice as well
-as previous research in optimal control of fisheries and other natural
-systems under uncertainty.  When population growth
-varies stochastically, any fixed annual harvest such as MSY is no longer optimal:
-instead, management must respond by dynamically harvesting more in years
-of high growth and less in years of low growth.  Reed's proof [@Reed1979]
-nevertheless demonstrated that the optimal harvest can still be captured
-by a simple rule: harvesting just enough to reduce the population back
-down to $B_{MSY}$.  This result is known as a "constant-escapement" policy 
-because the population escaping harvest is always the same, 
-even while the harvest itself varies dynamically with each stochastic shock. 
-Moreover, Reed shows that under some relatively generic assumptions, this escapement
-is identical to that predicted by the deterministic theory  ($B_{MSY}$).
-Because the optimal harvest policy in the deterministic theory follows directly
-from the growth rate (just find the stock size for which recruitment function
-is maximized, no need for dynamic programming), attention shifted to inferring
-that state equation from data.  Our focus is not on inferring this state
-equation, but on the subsequent step of: having done so, how does one determine
-the optimal harvest policy while also accounting for measurement uncertainty? 
-Previous attempts to address this issue over the past three decades
-have relied on various approximations which have usually led to the paradox
-that harvests should increase under additional measurement uncertainty, contrary 
-to our intuition that we should respond to greater uncertainty with greater caution.
-Using more recent optimal control algorithms first developed in the artificial
-intelligence literature, we provide a solution which avoids such assumptions
-and resolves this paradox in favor of common sense intuition.
-
-Despite it's generality, Reed's proof makes a subtle assumption which
-reduces the the impact uncertainty and underlies his conclusion
-that the deterministic intuition of an average MSY is still optimal in 
-the stochastic case.
-In Reed's construction, the manager measures the stock
-size without error during each time step before setting the harvest policy.
-An important consequence of this construction is that this model can
-thus never lead to accidental extinction of the stock.
-Researchers have long recognized the weakness of this assumption of
-perfect knowledge of the stock size and have sought to understand how it
-may impact the resulting management policy [@Ludwig1981].  Because there is no risk of
-accidental extinction in this scenario, it is possible that the optimal
-policy may be more aggressive than is warranted.  Unfortunately, due
-to the complexity of the analysis required to account for measurement
-uncertainty, this question has never been satisfactorily answered.
-Moreover, previous attempts at an approximate solution have consistently
-found the surprising conclusion that the Reed solution is usually less
-aggressive (dictating higher escapement or lower catch quotas) than
-attempts to account for measurement uncertainty.  We briefly review the
-main results in the ecological and resource economics literature that
-have attempted to address this through various approximations before
-presenting our solution, which draws on substantial algorithmic advances
-that have been made in other fields.
-
-The important way in which perfect observations diminish the role of
-uncertainty in the model of Reed [@Reed1979] was first recognized in the classic
-work of Clark & Kirkwood [@Clark1986].  They acknowledge the importance of measurement
-uncertainty to the optimal solution, while observing that if
-this assumption was relaxed, the "difficulty of the problem increases
-markedly." Instead, @Clark1986 consider a perfectly observed model in
-which harvest takes place after stochastic recruitment but before the
-stock is measured, making accidental extinction a possibility
-should realized growth fall sufficiently short of expectation.
-In this case, @Clark1986 demonstrate that the optimal solution deviates
-from the constant escapement rule, with the resulting optimal
-policy is often significantly less cautious than the constant escapement policy.
-In certain cases, @Clark1986 demonstrate that harvesting to extinction
-is actually optimal. Other recent work [@Holden2015; @Johnson2016] has taken 
-a similar approach to @Clark1986 in which measurement becomes uncertain due to
-stochastic stock dynamics occurring between measurement and harvest, which
-can also suggest less conservative (higher harvest) escapement rates.
-
-Rougharden [@Roughgarden1996] take a complementary approach, introducing measurement
-uncertainty but searching only for the optimal constant escapement,
-despite presenting no evidence (and indeed contrary to @Clark1986)
-that the optimal solution under this circumstance should be a
-constant escapement policy. By applying the rule-of-thumb assumption
-that the optimal policy can be expressed simply as a constant escapement
-(or equivalently, target-stock), the set of possible policies is reduced
-dramatically and the calculation is relatively trivial.  As expected,
-this restriction results in a policy that is always more conservative (lower harvest) than
-the Reed solution which ignores measurement error, but both economically
-inefficient and, as we shall see later, still less conservative than
-the truly optimal policy for large stocks.   Other studies 
-have likewise attempted to include uncertainties in model parameters and measurements
-by making this same assumption that a constant escapement solution is optimal and 
-thus avoiding solving the full sequential decision optimization problem [@Ludwig1981; @Moxnes2003; @Engen1997].
-
-Sethi et al. [@Sethi2005] seek to address
-the limitations of @Roughgarden1996 by searching for the true optimal
-policy without the strong assumption of constant escapement.
-Like @Clark1986 and others, @Sethi2005 finds support for the counter-intuitive result
-of an optimal policy that is significantly less cautious when accounting
-for measurement uncertainty than the corresponding constant escapement
-solution which ignores it. 
-Unfortunately, the approach taken by @Sethi2005 once again avoids solving
-the full optimal decision-making problem by means of a simplification
-which we might term the amnesic assumption: during each subsequent
-decision the manager forgets any previous knowledge of the system.
-This allows them to use the standard Stochastic Dynamic Programming
-(SDP) solution method used by @Reed1979 and @Clark1986. This method
-assumes that problem in question is a Markov Decision Process (MDP):
-one in which given the state $x$ and action $a$, the probability of a
-transition to state $x'$ is conditionally independent of all previous
-states and actions (known as the Markov property).  This holds only if the state
-is known without error -- once we introduce any amount of measurement
-uncertainty about the state $x$, the process with respect to the observed
-state is no longer Markovian but depends on all previous observations.
-
-This problem can be solved by considering
-the 'belief' state rather than the observed state, which restores
-the Markov property at the cost of much greater problem complexity:
-one must consider a (continuously-valued) probability
-that the system is in each of the $|X|$ possible states of such a 
-Partially Observed Markov Decision Process (POMDP). This is the
-key feature of the POMDP approach: all previous observations are
-collectively reflected in this belief distribution, and decisions
-are based not only on the most recent observation (as in MDP), but
-are also influenced by this prior belief.
-
-The mechanics of 
-updating these beliefs, while technical, have been well
-illustrated elsewhere even in conservation applications
-[e.g. @Lane1989; @Chades2008; @Haight2010; @Chades2011;
-@Williams2011; @Springborn2013; @Fackler2014; @Fackler22014; @Baggio2016].
-However, these previous studies have relied on relatively direct algorithms 
-which are not amenable to the large number of states and actions required by the optimal
-fisheries harvest problem.  By leveraging newer point-based algorithms developed in the artificial 
-intelligence community [@Kurniawati2008; @Shani2013; see SI; R code], we are able to solve
-the POMDP models for fisheries management directly. 
-These solutions illustrate how incorporating the information
-of all previous observations into a prior belief allows the
-POMDP approach to handle measurement error which would 
-cause a stock collapse under MDP or deterministic methods.
-We also evaluate the performance of proposed method
-on a real-world time series of population estimates of
-marine ecosystems from RAM Legacy Stock
-Assessment Database [@Ricard2011].
-
-## Model formulation
-
-Our formulation of the fisheries management problem closely follows that of 
-earlier work [e.g. @Sethi2005], which we summarize here. Alternate formulations are provided in the SI,
-illustrating that our conclusions are not constrained to these particular choices.
-
-The fisheries manager seeks to maximize the expected net present value of the fishery
-over an infinite time horizon under two sources of uncertainty: population dynamics (i.e. growth of the stock),
-and stock measurements, given an model estimate of those dynamics:
-
-\begin{equation}\label{optim}
-\max_{\{h_t\} \geq 0} \mathbf{E} \left[ \sum_0^{\infty} \gamma^t U(x_t, h_t) \right]
-\end{equation}
-
-The state of the system, $x_t$ corresponds to the (true) fish stock at time $t$. 
-The harvest is $h_t = \min \left(x_t,a_t\right)$, where $a_t$ is the attempted harvest at time step $t$.
-For simplicity, we assume the reward (utility) is directly proportional to the
-harvest, $U(x_t, h_t) = \min(h_t, x_t)$, with future rewards discounted by multiplicative discount factor $\gamma = 0.95$;
-alternate harvest functions and alternate discount rate are considered in the SI.
-
-The population dynamics is formulated as:
-
-\begin{equation}
-x_t = \epsilon_{g} f(s_{t-1})
-\end{equation}
-
-where $s_{t-1} = x_{t-1} - h_{t-1}$ is the escapement at time $t-1$, and
-$\epsilon_{g}$ corresponds to stochastic component of recruitment (growth noise).
-
-While it is theoretically possible to learn about this model and its parameters during the
-decision process itself (so-called adaptive management, e.g. @Johnson2015 for MDP case; 
-we discuss the potential for POMDP adaptive management in *Future Directions*), it is
-typical to consider that state-space dynamics $f$ have been previously estimated from data.  
-The goal of a decision process is to determine the optimal sequence of actions given this
-model. From the population dynamics state
-equation we derive a set of transition matrices $T(x_t, a_t, x_{t+1})$ describing the
-probability of a transition from $x_t$ to $x_{t+1}$ under action $a_t$ (SI Methods).
-While such transition matrices can be defined even when $f$ represents only a posterior
-distribution of model parameter values [e.g. @Boettiger2015] or an age-structured 
-model [e.g. @Holden2015], for simplicity of presentation we will focus on the case
-in which $f$ is a fixed model with known parameters, given by the Ricker equation [@Ricker1954]:
-
-\begin{equation}\label{ricker}
-f(s_t) = s_t \exp\left(r \left(1-\frac{s_t}{K}\right)\right),
-\end{equation}
-
-where $K$ denotes the carrying capacity of the system, and $r$ the growth rate.
-The optimal management of this model in the deterministic and growth-noise-only (MDP)
-case is already well understood and thus will provide a useful foil against
-which to examine the role of measurement uncertainty.  
-To demonstrate that our results are not unique to this choice of \eqref{ricker}, two alternate models,
-the Beverton-Holt model [@Beverton1957], and the Allen model [@Allen2005], are evaluated in the SI.
-
-The POMDP problem is distinguished by the introduction of measurement error, in which 
-the observed stock $z_t$ is measured with uncertainty, $z_t = \epsilon_{m} x_t$. 
-In general this is defined by observation matrices $O(z_t, x_t, a_t)$ giving the 
-probability of observing $z_t$ given a system in state $x_t$ and having taken action $a_t$,
-though in our formulation this probability is independent of the action.
-
-We will assume for simplicity that the stochastic growth noise $\epsilon_g$ and measurement error $\epsilon_m$ 
-are distributed log-normally with mean $1$ and scale parameter $\sigma_{g}$ 
-and $\sigma_m$ respectively. Larger $\sigma$ corresponds to higher variance,
-scaled to be comparable to the uniform noise distribution shown in the SI,
-(see *Materials and Methods*).  We evaluate the effect of
-different noise levels in both growth and measurements, $\sigma_{g},
-\sigma_{m} \in \{0, 0.1, 0.5\}$.
-
-As the numerical solutions of the MDP and POMDP problems require discrete states,
-we consider a system of 50 states on a uniform grid of 0 to 1 with $K = 0.66$ and
-$r = 1$; (alternative growth rates shown in the SI).   
-The transition matrices $T$, together with the observation matrix $O$ and the utility
-function, completely specify a POMDP problem which can be solved with a numerical
-algorithm (*Materials and Methods*).
-
-# Results {#results .unnumbered}
-
-To illustrate the impact of measurement error on management outcomes,
-we compare two scenarios: with measurement error ($\sigma_m = 0.5$,
-see *Materials and Methods*) and without measurement error (Fig \ref{sim}).
-In each scenario, we compute the optimal policy using three different
-solution methods: POMDP, MDP, and a policy we refer to as Offline POMDP.
-In POMDP, the optimal policy at each time point depends not only on the
-most recent observation, but on the manager's prior belief of the state,
-which depends on all prior observations. As such, the optimal POMDP
-policy for a particular observation cannot be computed ahead of time
-but must be calculated "online", as observations come in.  In contrast,
-in MDP, the optimal policy depends only on the most recent observation.
-The "Offline" policy uses the POMDP solution given a fixed, uniform
-prior belief, rather than updating the prior belief over states after
-each observation.  For each scenario (with and without measurement error), 
-we generate 100 simulations of  the fishery being managed under each 
-strategy (MDP, Offline, POMDP) for 100 time steps.  We compute the average and standard
-deviation of the stock size over these 100 replicates, Fig. \ref{sim},
-as well as the mean economic return of each case (Table \ref{Table1})
-and the fraction of simulated stocks that have collapsed by the end
-of the simulation (Table \ref{Table2}). While we perform these
-simulations over a range of models, parameter values and noise distributions, for
-simplicity of presentation we focus on simulations
-from a Ricker model with log-normally distributed noise (see *Materials and Methods*);
-comparable results under the alternative choices are presented in the SI.
-
-\begin{figure}[h]
-\centering
-\caption{Forward simulations of managing a marine ecosystem under POMDP and MDP models in the presence ( $\sigma_m = 0.5$, top panel) or absence ( $\sigma_m = 0$, bottom panel) of measurement error.
-Lines show the mean stock size over 100 replicate simulations, error bands show +/- 1 standard deviation.
-Many of the replicate simulations under MDP quickly collapse the stock by over-harvesting when measurement error is present.
-In contrast, the POMDP management is robust to the level of measurement error, avoiding stock collapse whether or not measurement error is actually present.
-The growth model is Ricker, the noise is log-normally distributed, the carrying capacity of the system is $K = 0.66$, the maximum per capita growth rate is $r = 1$, growth noise is $\sigma_g = 0.5$, and the discount factor is $\gamma = 0.95$.}\label{sim}
-\end{figure}
-
-Fig. \ref{sim}, top panel, illustrates that ignoring or underestimating
-the measurement error result in dramatic collapse of the stock and
-significant loss in revenue: the mean MDP state declines as subsequent
-replicate simulations collapse (94/100 collapse within 100 time steps)
-when measurement error leads to over-harvesting.  The POMDP method
-retains a steady, sustainable harvest that keeps the population below
-carrying capacity but results in minimal population collapse (4/100). The
-offline solution results in some stock collapse (34/100) as a result of
-not updating the manager's belief state over subsequent observations.
-Unsurprisingly, average economic returns are also much lower for MDP
-when measurement error is present, returning an average of 66% of the net present value returned by
-POMDP (Table \ref{Table1}), and would be lower still with less aggressive discounting of future
-rewards (Here $\gamma = 0.95$ makes later collapses less relevant to the total).
-Fig. S1 shows stock trajectories from several individual replicates, which 
-show relatively little decline before collapsing suddenly when managed under MDP
-when measurement error is present.  Averaging over an ensemble of such trajectories
-gives the appearance of a smooth decline.
-
-On the other hand, overestimating measurement error might result in a
-trivial loss in revenue but does not result in collapse of the stock,
-(Fig. \ref{sim}, lower panel). In this scenario, since no
-measurement error is present, the MDP solution is indeed optimal, but the
-POMDP policy which assumes a measurement error of $\sigma_m = 0.5$ and
-thus overestimates the measurement uncertainty performs nearly as well
-(average of 94% of the MDP net present value and likewise no collapses of the stock, Table \ref{Table2}).
-In the SI we report the same patterns for Ricker model
-with uniform noise (Fig. S2), Beverton-Holt model with log-normal noise
-(Fig. S3), and Allen model with log-normal noise (Fig. S4).
-
-<!-- Perhaps in-text quotations of these numbers is sufficient?-->
-
-\begin{table}
-\centering
-\caption{Ecological and economic performance when
-measurement error is present, (\(\sigma_m = 0.5\))}
-\begin{tabular}{lccc}
-& MDP & Offline & POMDP \\
-\midrule
-Fraction collapsed & 94\% & 34\% & 4\% \\
-Economic return & 66\% & 99\% & 100\% \\
-\bottomrule
-\end{tabular}
-\label{Table1}
-\end{table}
-
-\begin{table}
-\centering
-\caption{Ecological and economic performance when
-measurement are perfect, (\(\sigma_m = 0\))}
-\begin{tabular}{lccc}
-& MDP & Offline & POMDP \\
-\midrule
-Fraction collapsed & 0\% & 0\% & 0\% \\
-Economic return & 100\% & 96\% & 94\% \\
-\bottomrule
-\end{tabular}
-\label{Table2}
-\end{table}
-
-To facilitate comparison with previous research and management practice,
-we plot the corresponding optimal action in terms of *expected escapement*: the
-stock the manager expects will be left in the sea after harvest $z_t - h_t$.
-Recall that introducing measurement error means that this value can differ
-from realized escapement, $x_t - h_t$. Fig. \ref{policy} shows the optimal expected 
-escapement as a function of observed stock sizes over a range of growth 
-noise and observation error intensities, assuming a uniform prior belief over
-states $x_t \in (0,1)$ for $\sigma_m > 0$. These policies are determined numerically
-using the SARSOP algorithm for POMDP [ @Kurniawati2008], as discussed in *Materials and Methods* and
-illustrated in the supplementary code provided.
-
-In the MDP case ($\sigma_m = 0$),
-we see the classical result of constant escapement [@Reed1979], where the stock is harvested
-back down to the same escapement regardless of observed stock or growth noise.
-When growth noise is large ($\sigma_g = 0.5$) relative to measurement
-error ($\sigma_m = 0.1$), the policy converges towards the the constant
-optimal escapement solution of @Reed1979, as expected. These patterns
-hold independent of the population growth model (e.g. Beverton-Holt and
-Allen) and the shape of the noise distribution (Fig. S5--S6).
-The most obvious difference between the MDP and POMDP policies arises in the region of
-linear or nearly linear increases in expected escapement occurring at very large observed stock sizes,
-which arises from any policy whose harvest remains *constant* over that range. Not increasing
-harvest in response to higher observed stock sizes always result in a higher expected escapement, $z_t - h_t$.  
-This corresponds to a policy which believes such high observations are almost surely the 
-result of measurement error, and suggests that overestimated stock sizes are responsible for MDP
-over-harvesting and the subsequent collapse of the stock. At smaller stock sizes, the policies
-shown in Fig. \ref{policy} appear more similar over any measurement error, but in fact this is only a consequence of the
-uniform prior belief. While Fig. \ref{policy} completely specifies the management policy under MDP,
-this is not true for the POMDP solutions ($\sigma_m > 0$), which also depend on all
-prior observations through the influence of the belief state.
-
-\begin{figure}[h]
-\centering
-\caption{Optimal escapement policy as a function of measured stock for Ricker model, with different measurement errors and growth noise.
-Noises are log-normally distributed, and the prior belief state is uniform among all stocks.
-The carrying capacity of the system is $K = 0.66$, the growth rate is $r = 1$ and the discount factor is $\gamma = 0.95$.}\label{policy}
-\end{figure}
-
-To illustrate this key difference, we plot (Fig. \ref{bel})
-the POMDP policy under two alternative prior belief states: (1) Low
-Stock: where the prior belief is centered at $x = 0.2$ , and (2) High Stock:
-where the prior belief is centered at the location of carrying capacity $x =
-0.66$. This figure is generated in an identical manner to Fig. \ref{policy} 
-except for the different choice of prior belief.  Visualizations of these 
-prior distributions are shown in Fig. S12. When the prior belief skews towards low stocks, the
-probability that an observation of a high stock is an overestimate due to
-measurement error is greater, and consequently the policy becomes more conservative
-(higher escapement). This ability to incorporate the knowledge from prior
-observations accounts for the difference
-in performance between the POMDP solution and the Offline solution in
-Fig. \ref{sim} when measurement error is present. 
-The evolution of the belief state in response to subsequent observations
-over a given simulation is shown for the different models in Figs S14 - S16.
-Corresponding plots for alternate models are shown in Fig. S9.
-The patterns shown here are further robust to the reward function (including
-linear cost to harvest and a diminishing returns model, Fig. S8) and the
-discount rate Fig. S13.
-
-\begin{figure}[h]
-\centering
-\caption{Optimal escapement policy as a function of measured stock and different prior belief states: (1) Uniform: uniform prior belief over all stocks, (2) Low Stock: the prior belief is centered at $x = 0.2$, and (3) High Stock: the prior belief is centered at the location of carrying capacity $x =
-0.66$.
-In this figure, growth model is Ricker, measurement and growth noises are $\sigma_m = \sigma_g = 0.5$, and are log-normally distributed, the growth rate is $r = 1$, and the discount factor is $\gamma = 0.95$.}\label{bel}
-\end{figure}
-
-## Application to real-world fisheries
-
-Using historical time series of Argentine Hake on the Patagonian Shelf and
-European Plaice on the Celtic-Biscay Shelf from the the RAM Legacy Stock
-Assessment database [@Ricard2011], we compare the catch levels recommended
-by POMDP and MDP methods against historical values (Fig. \ref{RAM2}). We have selected
-these two examples for having both adequately long records and show very
-different trends in both population biomass and catch. We non-dimensionalize
-the data and then use Maximum
-Likelihood Estimation (MLE) through NIMBLE [@nimble-software:2016] to
-fit a Ricker model for both these examples (details of MLE fits can be
-found in the SI). Based on this model estimate, we then determine MDP
-and POMDP policies using the appropriate algorithms.  In the case of MDP,
-the optimal harvest is determined directly from that year's estimated
-stock abundance.  In the case of POMDP, the optimal harvest depends both
-on this stock estimate and the prior belief, which is sequentially updated
-based on each previous observation (stock estimate) and action (catch).
-
-Both examples illustrate the contrast between the ignoring (MDP) and accounting for measurement error (POMDP).
-The POMDP solution tends to recommend a lower harvest than the MDP solution
-in both cases, but not strictly so.  In Plaice example, we see the the POMDP solution is less 
-volatile, increasing harvest more modestly in response to observing stock 
-sizes while fishing pressure remains constant or increases (e.g. 1980-1990), 
-while also decreasing less if stocks decline under constant or decreasing fishing
-pressure (e.g. 1990-2000). Under the MDP model, only sharp changes in catch
-can drive these sharp shifts in stock levels, while the POMDP model permits that these
-can arise in part by measurement error and thus require less volatility in the harvest. 
-In the Hake time series, the continued decline of the stock leads both MDP
-and POMDP solutions to recommend smaller harvests than historically observed.
-Again the POMDP solution is less volatile than the MDP, starting with a more
-modest harvest at the beginning of the time-series (where a uniform prior over
-states is assumed), and at first decreasing more gradually but then retaining
-the moratorium on fishing when a small rebound (circa 2005) leads the MDP to a
-renewed harvest.
-
-\begin{figure}[h]
-\centering
-\caption{This figure compares the actual stock biomass and catch of European Plaice on the Celtic-Biscay Shelf and Argentine Hake on the Patagonian Shelf with actions suggested by MDP and POMDP frameworks.
-In the case of POMDP we have used measurement error of $\sigma_m = 0.5$.}\label{RAM2}
-\end{figure}
-
-# Discussion {#discussion .unnumbered}
-
-We have shown that fully accounting for measurement error in calculating
-the optimal fishing harvest leads to management that is usually more conservative
-(higher expected escapement) as the measurement error increases. This
-contrasts with the classic solution of constant escapement, in which
-larger estimates of the stock size justify proportionally larger harvests.
-The importance of our intuitively sensible conclusion is underscored
-by the contrast to both much of current fisheries management practice
-and previous research into measurement uncertainty.  Most fishing policy
-still reflects the results @Reed1979 driven in
-the absence of measurement error, and even much previous work on the role
-of measurement uncertainty  has either assumed constant escapement-type
-solutions [@Roughgarden1996; @Moxnes2003; @Engen1997] or relied on approximations which
-have argued that accounting for measurement uncertainty should result
-in an *increased* harvest (lower target stock size) [@Ludwig1981; @Clark1986; @Sethi2005].
-
-Using simulations (Fig. \ref{sim}, S2-S4), we have shown that failing to account for measurement error
-can lead to dramatic ecological and economic consequences. MDP solutions, which
-ignore measurement error, collapse 94% of simulated time series within 100 time
-steps. This does not mean that measurement error in the simulations is so large
-as to make informed management impossible -- in fact the optimal POMDP solution,
-which accounts for measurement error, experiences a collapse in only 4% of
-simulations shown here. While the precise degree of measurement uncertainty may
-itself be unknown, these results also show that it is better to overestimate
-than underestimate this uncertainty. Even when no measurement error is present,
-a POMDP solution which assumes a large measurement error ($\sigma_m = 0.5$) performs
-nearly as well as the optimal MDP solution in all of our simulations. Fig. \ref{policy}
-and \ref{bel} provide the intuition behind why MDP approach results in such a high probability
-of collapse: when measurement error results in a chance over-estimation of the 
-stock, the resulting harvest under MDP can be much larger than under POMDP. A single
-large or a few consecutive smaller over-estimations of the stock size can thus 
-quickly drive the stock to zero.  In contrast, the POMDP solution will harvest
-much less under the same observation error, particularly when it compares that
-observation against the accumulated prior belief that the current state should
-be lower, Fig \ref{bel}.
-
-We have shown that this dramatic reduction in the probability of a 
-collapse is not trivially the result of a more cautious strategy; e.g. one which always
-harvests less than MDP.  Rather, it is the result of a making full use
-of all of the data through the role of the belief state.
-The POMDP solution, unlike the MDP solution, depends on
-the entire history of previous observations, which inform the current
-belief state.  As shown in Fig \ref{bel}, this can correspond
-to higher harvests if prior observations give sufficient evidence that 
-the stock is higher than suggested by the most recent measurement. As a result,
-the POMDP policy can outperform the MDP policy economically as well as 
-ecologically.
-
-We have also demonstrated these patterns by comparing harvests recommended
-by MDP methods against those recommended by POMDP methods using two separate
-historical time series.  These examples provide a further illustration
-of how accounting for measurement uncertainty results in very different policy
-outcomes than ignoring it.  Note that this analysis does not seek to explain
-the differences between either the MDP and POMDP solutions from the historical
-catches actually observed.  While it is very possible that the observed harvests
-were governed by some sort of constant-escapement policy, actual catch quotas
-or other policy information is not available in this data. Fluctuations in historical
-harvest may reflect many other factors than such a policy, such as different stock
-models, external constraints such as the cost of adjusting policy [@Boettiger2016], 
-or other economic, political or environmental factors.  While we are free to
-make comparisons between MDP and POMDP solutions on the historical data, understanding
-why historical harvests differed from either of these would need an understanding of 
-these contextual issues that are beyond our data and our scope here.
-Given the stock-recruitment model for the fishery, the MDP optimum harvest depends only
-on the most recent stock estimate, while both the full time series of historical harvests
-and historical stock estimates influence the POMDP optimum, by altering the belief
-state at the time.  We have
-relied on estimating reasonable candidate models and parameters from the time series
-a priori (specifically MLE parameters to the Ricker model). Ideally model
-and parameters would also be inferred over the course of the sequential decisions,
-which would integrate over such uncertainty in models or parameters.
-While MDP methods to address these issues of model uncertainty
-and parameter uncertainty have been developed elsewhere [e.g. @Johnson2015] this remains
-a subject for future research in POMDP applications (see *Future directions*).
-
-We have shown that our result is generally robust to our assumptions:
-this pattern holds for differing models (Ricker, Beverton-Holt, and Allen),
-differing magnitudes and also different shapes of noise distributions
-(log-normal, as considered by @Clark1986 or uniform, as considered by
-@Sethi2005), the choice of the reward/profit equation, or the details of
-the discretization (see SI).  This robustness is not surprising, being consistent
-with sensitivity observed in previous work.  Indeed, in the absence of
-measurement error most of these scenarios were already covered by Reed's
-proof, (with the exception of the Allen model whose change in concavity
-violates the theorem and likewise shows deviations here).
-Thus it is perhaps all the more surprising that
-constant escapement is not robust to the assumption of measurement error.
-
-The optimization approach considered here can be distinguished from more heuristic
-approaches to address the uncertainty, such as pretty good yield (PGY)[@Hilborn2010; @Rindorf2016] or management strategy evaluation (MSE).
-Methods based on MSE have been applied in fishery management for providing the decision maker with a rationale to base their decisions given their specific optimization objectives [@Smith1994; @Mapstone2008; @Bunnefeld2011].
-The main difference between MSE and methods based on dynamic programming is that MSE does not solve the optimization problem to provide the optimal strategy, rather it evaluates consequences of a set of strategies defined based on management objectives and provides the outcome of those simulations to the decision maker.
-In that sense, MSE can be compared to the policy evaluation step of optimization in dynamic programming using policy iteration methods [@Sutton1998], while the policy improvement step is being ignored due to computational complexity.
-With advancements in efficient solving of dynamic programming optimization under several sources of uncertainty [@Shani2013], the traditional MSE approach can be interpreted as an internal step in the optimization process that tries to identify the best management strategy, searching through all viable and possible strategies.
-MSE can be also compared to Monte Carlo search algorithms within the space of possible policies that the manager can explore and evaluating the policies based on the outcome [@Smith1999].
-For example, @Smith1999 evaluates the effect of several sources of uncertainty such as structural and parameter uncertainty in the models, errors in data and observation systems, estimation uncertainty, and management implementation uncertainty in partnership with Australian Fisheries Management Authority.
-
-Our results have important implications for management of fisheries.
-The most evident of these is that given the unavoidable realities of measurement
-error in stock estimates, the wide use of management based on constant-escapement
-solutions and derivative concepts such as maximum sustainable yield (MSY), maximum
-sustainable biomass ($B_{MSY}$),
-which are central to scientific fisheries management and policy
-included the United Nations Law of the Sea [@UNCLOS1982], are insufficient to
-guarantee either maximum sustainable economic yield or persistence of fish stocks.
-Instead, target harvest or target escapement must reflect current and previous 
-stock estimates and the degree of measurement uncertainty present. While precise 
-estimates of the uncertainty in measurements, $\sigma_m$ may be difficult to obtain,
-we have shown that it better both ecologically (in the probability of stock collapse)
-and economically (in the profits from fishing) to overestimate than underestimate this
-quantity.
-
-This important qualitative impact that measurement error has on the
-optimal solution of an important and well-studied ecological management
-problem has more general implications.  The role of measurement error
-is most often acknowledged in the breach -- a mathematically
-convenient assumption only.  After all, inherent stochasticity has often been found
-to make only small differences from deterministic solutions, as @Reed1979
-proved in this case. Our result demonstrates how frail such an assumption
-regarding measurement uncertainty may be, particularly when determining
-optimal policy.  No doubt the mathematical and computational difficulty of
-accounting for measurement uncertainty has forestalled greater attention
-to this issue.  The growing wave of available data makes such concerns
-about measurement error all the more pressing, as imperfect but readily
-available observations make such data-driven optimization possible in more
-applications.
-
-These results have both tactical and strategic implications for ecological 
-modeling and fisheries management [sensu @May1973]. At a strategic level,
-this highlights the danger in ignoring the potentially asymmetric impact of
-measurement error, which can lead to significant over-harvesting and potential
-collapse of the stock.  At a tactical level, the precise amount by which
-harvest should be adjusted to account for a given measurement error to maximize
-the expected value of a stock will require a POMDP solution for the specific 
-fishery in question, which is now computationally feasible using the approach
-we have presented here. We hope that the recent advances in efficient and accurate
-POMDP solvers, together with greater attention to these challenges,
-will encourage a wider consideration of the role of measurement error
-in determining appropriately precautionary management policy.
-
-In conclusion, we have demonstrated the importance of measurement error
-in the long-standing question of optimal harvests in fisheries.  We have 
-shown that overestimating the measurement error can still result in policies
-that are nearly optimal both economically and ecologically, while ignoring or
-underestimating the measurement noise can cause dramatic collapses of fish
-stocks even when no tipping point is present.  This suggests that the failure
-to account for measurement error in this way in even the most advanced 
-fisheries management could contribute to long-term trends in declining
-stocks [e.g. @Costello2016]. We believe the approach taken
-here both highlights the importance of accounting for measurement error and
-provides an example of how this can be accomplished in the POMDP framework
-using modern algorithms on complex problems. We have also implemented the
-proposed method on a real-world problem of managing a marine ecosystem and show
-that the policy suggested by POMDP is significantly different with respect to the
-actual catch landings.
-
-## Future directions
-
-Several limitations that have been studied in MDP models, such as parameter 
-and model uncertainty and adaptive management [e.g. @Johnson2015] remain
-intriguing future challenges for POMDP problems. Future work could extend the examples
-shown here to include more realistic models, such as those with age structure, as
-has only recently been done for the MDP case [@Holden2015].
-There remains a need for further improvement of the proposed method to include such model uncertainty. 
-Previous attempts that adapt POMDP tools to include model uncertainty [i.e. @Chades2012] have had to
-assume full observability of state space, ignoring measurement uncertainty. 
-A second limiting assumption common to MDPs and POMDPs is that of stationary dynamics: 
-that the population dynamics equation itself is not changing over
-time. In reality, forces such as climate change and other forms of environmental variations violate this
-assumption. @Fackler22014 attempts to address this problem by 
-adaptation of Mixed Observability MDP [@Ong2010] framework, although the proposed algorithm
-for finding optimal solution to such non-stationary systems is intractable, except for 
-extremely simplified examples used in that study. Our future work involve relaxing
-these assumptions by proposing computationally feasible approximate algorithms within the 
-POMDP framework to deal both with model uncertainty and non-stationarity of system's dynamics.
-Lastly, the introduction of POMDP approaches raises a central question for managers: when is it
-worthwhile to actively reduce measurement error to derive a better policy, and when is it
-best simply to account for current measurement error?  This issue can be addressed through
-a Value of Information framework, as @Johnson2015 illustrates in the case of MDPs.
-
-# Materials and Methods {#methods .unnumbered}
-
-A Partially Observed Markov Decision Process (POMDP) [@Smallwood1973;
-@Sondik1978] is a generalization of a Markov Decision Process (MDP),
-which is a fundamental model for sequential decision making.  In the
-ecological literature, MDPs are often referred to by their solution
-method: Stochastic Dynamic Programming [SDP see @Marescot2013].
-The MDP approach has been widely used in behavioral ecology
-[@Mangel1988] as well as conservation ecology and resource economics
-[e.g. @Mangel1985]. Introductions to MDP and solving them through dynamic
-programming are provided in @Mangel1985, @Bertsekas1996 and @Sutton1998.
-Specifically, POMDPs overcome one of the main limitations of MDP which
-is the assumption about the full observability of the system's state.
-In POMDP, the exact state of the system cannot be
-observed directly, but can be inferred by indirect measurements.
-POMDPs can be represented as a probabilistic graphical models
-[@Barber2012] as in Fig. S11.  For further details regard POMDP framework,
-its parameters, and details of solving the optimization problem,
-refer to the SI.
-
-Our approach utilizes the SARSOP algorithm of @Kurniawati2008. Computational 
-requirements are still significantly more demanding than the equivalent MDP solution.
-Precise computational time and memory requirements depend not only on system size but on
-the precision requested of the approximate, point-based solution. Reasonable
-solutions for the problems considered here can be reached within an hour of computation,
-though were run for five hours each to confirm convergence. 
-As a supplement to this paper we provide example scripts for replicating
-the analyses shown here, which can be found at https://github.com/boettiger-lab/sarsop.

### 2017-12-13 manuscripts/scratch-early-revisions.Rmd
separate out old / scratch drafts from current working draft

Yet fisheries are not managed at their maximum ecological or economic potential.  
Taken as a whole, most ocean fisheries yield a net negative economic return
[@FAO2009], while some 68% of global fisheries are not in good biological condition [@Costello2016].
Nevertheless, fisheries management in both theory and practice are

In a deterministic world at equilibrium (without economic discounting), 
the policies of constant escapement and maximum
sustainable yield look identical: the stock is maintained at a fixed stock
size or biomass, called $B_{MSY}$ in the MSY/fisheries jargon, which is 
equivalent to the escapement $S^*$ in the optimal control jargon and equal
to half the carrying capacity, or $K/2$, in the familiar Graham-Schaefer 
(logistic growth followed by harvest) model. Likewise, the harvest is identical,
occuring after recruitment and bringing the population back to this optimal
level; an annual harvest of $H_{MSY} = rK/4$ in Graham-Schafer model for either policy.
This similarity disguises significant differences in how the policies behave
when the biomass is either above or below $B_{MSY}$.

The economically optimal amount to harvest when the population is below 
$B_{MSY}$ is zero -- no harvest.  This permits the population to return
as rapidly as possible to the stock size at which it generates the maximum
yield, and thus the maximum economic return.  This result holds even when
recruitment is stochastic and thus future stock sizes are uncertain, 
as rigorously proven by @Reed1979.  Perhaps surprisingly, this is not what
fisheries management thinks of as a maximum sustainable yield (MSY) policy,
and does not reflect how fisheries are managed today. Rather, an MSY policy
is intepreted as fishing at the *effort*, $F_{MSY} = H_{MSY} / B_{MSY}$ that would generate the maximum
sustainable yield if the population was at the optimal biomass, $B_{MSY}$,
*regardless* of the current stock size. It is important to realize this policy
is framed in terms of *effort* $F$: for example, the length of the year open to fishing,
or equivalently, the per-capita mortality rate of fish, and not in terms of harvest, $H$.
A population below $B_{MSY}$ that remains harvested at a constant $H_{MSY}$ will eventually be
driven to extinction, but harvested at a constant $H = F_{MSY} \cdot B$ for $B < B_{MSY}$,
the stock will eventually recover to $B_{MSY}$.  Consequently, this simple policy should
converge to the optimal policy in the long run, assuming there is no uncertainty in
recruitment, measurement, or implementation.

The economically optimal policy of constant-escapement also differs from MSY
when stocks are above $B_{MSY}$.  As before, the MSY policy continues to assume
a fixed mortality rate, $F_{MSY}$, which should eventually reduce the stock to
the level of $B_{MSY}$.  Also as before, the optimal policy is not so patient --
to derive maximum yield it seeks to return the population to the $B_{MSY}$ not
*eventually*, but as quickly as possible, and thus will recommend a harvest larger
than $H_{MSY}$ to accomplish this. These differences become more salient in the case
of a population under stochastic growth, since the resulting variation leads
to a population that frequently above or below $B_{MSY}$.  The greater responsiveness
of the constant-escapement policy to the current stock size thus necessarily 
results in both the faster recovery of overexploited stocks and higher economic
returns.

The variation in the harvest required to achieve this constant escapement policy
can seem impractical from a top-down management perspective.  However, it is 
important to realize that the application and implications of this policy are
not purely academic. Economists have for some time observed these inherent 
inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
Approaches such as individual transferable quotas (ITQs) would incentivise
fisheries behavior to converge towards the more dynamic harvests of a constant
escapement policy, as doing so would increase the value of an individual's share.
It is precisely this reasoning which explains the broad patterns of recovery
observed in the landmark paper @Costello2016, which compares future of global
fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
approach should achieve the economically optimal policy of constant-escapement,
which alone explains why RBFM should see both faster recovery and higher 
economic yields than MSY, as we have just described. The substantial detail their
analysis introduces with respect to the dynamics of over 4000 individual stocks
as well as additional economic knock-on effects from managing based on quotas
rather than fishing effort permit them to offer quantitatively precise 
estimates regarding the time required and economic value generated by such
recovery, but are not essential to understand the broad pattern that suggests
that nearly all stocks should recover under either an RBFM or MSY-style 
management, but that the recovery would be fastest and the economic benefits
greatest under the former.  That conclusion was already implicit in how
these models are defined.

### 2017-12-13 manuscripts/scratch-original.Rmd
separate out old / scratch drafts from current working draft

Measuring and understanding collapse and over-exploitation of fisheries
world-wide remains an essential issue for the ecological sciences
[@Costello2016].  Effective management practices in marine ecosystems also
inform conservation and resource management in less well-studied
systems.

<!-- Set up broad relevance of understanding this issue -->

Central to both fisheries research and
policy, including international agreements such as the United Nations
Law of the Sea [@UNCLOS1982], is the principle of  maximal sustainable yield (MSY).
MSY aims to maintain a population at a fixed population size which should result in
the greatest long-term harvest (often referred to as Biomass at MSY,
$B_{MSY}$). While the theory dates back to classic work
of 1950s such as Beverton [@Beverton1957] and Schaefer [@Schaefer1957]
in deterministic models, it is Reed 1979 paper [@Reed1979] which proves
that the intuitive result of the deterministic models remains the optimal
policy for maximizing expected value of a fishery subject to stochastic
growth (i.e. environmental noise). Reed's result has allowed management
to largely sidestep the complexity of optimal control of stochastic
systems.

This has long begged the question: if stochastic growth can be
safely ignored in determining the optimal harvest, is the same true of
measurement error?  Thanks to recent computational advances for solving
the decision theory problem in the case of imperfect observations
(measurement error), we show that -- contrary to both the intuition
from Reed's result and previous, approximate solutions -- ignoring
measurement error can result in significant over-harvesting, decreased
economic returns, and the potential for complete collapse of the stock.

Some historical perspective is necessary to put this result in the
appropriate context with respect to both management practice as well
as previous research in optimal control of fisheries and other natural
systems under uncertainty.  When population growth
varies stochastically, any fixed annual harvest such as MSY is no longer optimal:
instead, management must respond by dynamically harvesting more in years
of high growth and less in years of low growth.  Reed's proof [@Reed1979]
nevertheless demonstrated that the optimal harvest can still be captured
by a simple rule: harvesting just enough to reduce the population back
down to $B_{MSY}$.  This result is known as a "constant-escapement" policy 
because the population escaping harvest is always the same, 
even while the harvest itself varies dynamically with each stochastic shock. 
Moreover, Reed shows that under some relatively generic assumptions, this escapement
is identical to that predicted by the deterministic theory  ($B_{MSY}$).
Because the optimal harvest policy in the deterministic theory follows directly
from the growth rate (just find the stock size for which recruitment function
is maximized, no need for dynamic programming), attention shifted to inferring
that state equation from data.  Our focus is not on inferring this state
equation, but on the subsequent step of: having done so, how does one determine
the optimal harvest policy while also accounting for measurement uncertainty? 
Previous attempts to address this issue over the past three decades
have relied on various approximations which have usually led to the paradox
that harvests should increase under additional measurement uncertainty, contrary 
to our intuition that we should respond to greater uncertainty with greater caution.
Using more recent optimal control algorithms first developed in the artificial
intelligence literature, we provide a solution which avoids such assumptions
and resolves this paradox in favor of common sense intuition.

Despite it's generality, Reed's proof makes a subtle assumption which
reduces the the impact uncertainty and underlies his conclusion
that the deterministic intuition of an average MSY is still optimal in 
the stochastic case.
In Reed's construction, the manager measures the stock
size without error during each time step before setting the harvest policy.
An important consequence of this construction is that this model can
thus never lead to accidental extinction of the stock.
Researchers have long recognized the weakness of this assumption of
perfect knowledge of the stock size and have sought to understand how it
may impact the resulting management policy [@Ludwig1981].  Because there is no risk of
accidental extinction in this scenario, it is possible that the optimal
policy may be more aggressive than is warranted.  Unfortunately, due
to the complexity of the analysis required to account for measurement
uncertainty, this question has never been satisfactorily answered.
Moreover, previous attempts at an approximate solution have consistently
found the surprising conclusion that the Reed solution is usually less
aggressive (dictating higher escapement or lower catch quotas) than
attempts to account for measurement uncertainty.  We briefly review the
main results in the ecological and resource economics literature that
have attempted to address this through various approximations before
presenting our solution, which draws on substantial algorithmic advances
that have been made in other fields.

The important way in which perfect observations diminish the role of
uncertainty in the model of Reed [@Reed1979] was first recognized in the classic
work of Clark & Kirkwood [@Clark1986].  They acknowledge the importance of measurement
uncertainty to the optimal solution, while observing that if
this assumption was relaxed, the "difficulty of the problem increases
markedly." Instead, @Clark1986 consider a perfectly observed model in
which harvest takes place after stochastic recruitment but before the
stock is measured, making accidental extinction a possibility
should realized growth fall sufficiently short of expectation.
In this case, @Clark1986 demonstrate that the optimal solution deviates
from the constant escapement rule, with the resulting optimal
policy is often significantly less cautious than the constant escapement policy.
In certain cases, @Clark1986 demonstrate that harvesting to extinction
is actually optimal. Other recent work [@Holden2015; @Johnson2016] has taken 
a similar approach to @Clark1986 in which measurement becomes uncertain due to
stochastic stock dynamics occurring between measurement and harvest, which
can also suggest less conservative (higher harvest) escapement rates.

Rougharden [@Roughgarden1996] take a complementary approach, introducing measurement
uncertainty but searching only for the optimal constant escapement,
despite presenting no evidence (and indeed contrary to @Clark1986)
that the optimal solution under this circumstance should be a
constant escapement policy. By applying the rule-of-thumb assumption
that the optimal policy can be expressed simply as a constant escapement
(or equivalently, target-stock), the set of possible policies is reduced
dramatically and the calculation is relatively trivial.  As expected,
this restriction results in a policy that is always more conservative (lower harvest) than
the Reed solution which ignores measurement error, but both economically
inefficient and, as we shall see later, still less conservative than
the truly optimal policy for large stocks.   Other studies 
have likewise attempted to include uncertainties in model parameters and measurements
by making this same assumption that a constant escapement solution is optimal and 
thus avoiding solving the full sequential decision optimization problem [@Ludwig1981; @Moxnes2003; @Engen1997].

Sethi et al. [@Sethi2005] seek to address
the limitations of @Roughgarden1996 by searching for the true optimal
policy without the strong assumption of constant escapement.
Like @Clark1986 and others, @Sethi2005 finds support for the counter-intuitive result
of an optimal policy that is significantly less cautious when accounting
for measurement uncertainty than the corresponding constant escapement
solution which ignores it. 
Unfortunately, the approach taken by @Sethi2005 once again avoids solving
the full optimal decision-making problem by means of a simplification
which we might term the amnesic assumption: during each subsequent
decision the manager forgets any previous knowledge of the system.
This allows them to use the standard Stochastic Dynamic Programming
(SDP) solution method used by @Reed1979 and @Clark1986. This method
assumes that problem in question is a Markov Decision Process (MDP):
one in which given the state $x$ and action $a$, the probability of a
transition to state $x'$ is conditionally independent of all previous
states and actions (known as the Markov property).  This holds only if the state
is known without error -- once we introduce any amount of measurement
uncertainty about the state $x$, the process with respect to the observed
state is no longer Markovian but depends on all previous observations.

This problem can be solved by considering
the 'belief' state rather than the observed state, which restores
the Markov property at the cost of much greater problem complexity:
one must consider a (continuously-valued) probability
that the system is in each of the $|X|$ possible states of such a 
Partially Observed Markov Decision Process (POMDP). This is the
key feature of the POMDP approach: all previous observations are
collectively reflected in this belief distribution, and decisions
are based not only on the most recent observation (as in MDP), but
are also influenced by this prior belief.

The mechanics of 
updating these beliefs, while technical, have been well
illustrated elsewhere even in conservation applications
[e.g. @Lane1989; @Chades2008; @Haight2010; @Chades2011;
@Williams2011; @Springborn2013; @Fackler2014; @Fackler22014; @Baggio2016].
However, these previous studies have relied on relatively direct algorithms 
which are not amenable to the large number of states and actions required by the optimal
fisheries harvest problem.  By leveraging newer point-based algorithms developed in the artificial 
intelligence community [@Kurniawati2008; @Shani2013; see SI; R code], we are able to solve
the POMDP models for fisheries management directly. 
These solutions illustrate how incorporating the information
of all previous observations into a prior belief allows the
POMDP approach to handle measurement error which would 
cause a stock collapse under MDP or deterministic methods.
We also evaluate the performance of proposed method
on a real-world time series of population estimates of
marine ecosystems from RAM Legacy Stock
Assessment Database [@Ricard2011].

## Model formulation

Our formulation of the fisheries management problem closely follows that of 
earlier work [e.g. @Sethi2005], which we summarize here. Alternate formulations are provided in the SI,
illustrating that our conclusions are not constrained to these particular choices.

The fisheries manager seeks to maximize the expected net present value of the fishery
over an infinite time horizon under two sources of uncertainty: population dynamics (i.e. growth of the stock),
and stock measurements, given an model estimate of those dynamics:

\begin{equation}\label{optim}
\max_{\{h_t\} \geq 0} \mathbf{E} \left[ \sum_0^{\infty} \gamma^t U(x_t, h_t) \right]
\end{equation}

The state of the system, $x_t$ corresponds to the (true) fish stock at time $t$. 
The harvest is $h_t = \min \left(x_t,a_t\right)$, where $a_t$ is the attempted harvest at time step $t$.
For simplicity, we assume the reward (utility) is directly proportional to the
harvest, $U(x_t, h_t) = \min(h_t, x_t)$, with future rewards discounted by multiplicative discount factor $\gamma = 0.95$;
alternate harvest functions and alternate discount rate are considered in the SI.

The population dynamics is formulated as:

\begin{equation}
x_t = \epsilon_{g} f(s_{t-1})
\end{equation}

where $s_{t-1} = x_{t-1} - h_{t-1}$ is the escapement at time $t-1$, and
$\epsilon_{g}$ corresponds to stochastic component of recruitment (growth noise).

While it is theoretically possible to learn about this model and its parameters during the
decision process itself (so-called adaptive management, e.g. @Johnson2015 for MDP case; 
we discuss the potential for POMDP adaptive management in *Future Directions*), it is
typical to consider that state-space dynamics $f$ have been previously estimated from data.  
The goal of a decision process is to determine the optimal sequence of actions given this
model. From the population dynamics state
equation we derive a set of transition matrices $T(x_t, a_t, x_{t+1})$ describing the
probability of a transition from $x_t$ to $x_{t+1}$ under action $a_t$ (SI Methods).
While such transition matrices can be defined even when $f$ represents only a posterior
distribution of model parameter values [e.g. @Boettiger2015] or an age-structured 
model [e.g. @Holden2015], for simplicity of presentation we will focus on the case
in which $f$ is a fixed model with known parameters, given by the Ricker equation [@Ricker1954]:

\begin{equation}\label{ricker}
f(s_t) = s_t \exp\left(r \left(1-\frac{s_t}{K}\right)\right),
\end{equation}

where $K$ denotes the carrying capacity of the system, and $r$ the growth rate.
The optimal management of this model in the deterministic and growth-noise-only (MDP)
case is already well understood and thus will provide a useful foil against
which to examine the role of measurement uncertainty.  
To demonstrate that our results are not unique to this choice of \eqref{ricker}, two alternate models,
the Beverton-Holt model [@Beverton1957], and the Allen model [@Allen2005], are evaluated in the SI.

The POMDP problem is distinguished by the introduction of measurement error, in which 
the observed stock $z_t$ is measured with uncertainty, $z_t = \epsilon_{m} x_t$. 
In general this is defined by observation matrices $O(z_t, x_t, a_t)$ giving the 
probability of observing $z_t$ given a system in state $x_t$ and having taken action $a_t$,
though in our formulation this probability is independent of the action.

We will assume for simplicity that the stochastic growth noise $\epsilon_g$ and measurement error $\epsilon_m$ 
are distributed log-normally with mean $1$ and scale parameter $\sigma_{g}$ 
and $\sigma_m$ respectively. Larger $\sigma$ corresponds to higher variance,
scaled to be comparable to the uniform noise distribution shown in the SI,
(see *Materials and Methods*).  We evaluate the effect of
different noise levels in both growth and measurements, $\sigma_{g},
\sigma_{m} \in \{0, 0.1, 0.5\}$.

As the numerical solutions of the MDP and POMDP problems require discrete states,
we consider a system of 50 states on a uniform grid of 0 to 1 with $K = 0.66$ and
$r = 1$; (alternative growth rates shown in the SI).   
The transition matrices $T$, together with the observation matrix $O$ and the utility
function, completely specify a POMDP problem which can be solved with a numerical
algorithm (*Materials and Methods*).

# Results {#results .unnumbered}

To illustrate the impact of measurement error on management outcomes,
we compare two scenarios: with measurement error ($\sigma_m = 0.5$,
see *Materials and Methods*) and without measurement error (Fig \ref{sim}).
In each scenario, we compute the optimal policy using three different
solution methods: POMDP, MDP, and a policy we refer to as Offline POMDP.
In POMDP, the optimal policy at each time point depends not only on the
most recent observation, but on the manager's prior belief of the state,
which depends on all prior observations. As such, the optimal POMDP
policy for a particular observation cannot be computed ahead of time
but must be calculated "online", as observations come in.  In contrast,
in MDP, the optimal policy depends only on the most recent observation.
The "Offline" policy uses the POMDP solution given a fixed, uniform
prior belief, rather than updating the prior belief over states after
each observation.  For each scenario (with and without measurement error), 
we generate 100 simulations of  the fishery being managed under each 
strategy (MDP, Offline, POMDP) for 100 time steps.  We compute the average and standard
deviation of the stock size over these 100 replicates, Fig. \ref{sim},
as well as the mean economic return of each case (Table \ref{Table1})
and the fraction of simulated stocks that have collapsed by the end
of the simulation (Table \ref{Table2}). While we perform these
simulations over a range of models, parameter values and noise distributions, for
simplicity of presentation we focus on simulations
from a Ricker model with log-normally distributed noise (see *Materials and Methods*);
comparable results under the alternative choices are presented in the SI.

\begin{figure}[h]
\centering
\caption{Forward simulations of managing a marine ecosystem under POMDP and MDP models in the presence ( $\sigma_m = 0.5$, top panel) or absence ( $\sigma_m = 0$, bottom panel) of measurement error.  Lines show the mean stock size over 100 replicate simulations, error bands show +/- 1 standard deviation.  Many of the replicate simulations under MDP quickly collapse the stock by over-harvesting when measurement error is present.  In contrast, the POMDP management is robust to the level of measurement error, avoiding stock collapse whether or not measurement error is actually present.  The growth model is Ricker, the noise is log-normally distributed, the carrying capacity of the system is $K = 0.66$, the maximum per capita growth rate is $r = 1$, growth noise is $\sigma_g = 0.5$, and the discount factor is $\gamma = 0.95$.}\label{sim}
\end{figure}

Fig. \ref{sim}, top panel, illustrates that ignoring or underestimating
the measurement error result in dramatic collapse of the stock and
significant loss in revenue: the mean MDP state declines as subsequent
replicate simulations collapse (94/100 collapse within 100 time steps)
when measurement error leads to over-harvesting.  The POMDP method
retains a steady, sustainable harvest that keeps the population below
carrying capacity but results in minimal population collapse (4/100). The
offline solution results in some stock collapse (34/100) as a result of
not updating the manager's belief state over subsequent observations.
Unsurprisingly, average economic returns are also much lower for MDP
when measurement error is present, returning an average of 66% of the net present value returned by
POMDP (Table \ref{Table1}), and would be lower still with less aggressive discounting of future
rewards (Here $\gamma = 0.95$ makes later collapses less relevant to the total).
Fig. S1 shows stock trajectories from several individual replicates, which 
show relatively little decline before collapsing suddenly when managed under MDP
when measurement error is present.  Averaging over an ensemble of such trajectories
gives the appearance of a smooth decline.

On the other hand, overestimating measurement error might result in a
trivial loss in revenue but does not result in collapse of the stock,
(Fig. \ref{sim}, lower panel). In this scenario, since no
measurement error is present, the MDP solution is indeed optimal, but the
POMDP policy which assumes a measurement error of $\sigma_m = 0.5$ and
thus overestimates the measurement uncertainty performs nearly as well
(average of 94% of the MDP net present value and likewise no collapses of the stock, Table \ref{Table2}).
In the SI we report the same patterns for Ricker model
with uniform noise (Fig. S2), Beverton-Holt model with log-normal noise
(Fig. S3), and Allen model with log-normal noise (Fig. S4).

<!-- Perhaps in-text quotations of these numbers is sufficient?-->

\begin{table}
\centering
\caption{Ecological and economic performance when
measurement error is present, (\(\sigma_m = 0.5\))}
\begin{tabular}{lccc}
& MDP & Offline & POMDP \\
\midrule
Fraction collapsed & 94\% & 34\% & 4\% \\
Economic return & 66\% & 99\% & 100\% \\
\bottomrule
\end{tabular}
\label{Table1}
\end{table}

\begin{table}
\centering
\caption{Ecological and economic performance when
measurement are perfect, (\(\sigma_m = 0\))}
\begin{tabular}{lccc}
& MDP & Offline & POMDP \\
\midrule
Fraction collapsed & 0\% & 0\% & 0\% \\
Economic return & 100\% & 96\% & 94\% \\
\bottomrule
\end{tabular}
\label{Table2}
\end{table}

To facilitate comparison with previous research and management practice,
we plot the corresponding optimal action in terms of *expected escapement*: the
stock the manager expects will be left in the sea after harvest $z_t - h_t$.
Recall that introducing measurement error means that this value can differ
from realized escapement, $x_t - h_t$. Fig. \ref{policy} shows the optimal expected 
escapement as a function of observed stock sizes over a range of growth 
noise and observation error intensities, assuming a uniform prior belief over
states $x_t \in (0,1)$ for $\sigma_m > 0$. These policies are determined numerically
using the SARSOP algorithm for POMDP [ @Kurniawati2008], as discussed in *Materials and Methods* and
illustrated in the supplementary code provided.

In the MDP case ($\sigma_m = 0$),
we see the classical result of constant escapement [@Reed1979], where the stock is harvested
back down to the same escapement regardless of observed stock or growth noise.
When growth noise is large ($\sigma_g = 0.5$) relative to measurement
error ($\sigma_m = 0.1$), the policy converges towards the the constant
optimal escapement solution of @Reed1979, as expected. These patterns
hold independent of the population growth model (e.g. Beverton-Holt and
Allen) and the shape of the noise distribution (Fig. S5--S6).
The most obvious difference between the MDP and POMDP policies arises in the region of
linear or nearly linear increases in expected escapement occurring at very large observed stock sizes,
which arises from any policy whose harvest remains *constant* over that range. Not increasing
harvest in response to higher observed stock sizes always result in a higher expected escapement, $z_t - h_t$.  
This corresponds to a policy which believes such high observations are almost surely the 
result of measurement error, and suggests that overestimated stock sizes are responsible for MDP
over-harvesting and the subsequent collapse of the stock. At smaller stock sizes, the policies
shown in Fig. \ref{policy} appear more similar over any measurement error, but in fact this is only a consequence of the
uniform prior belief. While Fig. \ref{policy} completely specifies the management policy under MDP,
this is not true for the POMDP solutions ($\sigma_m > 0$), which also depend on all
prior observations through the influence of the belief state.

\begin{figure}[h]
\centering
\caption{Optimal escapement policy as a function of measured stock for Ricker model, with different measurement errors and growth noise. Noises are log-normally distributed, and the prior belief state is uniform among all stocks. The carrying capacity of the system is $K = 0.66$, the growth rate is $r = 1$ and the discount factor is $\gamma = 0.95$.}\label{policy}
\end{figure}

To illustrate this key difference, we plot (Fig. \ref{bel})
the POMDP policy under two alternative prior belief states: (1) Low
Stock: where the prior belief is centered at $x = 0.2$ , and (2) High Stock:
where the prior belief is centered at the location of carrying capacity $x =
0.66$. This figure is generated in an identical manner to Fig. \ref{policy} 
except for the different choice of prior belief.  Visualizations of these 
prior distributions are shown in Fig. S12. When the prior belief skews towards low stocks, the
probability that an observation of a high stock is an overestimate due to
measurement error is greater, and consequently the policy becomes more conservative
(higher escapement). This ability to incorporate the knowledge from prior
observations accounts for the difference
in performance between the POMDP solution and the Offline solution in
Fig. \ref{sim} when measurement error is present. 
The evolution of the belief state in response to subsequent observations
over a given simulation is shown for the different models in Figs S14 - S16.
Corresponding plots for alternate models are shown in Fig. S9.
The patterns shown here are further robust to the reward function (including
linear cost to harvest and a diminishing returns model, Fig. S8) and the
discount rate Fig. S13.

\begin{figure}[h]
\centering
\caption{Optimal escapement policy as a function of measured stock and different prior belief states: (1) Uniform: uniform prior belief over all stocks, (2) Low Stock: the prior belief is centered at $x = 0.2$, and (3) High Stock: the prior belief is centered at the location of carrying capacity $x =
0.66$. In this figure, growth model is Ricker, measurement and growth noises are $\sigma_m = \sigma_g = 0.5$, and are log-normally distributed, the growth rate is $r = 1$, and the discount factor is $\gamma = 0.95$.}\label{bel}
\end{figure}

## Application to real-world fisheries

Using historical time series of Argentine Hake on the Patagonian Shelf and
European Plaice on the Celtic-Biscay Shelf from the the RAM Legacy Stock
Assessment database [@Ricard2011], we compare the catch levels recommended
by POMDP and MDP methods against historical values (Fig. \ref{RAM2}). We have selected
these two examples for having both adequately long records and show very
different trends in both population biomass and catch. We non-dimensionalize
the data and then use Maximum
Likelihood Estimation (MLE) through NIMBLE [@nimble-software:2016] to
fit a Ricker model for both these examples (details of MLE fits can be
found in the SI). Based on this model estimate, we then determine MDP
and POMDP policies using the appropriate algorithms.  In the case of MDP,
the optimal harvest is determined directly from that year's estimated
stock abundance.  In the case of POMDP, the optimal harvest depends both
on this stock estimate and the prior belief, which is sequentially updated
based on each previous observation (stock estimate) and action (catch).

Both examples illustrate the contrast between the ignoring (MDP) and accounting for measurement error (POMDP).
The POMDP solution tends to recommend a lower harvest than the MDP solution
in both cases, but not strictly so.  In Plaice example, we see the the POMDP solution is less 
volatile, increasing harvest more modestly in response to observing stock 
sizes while fishing pressure remains constant or increases (e.g. 1980-1990), 
while also decreasing less if stocks decline under constant or decreasing fishing
pressure (e.g. 1990-2000). Under the MDP model, only sharp changes in catch
can drive these sharp shifts in stock levels, while the POMDP model permits that these
can arise in part by measurement error and thus require less volatility in the harvest. 
In the Hake time series, the continued decline of the stock leads both MDP
and POMDP solutions to recommend smaller harvests than historically observed.
Again the POMDP solution is less volatile than the MDP, starting with a more
modest harvest at the beginning of the time-series (where a uniform prior over
states is assumed), and at first decreasing more gradually but then retaining
the moratorium on fishing when a small rebound (circa 2005) leads the MDP to a
renewed harvest.

\begin{figure}[h]
\centering
\caption{This figure compares the actual stock biomass and catch of European Plaice on the Celtic-Biscay Shelf and Argentine Hake on the Patagonian Shelf with actions suggested by MDP and POMDP frameworks. In the case of POMDP we have used measurement error of $\sigma_m = 0.5$.}\label{RAM2}
\end{figure}

# Discussion {#discussion .unnumbered}

We have shown that fully accounting for measurement error in calculating
the optimal fishing harvest leads to management that is usually more conservative
(higher expected escapement) as the measurement error increases. This
contrasts with the classic solution of constant escapement, in which
larger estimates of the stock size justify proportionally larger harvests.
The importance of our intuitively sensible conclusion is underscored
by the contrast to both much of current fisheries management practice
and previous research into measurement uncertainty.  Most fishing policy
still reflects the results @Reed1979 driven in
the absence of measurement error, and even much previous work on the role
of measurement uncertainty  has either assumed constant escapement-type
solutions [@Roughgarden1996; @Moxnes2003; @Engen1997] or relied on approximations which
have argued that accounting for measurement uncertainty should result
in an *increased* harvest (lower target stock size) [@Ludwig1981; @Clark1986; @Sethi2005].

Using simulations (Fig. \ref{sim}, S2-S4), we have shown that failing to account for measurement error
can lead to dramatic ecological and economic consequences. MDP solutions, which
ignore measurement error, collapse 94% of simulated time series within 100 time
steps. This does not mean that measurement error in the simulations is so large
as to make informed management impossible -- in fact the optimal POMDP solution,
which accounts for measurement error, experiences a collapse in only 4% of
simulations shown here. While the precise degree of measurement uncertainty may
itself be unknown, these results also show that it is better to overestimate
than underestimate this uncertainty. Even when no measurement error is present,
a POMDP solution which assumes a large measurement error ($\sigma_m = 0.5$) performs
nearly as well as the optimal MDP solution in all of our simulations. Fig. \ref{policy}
and \ref{bel} provide the intuition behind why MDP approach results in such a high probability
of collapse: when measurement error results in a chance over-estimation of the 
stock, the resulting harvest under MDP can be much larger than under POMDP. A single
large or a few consecutive smaller over-estimations of the stock size can thus 
quickly drive the stock to zero.  In contrast, the POMDP solution will harvest
much less under the same observation error, particularly when it compares that
observation against the accumulated prior belief that the current state should
be lower, Fig \ref{bel}.

We have shown that this dramatic reduction in the probability of a 
collapse is not trivially the result of a more cautious strategy; e.g. one which always
harvests less than MDP.  Rather, it is the result of a making full use
of all of the data through the role of the belief state.
The POMDP solution, unlike the MDP solution, depends on
the entire history of previous observations, which inform the current
belief state.  As shown in Fig \ref{bel}, this can correspond
to higher harvests if prior observations give sufficient evidence that 
the stock is higher than suggested by the most recent measurement. As a result,
the POMDP policy can outperform the MDP policy economically as well as 
ecologically.

We have also demonstrated these patterns by comparing harvests recommended
by MDP methods against those recommended by POMDP methods using two separate
historical time series.  These examples provide a further illustration
of how accounting for measurement uncertainty results in very different policy
outcomes than ignoring it.  Note that this analysis does not seek to explain
the differences between either the MDP and POMDP solutions from the historical
catches actually observed.  While it is very possible that the observed harvests
were governed by some sort of constant-escapement policy, actual catch quotas
or other policy information is not available in this data. Fluctuations in historical
harvest may reflect many other factors than such a policy, such as different stock
models, external constraints such as the cost of adjusting policy [@Boettiger2016], 
or other economic, political or environmental factors.  While we are free to
make comparisons between MDP and POMDP solutions on the historical data, understanding
why historical harvests differed from either of these would need an understanding of 
these contextual issues that are beyond our data and our scope here.
Given the stock-recruitment model for the fishery, the MDP optimum harvest depends only
on the most recent stock estimate, while both the full time series of historical harvests
and historical stock estimates influence the POMDP optimum, by altering the belief
state at the time.  We have
relied on estimating reasonable candidate models and parameters from the time series
a priori (specifically MLE parameters to the Ricker model). Ideally model
and parameters would also be inferred over the course of the sequential decisions,
which would integrate over such uncertainty in models or parameters.
While MDP methods to address these issues of model uncertainty
and parameter uncertainty have been developed elsewhere [e.g. @Johnson2015] this remains
a subject for future research in POMDP applications (see *Future directions*).

We have shown that our result is generally robust to our assumptions:
this pattern holds for differing models (Ricker, Beverton-Holt, and Allen),
differing magnitudes and also different shapes of noise distributions
(log-normal, as considered by @Clark1986 or uniform, as considered by
@Sethi2005), the choice of the reward/profit equation, or the details of
the discretization (see SI).  This robustness is not surprising, being consistent
with sensitivity observed in previous work.  Indeed, in the absence of
measurement error most of these scenarios were already covered by Reed's
proof, (with the exception of the Allen model whose change in concavity
violates the theorem and likewise shows deviations here).
Thus it is perhaps all the more surprising that
constant escapement is not robust to the assumption of measurement error.

The optimization approach considered here can be distinguished from more heuristic
approaches to address the uncertainty, such as pretty good yield (PGY)[@Hilborn2010; @Rindorf2016] or management strategy evaluation (MSE).  Methods based on MSE have been applied in fishery management for providing the decision maker with a rationale to base their decisions given their specific optimization objectives [@Smith1994; @Mapstone2008; @Bunnefeld2011]. The main difference between MSE and methods based on dynamic programming is that MSE does not solve the optimization problem to provide the optimal strategy, rather it evaluates consequences of a set of strategies defined based on management objectives and provides the outcome of those simulations to the decision maker. In that sense, MSE can be compared to the policy evaluation step of optimization in dynamic programming using policy iteration methods [@Sutton1998], while the policy improvement step is being ignored due to computational complexity. With advancements in efficient solving of dynamic programming optimization under several sources of uncertainty [@Shani2013], the traditional MSE approach can be interpreted as an internal step in the optimization process that tries to identify the best management strategy, searching through all viable and possible strategies. MSE can be also compared to Monte Carlo search algorithms within the space of possible policies that the manager can explore and evaluating the policies based on the outcome [@Smith1999]. For example, @Smith1999 evaluates the effect of several sources of uncertainty such as structural and parameter uncertainty in the models, errors in data and observation systems, estimation uncertainty, and management implementation uncertainty in partnership with Australian Fisheries Management Authority.

Our results have important implications for management of fisheries.
The most evident of these is that given the unavoidable realities of measurement
error in stock estimates, the wide use of management based on constant-escapement
solutions and derivative concepts such as maximum sustainable yield (MSY), maximum
sustainable biomass ($B_{MSY}$),
which are central to scientific fisheries management and policy
included the United Nations Law of the Sea [@UNCLOS1982], are insufficient to
guarantee either maximum sustainable economic yield or persistence of fish stocks.
Instead, target harvest or target escapement must reflect current and previous 
stock estimates and the degree of measurement uncertainty present. While precise 
estimates of the uncertainty in measurements, $\sigma_m$ may be difficult to obtain,
we have shown that it better both ecologically (in the probability of stock collapse)
and economically (in the profits from fishing) to overestimate than underestimate this
quantity.

This important qualitative impact that measurement error has on the
optimal solution of an important and well-studied ecological management
problem has more general implications.  The role of measurement error
is most often acknowledged in the breach -- a mathematically
convenient assumption only.  After all, inherent stochasticity has often been found
to make only small differences from deterministic solutions, as @Reed1979
proved in this case. Our result demonstrates how frail such an assumption
regarding measurement uncertainty may be, particularly when determining
optimal policy.  No doubt the mathematical and computational difficulty of
accounting for measurement uncertainty has forestalled greater attention
to this issue.  The growing wave of available data makes such concerns
about measurement error all the more pressing, as imperfect but readily
available observations make such data-driven optimization possible in more
applications.

These results have both tactical and strategic implications for ecological 
modeling and fisheries management [sensu @May1973]. At a strategic level,
this highlights the danger in ignoring the potentially asymmetric impact of
measurement error, which can lead to significant over-harvesting and potential
collapse of the stock.  At a tactical level, the precise amount by which
harvest should be adjusted to account for a given measurement error to maximize
the expected value of a stock will require a POMDP solution for the specific 
fishery in question, which is now computationally feasible using the approach
we have presented here. We hope that the recent advances in efficient and accurate
POMDP solvers, together with greater attention to these challenges,
will encourage a wider consideration of the role of measurement error
in determining appropriately precautionary management policy.

In conclusion, we have demonstrated the importance of measurement error
in the long-standing question of optimal harvests in fisheries.  We have 
shown that overestimating the measurement error can still result in policies
that are nearly optimal both economically and ecologically, while ignoring or
underestimating the measurement noise can cause dramatic collapses of fish
stocks even when no tipping point is present.  This suggests that the failure
to account for measurement error in this way in even the most advanced 
fisheries management could contribute to long-term trends in declining
stocks [e.g. @Costello2016]. We believe the approach taken
here both highlights the importance of accounting for measurement error and
provides an example of how this can be accomplished in the POMDP framework
using modern algorithms on complex problems. We have also implemented the
proposed method on a real-world problem of managing a marine ecosystem and show
that the policy suggested by POMDP is significantly different with respect to the
actual catch landings.

## Future directions

Several limitations that have been studied in MDP models, such as parameter 
and model uncertainty and adaptive management [e.g. @Johnson2015] remain
intriguing future challenges for POMDP problems. Future work could extend the examples
shown here to include more realistic models, such as those with age structure, as
has only recently been done for the MDP case [@Holden2015].
There remains a need for further improvement of the proposed method to include such model uncertainty. 
Previous attempts that adapt POMDP tools to include model uncertainty [i.e. @Chades2012] have had to
assume full observability of state space, ignoring measurement uncertainty. 
A second limiting assumption common to MDPs and POMDPs is that of stationary dynamics: 
that the population dynamics equation itself is not changing over
time. In reality, forces such as climate change and other forms of environmental variations violate this
assumption. @Fackler22014 attempts to address this problem by 
adaptation of Mixed Observability MDP [@Ong2010] framework, although the proposed algorithm
for finding optimal solution to such non-stationary systems is intractable, except for 
extremely simplified examples used in that study. Our future work involve relaxing
these assumptions by proposing computationally feasible approximate algorithms within the 
POMDP framework to deal both with model uncertainty and non-stationarity of system's dynamics.
Lastly, the introduction of POMDP approaches raises a central question for managers: when is it
worthwhile to actively reduce measurement error to derive a better policy, and when is it
best simply to account for current measurement error?  This issue can be addressed through
a Value of Information framework, as @Johnson2015 illustrates in the case of MDPs.

# Materials and Methods {#methods .unnumbered}

A Partially Observed Markov Decision Process (POMDP) [@Smallwood1973;
@Sondik1978] is a generalization of a Markov Decision Process (MDP),
which is a fundamental model for sequential decision making.  In the
ecological literature, MDPs are often referred to by their solution
method: Stochastic Dynamic Programming [SDP see @Marescot2013].
The MDP approach has been widely used in behavioral ecology
[@Mangel1988] as well as conservation ecology and resource economics
[e.g. @Mangel1985]. Introductions to MDP and solving them through dynamic
programming are provided in @Mangel1985, @Bertsekas1996 and @Sutton1998.
Specifically, POMDPs overcome one of the main limitations of MDP which
is the assumption about the full observability of the system's state.
In POMDP, the exact state of the system cannot be
observed directly, but can be inferred by indirect measurements.
POMDPs can be represented as a probabilistic graphical models
[@Barber2012] as in Fig. S11.  For further details regard POMDP framework,
its parameters, and details of solving the optimization problem,
refer to the SI.

Our approach utilizes the SARSOP algorithm of @Kurniawati2008. Computational 
requirements are still significantly more demanding than the equivalent MDP solution.
Precise computational time and memory requirements depend not only on system size but on
the precision requested of the approximate, point-based solution. Reasonable
solutions for the problems considered here can be reached within an hour of computation,
though were run for five hours each to confirm convergence. 
As a supplement to this paper we provide example scripts for replicating
the analyses shown here, which can be found at https://github.com/boettiger-lab/sarsop.

### 2017-12-14 manuscripts/manuscript.Rmd
refs

@@
 
-In the case of no uncertainty, the optimal harvest strategy is to reduce the population to the size at which it obtains the maximum growth rate, known as Biomass at Maximum Sustainable Yield, $B_{MSY}$ (this is adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @ClarkTextbook).
+In the case of no uncertainty, the optimal harvest strategy is to reduce the population to the size at which it obtains the maximum growth rate, known as Biomass at Maximum Sustainable Yield, $B_{MSY}$ (this is adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
 For the Graham-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$, with a corresponding annual harvest of $H_{MSY} = r K / 4$.
@@
 
-The resulting solutions only deepened the paradox, rather than recovering a more cautious policy in response to the uncertainty, @Clark1987 finds policies that are often *less* cautious:
+The resulting solutions only deepened the paradox, rather than recovering a more cautious policy in response to the uncertainty, @Clark1986 finds policies that are often *less* cautious:
 
@@
 As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms.
-While not unknown to the conservation literature [@Williams2009], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
+While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
 @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
@@
 Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
-For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop, which can be used to replicate any of the results presented here and explore other variations.
+For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop-pkg, which can be used to replicate any of the results presented here and explore other variations.
 

### 2017-12-22 appendix/appendix.Rmd
Appendix example based on f(x - h); recruit before harvest.

@@
 
-write_csv(sims, "sims.csv")
+write_csv(sims, "appendix_files/sims.csv")
 ```
@@
 
+```{r}
+policy_table <- tibble(state = 1:length(states)) %>%
+  bind_cols(policies) %>%
+  gather(policy, harvest, -state) %>%
+  mutate(harvest = actions[harvest], state = states[state])
+
+write_csv(policy_table, "appendix_files/policy_table.csv")
+```
+
 Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (deterministic) recruitment would land us right at it $B_{MSY}$.
@@
 ```{r}
-policy_table <- tibble(state = 1:length(states)) %>%
-  bind_cols(policies) %>%
-  gather(policy, harvest, -state) %>%
-  mutate(harvest = actions[harvest], state = states[state])
-
 policy_table %>% 
@@
   coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-
 ```
 
-Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles.
+Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.
+The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles when we plot in terms of escapement instead of harvest (as is common in the optimal control literature in fisheries, e.g.
+@Sethi2005).
+
+Note that under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population
+is harvested back down to that size.
+Note that even stocks observed at states slightly below $B_{MSY} = 0.5$ achieve this target escapement, since we are following the classic Graham Shaeffer formulation here where we observe first, then recruitment happens before harvest, and thus we see the population at smaller size than we harvest it.
+In classical escapement analysis, observations are usually indexed instead to occur immediately before harvests, and this inflection point occurs right at $B_{MSY}$.
 
@@
 
-## Historical data estimation
+In the above plots for the MSY
 
-## Historical data POMDP & other solutions
+```{r}
+unif_prior = rep(1, length(states)) / length(states) # initial belief
+unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)
 
-## Historical combined plot
+```
+
+```{r}
+i_star <- which.min(abs(states - S_star))
+ i_low <- which.min(abs(states - 0.5 * S_star))
+ i_high <- which.min(abs(states - 1.5 * S_star))
+#i_low <- 2
+#i_high <- which.min(abs(states - K))
+
+prior_star <- m$observation[,i_star,1]
+prior_low <- m$observation[,i_low,1]
+prior_high <- m$observation[,i_high,1]
+
+star <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_star)
+low <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_low)
+high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)
+
+```

### 2017-12-22 appendix/appendixB.Rmd
Appendix example based on f(x - h); recruit before harvest.

@@
 
-```{r libraries}
+```{r libraries, message=FALSE}
 # devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
@@
 ```{r}
+if(!file.exists("ramlegacy.zip")){
 download.file("https://depts.washington.edu/ramlegac/wordpress/databaseVersions/RLSADB_v3.0_(assessment_data_only)_excel.zip",
               "ramlegacy.zip")
+}
 ```
@@
   mutate(model = recode(sigma_m, 
-                        "0" = "reed",
-                        "0.1" = "pomdp_0.1", 
-                        "0.2" = "pomdp_0.2")) %>%
+                        "0" = "Reed",
+                        "0.1" = "POMDP", 
+                        "0.2" = "POMDP_0.2")) %>%
   select(-sigma_m) %>%
   bind_rows(historical)
+
+write_csv(sims,"appendixB_files/sims.csv")
 ```
@@
 sims %>%
-  filter(model %in% c("biomass", "catch", "pomdp_0.1", "reed", "pgy")) %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
   ggplot(aes(year, stock, col=model)) +

### 2017-12-22 appendix/appendix_harvestfirst.Rmd
Appendix example based on f(x - h); recruit before harvest.

--- 2017-12-13 appendix/appendix.Rmd
@@
 ggplot2::theme_set(ggplot2::theme_bw())
-```
-
-```{r helper_routines, include = FALSE}
-## map cts values to nearest grid indices 
-index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
-
-det_simulations <- function(x0 = 1/6, Tmax = 20, 
-                              B_MSY = K / 2, 
-                              F_MSY = r / 2 , 
-                              F_PGY = 0.8 * F_MSY){
-
-  msy_policy <- function(x) F_MSY * x
-  pgy_policy <- function(x) F_PGY * x
-  det_policy <- function(x) pmax(f(x,0) - B_MSY,0) ## ASSUMES harvest takes place *after* recruitment
-  reed_policy  <- function(x) pmax(x - B_MSY,0) ## Optimal when harvest takes place *before* recruitment, assumes we are seeing a pre-harvest size.
-  ## consequently, if we use such a policy when we are seeing a the post-harvest size, we give too much buffer.
-
-  # a simulation function
-  det_sim <- function(policy, f, x0, Tmax){
-    action <- state <- obs <- as.numeric(rep(NA,Tmax))
-    state[1] <- x0
-    for(t in 1:(Tmax-1)){
-      obs[t] <- state[t]
-
-      action[t] <- policy(obs[t])
-      state[t+1] <- f(state[t], action[t]) 
-    }
-    data.frame(time = 1:Tmax, state, action, obs)
-  }
-
-  # simulate each policy, returning all results in a single unified data frame 
-  sims <- 
-    list(msy = msy_policy, 
-         det = det_policy, 
-         pgy = pgy_policy
-         #opt = reed_policy
-         ) %>% 
-    map_df(det_sim, 
-           f, x0, Tmax, 
-           .id = "method")
-
-  sims
-}
 ```
@@
 
-## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
+## Unlike classic Graham-Schaefer, this assumes harvest occurs right after assessment, before recruitment
 f <- function(x, h){ 
-    x + x * r * (1 - x / K) - pmin(x,h)
+    s <- pmax(x-h, 0) 
+    s + s * r * (1 - s / K) 
 }
@@
 
-If we observe->recruit->harvest, $X_{t+1} = f(X_t) - h_t$, then we see the stock right after a harvest and before recuitment, when it is at $X_t ~ $B_{MSY}$.  
-If we instead did observe->harvest->recruit, we would see the stock at its pre-harvest size, $X_t ~ B_{MSY} + $H_{MSY}$.
+Since we `observe` -> `harvest` -> `recruit`, we would observe the stock at its pre-harvest size, $X_t \sim B_{MSY} + H_{MSY}$.
 
 ```{r}
-B_MSY <- S_star
-MSY <- f(B_MSY,0) - B_MSY
+#B_MSY <- S_star     # recruit first, as in classic Graham-Schaefer
+B_MSY <- f(S_star,0) # harvest first
+
+#MSY <- f(B_MSY,0) - B_MSY  # recruit first
+MSY <- B_MSY - S_star  # harvest first
 
@@
 F_PGY = 0.8 * F_MSY
-```
-
-```{r optional_checks}
-## Make sure these match the known analytic solutions for Graham-Schaefer model (only true of discount = 1)
-if(discount == 1){
-  testthat::expect_equal(B_MSY, K / 2)
-  testthat::expect_equal(MSY, r * K / 4)
-  testthat::expect_equal(F_MSY, r / 2)
-}
 ```
@@
 ```{r}
-det_simulations(x0 = 1/6, Tmax = 20, B_MSY, F_MSY, F_PGY) %>%
+
+msy_policy <- function(x) F_MSY * x
+pgy_policy <- function(x) F_PGY * x
+
+## ASSUMES harvest takes place *before* recruitment, f(x-h), not after.
+escapement_policy <- function(x) pmax(x - S_star,0)  
+x0 <- K/6
+Tmax = 20
+det_sim <- function(policy, f, x0, Tmax){
+    action <- state <- obs <- as.numeric(rep(NA,Tmax))
+    state[1] <- x0
+
+    for(t in 1:(Tmax-1)){
+      action[t] <- policy(state[t])
+      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
+      state[t+1] <- f(state[t], action[t]) 
+    }
+    data.frame(time = 1:Tmax, state, action, obs)
+  }
+
+sims <- 
+  list(msy = msy_policy, 
+       pgy = pgy_policy,
+       det = escapement_policy) %>% 
+  map_df(det_sim, 
+         f, x0, Tmax, 
+         .id = "method") 
+sims %>%
   ggplot(aes(time, state, col=method)) + 
@@
   geom_hline(aes(yintercept=B_MSY), lwd=1)
+
 ```
@@
 
+write_csv(policy_table, "appendix_files/policy_table.csv")
+```
+
+```{r}
 policy_table %>% 
@@
   coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-
 ```
@@
 ### Adding the POMDP policy solution
-
-## Historical data estimation
-
-## Historical data POMDP & other solutions
-
-## Historical combined plot

### 2017-12-22 manuscripts/manuscript.Rmd
Appendix example based on f(x - h); recruit before harvest.

@@
 output: rticles::elsevier_article
+layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column
+
 ---
+
+```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
+library(tidyverse)
+library(gridExtra)
+library(ggthemes)
+colours <- ptol_pal()(2)
+```
+
+```{r suggested_libs}
+library(hrbrthemes)
+library(extrafont)
+##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
+hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
+extrafont::loadfonts()
+ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+#ggplot2::theme_set(theme_bw()) # simpler theme if you do not have these fonts
+```
 
@@
 
-- Replicate simulations
+### Replicate simulations
 
@@
 
-- historical data.
+### Comparisons to historical data
 
 ```{r}
+sims <- read_csv("../appendix/appendixB_files/sims.csv")
+sims %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
+  ggplot(aes(year, stock, col=model)) +
+  geom_line(lwd=1) +
+  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
+  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
 ```
@@
 -->
+
+# References

### 2017-12-23 appendix/appendix.Rmd
pomdp policies not smooth

@@
 
-```{r eval=FALSE}
+```{r eval=TRUE}
 dir.create(log_dir)
@@
              precision = 0.00000002,
-             timeout = 10000,
+             timeout = 20000,
              log_dir = log_dir,
@@
   gather(policy, harvest, -state) %>%
-  mutate(harvest = actions[harvest], state = states[state])
+  mutate(harvest = actions[harvest], state = states[state]) %>%
+  mutate(escapement = f(state,harvest))
 
@@
 policy_table   %>% 
-  mutate(escapement = f(state,harvest))  %>% 
   ggplot(aes(state, escapement, col=policy)) + 
@@
 
-### Adding the POMDP policy solution
+### Policies under uncertainty
 
-In the above plots for the MSY
+In strategies whose policies are shown in the above plots all ignore both stochasticity and measurement error.
+If want to compare these to an MDP or POMDP policy, we must specify the level of uncertainty.
+
+In the absence of measurement uncertainty this is straight forward.
+@Reed1979 essentially tells us that for small growth noise (satisfying or approximately satisfying Reed's self-sustaining condition) that the stochastic optimal policy is equal to the deterministic optimal policy.
+We can confirm this numerically as follows.
+
+First we grab the transition matrix we have already defined for small `sigma_g`:
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.02, sigma_m ==0) %>% pull(scenario) %>% as.integer()
+m <- models[[i]]
+```
+
+With no observation uncertainty, we can solve numerically for the optimal policy with stochastic dynamic programming
+
+```{r}
+mdp <- MDPtoolbox::mdp_policy_iteration(m$transition, m$reward, discount) 
+```
+
+Adding this to the plot we see the result is identical to the deterministic case:
+
+```{r}
+bind_rows(policy_table,
+  data.frame(state = states, 
+             policy = "sdp 0.02",
+             harvest = actions[mdp$policy],
+             stringsAsFactors = FALSE) %>%
+  mutate(escapement = f(state,harvest))) %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```
+
+Repeating this for larger stochasticity, we get a slightly more conservative result:
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0) %>% pull(scenario) %>% as.integer()
+m <- models[[i]]
+mdp <- MDPtoolbox::mdp_policy_iteration(m$transition, m$reward, discount) 
+```
+
+```{r}
+bind_rows(policy_table,
+  data.frame(state = states, 
+             policy = "sdp 0.1",
+             harvest = actions[mdp$policy],
+             stringsAsFactors = FALSE)) %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```
+
+### Comparing POMDP Policies
+
+The comparison of POMDP policy is yet more complicated, but the POMDP policy cannot be expressed merely in terms of a target harvest (or escapement) level given an estimation of the stock size (state).
+The optimal solution for the partially observed system must also reflect all prior observations of the system, not merely the most recent observation, as the system is not Markovian in the observed state variable.
+We summarize this history as a prior "belief"" about the state, which is updated according to Bayes rule after each observation.
+(Note that @Sethi2005 fails to realize this and plots solutions with measurement uncertainty without reference to the prior, which explains their counter-intuitive finding that increased uncertainty should result in increased harvest rates).
+
+Let us look at the POMDP solutions under various priors focusing on the case of moderate uncertainty, $\sigma_g = \sigma_m = 0.1$.
+(Recall we have already solved the POMDP solution for this model in the simulations above, as defined by the `alpha` vectors, so we can quickly load that solution now.)
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.1) %>% pull(scenario) %>% as.integer()
+m <- models[[i]]
+alpha <- alphas[[i]] # we need the corresponding alpha vectors
+```
+
+We will consider what the POMDP solution looks like under a few different prior beliefs.
+A uniform prior sounds like a conservative assumption, but it is not: it puts significantly more weight on improbably large stock values than other priors.
+(Loading the $\alpha$ vectors from our POMDP solution computed earlier, we can then compute a POMDP given these $\alpha$, the matrices for transition, observation, and reward, and the prior we are using)
 
@@
 
+For more realistic set of priors, we will consider priors centered at the target $B_{MSY}$ size (or $S^*$ in the language of Reed), at half $B_{MSY}$, and at 1.5 times $B_{MSY}$, each with a standard deviation of $\sigma_m = 0.1$ (i.e.
+the uncertainty around a single observation of a stock at that size.)
+
 ```{r}
@@
 i_high <- which.min(abs(states - 1.5 * S_star))
-#i_low <- 2
-#i_high <- which.min(abs(states - K))
 
@@
 ```
+
+We gather these solutions into a single data frame and convert from grid indices to continuous values
+
+```{r}
+df <- unif
+df$medium <- star$policy
+df$low <- low$policy
+df$high <- high$policy
+
+pomdp_policies <- df %>% 
+  select(-value) %>% 
+  rename(uniform = policy) %>% 
+  gather(policy, harvest, -state) %>%
+  mutate(state = states[state], 
+         harvest = actions[harvest]) %>%
+  mutate(escapement = f(state,harvest)) %>%
+  select(state, policy, harvest, escapement)
+
+```
+
+```{r}
+pomdp_policies %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```

### 2017-12-23 appendix/appendix.Rmd
fancy plot format

--- 2017-12-22 appendix/appendix.Rmd
@@
 
-```{r aethetics, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE)
-ggplot2::theme_set(ggplot2::theme_bw())
+```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
+library(tidyverse)
+library(gridExtra)
+library(ggthemes)
+colours <- ptol_pal()(2)
+```
+
+```{r suggested_libs}
+library(hrbrthemes)
+library(extrafont)
+##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
+hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
+extrafont::loadfonts()
+ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+#ggplot2::theme_set(theme_bw()) # simpler theme if you do not have these fonts
 ```
@@
 
-This uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.
+This uses a generic optimization routine to find the stock size at which the maximum
+growth rate is achieved with respect to discount, i.e. $$\max_x \lbrace f(x) - x/\delta \rbrace$$
 

### 2017-12-23 appendix/appendix_harvestfirst.Rmd
harvest_first

@@
 #B_MSY <- S_star     # recruit first, as in classic Graham-Schaefer
-B_MSY <- f(S_star,0) # harvest first
+B_MSY <- f(S_star,0) # harvest first, we observe the population at B_msy + h
 
@@
 ```{r}
-meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
-                    sigma_m = c(0, 0.1, 0.2),
+meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.15), 
+                    sigma_m = c(0, 0.1, 0.15),
                     stringsAsFactors = FALSE) %>%
         mutate(scenario  = as.character(1:length(sigma_m)))
-log_dir <- "appendix_alphas" # Store the computed solution files here
+log_dir <- "appendix_alphas_harvestfirst" # Store the computed solution files here
 meta
@@
              precision = 0.00000002,
-             timeout = 10000,
+             timeout = 15000,
              log_dir = log_dir,
@@
 
-Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.  The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles.
-
 ```{r}

### 2017-12-23 appendix/appendix_harvestfirst.Rmd
harvest_first policies pomdp

@@
 
-write_csv(sims, "appendix_files/sims.csv")
+write_csv(sims, "appendix_harvestfirst_files/sims.csv")
 ```
@@
 
+```{r}
+policy_table <- tibble(state = 1:length(states)) %>%
+  bind_cols(policies) %>%
+  gather(policy, harvest, -state) %>%
+  mutate(harvest = actions[harvest], state = states[state]) %>%
+  mutate(escapement = state - harvest)
+
+write_csv(policy_table, "appendix_files/policy_table.csv")
+```
+
 Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (deterministic) recruitment would land us right at it $B_{MSY}$.
@@
 ```{r}
-policy_table <- tibble(state = 1:length(states)) %>%
-  bind_cols(policies) %>%
-  gather(policy, harvest, -state) %>%
-  mutate(harvest = actions[harvest], state = states[state])
-
-write_csv(policy_table, "appendix_files/policy_table.csv")
-```
-
-```{r}
 policy_table %>% 
@@
 
+Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.
+The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles when we plot in terms of escapement instead of harvest (as is common in the optimal control literature in fisheries, e.g.
+@Sethi2005).
+
+Note that under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population
+is harvested back down to that size.
+Note that even stocks observed at states slightly below $B_{MSY} = 0.5$ achieve this target escapement, since we are following the classic Graham Shaeffer formulation here where we observe first, then recruitment happens before harvest, and thus we see the population at smaller size than we harvest it.
+In classical escapement analysis, observations are usually indexed instead to occur immediately before harvests, and this inflection point occurs right at $B_{MSY}$.
+
 ```{r}
 policy_table   %>% 
-  mutate(escapement = f(state,harvest))  %>% 
   ggplot(aes(state, escapement, col=policy)) + 
@@
 
-### Adding the POMDP policy solution
+### Policies under uncertainty
+
+In strategies whose policies are shown in the above plots all ignore both stochasticity and measurement error.
+If want to compare these to an MDP or POMDP policy, we must specify the level of uncertainty.
+
+In the absence of measurement uncertainty this is straight forward.
+@Reed1979 essentially tells us that for small growth noise (satisfying or approximately satisfying Reed's self-sustaining condition) that the stochastic optimal policy is equal to the deterministic optimal policy.
+We can confirm this numerically as follows.
+
+First we grab the transition matrix we have already defined for small `sigma_g`:
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.02, sigma_m ==0) %>% pull(scenario) %>% as.integer()
+m <- models[[i]]
+```
+
+With no observation uncertainty, we can solve numerically for the optimal policy with stochastic dynamic programming
+
+```{r}
+mdp <- MDPtoolbox::mdp_policy_iteration(m$transition, m$reward, discount) 
+```
+
+Adding this to the plot we see the result is identical to the deterministic case:
+
+```{r}
+bind_rows(policy_table,
+  data.frame(state = states, 
+             policy = "sdp 0.02",
+             harvest = actions[mdp$policy],
+             stringsAsFactors = FALSE) %>%
+  mutate(escapement = state- harvest)) %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```
+
+Repeating this for larger stochasticity, we get a slightly more conservative result:
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0) %>% pull(scenario) %>% as.integer()
+m <- models[[i]]
+mdp <- MDPtoolbox::mdp_policy_iteration(m$transition, m$reward, discount) 
+```
+
+```{r}
+bind_rows(policy_table,
+  data.frame(state = states, 
+             policy = "sdp 0.1",
+             harvest = actions[mdp$policy],
+             stringsAsFactors = FALSE)) %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```
+
+### Comparing POMDP Policies
+
+The comparison of POMDP policy is yet more complicated, but the POMDP policy cannot be expressed merely in terms of a target harvest (or escapement) level given an estimation of the stock size (state).
+The optimal solution for the partially observed system must also reflect all prior observations of the system, not merely the most recent observation, as the system is not Markovian in the observed state variable.
+We summarize this history as a prior "belief"" about the state, which is updated according to Bayes rule after each observation.
+(Note that @Sethi2005 fails to realize this and plots solutions with measurement uncertainty without reference to the prior, which explains their counter-intuitive finding that increased uncertainty should result in increased harvest rates).
+
+Let us look at the POMDP solutions under various priors focusing on the case of moderate uncertainty, $\sigma_g = \sigma_m = 0.1$.
+(Recall we have already solved the POMDP solution for this model in the simulations above, as defined by the `alpha` vectors, so we can quickly load that solution now.)
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.1) %>% pull(scenario) %>% as.integer()
+m <- models[[i]]
+alpha <- alphas[[i]] # we need the corresponding alpha vectors
+```
+
+We will consider what the POMDP solution looks like under a few different prior beliefs.
+A uniform prior sounds like a conservative assumption, but it is not: it puts significantly more weight on improbably large stock values than other priors.
+(Loading the $\alpha$ vectors from our POMDP solution computed earlier, we can then compute a POMDP given these $\alpha$, the matrices for transition, observation, and reward, and the prior we are using)
+
+```{r}
+unif_prior = rep(1, length(states)) / length(states) # initial belief
+unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)
+
+```
+
+For more realistic set of priors, we will consider priors centered at the target $B_{MSY}$ size (or $S^*$ in the language of Reed), at half $B_{MSY}$, and at 1.5 times $B_{MSY}$, each with a standard deviation of $\sigma_m = 0.1$ (i.e.
+the uncertainty around a single observation of a stock at that size.)
+
+```{r}
+i_star <- which.min(abs(states - S_star))
+i_low <- which.min(abs(states - 0.5 * S_star))
+i_high <- which.min(abs(states - 1.5 * S_star))
+
+prior_star <- m$observation[,i_star,1]
+prior_low <- m$observation[,i_low,1]
+prior_high <- m$observation[,i_high,1]
+
+star <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_star)
+low <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_low)
+high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)
+
+```
+
+We gather these solutions into a single data frame and convert from grid indices to continuous values
+
+```{r}
+df <- unif
+df$medium <- star$policy
+df$low <- low$policy
+df$high <- high$policy
+
+pomdp_policies <- df %>% 
+  select(-value) %>% 
+  rename(uniform = policy) %>% 
+  gather(policy, harvest, -state) %>%
+  mutate(state = states[state], 
+         harvest = actions[harvest]) %>%
+  mutate(escapement = state-harvest) %>%
+  select(state, policy, harvest, escapement)
+
+```
+
+```{r}
+pomdp_policies %>% 
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```

### 2017-12-28 appendix/appendix.Rmd
run longer

--- 2017-12-23 appendix/appendix.Rmd
@@
 ```{r}
-meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.2), 
-                    sigma_m = c(0, 0.1, 0.2),
+meta <- expand.grid(sigma_g = c(0.02, 0.1, 0.15), 
+                    sigma_m = c(0, 0.1, 0.15),
                     stringsAsFactors = FALSE) %>%
@@
 
-```{r eval=TRUE}
+```{r eval=FALSE}
 dir.create(log_dir)
+options(mc.cores=2)
 
@@
 ```{r}
-meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% left_join(meta) %>% arrange(scenario)
+meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% 
+  mutate(scenario = as.character(scenario)) %>%
+  left_join(meta) %>% arrange(scenario)
 alphas <- alphas_from_log(meta, log_dir)

### 2017-12-28 appendix/appendix.Rmd
harvest & escapement versions
run even longer

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
+devtools::install("..")
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
@@
 
-```{r eval=FALSE}
+```{r eval=TRUE}
 dir.create(log_dir)
-options(mc.cores=2)
+options(mc.cores=9)
 
@@
              precision = 0.00000002,
-             timeout = 20000,
+             timeout = 60000,
              log_dir = log_dir,
@@
 
+### Harvest
+
 ```{r}
@@
 ```
+
+### Escapement
+
+```{r}
+pomdp_policies %>% 
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```

### 2017-12-28 appendix/appendix_harvestfirst.Rmd
index fun
larger discount for faster convergence

@@
 
-Utility (reward) function:
+Utility (reward) function.  (Note that setting a smaller discount will take longer to converge to smooth POMDP solution, but result in `S_star` closer to the simple `B_MSY`.)
 
@@
 reward_fn <- function(x,h) pmin(x,h)
-discount <- 0.99
+discount <- 0.95
 ```
@@
 This uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.
+(Note that this also depends on the discount rate since future profits are worth proportionally less than current profits).
 
@@
                    c(0, 2*K))$minimum
+S_star
 ```
@@
 ```{r}
+index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
 
@@
 
-## POMDP solution (slow, >10,000 seconds per scenario, & memory intensive)
+## POMDP solution 
+## (slow, >10,000 seconds per scenario for discount ~ 0.95;
+## needs much longer for discount ~ 0.99)
 system.time(
@@
 ```{r}
-meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% left_join(meta) %>% arrange(scenario)
+meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% 
+  mutate(scenario = as.character(scenario)) %>%
+  left_join(meta) %>% 
+  arrange(scenario)
 alphas <- alphas_from_log(meta, log_dir)
@@
 ```
+
+```{r}
+pomdp_policies %>% 
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```

### 2017-12-28 appendix/appendix.Rmd
tidy code blocks up a bit

@@
 
-```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-devtools::install("..")
+```{r message=FALSE}
+# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
+library(sarsop)       # the main POMDP package
+library(tidyverse)    # for munging and plotting
+library(parallel)
+
+ggplot2::theme_set(theme_bw())
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
+
+options(mc.cores=parallel::detectCores())
+```
+
+```{r suggested_libs, include=FALSE, eval=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
-library(tidyverse)
+devtools::install("..", dep=TRUE)
 library(gridExtra)
@@
 colours <- ptol_pal()(2)
-```
-
-```{r suggested_libs}
 library(hrbrthemes)
@@
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
-#ggplot2::theme_set(theme_bw()) # simpler theme if you do not have these fonts
-```
-
-```{r helper_routines, include = FALSE}
-## map cts values to nearest grid indices 
-index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
-
-det_simulations <- function(x0 = 1/6, Tmax = 20, 
-                              B_MSY = K / 2, 
-                              F_MSY = r / 2 , 
-                              F_PGY = 0.8 * F_MSY){
-
-  msy_policy <- function(x) F_MSY * x
-  pgy_policy <- function(x) F_PGY * x
-  det_policy <- function(x) pmax(f(x,0) - B_MSY,0) ## ASSUMES harvest takes place *after* recruitment
-  reed_policy  <- function(x) pmax(x - B_MSY,0) ## Optimal when harvest takes place *before* recruitment, assumes we are seeing a pre-harvest size.
-  ## consequently, if we use such a policy when we are seeing a the post-harvest size, we give too much buffer.
-
-  # a simulation function
-  det_sim <- function(policy, f, x0, Tmax){
-    action <- state <- obs <- as.numeric(rep(NA,Tmax))
-    state[1] <- x0
-    for(t in 1:(Tmax-1)){
-      obs[t] <- state[t]
-
-      action[t] <- policy(obs[t])
-      state[t+1] <- f(state[t], action[t]) 
-    }
-    data.frame(time = 1:Tmax, state, action, obs)
-  }
-
-  # simulate each policy, returning all results in a single unified data frame 
-  sims <- 
-    list(msy = msy_policy, 
-         det = det_policy, 
-         pgy = pgy_policy
-         #opt = reed_policy
-         ) %>% 
-    map_df(det_sim, 
-           f, x0, Tmax, 
-           .id = "method")
-
-  sims
-}
-```
-
-```{r message=FALSE}
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
-library(sarsop)       # the main POMDP package
-library(tidyverse)    # for munging and plotting
-library(parallel)
-```
-
-```{r}
-options(mc.cores=parallel::detectCores())
 ```
@@
 
+```{r helper_routines}
+## map cts values to nearest grid indices
+
+det_simulations <- function(x0 = 1/6, Tmax = 20, 
+                              B_MSY = K / 2, 
+                              F_MSY = r / 2 , 
+                              F_PGY = 0.8 * F_MSY){
+  msy_policy <- function(x) F_MSY * x
+  pgy_policy <- function(x) F_PGY * x
+  det_policy <- function(x) pmax(f(x,0) - B_MSY,0) ## ASSUMES harvest takes place *after* recruitment
+
+  det_sim <- function(policy, f, x0, Tmax){
+    action <- state <- obs <- as.numeric(rep(NA,Tmax))
+    state[1] <- x0
+    for(t in 1:(Tmax-1)){
+      obs[t] <- state[t]
+      action[t] <- policy(obs[t])
+      state[t+1] <- f(state[t], action[t]) 
+    }
+    data.frame(time = 1:Tmax, state, action, obs)
+  }
+
+  # simulate each policy, returning all results in a single unified data frame 
+  list(msy = msy_policy, det = det_policy, pgy = pgy_policy) %>% 
+    map_df(det_sim, f, x0, Tmax, .id = "method") 
+}
+```
+
 ```{r}
@@
 ```{r}
+# map to grid
+index <- function(x, grid) map_int(x, ~ which.min(abs(.x - grid)))
 

### 2017-12-28 appendix/appendix.Rmd
discount at 0.95

@@
              precision = 0.00000002,
-             timeout = 60000,
+             timeout = 15000,
              log_dir = log_dir,

### 2017-12-28 appendix/appendixB.Rmd
discount at 0.95

@@
 reward_fn <- function(x,h) pmin(x,h)
-discount <- .99
+discount <- .95
 ```

### 2017-12-29 appendix/appendixB_99.Rmd
script

--- 2017-12-22 appendix/appendixB.Rmd
@@
 ```{r}
-options(mc.cores = parallel::detectCores()) # Reserve ~ 10 GB per core
+options(mc.cores = 3) # Reserve ~ 10 GB per core
 log_dir <- "pomdp_historical"
@@
              precision = 2e-6,
-             timeout = 20000,
+             timeout = 70000,
              log_dir = log_dir,

### 2017-12-29 appendix/appendix_harvestfirst_99.Rmd
script

--- 2017-12-28 appendix/appendix_harvestfirst.Rmd
@@
 reward_fn <- function(x,h) pmin(x,h)
-discount <- 0.95
+discount <- 0.99
 ```
@@
 
+```{r}
+options(mc.cores = 3)
+```
+
 ```{r eval=TRUE}
@@
              precision = 0.00000002,
-             timeout = 15000,
+             timeout = 60000,
              log_dir = log_dir,

### 2017-12-29 appendix/solve_pomdp.R

@@ -1,69 +1,10 @@
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
-library(sarsop)       # the main POMDP package
-library(tidyverse)    # for munging and plotting
-library(parallel)
-options(mc.cores = 4) # Reserve ~ 10 GB per core
-log_dir <- "pomdp_intro"
 
-r <- 0.75
-K <- 1
-## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
-f <- function(x, h){ 
-  x + x * r * (1 - x / K) - pmin(x,h)
-}
-reward_fn <- function(x,h) pmin(x,h)
-discount <- 0.99
-
-## Discretize space
-states <- seq(0,2, length=100)
-actions <- states
-observations <- states
-
-## All parameter values combinations for which we want solutions
-meta <- expand.grid(sigma_g = c(0.01, 0.05, 0.1), 
-                    sigma_m = c(0, 0.1, 0.2),
-                    stringsAsFactors = FALSE)
-
-## Create the models
-models <- 
-  parallel::mclapply(1:dim(meta)[1], 
-                     function(i){
-                       fisheries_matrices(
-                         states = states,
-                         actions = actions,
-                         observed_states = observations,
-                         reward_fn = reward_fn,
-                         f = f,
-                         sigma_g = meta[i,"sigma_g"][[1]],
-                         sigma_m = meta[i,"sigma_m"][[1]],
-                         noise = "normal")
-                     })
-
-## log metadata associated with each run
-
-dir.create(log_dir)
-
-## POMDP solution (slow, >10,000 seconds per loop memory intensive)
-system.time(
-  alphas <- 
-    parallel::mclapply(1:length(models), 
-    function(i){
-      log_data <- data.frame(model = "gs", 
-                             r = r, 
-                             K = K, 
-                             sigma_g = meta[i,"sigma_g"][[1]], 
-                             sigma_m = meta[i,"sigma_m"][[1]], 
-                             noise = "normal")
-      
-      sarsop(models[[i]]$transition,
-             models[[i]]$observation,
-             models[[i]]$reward,
-             discount = discount,
-             precision = 2e-6,
-             timeout = 10000,
-             log_dir = log_dir,
-             log_data = log_data)
-    })
-)
+rmarkdown::render("appendix/appendix_harvestfirst.Rmd")
[…]

### 2017-12-29 appendix/solve_pomdp.R
clear environment to free space as we go

@@ -1,10 +1,15 @@
 
 rmarkdown::render("appendix/appendix_harvestfirst.Rmd")
+rm(list=ls())
 rmarkdown::render("appendix/appendixB.Rmd")
+rm(list=ls())
 #rmarkdown::render("appendix/appendix.Rmd")      ## running on chameleon
-
+rm(list=ls())
 
 ## super slow
 #rmarkdown::render("appendix/appendix_99.Rmd")   ## running on 24-core
+rm(list=ls())
 rmarkdown::render("appendix/appendix_harvestfirst_99.Rmd")
+rm(list=ls())
 rmarkdown::render("appendix/appendixB_99.Rmd")
+rm(list=ls())

### 2017-12-29 appendix/appendix.Rmd
appendix discount 99, missing 1 scenario

--- 2017-12-28 appendix/appendix.Rmd
@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-devtools::install("..")
+#devtools::install("..", dep=TRUE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
+library(tidyverse)
+ggplot2::theme_set(theme_bw()) # simpler theme if you do not have these fonts
+```
+
+```{r suggested_libs, eval=FALSE}
 knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
-library(tidyverse)
+library(hrbrthemes)
+library(extrafont)
 library(gridExtra)
@@
 colours <- ptol_pal()(2)
-```
-
-```{r suggested_libs}
-library(hrbrthemes)
-library(extrafont)
 ##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
@@
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
-#ggplot2::theme_set(theme_bw()) # simpler theme if you do not have these fonts
 ```

### 2017-12-29 appendix/appendix.Rmd
longer runs at 0.95 for smoother results

--- 2017-12-28 appendix/appendix.Rmd
@@
 
-```{r eval=TRUE}
+```{r}
+options(mc.cores=1)
+```
+
+```{r eval=FALSE}
 dir.create(log_dir)
-options(mc.cores=9)
 
@@
              precision = 0.00000002,
-             timeout = 15000,
+             timeout = 25000,
              log_dir = log_dir,
@@
 
+set.seed(12345)
 static_sims <- 
@@
 unif_prior <- rep(1, length(states)) / length(states)
+set.seed(12345)
+
 pomdp_sims <- 
-  map2_dfr(models, alphas, function(.x, .y){
+  map2_dfr(models[as.integer(meta$scenario)], alphas, function(.x, .y){
              sim_pomdp(.x$transition, .x$observation, .x$reward, discount, 
@@
 ```{r}
-i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0) %>% pull(scenario) %>% as.integer()
+i <- meta %>% filter(sigma_g == 0.15, sigma_m ==0) %>% pull(scenario) %>% as.integer()
 m <- models[[i]]
@@
   data.frame(state = states, 
-             policy = "sdp 0.1",
+             policy = "sdp 0.15",
              harvest = actions[mdp$policy],

### 2018-01-03 appendix/solve_pomdp.R
tidy up a bit

@@ -1,69 +1,0 @@
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
-library(sarsop)       # the main POMDP package
-library(tidyverse)    # for munging and plotting
-library(parallel)
-options(mc.cores = 4) # Reserve ~ 10 GB per core
-log_dir <- "pomdp_intro"
-
-r <- 0.75
-K <- 1
-## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
-f <- function(x, h){ 
-  x + x * r * (1 - x / K) - pmin(x,h)
-}
-reward_fn <- function(x,h) pmin(x,h)
-discount <- 0.99
-
-## Discretize space
-states <- seq(0,2, length=100)
-actions <- states
-observations <- states
-
-## All parameter values combinations for which we want solutions
-meta <- expand.grid(sigma_g = c(0.01, 0.05, 0.1), 
-                    sigma_m = c(0, 0.1, 0.2),
-                    stringsAsFactors = FALSE)
-
-## Create the models
-models <- 
-  parallel::mclapply(1:dim(meta)[1], 
-                     function(i){
-                       fisheries_matrices(
-                         states = states,
-                         actions = actions,
-                         observed_states = observations,
-                         reward_fn = reward_fn,
-                         f = f,
-                         sigma_g = meta[i,"sigma_g"][[1]],
-                         sigma_m = meta[i,"sigma_m"][[1]],
-                         noise = "normal")
-                     })
-
-## log metadata associated with each run
-
-dir.create(log_dir)
-
-## POMDP solution (slow, >10,000 seconds per loop memory intensive)
-system.time(
-  alphas <- 
-    parallel::mclapply(1:length(models), 
-    function(i){
-      log_data <- data.frame(model = "gs", 
-                             r = r, 
-                             K = K, 
-                             sigma_g = meta[i,"sigma_g"][[1]], 
-                             sigma_m = meta[i,"sigma_m"][[1]], 
-                             noise = "normal")
-      
-      sarsop(models[[i]]$transition,
-             models[[i]]$observation,
-             models[[i]]$reward,
-             discount = discount,
-             precision = 2e-6,
-             timeout = 10000,
-             log_dir = log_dir,
-             log_data = log_data)
-    })
-)
-
-

### 2018-01-04 appendix/appendix_harvestfirst.Rmd
discount at 0.95
WIP

@@
 options(mc.cores=parallel::detectCores())
+log_dir <- "../data/observe_harvest_recruit" # Store the computed solution files here
+
 ```
@@
 escapement_policy <- function(x) pmax(x - S_star,0)
+
 x0 <- K/6
 Tmax = 20
-det_sim <- function(policy, f, x0, Tmax){
+do_det_sim <- function(policy, f, x0, Tmax){
     action <- state <- obs <- as.numeric(rep(NA,Tmax))
@@
 
-sims <- 
+det_sims <- 
   list(msy = msy_policy, 
@@
        det = escapement_policy) %>% 
-  map_df(det_sim, 
+  map_df(do_det_sim, 
          f, x0, Tmax, 
          .id = "method")
-sims %>%
+
+```
+
+With no stochasticity, MSY leads to the same long-term stock size as under the constant escapement rule, but takes longer to get there.
+This level is essentially $B_{MSY}$, though because the model considered here implements events in the order: *observe, harvest, recruit*; rather than *observe, recruit, harvest*, we see the stock at the pre-harvest size of $B_{MSY} + H_{MSY}$ (that is, $K/2 + rK/4$).
+More conservative rules, such as a harvest set to 80% of MSY result in faster recovery of the stock than under MSY, but slower than under constant escapement.
+Due to the reduced maximum harvest rate, such rules lead to stock returning to a value higher than $B_${MSY}$.
+
+```{r}
+det_sims %>%
   ggplot(aes(time, state, col=method)) + 
@@
 policies <- data.frame(
-  det = index(pmax(f(states,0) - S_star,0), actions), # assumes obs->recruit->harvest, f(x_t) - h_t
- # det = index(pmax(states - S_star,0),     actions), # assumes obs->harvest->recruit, f(x_t - h_t)
-  msy = index(states * F_MSY,               actions),
-  pgy = index(states * F_PGY,               actions))
+#  det = index(pmax(f(states,0) - S_star,0), actions), # obs,recruit,harv, f(x_t) - h_t
+  det = index(escapement_policy(states),     actions), # obs,harv,recruit, f(x_t - h_t)
+  msy = index(msy_policy(states),               actions),
+  pgy = index(pgy_policy(states),               actions))
 ```
@@
         mutate(scenario  = as.character(1:length(sigma_m)))
-log_dir <- "appendix_alphas_harvestfirst" # Store the computed solution files here
 meta
@@
 
-```{r eval=TRUE}
+```{r eval=FALSE}
 dir.create(log_dir)
@@
 ```{r}
-meta <- meta_from_log(data.frame(model="gs"), log_dir) %>% 
+meta <- meta_from_log(data.frame(model="gs", discount=discount), log_dir) %>% 
   mutate(scenario = as.character(scenario)) %>%
@@
 ```{r static_sims}
+set.seed(12345)
+
 Tmax <- 100
@@
 reps <- 100
-
 static_sims <- 
@@
 ```{r pomdp_sims}
-# slower!
+set.seed(12345)
+
 unif_prior <- rep(1, length(states)) / length(states)
@@
 
-write_csv(sims, "appendix_harvestfirst_files/sims.csv")
+sims <- sims %>% select(time, state, rep, method, sigma_m, sigma_g)
+
+sim_col_types <- paste0(gsub("^(\\w).*", "\\1", sapply(unname(sims), class)), collapse = "")
+write_csv(sims, file.path(log_dir, "sims.csv"))
 ```
@@
 ```{r}
+sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = sim_col_types)
+
 sims %>%
@@
   geom_line() + 
-  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.2) +
-  facet_grid(sigma_m ~ sigma_g)
+  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
+  facet_grid(sigma_m ~ sigma_g, 
+             labeller = label_bquote(sigma[m] == .(sigma_m),
+                                     sigma[g] == .(sigma_g))) + 
+  coord_cartesian(ylim = c(0, 1))
+
 ```
@@
 
-write_csv(policy_table, "appendix_files/policy_table.csv")
 ```
@@
 pomdp_policies <- df %>% 
-  select(-value) %>% 
-  rename(uniform = policy) %>% 
+  select(state, uniform = policy, low, medium, high) %>% 
   gather(policy, harvest, -state) %>%
@@
   select(state, policy, harvest, escapement)
+
+all_policies <- bind_rows(policy_table, pomdp_policies)
+
+write_csv(all_policies, file.path(log_dir, "all_policies.csv"))
 
@@
 ```
+
+```{r}
+all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
+all_policies %>% 
+  filter(policy %in% c("low", "medium", "det", "msy")) %>%
+  mutate(policy = fct_recode(policy,
+                             "MSY" = "msy",
+                             "Constant escapement" = "det",
+                             "POMDP: low prior" = "low",
+                             "POMDP: medium prior" = "medium"
+                             )) %>%
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```
+
+```{r}
+all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
+all_policies %>% 
+  filter(policy %in% c("low", "medium", "det", "msy")) %>%
+  mutate(policy = fct_recode(policy,
+                             "MSY" = "msy",
+                             "Constant escapement" = "det",
+                             "POMDP: low prior" = "low",
+                             "POMDP: medium prior" = "medium"
+                             )) %>%
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line()  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+```

### 2018-01-04 manuscripts/manuscript.Rmd
WIP

--- 2017-12-14 manuscripts/manuscript.Rmd
@@
 output: rticles::elsevier_article
+layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column
+
 ---
+
+```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
+library(tidyverse)
+library(gridExtra)
+library(ggthemes)
+colours <- ptol_pal()(4)
+berkeley_colors <- c("#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")
+```
+
+```{r suggested_libs}
+library(hrbrthemes)
+library(extrafont)
+##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
+extrafont::loadfonts()
+hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
+ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+```
+
+```{r}
+log_dir <- "../data/observe_harvest_recruit" # solution files here
+```
 
@@
 
-- Replicate simulations
+<!--
+PGY, MSY, Det, MDP (sigma_g), POMDP (sigma_g, sigma_m)
+-->
 
-Moderate / modest sigma_g, sigma_m
+### Replicate simulations
 
-- non-pomdp policies
+Figure 1 shows average fish biomass
 
-- pomdp policies
+```{r fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
+Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement, maximum sustainable yield (MSY), and the porposed partially observed Markov decision process (POMDP) method.
+Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.
+Environmental stochasticity is fixed a standard deviation of 0.15 in each panel.
+Carrying capacity K normalized to 1, r = 0.75.
+Additional environmental noise levels shown in supplement."}
+sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnn")
+sims %>%
+  filter(method != "pgy", sigma_g == "0.15") %>%
+  mutate(method = fct_recode(method, 
+                             "Constant escapement" = "det",
+                             "MSY" = "msy",
+                             "POMDP" = "pomdp")) %>%
+  group_by(time, method, sigma_m) %>%
+  summarise(mean = mean(state), sd = sd(state)) %>%
+  ggplot(aes(time, mean, col=method, fill=method)) + 
+  geom_line() + 
+  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
+  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
+  coord_cartesian(ylim = c(0, 1)) + 
+  theme(legend.position = "bottom") + 
+  ylab("Mean biomass")
+```
 
-- historical data.
+POMDP policies depend on the prior belief.
 
 ```{r}
+all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
+all_policies %>% 
+  filter(policy %in% c("low", "medium", "det", "msy")) %>%
+  mutate(policy = fct_recode(policy,
+                             "MSY" = "msy",
+                             "Constant escapement" = "det",
+                             "POMDP: low prior" = "low",
+                             "POMDP: medium prior" = "medium"
+                             )) %>%
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line(lwd=1)  + 
+  coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
+```
+
+### Comparisons to historical data
+
+```{r eval=FALSE}
+sims %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
+  ggplot(aes(year, stock, col=model)) +
+  geom_line(lwd=1) +
+  scale_color_manual(values = c("black", "grey", berkeley_colors) ) +
+  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
 ```

### 2018-01-04 manuscripts/manuscript.Rmd
tweaking extrafont issues

@@
 library(hrbrthemes)
-library(extrafont)
+#library(extrafont)
 ##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
-extrafont::loadfonts()
+#extrafont::loadfonts()
 hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 

### 2018-01-04 appendix/appendix.Rmd
theme should work once container updates

@@
 options(mc.cores=parallel::detectCores())
-```
-
-```{r suggested_libs, include=FALSE, eval=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
-devtools::install("..", dep=TRUE)
-library(gridExtra)
-library(ggthemes)
-colours <- ptol_pal()(2)
-library(hrbrthemes)
-library(extrafont)
-##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
-hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
-extrafont::loadfonts()
-ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
 ```

### 2018-01-04 manuscripts/manuscript.Rmd
theme should work once container updates

--- 2018-01-04 manuscripts/manuscript.Rmd
@@
 library(hrbrthemes)
+library(ggplot2)
+library(Cairo)
 library(extrafont)
-##extrafont::font_import(prompt=FALSE)    # One-time use to set up fonts, can be slow.
 extrafont::loadfonts()
-hrbrthemes::import_roboto_condensed()     # One-time use to set up fonts 
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
@@
   mutate(policy = fct_recode(policy,
+                             "Constant escapement" = "det",
                               "MSY" = "msy",
-                             "Constant escapement" = "det",
                              "POMDP: low prior" = "low",
@@
   geom_line(lwd=1)  +
+  scale_color_manual(values = c("black", berkeley_colors) ) +
   coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
@@
 sims %>%
-  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed")) %>%
   ggplot(aes(year, stock, col=model)) +

### 2018-01-04 manuscripts/manuscript.Rmd
WIP

@@
   We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.
-Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to the intuition of the precautionary principle.
-Through the use of cutting-edge point-based POMDP methods, we demonstrate that the standard management theory which ignores measurement uncertainty may lead to over-exploitation of fisheries and increased probability of economic and ecological collapse.
-In contrast, POMDP solutions which account for measurement error are overall more conservative, avoiding over-exploitation while generating higher economic value.
+Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to intuition and the precautionary principle.
+Through the use of cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.
+In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning.
 
@@
 
-```{r suggested_libs}
+```{r suggested_libs, message=FALSE}
 library(hrbrthemes)
@@
 
-The area of fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.
+Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.
 Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.
+In particular, the approach here illustrates the limitations of both over-simple optimal control approaches and rule-of-thumb precautionary approaches, and offers instead a more general strategy for dealing with measurement uncertainty adaptively.
+
 We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.
@@
 We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.
-To facilitate tractability and intepretation, we will consider a simple and well-studied model of population growth (stock recruitment) in marine fisheries, the Graham-Schaefer model:
+To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1974; @Gordon1974] of logistic population growth (stock recruitment),
 
-Though this and very similar models are widely used in large scale analyses across diverse stocks (e.g.
-[@Costello2016; @Britten2017]), and common in other conservation contexts where richer models are unavailable, neither the existing approaches nor the one we introduce here are limited to such one-dimensional stochastic models with known parameters.
-In practice, this model may include age or stage structure and will reflect uncertainty not only in the stochastic recruitment but in the parameters themselves, usually as the result of a heirarchical model estimation process.
-All that is required of the model is that it defines the probability of transitioning to a new state given the current state (here, simply the stock size) and the action taken (here, just the harvest).
-Uncertainty either in the parameters, in the model, or inherent in the stochastic process can all be reflected in such probabilities.
+$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$
 
+where $x_t$ is the current stock size, $h_t$ the harvest chosen that year, with parameters $r$ giving the individual growth rate, and $K$ the carrying capacity and $\varepsilon$ representing stochastic recruitment in a variable environment.
+We will assume for simplicity $\varepsilon \sim \mathcal{N}(1, \sigma_g)$.
+
+This and similar models are widely used in large scale analyses across diverse stocks (e.g.
+[@Costello2016; @Britten2017]), and form the basis for much of bioeconomic theory [@Clark1990].
+As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arsing only because of some particular subtle assumption hidden in a more complex model.
+
+However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
+Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
+@].
+As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.
+
+## The decision problem
+
+We seek to determine the sequence of actions $h_t$ for $t \in [1, ...
+\infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.
+We will denote the discount rate $\delta$.
+For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax.
 Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,
 
-$$y_t = Z_m x_t$$
-for some measurement error $Z_m$.
-Given the estimated recruitment model along with this estimate of the stock size, the manager must set a harvest quota $h_t$ for that year.
-The methods we consider here all differ primarily in their handling of these uncertainties.
-If this system were completely deterministic so that there was no uncertainty, all would produce identical long term policies and thus achieve identical outcomes.
+$$y_t = \xi_t x_t$$
 
-In the case of no uncertainty, the optimal harvest strategy is to reduce the population to the size at which it obtains the maximum growth rate, known as Biomass at Maximum Sustainable Yield, $B_{MSY}$ (this is adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
-For the Graham-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$, with a corresponding annual harvest of $H_{MSY} = r K / 4$.
+Given an estimated recruitment model along with this measurement of the stock size, the manager must set a harvest quota $h_t$ for that year in order to maximize their expected long term (net) present value:
 
-This equilibrium analysis says nothig about the appropriate harvest at stock sizes above or below $B_{MSY}$, which highlights some essential differences between the optimal control policy and more typical fisheries management.
-The optimal policy is impatient: it seeks to achieve the optimal yield as quickly as possible.
-Below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, a harvest greater than $H_{MSY}$ can be optimal for that year.
-This strategy is known as "constant escapement," since a constant sock size $B_{MSY}$ escapes harvest each year.
-In practice, it is more common to pursue a policy of constant fishing effort (fishing mortality), $F_{MSY} = H_{MSY} / B_{MSY}$.
-Continuing to harvest at $F_{MSY}$ when a stock is above or below $B_{MSY}$ is economically inefficient and causes depleted stocks to rebuild more slowly, but, in the absense of uncertainty, ensures a return in the long run to stock sizes of $B_{MSY}$ and harvests of $H_{MSY}$.
-These differences will be important in understanding how these approaches perform once uncertainty is introduced.
+$$\max_{ \lbrace h_t \rbrace} \sum_{t=0}^{\infty} \delta x_t $$
+
+A few subtlies arise in this seemlingly simple problem statement that we must address.
+First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?
+Following convention in the optimal control literature (e.g.
+[@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.
+
+A second subtly is in the nature of the decision problem itself: the optimization problem must always consider the set ${h_t}$ all future actions simultaneously, not merely the next action.
+This is a game of chess, where it can be optimal to make costly short term actions (such as closing the fishery) if they pay off in the long run.
+This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a single harvest quota $H_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.
+Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.
+
+## Current theory and current practice
+
+@Gordon1954 and @Schaefer1954 independently showed in the same year that the maximum sustainable yield for the model bearing their names is achieved by reducing the stock to the  population to the size at which it obtains the maximum growth rate which is known as Biomass at Maximum Sustainable Yield, $B_{MSY}$.
+For the Gordon-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$.
+They observed that fishing at a *constant yield* (that is, an individual fish mortality, $F$, such that harvest $H = F \cdot B$) of $F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}$)  will eventually lead to a population that converges to the biomass $B_{MSY}$ and produces the maximum sustainable harvest, $H_{MSY} = r K / 4$ in this model.
+From here, theory and practice diverge.
+
+This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis, and does not solve the time-dependent optimization problem above.
+In particular, maximum sustainable yield will not be optimal whenever the stock is away from $B_{MSY}$.
+@Clark1973 shows that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*.
+The time-depednent optimal stategy is impatient: it seeks to achieve the optimal yield as quickly as possible.
+Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, mortalities greater than $F_{MSY}$ can be optimal to bring the stock back to $B_{MSY}$ more quickly.
+(This intution must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
+This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.
+While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.
 
@@
 
-As discussed above, absent the short-cut provided by Reed's theorems, solving for the optimal management policy in a stochastically fluctuating population involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP).
+As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP).
 Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).

### 2018-01-06 appendix/appendixB.Rmd
updated appendix B alphas

@@
 options(mc.cores = parallel::detectCores()) # Reserve ~ 10 GB per core
-log_dir <- "pomdp_historical"
+log_dir <- "appendixB_alphas"
 ```
@@
 
-```{r eval=FALSE}
+```{r eval=TRUE}
 dir.create(log_dir)
+options(mc.cores = 1)
 
@@
              precision = 2e-6,
-             timeout = 20000,
+             timeout = 25000,
              log_dir = log_dir,

### 2018-01-06 manuscripts/manuscript.Rmd

@@
 bibliography: citations.bib
+csl: the-american-naturalist.csl
 output: rticles::elsevier_article
@@
 
-Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1957], thanks to its global relevance, long history, and readily available data.
+Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.
 Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.
@@
 
-## Model
+# Model
 
@@
 
-## The decision problem
+# The Decision Problem
 
@@
 \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.
-We will denote the discount rate $\delta$.
+We will denote the discount rate $\gamma$.
 For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax.
@@
 
-$$\max_{ \lbrace h_t \rbrace} \sum_{t=0}^{\infty} \delta x_t $$
+$$\max_{ \lbrace h_t \rbrace} \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$
 
@@
 
-## Current theory and current practice
+<!-- MDP, POMDP -->
+
+# Current theory and practice
 
@@
 
-## Pardoxes of uncertainty
-
 The literature of optimal harvests under uncertainty is one of paradoxes, which can only be resolved by a more comprehensive treatment of measurement uncertainty.
 
-### Reed's Paradox: $S = D$
+## Reed's Paradox: $S = D$
 
@@
 
-### Clark's Paradox: $M < D$
+## Clark's Paradox: $M < D$
 
@@
 > For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
-of the problem increases markedly if this assumption is
-relaxed.)
+of the problem increases markedly if this assumption is relaxed.)
 
-The resulting solutions only deepened the paradox, rather than recovering a more cautious policy in response to the uncertainty, @Clark1986 finds policies that are often *less* cautious:
+By leveraging algorithms (and also computational hardware) developed over two decades later are we able to crack this mathematical difficulty Clark refers to directly without this "simplyfying assumption."  Unforunately for @Clark1986, this "simplifying assumption" they made instead only deepened the paradox, leading them to find policies that become *less* cautious as uncertainty increases:
 
@@
 
-Using different but still implicitly approximate results nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observe with some concern:
+Using a different but still flawed assumption nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observed with some concern:
 
@@
 Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.
+
 First we will introduce the more heuristic alternative to addressing uncertainty, which more accurately reflects both intuition and actual management.
 
-## Uncertainty in practice
+## Uncertainty in Practice
 
@@
 Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
-For those interested in exploring or applying these methods, we have developed a preliminary R package, @sarsop-pkg, which can be used to replicate any of the results presented here and explore other variations.
+For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.
 
-## Results
+# Results
 
@@
 -->
-
-### Replicate simulations
 
@@
 
-### Comparisons to historical data
-
 ```{r eval=FALSE}
@@
 
-## Discussion
+# Discussion
 
@@
 
-### A practical implementation?
+## A practical implementation?
 

### 2018-01-06 appendix/appendixB.Rmd
updated figures under appendix B

@@
 
-```{r}
+```{r message=FALSE}
 meta <- meta_from_log(data.frame(model="gs"), log_dir)  %>% 
+  mutate(scenario = as.character(scenario)) %>%
   left_join(
@@
     select(meta, sigma_m, commonname, scenario),   
-    by = c("sigma_m", "commonname")) %>% 
+    by = c("scenario", "sigma_m", "commonname")) %>% 
   arrange(scenario)
@@
 ```{r}
+set.seed(123456)
 pomdp_sims <- 

### 2018-01-07 manuscripts/manuscript.Rmd
WIP

@@
 ```{r}
-log_dir <- "../data/observe_harvest_recruit" # solution files here
+log_dir <- "../data" # solution files here
 ```
@@
 
-## Uncertainty in Practice
+## Uncertainty in Practice: TAC
 
 Fortunately, management practice has so far ignored this paradoxical suggestion that uncertainty should be met with greater rather than reduced fishing quotas.
-Instead, most well-managaged fisheries operate on rules that are more the result of common sense than mathematical optimization.
-While details differ, the basic approach can be thought of as starting with the deterministic notion of $B_{MSY}$ and corresponding optimal fishing level $F_{MSY}$ as a way of determining an Overfishing Limit, (OFL), which is then adjusted to some lower level, Allowable Biological Catch (ABC) to reflect uncertainty in the stock assessment.
-ABC is then adjusted lower once again, to determine a Total Allowable Catch, (TAC), reflecting uncertainty in reported catch or other uncertainty.
-To capture this process, we will consider a policy in which the harvest quota is set at a TAC of 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
-We will refer to both MSY policies and this more cautious adjustment of the MSY policy as "constant mortality" policies, since they are defined by a constant fishing mortality, $F$, rather than by a constant escapement, $S$.
-
-Note that compared to Reed's constant escapement, this policy still permits some level of harvest even when stocks are measured below $B_{MSY}$.
+Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1989].
+Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Botsford1997].
+Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.
+Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Overfishing Limit, (OFL).
+Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment.
+To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.
+To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
+To distinguish this approach from MSY, we will refer to this approach as a TAC policy.
+This more closely represents a heuristic [e.g.
+@Hilborn2010] or resilience-based approach [sensu @Fischer2012] than an optimization-based policy.
+Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.
+Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.
 
@@
 <!--
-PGY, MSY, Det, MDP (sigma_g), POMDP (sigma_g, sigma_m)
+TAC, MDP, POMDP (sigma_g, sigma_m)
 -->
@@
 Additional environmental noise levels shown in supplement."}
-sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnn")
+sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
 sims %>%
-  filter(method != "pgy", sigma_g == "0.15") %>%
+  filter(method != "msy", sigma_g == "0.15") %>%
   mutate(method = fct_recode(method, 
                              "Constant escapement" = "det",
-                             "MSY" = "msy",
+                             "TAC" = "pgy",
                              "POMDP" = "pomdp")) %>%
@@
 ```{r}
-all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
+all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
 all_policies %>% 
-  filter(policy %in% c("low", "medium", "det", "msy")) %>%
+  filter(policy %in% c("low", "medium", "det", "pgy")) %>%
   mutate(policy = fct_recode(policy,
                              "Constant escapement" = "det",
-                              "MSY" = "msy",
+                             "TAC" = "pgy",
                              "POMDP: low prior" = "low",
@@
 
-```{r eval=FALSE}
-sims %>%
-  filter(model %in% c("biomass", "catch", "POMDP", "Reed")) %>%
+```{r message=FALSE, warning=FALSE}
+read_csv(file.path(log_dir, "ram_ex/ram_ex.csv")) %>% 
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
+   mutate(model = fct_recode(model,
+                             "Constant escapement" = "Reed",
+                             "TAC" = "pgy"
+                             )) %>%
   ggplot(aes(year, stock, col=model)) +
   geom_line(lwd=1) +
-  scale_color_manual(values = c("black", "grey", berkeley_colors) ) +
+  scale_color_manual(values = c("black", "grey", berkeley_colors)) +
   facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  

### 2018-01-08 manuscripts/manuscript.Rmd
drone test

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, dev="CairoPDF", fig.width=7, fig.height=7)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7)
 library(tidyverse)
@@
 ```{r suggested_libs, message=FALSE}
+#knitr::opts_chunk$set( dev="CairoPDF")
+knitr::opts_chunk$set( dev="cairo_pdf")
 library(hrbrthemes)
+hrbrthemes::import_roboto_condensed() 
 library(ggplot2)
@@
 ```{r message=FALSE, warning=FALSE}
-read_csv(file.path(log_dir, "ram_ex/ram_ex.csv")) %>% 
+ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
+ram_ex %>%
   filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%

### 2018-01-12 manuscripts/manuscript.Rmd
a few more citations

@@
 We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.
-To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1974; @Gordon1974] of logistic population growth (stock recruitment),
+To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),
 
@@
 Fortunately, management practice has so far ignored this paradoxical suggestion that uncertainty should be met with greater rather than reduced fishing quotas.
-Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1989].
-Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Botsford1997].
+Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].
+Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].
 Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.
@@
 This more closely represents a heuristic [e.g.
-@Hilborn2010] or resilience-based approach [sensu @Fischer2012] than an optimization-based policy.
+@Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.
 Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.
@@
 
-```{r fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
+```{r fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
 Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement, maximum sustainable yield (MSY), and the porposed partially observed Markov decision process (POMDP) method.
@@
 Carrying capacity K normalized to 1, r = 0.75.
-Additional environmental noise levels shown in supplement."}
+Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material."}
 sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
@@
 
-```{r}
+```{r fig.width=3.5,fig.height=3.5, fig.cap=""}
 all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE, fig.width=3.5, fig.height = 7, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
+POMDP solution assumes a measurement error of 10%."}
 ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))

### 2018-01-12 appendix/appendixB.Rmd
results discussion WIP tweaking plots

@@
 knitr::opts_chunk$set(message=FALSE, comment="", cache=TRUE)
+```
+
+```{r plotting}
+library(gridExtra)
+library(ggthemes)
+library(ggplot2)
+library(Cairo)
+library(extrafont)
+library(hrbrthemes)
+
+extrafont::loadfonts()
+knitr::opts_chunk$set( dev="cairo_pdf")
+hrbrthemes::import_roboto_condensed() 
+ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+
+palette <- ptol_pal()(6)
+palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
+
+colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
+                    c("TAC", "POMDP", "MSY", "CE", 
+                      "POMDP: low prior",
+                      "POMDP: medium prior",
+                      "biomass",
+                      "catch"))
 ```
@@
 options(mc.cores = parallel::detectCores()) # Reserve ~ 10 GB per core
-log_dir <- "appendixB_alphas"
+log_dir <- "../data/ram_ex"
 ```
@@
 meta <- expand.grid(commonname = pars$commonname, 
-                    sigma_m = c(0, 0.1, 0.2),
+                    sigma_m = c(0, 0.1, 0.15),
                     stringsAsFactors = FALSE) %>%
@@
                         "0.1" = "POMDP", 
-                        "0.2" = "POMDP_0.2")) %>%
+                        "0.15" = "POMDP_0.15")) %>%
   select(-sigma_m) %>%
@@
 
-write_csv(sims,"appendixB_files/sims.csv")
+write_csv(sims,file.path(log_dir, "ram_ex.csv"))
 ```

### 2018-01-12 appendix/appendix_harvestfirst.Rmd
results discussion WIP tweaking plots

@@
 library(parallel)
+```
+
+```{r plotting}
+library(gridExtra)
+library(ggthemes)
+library(ggplot2)
+library(Cairo)
+library(extrafont)
+library(hrbrthemes)
+
+extrafont::loadfonts()
+knitr::opts_chunk$set( dev="cairo_pdf")
+hrbrthemes::import_roboto_condensed() 
+ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+
+palette <- ptol_pal()(6)
+palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
+
+colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
+                    c("TAC", "POMDP", "MSY", "CE", 
+                      "POMDP: low prior",
+                      "POMDP: medium prior",
+                      "biomass",
+                      "catch"))
 ```
@@
 x0 <- K/6
-Tmax = 20
+Tmax = 100
 do_det_sim <- function(policy, f, x0, Tmax){
@@
 
+write_csv(det_sims, "../data/observe_harvest_recruit/det_sims.csv")
 ```
@@
 
-```{r}
+```{r fig.width=3.5, fig.height=3.5}
 det_sims %>%
+  mutate(method = fct_recode(method, 
+                             "CE" = "det",
+                             "TAC" = "pgy",
+                             "MSY" = "msy")) %>%
   ggplot(aes(time, state, col=method)) + 
-  geom_line() + 
-  geom_hline(aes(yintercept=B_MSY), lwd=1)
+  geom_line(lwd=1) + 
+  scale_color_manual(values = colors) +
+  coord_cartesian(ylim = c(0, 1)) + 
+  theme(legend.position = "bottom") + 
+  ylab("Mean biomass")
+```
 

### 2018-01-12 manuscripts/manuscript.Rmd
results discussion WIP tweaking plots

@@
 library(ggthemes)
-colours <- ptol_pal()(4)
-berkeley_colors <- c("#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")
+palette <- ptol_pal()(6)
+palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
 ```
@@
 
-Figure 1 shows average fish biomass
+```{r load_data, message=FALSE, warning=FALSE}
+sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
+det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
+all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
+ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
+
+colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
+                    c("TAC", "POMDP", "MSY", "CE", 
+                      "POMDP: low prior",
+                      "POMDP: medium prior",
+                      "biomass",
+                      "catch"))
+
+```
+
+Figure 1 shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
+Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.
+Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.
+In each panel shown in Figure 1 stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
+In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.
+The first panel of Figure 1 shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.
+Without mesasurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
+As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.
+Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.
+
+Figure 1 confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.
+In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.
+Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC stategy rather than the CE strategy that lead to the highest overfishing rate.
+In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.
+
+<!-- Show results of POMDP when over-estimating! -->
 
 ```{r fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
-Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement, maximum sustainable yield (MSY), and the porposed partially observed Markov decision process (POMDP) method.
+Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement (CE), Total allowable catch (TAC = 80% MSY), and the proposed partially observed Markov decision process (POMDP) method.
 Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.
@@
 Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material."}
-sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
 sims %>%
@@
   mutate(method = fct_recode(method, 
-                             "Constant escapement" = "det",
+                             "CE" = "det",
                              "TAC" = "pgy",
@@
   summarise(mean = mean(state), sd = sd(state)) %>% 
+  mutate(sigma_m = as.character(sigma_m)) ->
+fig1_df
+
+fig1_df %>%
   ggplot(aes(time, mean, col=method, fill=method)) + 
-  geom_line() + 
+  geom_line(lwd=1) + 
   geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
   facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
+  scale_color_manual(values = colors) +
+  scale_fill_manual(values = colors) +
   coord_cartesian(ylim = c(0, 1)) + 
@@
 
-```{r fig.width=3.5,fig.height=3.5, fig.cap=""}
-all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
+```{r fig.width=7,fig.height=3.5, fig.cap=""}
 all_policies %>% 
@@
   mutate(policy = fct_recode(policy,
-                             "Constant escapement" = "det",
+                             "CE" = "det",
                              "TAC" = "pgy",
@@
                              "POMDP: medium prior" = "medium"
-                             )) %>%
+                             )) -> policy_df
+policy_df %>%
   ggplot(aes(state, harvest, col=policy)) + 
   geom_line(lwd=1)  +
-  scale_color_manual(values = c("black", berkeley_colors) ) +
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
+```
+```{r fig.width=7,fig.height=3.5, fig.cap=""}
+policy_df %>%
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line(lwd=1)  +
+  scale_color_manual(values = colors) +
   coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
@@
 
-```{r message=FALSE, warning=FALSE, fig.width=3.5, fig.height = 7, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
+```{r message=FALSE, warning=FALSE, fig.width=7, fig.height = 3.5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
 POMDP solution assumes a measurement error of 10%."}
-ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
 ram_ex %>%
@@
    mutate(model = fct_recode(model,
-                             "Constant escapement" = "Reed",
+                             "CE" = "Reed",
                              "TAC" = "pgy"
@@
   geom_line(lwd=1) +
-  scale_color_manual(values = c("black", "grey", berkeley_colors)) +
+  scale_color_manual(values = colors) +
   facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  

### 2018-01-16 appendix/appendix_harvestfirst.Rmd
WIP

@@
 sims <- bind_rows(static_sims, pomdp_sims) %>%
-    left_join(meta) %>%    ## combine output with scenarios
-    mutate(state = states[state], action = actions[action]) # index -> value
-
-sims <- sims %>% select(time, state, rep, method, sigma_m, sigma_g)
+    left_join(meta) %>%  ## include scenario information (sigmas; etc)
+    mutate(state = states[state], action = actions[action]) %>%
+    select(time, state, rep, method, sigma_m, sigma_g, value)
 

### 2018-01-16 manuscripts/manuscript.Rmd
WIP

@@
 library(gridExtra)
-library(ggthemes)
-palette <- ptol_pal()(6)
-palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
 ```
 
-```{r suggested_libs, message=FALSE}
-#knitr::opts_chunk$set( dev="CairoPDF")
-knitr::opts_chunk$set( dev="cairo_pdf")
-library(hrbrthemes)
-hrbrthemes::import_roboto_condensed() 
+```{r optional_plot_fonts, message=FALSE}
 library(ggplot2)
@@
 library(extrafont)
+library(hrbrthemes)
+
 extrafont::loadfonts()
+knitr::opts_chunk$set( dev="cairo_pdf")
+hrbrthemes::import_roboto_condensed() 
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
@@
 
-```{r}
+```{r plot_colors}
+#palette <- ggthemes::ptol_pal()(6)
+palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
+
+colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
+                    c("TAC", "POMDP", "MSY", "CE", 
+                      "POMDP: low prior",
+                      "POMDP: medium prior",
+                      "biomass",
+                      "catch"))
+
+```
+
+```{r load_data, message=FALSE, warning=FALSE}
 log_dir <- "../data" # solution files here
+
+sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
+det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
+all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
+ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
+
 ```
@@
 
-```{r load_data, message=FALSE, warning=FALSE}
-sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
-det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
-all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
-ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
-
-colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
-                    c("TAC", "POMDP", "MSY", "CE", 
-                      "POMDP: low prior",
-                      "POMDP: medium prior",
-                      "biomass",
-                      "catch"))
-
-```
-
 Figure 1 shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
@@
 
-```{r fig.width=7,fig.height=3.5, fig.cap=""}
+```{r fig.width=7,fig.height=7, fig.cap=""}
 all_policies %>% 
@@
                              )) -> policy_df
-policy_df %>%
+harvest <- policy_df %>%
   ggplot(aes(state, harvest, col=policy)) + 
@@
   scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
-```
-```{r fig.width=7,fig.height=3.5, fig.cap=""}
-policy_df %>%
+  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
+  ggtitle("Harvest view") + 
+  theme(legend.position="none")
+
+escapement <- policy_df %>%
   ggplot(aes(state, escapement, col=policy)) + 
@@
   scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0, .8))
+  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
+  ggtitle("Escapement view") + 
+  theme(legend.position = "bottom")
+
+gridExtra::grid.arrange(harvest, escapement, 
+                        layout_matrix = rbind(c(1,1), c(2,2)))
+
 ```

### 2018-01-16 appendix/appendix_harvestfirst.Rmd
WIP

@@
 
-##  Figure S1
+##  Figure S2
 
@@
   coord_cartesian(ylim = c(0, 1))
+
+```
+
+##  Figure S3
+
+Economic Value
+
+```{r}
+sims %>%
+  select(time, value, rep, method, sigma_m, sigma_g) %>%
+  filter(sigma_g %in% c("0.1", "0.15")) %>%
+  group_by(rep, method, sigma_m, sigma_g) %>%
+  summarise(npv = sum(value)) %>%  
+  group_by(method, sigma_m, sigma_g) %>%
+  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
+  ungroup() %>%
+  mutate(method = fct_recode(method,
+                             "MSY" = "msy",
+                             "CE" = "det",
+                             "TAC" = "pgy",
+                             "POMDP" = "pomdp")) %>%
+
+  ggplot(aes(method, net_value, ymin=net_value-se, ymax=net_value+se, fill=method)) + 
+  geom_bar(position=position_dodge(), stat="identity") + 
+  geom_errorbar(size=.3,width=.2,
+                position=position_dodge(.9)) +
+  facet_grid(sigma_m~sigma_g, 
+             labeller = label_bquote(sigma[m] == .(sigma_m),
+                                     sigma[g] == .(sigma_g))) +
+  theme(legend.position = "none") + 
+  scale_color_manual(values = colors) +
+  scale_fill_manual(values = colors) +
+  ylab("expected net present value")
 

### 2018-01-16 manuscripts/manuscript.Rmd
WIP

@@
 
-sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnn")
+sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnnn")
 det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
@@
 
-<!-- Show results of POMDP when over-estimating! -->
-
 ```{r fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
@@
 sims %>%
+  select(-value) %>%
   filter(method != "msy", sigma_g == "0.15") %>%
@@
 
+This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also bourne out in terms of economic performance.
+Figure 2 shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.
+As before, environmental stochasticity is set at $\sigma_g = 0.15$.
+In the absense of measurement uncertainty, constant escapement (CE) is optimal, though POMDP performs nearly as well, while TAC underperforms.
+As measurement uncertainty increases, the economic value of CE declines, remaining better than TAC at modest measurement error $\sigma_m = 0.1$ but worse at higher uncertainty.
+Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure 1), these have little impact on the economic value due to the discount rate.
+A smaller discount rate would penalize unlikely but not improbable collapses more, since a significant amount of time is required to realize those rare events.
+Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.
+
+```{r fig.width=7, fig.height=3.5, fig.cap=""}
+sims %>%
+  select(time, value, rep, method, sigma_m, sigma_g) %>%
+  filter(sigma_g == "0.15", method != "msy") %>%
+  group_by(rep, method, sigma_m, sigma_g) %>%
+  summarise(npv = sum(value)) %>%  
+  group_by(method, sigma_m, sigma_g) %>%
+  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
+  ungroup() %>%
+  mutate(method = fct_recode(method,
+                             "CE" = "det",
+                             "TAC" = "pgy",
+                             "POMDP" = "pomdp")) %>%
+
+  ggplot(aes(method, net_value, ymin=net_value-se, ymax=net_value+se, fill=method)) + 
+  geom_bar(position=position_dodge(), stat="identity") + 
+  geom_errorbar(size=.3,width=.2,
+                position=position_dodge(.9)) +
+  facet_wrap(~sigma_m, 
+             labeller = label_bquote(sigma[m] == .(sigma_m))) +
+  theme(legend.position = "none") + 
+  scale_color_manual(values = colors) +
+  scale_fill_manual(values = colors) +
+  ylab("expected net present value")
+```
+
+<!-- Show results of POMDP when over-estimating! -->
+
+To better understand the differences in performance of these strategies and resolve the paradox of uncertainty, we must take a closer look at how the specific action recommended by each policy compares given the same observation.
+Figure 3 shows the action taken by each strategy in response to a measurement of the stock size.
+The first panel shows the policy in terms of harvest $H$, that is, the catch quota that would be set by each strategy.
+For reference, the second panel shows the same policies but in terms of expected escapement, $S = Y - H$ for observed stock $Y$, which is more common in the control literature [e.g.
+@Reed1979; @Clark1987; @Sethi2005] than in fisheries literature.
+The first panel clearly illustrates the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.
+Viewing the same policies in terms of escapement, second panel, makes it clear that only CE strategy corresponds to keeping escapement constant.
+This is important, as theoretical work ever since the proofs of @Reed1979 often attempts to simplify the calculations by assuming the optimal policy must be a constant-escapement type strategy [e.g.
+@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
+
+While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 2, this is not the case for the POMDP approach.
+This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.
+The POMDP policy cannot be specified by the most recent observation alone.
+Instead, the POMDP policy depends on all prior observations, not just the most recent.
+The reason for this complexity comes from the Markov property.
+Observations of the state in the perfectly observed system satisfy the Markov property: once we have measured the current biomass exactly, we cannot get any better estimate of the current stock size by studying older measurements.
+When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.
+POMDP formalizes this intutition by capturing the information from all previous observations into a *prior belief*.
+This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.
+The mechanics of this process are well documented in the extensive literature on POMDPs [e.g.
+], but for our purposes
+
+The other approaches exploit this property, and the simplifying assumptions introduced in previous work attempting to address measurement uncertainty [@Clark1987; @Sethi2005] serve to restore or assert this.
+The POMDP
+
+It is important to bear in mind several issues when intepreting these figures.
+First, the horizontal axis reflects the measured stock size, which may differ from the (unobservable) true stock size when measurement uncertainty is present.
+
+Consequently, the realized escapement may differ from the target level.
+
 POMDP policies depend on the prior belief.
 
-```{r fig.width=7,fig.height=7, fig.cap=""}
+```{r fig.width=7, fig.height=7, fig.cap=""}
 all_policies %>% 
@@
   ggtitle("Harvest view") + 
-  theme(legend.position="none")
+  theme(legend.position="none") + 
+  xlab("Estimated stock biomass")
 
@@
   coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
+  xlab("Estimated stock biomass") +
   ggtitle("Escapement view") + 

### 2018-01-18 appendix/appendix_harvestfirst.Rmd
WIP

@@
                       "POMDP: low prior",
-                      "POMDP: medium prior",
+                      "POMDP: high prior",
                       "biomass",
@@
   ylab("Mean biomass")
-```
-
 ```
@@
 ```{r}
+sim_col_types <-  "inicnnn"
 sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = sim_col_types)
@@
 ```{r}
-i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.1) %>% pull(scenario) %>% as.integer()
+i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.15) %>% pull(scenario) %>% as.integer()
 m <- models[[i]]
@@
 unif <- compute_policy(alpha, m$transition, m$observation, m$reward,  unif_prior)
-
 ```
@@
 high <- compute_policy(alpha, m$transition, m$observation, m$reward,  prior_high)
-
 ```
@@
 write_csv(all_policies, file.path(log_dir, "all_policies.csv"))
+```
+
+```{r}
+priors <- 
+data_frame(states, 
+           prior_low,
+           prior_star,
+           unif_prior,
+           prior_high) 
+write_csv(priors, file.path(log_dir, "priors.csv"))
+```
+
+```
+
+```
+
+POMDP policies depend on the prior belief.
+
+```{r fig.width=5, fig.height=7, fig.cap=""}
+priors <- 
+  read_csv(file.path(log_dir, "priors.csv")) %>%
+  select(states, 
+         "POMDP: low prior" = prior_low, 
+         "POMDP: high prior" = prior_high) %>%
+  gather(prior, belief, -states)
+
+policy_df <- all_policies %>% 
+  filter(policy %in% c("low", "high", "det", "pgy")) %>%
+  mutate(policy = fct_recode(policy,
+                             "CE" = "det",
+                             "TAC" = "pgy",
+                             "POMDP: low prior" = "low",
+                             "POMDP: high prior" = "high"
+                             )) 
+prior <- 
+  ggplot(priors,aes(states, belief, color = prior)) + 
+  geom_line(lwd=1)  + 
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1)) +
+  ggtitle("POMDP priors") + 
+  xlab("Estimated stock biomass") + 
+  theme(legend.position = "none")
+
+harvest <- policy_df %>%
+  ggplot(aes(state, harvest, col=policy)) + 
+  geom_line(lwd=1)  +
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
+  ggtitle("Harvest view") + 
+  theme(legend.position="none") + 
+  xlab("Estimated stock biomass")
+
+escapement <- policy_df %>%
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line(lwd=1)  +
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
+  xlab("Estimated stock biomass") +
+  ggtitle("Escapement view") + 
+  theme(legend.position = "bottom")
+
+gridExtra::grid.arrange(harvest, escapement, prior,
+  layout_matrix = rbind(1,1,2,2, 3))
 
@@
 ```{r}
-pomdp_policies %>% 
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+
 ```
-
-```{r}
-pomdp_policies %>% 
-  ggplot(aes(state, escapement, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-```
-
-```{r}
-all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
-all_policies %>% 
-  filter(policy %in% c("low", "medium", "det", "msy")) %>%
-  mutate(policy = fct_recode(policy,
-                             "MSY" = "msy",
-                             "Constant escapement" = "det",
-                             "POMDP: low prior" = "low",
-                             "POMDP: medium prior" = "medium"
-                             )) %>%
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-```
-
-```{r}
-all_policies <- read_csv(file.path(log_dir, "all_policies.csv"))
-all_policies %>% 
-  filter(policy %in% c("low", "medium", "det", "msy")) %>%
-  mutate(policy = fct_recode(policy,
-                             "MSY" = "msy",
-                             "Constant escapement" = "det",
-                             "POMDP: low prior" = "low",
-                             "POMDP: medium prior" = "medium"
-                             )) %>%
-  ggplot(aes(state, escapement, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-```

### 2018-01-18 manuscripts/manuscript.Rmd
WIP

@@
 
+Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
 While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 2, this is not the case for the POMDP approach.
@@
 The mechanics of this process are well documented in the extensive literature on POMDPs [e.g.
-], but for our purposes
+@pomdpstuff; @stuff; @more], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.
+
+Figure 2 shows two separate policy curves for the same POMDP solution given two different prior belief distributions. The lower prior
 
@@
 ```{r fig.width=7, fig.height=7, fig.cap=""}
-all_policies %>% 
-  filter(policy %in% c("low", "medium", "det", "pgy")) %>%
+priors <- 
+  read_csv(file.path(log_dir, "priors.csv")) %>%
+  select(states, 
+         "POMDP: low prior" = prior_low, 
+         "POMDP: high prior" = prior_high) %>%
+  gather(prior, belief, -states)
+
+policy_df <- all_policies %>% 
+  filter(policy %in% c("low", "high", "det", "pgy")) %>%
   mutate(policy = fct_recode(policy,
@@
                              "POMDP: low prior" = "low",
-                             "POMDP: medium prior" = "medium"
-                             )) -> policy_df
+                             "POMDP: high prior" = "high"
+                             )) 
+prior <- 
+  ggplot(priors,aes(states, belief, color = prior)) + 
+  geom_line(lwd=1)  + 
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1)) +
+  ggtitle("POMDP priors") + 
+  xlab("Estimated stock biomass") + 
+  theme(legend.position = "none")
+
 harvest <- policy_df %>%
@@
 
-gridExtra::grid.arrange(harvest, escapement, 
-                        layout_matrix = rbind(c(1,1), c(2,2)))
-
+gridExtra::grid.arrange(harvest, escapement, prior,
+  layout_matrix = rbind(1,1,2,2, 3))
 ```

### 2018-01-18 manuscripts/manuscript.Rmd
fix path

@@
 all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
+priors <- read_csv(file.path(log_dir, "observe_harvest_recruit/priors.csv"))
 ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
@@
 priors <- 
-  read_csv(file.path(log_dir, "priors.csv")) %>%
+  priors %>%
   select(states, 

### 2018-01-18 appendix/appendixB.Rmd
WIP

@@
 ```{r}
-sims %>%
-  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
+ram_ex %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy", "msy")) %>%
+   mutate(model = fct_recode(model,
+                             "MSY" = "msy",
+                             "CE" = "Reed",
+                             "TAC" = "pgy"
+                             )) %>%
   ggplot(aes(year, stock, col=model)) +
   geom_line(lwd=1) +
-  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
-  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
+  scale_color_manual(values = colors) +
+  facet_wrap(~commonname, scales = "free", ncol=1)  
 ```

### 2018-01-18 appendix/appendix_harvestfirst.Rmd
WIP

@@
 ```{r}
-i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0.15) %>% pull(scenario) %>% as.integer()
+i <- meta %>% filter(sigma_g == 0.15, sigma_m ==0.1) %>% pull(scenario) %>% as.integer()
 m <- models[[i]]
@@
 i_star <- which.min(abs(states - S_star))
-i_low <- which.min(abs(states - 0.5 * S_star))
-i_high <- which.min(abs(states - 1.5 * S_star))
+i_low <- which.min(abs(states - 0.25 * K))
+i_high <- which.min(abs(states - 0.75 * K))
 
@@
 
-```{r}
+POMDP policies depend on the prior belief.
+
+```{r fig.width=5, fig.height=7, fig.cap=""}
 priors <- 
@@
              unif_prior,
-           prior_high) 
-write_csv(priors, file.path(log_dir, "priors.csv"))
-```
-
-```
-
-```
-
-POMDP policies depend on the prior belief.
-
-```{r fig.width=5, fig.height=7, fig.cap=""}
-priors <- 
-  read_csv(file.path(log_dir, "priors.csv")) %>%
-  select(states, 
+             prior_high) %>%
+  select(state = states, 
          "POMDP: low prior" = prior_low, 
          "POMDP: high prior" = prior_high) %>%
-  gather(prior, belief, -states)
+  gather(policy, prior, -state) %>%
+  mutate(prior = prior*10)
 
@@
                              ))
+
+policies_w_priors <- right_join(priors, policy_df, by = c("state", "policy"))
+
+write_csv(policies_w_priors, file.path(log_dir, "policies_w_priors.csv"))
+
+```
+
+```{r fig.width=5, fig.height=7, fig.cap=""}
+policies_w_priors %>% 
+  gather(panel, value, -state, -policy) %>%
+  ggplot(aes(state, value, col=policy)) + 
+  geom_line(lwd=1)  +
+  facet_wrap(~panel, ncol=1, scales = "free_y") +
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1), ylim = c(0,.8)) + 
+  theme(legend.position = "bottom")
+```
+
+```{r  fig.width=5, fig.height=7, fig.cap=""}
 prior <- 
-  ggplot(priors,aes(states, belief, color = prior)) + 
+  ggplot(priors,aes(state, belief, color = policy)) + 
   geom_line(lwd=1)  + 

### 2018-01-18 manuscripts/manuscript.Rmd
WIP

--- 2018-01-18 manuscripts/manuscript.Rmd
@@
 det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
-all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
+#all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
+#prior_dat <- read_csv(file.path(log_dir, "observe_harvest_recruit/priors.csv"))
+policies_w_priors <- read_csv(file.path(log_dir, "observe_harvest_recruit/policies_w_priors.csv"))
 ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
@@
 Figure 3 shows the action taken by each strategy in response to a measurement of the stock size.
-The first panel shows the policy in terms of harvest $H$, that is, the catch quota that would be set by each strategy.
-For reference, the second panel shows the same policies but in terms of expected escapement, $S = Y - H$ for observed stock $Y$, which is more common in the control literature [e.g.
-@Reed1979; @Clark1987; @Sethi2005] than in fisheries literature.
-The first panel clearly illustrates the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.
-Viewing the same policies in terms of escapement, second panel, makes it clear that only CE strategy corresponds to keeping escapement constant.
+The first panel shows the policy in terms of expected escapement, $S = Y - H$ for observed stock $Y$ and harvest $H$, consistent with most of the relevant control and bioeconomics literature [e.g.
+@Reed1979; @Clark1986; @Sethi2005].
+For comparison, the second panel shows the same policies but in terms of harvest $H$.
+These plots clearly illustrate the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.
+Viewing the same policies in terms of escapement, makes it clear that only CE strategy corresponds to keeping escapement constant.
 This is important, as theoretical work ever since the proofs of @Reed1979 often attempts to simplify the calculations by assuming the optimal policy must be a constant-escapement type strategy [e.g.
 @Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
+Note that both CE and TAC policies are independent of the levels of uncertainty $\sigma_m$ and $\sigma_g$.
+TAC implicitly accounts for some uncertainty through its 80% MSY definition -- comparable plots with MSY can be found in the Supplemental Material.
+In principle, the CE policy could depend on $\sigma_g$, but as @Reed1979 proved, the constant escapement level with stochasticity is the same as in the deterministic case, S = D, unless the noise level is quite high (comparison plots in Supplemental material.)  In contrast to this, our POMDP plots depend very much on both the choice of $\sigma_m$ and $\sigma_g$.
+If $\sigma_m = 0$, they reduce exactly to the CE solution.
+Here we show the corresponding policies for POMDP solutions focusing on $\sigma_g = 0.15$ as above, with a modest measurement uncertainty $\sigma_m = 0.1$.
+Alternate combinations of measurement uncertainty can be found in the appendix but do not change the general pattern.
+In addition to an explicit dependence on the measurement error, our POMDP solutions depend on another piece of information: any prior observations of the stock size.
 
 Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
-While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 2, this is not the case for the POMDP approach.
+While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 3, this is not the case for the POMDP approach.
 This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.
@@
 
-Figure 2 shows two separate policy curves for the same POMDP solution given two different prior belief distributions. The lower prior
+Figure 3 shows two separate policy curves (in terms of harvest and escapement) for the same POMDP solution given two different prior belief distributions (panel 3, priors).
+While both priors shown express considerable uncertainty about the precise stock size prior to the most recent observation, the lower prior is centered at a value $\tfrac{1}{4}K$ stock size, while the high prior is centered at a value of $\tfrac{3}{4}K$, relative to a (post-harvest, before observation) target size of $B_{MSY} = K/2$.
+The POMDP policy is determined by the combination of this prior information and the most recent observation, as indicated by the two different POMDP curves for harvest/escapement shown corresponding to the different priors.
+As with the other policies, the higher the most recent observation (x axis) the higher the POMDP recommended harvest.
+Yet unlike the alternative strategies, the POMDP solution always reflects the prior information.
+Consequently, relative to constant escapement (that is, no measurement error), the POMDP with low prior starts harvesting only at higher stock sizes and always harvests less.
+In contrast under the high prior, the POMDP always harvests at the same or higher level than the constant escapement solution.
 
-The other approaches exploit this property, and the simplifying assumptions introduced in previous work attempting to address measurement uncertainty [@Clark1987; @Sethi2005] serve to restore or assert this.
-The POMDP
-
-It is important to bear in mind several issues when intepreting these figures.
-First, the horizontal axis reflects the measured stock size, which may differ from the (unobservable) true stock size when measurement uncertainty is present.
-
-Consequently, the realized escapement may differ from the target level.
-
-POMDP policies depend on the prior belief.
-
-```{r fig.width=7, fig.height=7, fig.cap=""}
-priors <- 
-  read_csv(file.path(log_dir, "priors.csv")) %>%
-  select(states, 
-         "POMDP: low prior" = prior_low, 
-         "POMDP: high prior" = prior_high) %>%
-  gather(prior, belief, -states)
-
-policy_df <- all_policies %>% 
-  filter(policy %in% c("low", "high", "det", "pgy")) %>%
-  mutate(policy = fct_recode(policy,
-                             "CE" = "det",
-                             "TAC" = "pgy",
-                             "POMDP: low prior" = "low",
-                             "POMDP: high prior" = "high"
-                             )) 
-prior <- 
-  ggplot(priors,aes(states, belief, color = prior)) + 
+```{r fig.width=5, fig.height=7, fig.cap="", message=FALSE, warning=FALSE}
+policies_w_priors %>% 
+  gather(panel, value, -state, -policy) %>%
+  ggplot(aes(state, value, col=policy)) + 
   geom_line(lwd=1)  +
+  facet_wrap(~panel, ncol=1, scales = "free_y") +
   scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1)) +
-  ggtitle("POMDP priors") + 
-  xlab("Estimated stock biomass") + 
-  theme(legend.position = "none")
-
-harvest <- policy_df %>%
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line(lwd=1)  +
-  scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
-  ggtitle("Harvest view") + 
-  theme(legend.position="none") + 
-  xlab("Estimated stock biomass")
-
-escapement <- policy_df %>%
-  ggplot(aes(state, escapement, col=policy)) + 
-  geom_line(lwd=1)  +
-  scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
-  xlab("Estimated stock biomass") +
-  ggtitle("Escapement view") + 
+  coord_cartesian(xlim = c(0,1), ylim = c(0,.8)) + 
   theme(legend.position = "bottom")
-
-gridExtra::grid.arrange(harvest, escapement, prior,
-  layout_matrix = rbind(1,1,2,2, 3))
 ```
 
-```{r message=FALSE, warning=FALSE, fig.width=7, fig.height = 3.5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
+**A resolution to the paradox**
+Herein lies our resolution to the paradox of uncertainty.
+Previous work created this paradox by suggesting that increased harvest rates (decreased target escapement) would often be the rational response to increased uncertainty.
+The exact solutions from the POMDP reveal that this is only an accident of the assumptions: it is indeed true that under certain circumstances, harvest levels should increase relative to the case of no uncertainty, but *only when prior knowledge suggests the stock size should be much higher than the most recent estimate would suggest*.
+In the POMDP solution, all information must be put into its historical (and constantly updated) context.
+When a measurement roughly matches the expectation of this prior context, Figure 3 shows that the optimal response from POMDP is roughly comparable or slightly more cautious to the harvest under no uncertainty, and not more aggressive as the paradox would suggest.
+Measurements that exceed expectations are tempered with some skepticism: while the CE solution is willing to meet a high stock measurement with a large harvest, the POMDP solutions increase harvest more cautiously.
+
+This difference between underestimating and overestimating accounts for the poor performance of the CE solution under large measurement uncertainty, where it over-harvests whenever measurements are too large.
+Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where chance local extinction becomes possible.
+It is precisely this assymetry: that too much overharvesting leads to an irreversible state of extinction, while too much underharvesting is always reversable (modulo some lost revenue) that lay behind Clark's original intutition that there was something fishy about Reed's result that S = D, that uncertainty required no extra caution.
+Constant escapement is particularly suscpetible to overharvesting starting from stock sizes much higher than $B_{MSY}$ since its bang-bang optimal solution attempts to bring stocks back to target level as fast as possible.
+CE is not so vulnerable to collapse from small stock sizes, since it shuts down all harvests once estimates fall below $B_{MSY}$.
+Stock collapse under TAC can also be driven by such a string of heads, but is unlikely at high stock sizes since harvest never exceeds $H_{MSY}$.
+In both cases, some measurement uncertainty interacts with inherent stochasticity, which provides a continued source of variation to sizes above $B_{MSY}$, where CE strategy is most vulnerable, and below $B_{MSY}$, where TAC is most vulnerable.
+
+```{r message=FALSE, warning=FALSE, fig.width=7, fig.height = 5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
 POMDP solution assumes a measurement error of 10%."}
@@
   scale_color_manual(values = colors) +
-  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
+  facet_wrap(~commonname, scales = "free", ncol=1)  
 ```
 
-**A POMDP policy reflects all prior observations**.
-The essential difference in a POMDP policy compared to either that of constant escapement $S$ or constant mortality $F$ is that nothing is just a constant -- instead, the harvest quota recommended by POMDP will always depend not only on the most recent observation of the stock size, but on all previous stock sizes.
-This dependency can be expressed in terms of a *prior belief* -- if all prior observations have led one to expect that the stock should be relatively small, a single observation of a very large stock is met by some skepticism under a POMDP, and the resulting harvest recommended by POMDP can differ substantially from that recommended by constant escapement or constant mortaility, which by definition consider only the most recent observation.
-If the observation is largely consistent with prior expectations, a POMDP policy will be more similar to that of the constant escapement policy, though with a natural level of precaution built in to allow for the possibility of error in measurement coupled with a negative shock to growth.
+**Historical examples and implementation**.
 
-# Discussion
+So far we have focused on simulation and an examination of the relative policies under different prior information.
+Figure 4 compares the harvest level that would have been recommmended by each of the strategies we have considered here against the historically observed catch recorded for two commericially fished stocks: Argentine Hake and European Plaice.
+Historical estimates of biomass and catch are taken from the R.A.
+Myers Legacy Stock Assessment database [@RAM].
+Posterior distributions for parameters for the Graham-Schaefer model are estimated with uncertainty through Markov Chain Monte Carlo using `nimble` [@nimble] on the historical data, illustrating how this process might be done in more complex models as well (details and code in Supplement, Appendix B).
+$B_{MSY} and corresponding TAC and CE policies are calculated based on posterior mean estimates of the model parameters, along with POMDP solution assuming a 10% measurement error rate.
+Though measurement error could be estimated directly from the raw data, this would not reflect the true measurement uncertainty arising from the stock assessment process.
+Each strategy is then compared to historical observations to determine a recommended harvest.
+As we have seen, the TAC and CE harvest policies are uniquely determined by the observed stock size relative to $B_{MSY}$, but the POMDP policy must be re-calculated each time step to reflect both the prior observation and prior action.
+Note that since we cannot replay history under these alternative policies, so the prior action taken is always the historical catch for that year and not the catch recommended by the POMDP policy.
 
-Is it surprising that POMDP model should do better?  Recall that Reed found stochasticity to make no difference, the deterministic model performs identically.
+These historical examples provide another useful lens to compare how the different strategies respond in the face of fluctuations in real data, rather than model simulations.
+In both stocks, historical catch almost always exceeds that recommended by any of our strategies, falling closer to the MSY value estimated (see appendix).
+In Argentine Hake, the persistent declines result in both POMDP and CE strategies quicly closing down the fishery in an attempt to let the stock recover to a more productive target biomass, while TAC persists with merely a reduced harvest.
+A small recovery of biomass in 1995 is met by an immediate uptick in harvest under CE, while the POMDP response is more conservative.
+The example of European Plaice illustrates more volatility in the stock size, revealing further differences in the strategies.
+Despite this volatility, the TAC level remains relatively level across all five decades, rising or dipping only slightly with changes in stock size.
+In contrast, the CE harvest tracks this volatility almost in perfect mirror.
+Here, the POMDP solution falls in between these extremes, almost mirroring which-ever of the two policies is more conservative: the POMDP solution matches the dips in harvest taken by CE to allow the stock to recover most quickly, but does not track the doubling of harvests recommended by CE in the mid 1980, increasingly only to the more modest harvest recommended by the TAC policy.
 
-## A practical implementation?
+Once again, we see that POMDP solution provides a more consistently precautionary policy than either the over-simplified optimal control solution of CE or the more rule-of-thumb approach represented by TAC.
+Though we have seen that the POMDP solution need not always be more conservative (lower harvest) than these strategies, it makes use of the all prior observations to tune the level of caution appropriately.
+
+## Discussion
+
+Using modern algorithms we have been able to crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk a long-standing paradox that increased measurement uncertainty should be met with larger harvests.
+We have illustrated that the POMDP approach introduced here performs consistently better across a range of stochastic growth and measurement error levels, while previous solutions based on either classic optimal control solutions or heuristic rules-of-thumb give significantly worse outcomes in both ecological and economic terms.
+Nevertheless, it may be tempting to see these results as either trivial or impractical, but a closer inspection reveals they are neither.
+
+First, It may seem obvious that POMDP would out-proform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.
+Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accomidate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.
+If stochasticity doesn't matter, it is easy to believe that measurement error would be minor at best.
+In fact, we find that once we account for measurement error, the effects of both stochasticity and measurement have a significant influence over the optimal policy.
+It is similarly tempting to dismiss the better performance of POMDP against a more rule-of-thumb[^1] based strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.
+While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to eek out both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.
+These results highlight the difficulty of finding effective heuristic rules and underscore the power and utility of more sophisticated strategies such as POMDP which can make the best use of all available information.
+
+^1: Given the sophisticated policies often in place to determine actual TAC levels, one might object to our characterization of this as merely a rule-of-thumb.
+We use this term merely to distinguish it from the optimal control approach -- that despite technical precision, TAC adjustments are set by optimal control
+
+If it is not obvious that POMDP solution would do significantly better, one might still ask if such approaches are practical.
+The importance of simple and effective rules of thumb in conservation management has been well documented [e.g.
+@Possingham].
+Against this we make the following observations.
+First, the strategies we have compared against are not purely of academic interest but underpin both current practice and management aspirations.
+Together, MSY and TAC-based management reflects much of modern best practice as it is understood in fisheries management and elsewhere; US salmon stocks are one of few examples using a constant escapement-based approach instead.
+While real-world application of these approaches frequently involves more complex models, the models we have considered here are the ones under which these strategies were designed and tested, and there is little reason to believe they would perform better under more complex cases such as age-structured models that do not always meet those assumptions [e.g.
+@Holden2015].
+The relatively poor ecological and economic performance of these common best-practices relative to the alternative approach of POMDP is thus reason enough for concern.
+This issue is further underscored by the reality of both long-running ecological detioriation [@Worm2006; @Worm2009] and economic costs of global fisheries.
+
+**Are moritoriums pratical?**
+Fisheries today are not managed at either their maximum ecological or economic potential.  
+Taken as a whole, most ocean fisheries yield a net negative economic return
+[@FAO2009], with fishing persisting only with the support of government subsidies [@Arnason2012], 
+meanwhile some 68% of global fisheries are in poor biological condition [@Costello2016].
+While the causes for this are both complex and numerous, we have seen that both the economic
+inefficiency and ecological health suffer under TAC/MSY and CE based management.
+
+**Is computationally intensive management feasible?**
+Lastly, one may dismiss the potential of POMDP approaches as being too computationally infeasible for everyday management practices.
+In many areas of natural resource management, even simple, rule-of-thumb based methods struggle to take root.
+In such a world, how can complex artificial intelligence algorithms requiring significant computational resources to execute play any role?
+And yet, in that same world far more complex and demanding algorithms are run every day to provide features such as face recognition to smart phone applications.
+If such resources and expertise can now facilitate every day social decisions with only a tap, surely we can one day harness the same complexity and power to facilitate conservation and resource management.
 
@@
 However, the economist's perspective reveals that this may not be as unrealistic as it may first appear.
-Fisheries today are not managed at either their maximum ecological or economic potential.  
+
 Taken as a whole, most ocean fisheries yield a net negative economic return

### 2018-01-23 appendix/appendix_harvestfirst.Rmd
overestimate

@@
 
-## Simulating the POMDP policies under uncertiainty
+## Simulating the POMDP policies under uncertainty
 
@@
 
+```
+
+-----
+
+## POMDP simulations overestimating measurement uncertainty
+
+```{r pomdp_sims}
+set.seed(12345)
+
+pomdp_overest_sims <- 
+  map2_dfr(models, alphas, function(.x, .y){
+    pomdp_overestimates(transition = .x$transition, 
+            model_observation = .x$observation, 
+            reward = .x$reward, 
+            discount = discount, 
+            true_observation <- models[[true]]$observation,
+            x0 = x0,
+            Tmax = Tmax,
+            alpha = .y,
+            reps = reps)$df %>% 
+              mutate(method = "pomdp") # include a column labeling method
+           },
+           .id = "scenario"
+  )
 ```

### 2018-01-23 appendix/pomdp_overestimate.R

@@ -1,0 +1,95 @@
+## parallel version
+pomdp_overestimates <-
+  function (transition, 
+            model_observation, 
+            reward, 
+            discount, 
+            true_observation,
+            x0, 
+            a0 = 1, 
+            Tmax = 20,
+            alpha,
+            reps = 100) 
+  {
+  sims <- parallel::mclapply(1:reps, function(i) {
+    sim <- pomdp_overestimate(transition, 
+                              model_observation, 
+                              reward, 
+                              discount,           
+                              true_observation, 
+                              x0, 
+                              a0, 
+                              Tmax, 
+                              alpha)
+    sim$df$rep <- i
+    sim
+  }, ...)
+  list(df = do.call(rbind, lapply(sims, `[[`, "df")), 
+       state_posterior = do.call(rbind, lapply(sims, `[[`, "state_posterior")))
+
+}
+
+pomdp_overestimate <-
+function (transition, 
+          model_observation, 
+          reward, 
+          discount, 
+          true_observation,
+          x0, 
+          a0 = 1, 
+          Tmax = 20,
+          alpha) 
+{
+  observation <- model_observation
+  
+  stopifnot(identical(dim(true_observation), dim(model_observation)))
+
+  n_states <- dim(observation)[1]
+  n_obs <- dim(observation)[2]
+  value <- obs <- action <- state <- numeric(Tmax + 1)
+  state_posterior <- array(NA, dim = c(Tmax + 1, n_states))
+  state_prior <- rep(1, n_states)/n_states # assume unif prior
+  
+  state[2] <- x0
+  action[1] <- a0
+  state_posterior[2, ] <- state_prior
+  
+  for (t in 2:Tmax) {
+      ## Policy based on model
+      out <- compute_policy(alpha, 
+                            transition, 
+                            model_observation, 
+                            reward, 
+                            state_posterior[t, ], 
+                            action[t - 1])
+      
+    ## Simulation based on True
+    obs[t] <- sample(1:n_obs, 
+                     1, 
+                     prob = 
+              true_observation[state[t],, action[t - 1]])
+    
+    action[t] <- out$policy[obs[t]]
+    value[t] <- reward[state[t], action[t]] * discount^(t - 1)
+    state[t + 1] <- sample(1:n_states, 
+                           1, 
+                prob = transition[state[t], , action[t]])
+    
+    ## Update belief based on model
+    state_posterior[t + 1, ] <- 
+      update_belief(state_posterior[t,], 
[…]

### 2018-01-24 appendix/appendix_harvestfirst.Rmd
add pomdp overestimate analysis

--- 2018-01-18 appendix/appendix_harvestfirst.Rmd
@@
   ylab("expected net present value")
+
+```
+
+-----
+
+## POMDP simulations overestimating measurement uncertainty
+
+While the POMDP approach requires an estimate of the measurement error, the precise distribution of measurement errors will itself be unknown in most cases.
+However, the POMDP approach is quite robust to overestimation of the measurement error.
+As an extreme example of this, we consider the case where the POMDP solution assumes the largest measurement error level considered here, $\sigma_m = 0.15$, while performing simulations in which measurements occur without error.
+In such a scenario, the
+
+```{r pomdp_sims}
+set.seed(12345)
+true <- 3 # sigma_g = 0.15, sigma_m = 0
+source("../appendix/pomdp_overestimate.R")
+pomdp_overest_sims <- 
+  map2_dfr(models, alphas, function(.x, .y){
+    pomdp_overestimates(transition = .x$transition, 
+            model_observation = .x$observation, 
+            reward = .x$reward, 
+            discount = discount, 
+            true_observation = models[[true]]$observation,
+            x0 = x0,
+            Tmax = Tmax,
+            alpha = .y,
+            reps = reps)$df %>% 
+              mutate(method = "pomdp") # include a column labeling method
+           },
+           .id = "scenario"
+  )
+```
+
+Combine the resulting data frames
+
+```{r}
+
+overest <-
+bind_rows(static_sims, pomdp_overest_sims) %>%
+  left_join(meta, by = "scenario") %>%  ## include scenario information (sigmas; etc)
+  mutate(state = states[state], action = actions[action]) %>%
+  select(time, state, rep, method, sigma_m, sigma_g, value) %>%
+  group_by(time, method, sigma_m, sigma_g) %>%
+  summarise(mean = mean(state), sd = sd(state)) %>%
+  filter(sigma_g == "0.15") %>% 
+  ungroup()
+
+overest %>% filter(method != "pomdp", sigma_m == "0")  %>%
+  bind_rows(
+  overest %>% filter(method == "pomdp", sigma_m == "0.15")) %>%
+  select(-sigma_m, -sigma_g) %>%
+  mutate(method = fct_recode(method,
+                             "MSY" = "msy",
+                             "CE" = "det",
+                             "TAC" = "pgy",
+                             "POMDP" = "pomdp")) %>%
+  ggplot(aes(time, mean, col=method, fill=method)) + 
+  geom_line(lwd = 1) + 
+  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
+  coord_cartesian(ylim = c(0, 1))
 

### 2018-01-24 appendix/pomdp_overestimate.R

@@ -23,7 +23,7 @@ pomdp_overestimates <-
                               alpha)
     sim$df$rep <- i
     sim
-  }, ...)
+  })
   list(df = do.call(rbind, lapply(sims, `[[`, "df")), 
        state_posterior = do.call(rbind, lapply(sims, `[[`, "state_posterior")))
 
@@ -56,7 +56,7 @@ function (transition,
   
   for (t in 2:Tmax) {
       ## Policy based on model
-      out <- compute_policy(alpha, 
+      out <- sarsop::compute_policy(alpha, 
                             transition, 
                             model_observation, 
                             reward, 
@@ -77,7 +77,7 @@ function (transition,
     
     ## Update belief based on model
     state_posterior[t + 1, ] <- 
-      update_belief(state_posterior[t,], 
+      sarsop:::update_belief(state_posterior[t,], 
                     transition, 
                     model_observation, 
                     obs[t], 

### 2018-01-24 appendix/appendix_harvestfirst.Rmd
overest sims

@@
                              "POMDP" = "pomdp")) %>%
+  write_csv(file.path(log_dir, "overest_sims.csv"))
+```
+
+```{r}
+read_csv(file.path(log_dir, "overest_sims.csv")) %>%
   ggplot(aes(time, mean, col=method, fill=method)) + 

### 2018-01-24 manuscripts/manuscript.Rmd
text on overest

@@
 ```{r load_data, message=FALSE, warning=FALSE}
-log_dir <- "../data" # solution files here
+log_dir <- "../data/observe_harvest_recruit" # solution files here
 
-sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnnn")
-det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
-#all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
-#prior_dat <- read_csv(file.path(log_dir, "observe_harvest_recruit/priors.csv"))
-policies_w_priors <- read_csv(file.path(log_dir, "observe_harvest_recruit/policies_w_priors.csv"))
-ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))
+## Fig1, Fig2
+sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnnn")
+
+## Fig 3
+overest_sims <- read_csv(file.path(log_dir, "overest_sims.csv"))
+
+## Fig 4
+policies_w_priors <- read_csv(file.path(log_dir, "policies_w_priors.csv"))
+
+## Fig 5
+ram_ex <- read_csv(file.path(log_dir, "../ram_ex/ram_ex.csv"))
 
@@
 
+<!-- Should TAC go here? -->
+
 The literature of optimal harvests under uncertainty is one of paradoxes, which can only be resolved by a more comprehensive treatment of measurement uncertainty.
@@
 Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.
-
-First we will introduce the more heuristic alternative to addressing uncertainty, which more accurately reflects both intuition and actual management.
+In addition to resolving this paradox, we will demonstrate this direct solution also out-performs current more heuristic rules of thumb such as those currently used in fisheries management to account for uncertainty.
 
@@
 Fortunately, management practice has so far ignored this paradoxical suggestion that uncertainty should be met with greater rather than reduced fishing quotas.
-Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].
+Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].
 Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].
@@
 Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.
-
 Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.
@@
 
-<!--
-TAC, MDP, POMDP (sigma_g, sigma_m)
--->
-
-Figure 1 shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
-Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.
-Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.
-In each panel shown in Figure 1 stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
-In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.
-The first panel of Figure 1 shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.
-Without mesasurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
-As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.
-Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.
-
-Figure 1 confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.
-In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.
-Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC stategy rather than the CE strategy that lead to the highest overfishing rate.
-In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.
-
-```{r fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
+```{r sims, fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty.
 Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement (CE), Total allowable catch (TAC = 80% MSY), and the proposed partially observed Markov decision process (POMDP) method.
@@
 Carrying capacity K normalized to 1, r = 0.75.
-Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material."}
+Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material.
+\\label{sims}"}
 sims %>%
@@
 
-This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also bourne out in terms of economic performance.
-Figure 2 shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.
-As before, environmental stochasticity is set at $\sigma_g = 0.15$.
-In the absense of measurement uncertainty, constant escapement (CE) is optimal, though POMDP performs nearly as well, while TAC underperforms.
-As measurement uncertainty increases, the economic value of CE declines, remaining better than TAC at modest measurement error $\sigma_m = 0.1$ but worse at higher uncertainty.
-Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure 1), these have little impact on the economic value due to the discount rate.
-A smaller discount rate would penalize unlikely but not improbable collapses more, since a significant amount of time is required to realize those rare events.
-Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.
+Figure \ref{sims} shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
+Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.
+Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.
+In each panel shown in Figure \ref{sims} stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
+In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.
+The first panel of Figure \ref{sims} shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.
+Without mesasurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
+As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.
+Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.
 
-```{r fig.width=7, fig.height=3.5, fig.cap=""}
+Figure \ref{sims} confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.
+In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.
+Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC stategy rather than the CE strategy that lead to the highest overfishing rate.
+In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.
+
+```{r econ, fig.width=7, fig.height=3.5, fig.cap="\\label{econ}"}
 sims %>%
@@
 
+This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also bourne out in terms of economic performance.
+Figure \ref{econ} shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.
+As before, environmental stochasticity is set at $\sigma_g = 0.15$.
+In the absense of measurement uncertainty, constant escapement (CE) is optimal, though POMDP performs nearly as well, while TAC underperforms.
+As measurement uncertainty increases, the economic value of CE declines, remaining better than TAC at modest measurement error $\sigma_m = 0.1$ but worse at higher uncertainty.
+Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure \ref{sims}), these have little impact on the economic value due to the discount rate.
+A smaller discount rate would penalize unlikely but not improbable collapses more, since a significant amount of time is required to realize those rare events.
+Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.
+
 <!-- Show results of POMDP when over-estimating! -->
 
+```{r overest, fig.width=4, fig.height=3.5, fig.cap = "\\label{overest}"}
+overest_sims %>%
+  filter(method != "MSY") %>%
+  ggplot(aes(time, mean, col=method, fill=method)) + 
+  geom_line(lwd = 1) + 
+  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
+  coord_cartesian(ylim = c(0, 1)) 
+```
+
+Figure \ref{overest} summarizes the results of the same simulations under the scenario in which the POMDP approach assumes a measurement error of $\sigma_m = 0.15$, when in fact all simulated measurements are made without error ($\sigma_m = 0$).
+This represents an extreme case of ovestimating the measurement error.
+Stochastic growth remains the same as before, $\sigma_g = 0.15$, and simulations under TAC and CE policies are shown for comparison.
+When measurement error is absent, Reed's proofs hold that the CE solution is optimal -- interestingly, the POMDP solutions track almost exactly the CE solution despite the misplaced assumption that measurements are quite poor.
+As we have already seen, TAC solution is insufficiently cautious for this level of stochasticicty in recruitment, resulting in overexploitation and long-term decline.
+The ability of the POMDP solution to perform nearly optimally even when significantly overestimating the level of measurement uncertainty contrasts sharply to the significant declines from ignoring measurement uncertatinty seen in the CE solutions in Figure \ref{sim}.
+This demonstrates that while we may not know precisely the level of measurement error, we achieve far better outcomes overestimating measurement uncertainty than underestimating it.
+This also underscores the observation that POMDP policy is quite robust to the details of the uncertainty.
+We can get a better understanding for the performance of POMDP in these simulations by looking more closely at how any individual decision under a POMDP strategy compares to the action chosen by the current alternatives.
+
+```{r policy, fig.width=5, fig.height=7, fig.cap="\\label{policy}", message=FALSE, warning=FALSE}
+policies_w_priors %>% 
+  gather(panel, value, -state, -policy) %>%
+  ggplot(aes(state, value, col=policy)) + 
+  geom_line(lwd=1)  +
+  facet_wrap(~panel, ncol=1, scales = "free_y") +
+  scale_color_manual(values = colors) +
+  coord_cartesian(xlim = c(0,1), ylim = c(0,.8)) + 
+  theme(legend.position = "bottom")
+```
+
 To better understand the differences in performance of these strategies and resolve the paradox of uncertainty, we must take a closer look at how the specific action recommended by each policy compares given the same observation.
-Figure 3 shows the action taken by each strategy in response to a measurement of the stock size.
+Figure \ref{policy} shows the action taken by each strategy in response to a measurement of the stock size.
 The first panel shows the policy in terms of expected escapement, $S = Y - H$ for observed stock $Y$ and harvest $H$, consistent with most of the relevant control and bioeconomics literature [e.g.
@@
 Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
-While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 3, this is not the case for the POMDP approach.
+While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach.
 This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.
@@
 
-Figure 3 shows two separate policy curves (in terms of harvest and escapement) for the same POMDP solution given two different prior belief distributions (panel 3, priors).
+Figure \ref{policy} shows two separate policy curves (in terms of harvest and escapement) for the same POMDP solution given two different prior belief distributions (panel 3, priors).
 While both priors shown express considerable uncertainty about the precise stock size prior to the most recent observation, the lower prior is centered at a value $\tfrac{1}{4}K$ stock size, while the high prior is centered at a value of $\tfrac{3}{4}K$, relative to a (post-harvest, before observation) target size of $B_{MSY} = K/2$.
@@
 
-```{r fig.width=5, fig.height=7, fig.cap="", message=FALSE, warning=FALSE}
-policies_w_priors %>% 
-  gather(panel, value, -state, -policy) %>%
-  ggplot(aes(state, value, col=policy)) + 
-  geom_line(lwd=1)  +
-  facet_wrap(~panel, ncol=1, scales = "free_y") +
-  scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0,.8)) + 
-  theme(legend.position = "bottom")
-```
-
 **A resolution to the paradox**
@@
 In the POMDP solution, all information must be put into its historical (and constantly updated) context.
-When a measurement roughly matches the expectation of this prior context, Figure 3 shows that the optimal response from POMDP is roughly comparable or slightly more cautious to the harvest under no uncertainty, and not more aggressive as the paradox would suggest.
+When a measurement roughly matches the expectation of this prior context, Figure \ref{policy} shows that the optimal response from POMDP is roughly comparable or slightly more cautious to the harvest under no uncertainty, and not more aggressive as the paradox would suggest.
 Measurements that exceed expectations are tempered with some skepticism: while the CE solution is willing to meet a high stock measurement with a large harvest, the POMDP solutions increase harvest more cautiously.
@@
 Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where chance local extinction becomes possible.
-It is precisely this assymetry: that too much overharvesting leads to an irreversible state of extinction, while too much underharvesting is always reversable (modulo some lost revenue) that lay behind Clark's original intutition that there was something fishy about Reed's result that S = D, that uncertainty required no extra caution.
+It is precisely this assymetry: that too much overharvesting leads to an irreversible state of extinction, while too much underharvesting is always reversable (modulo some lost revenue) that lay behind Clark's original intutition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.
 Constant escapement is particularly suscpetible to overharvesting starting from stock sizes much higher than $B_{MSY}$ since its bang-bang optimal solution attempts to bring stocks back to target level as fast as possible.
@@
 
-```{r message=FALSE, warning=FALSE, fig.width=7, fig.height = 5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
-POMDP solution assumes a measurement error of 10%."}
+```{r historical, message=FALSE, warning=FALSE, fig.width=7, fig.height = 5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.
+POMDP solution assumes a measurement error of 10%.
+\\label{historical}"}
 ram_ex %>%
@@
 So far we have focused on simulation and an examination of the relative policies under different prior information.
-Figure 4 compares the harvest level that would have been recommmended by each of the strategies we have considered here against the historically observed catch recorded for two commericially fished stocks: Argentine Hake and European Plaice.
+Figure \ref{histroical} compares the harvest level that would have been recommmended by each of the strategies we have considered here against the historically observed catch recorded for two commericially fished stocks: Argentine Hake and European Plaice.
 Historical estimates of biomass and catch are taken from the R.A.
@@
 
-<!-- True whether or not it is implemented in effort-based control or as a TAC-->
-
 <!--

### 2018-01-26 appendix/appendixB.Rmd
plotting

--- 2018-01-12 appendix/appendixB.Rmd
@@
 ---
-output: github_document
+output: pdf_document
 ---
@@
 
+```{r libraries, message=FALSE}
+# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
+library(tidyverse)
+library(sarsop)
+library(nimble)
+library(parallel)
+library(gridExtra)
+```
+
 ```{r plotting}
-library(gridExtra)
+## This chunk is optional, just tweaks plot appearance / fonts
 library(ggthemes)
-library(ggplot2)
 library(Cairo)
@@
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+#palette <- ptol_pal()(6)
+```
 
-palette <- ptol_pal()(6)
+```{r}
 palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
-
 colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
@@
                       "catch"))
-```
-
-```{r libraries, message=FALSE}
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
-library(tidyverse)
-library(sarsop)
-library(nimble)
-library(parallel)
 ```
@@
 ```{r}
-sims %>%
-  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
+ram_ex <- read_csv(file.path(log_dir, "ram_ex.csv"))
+
+ram_ex %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy", "msy")) %>%
+   mutate(model = fct_recode(model,
+                             "MSY" = "msy",
+                             "CE" = "Reed",
+                             "TAC" = "pgy"
+                             )) %>%
   ggplot(aes(year, stock, col=model)) +
   geom_line(lwd=1) +
-  scale_color_manual(values = c("black", "grey", "#D9661F", "#3B7EA1", "#FDB515", "#6C3302",  "#00B0DA",  "#CFDD45")) +
-  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
+  scale_color_manual(values = colors) +
+  facet_wrap(~commonname, scales = "free", ncol=1)  
 ```

### 2018-01-26 appendix/appendix_harvestfirst.Rmd
plotting

@@
 knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE)
-ggplot2::theme_set(ggplot2::theme_bw())
 ```
 
-```{r message=FALSE}
+```{r library, message=FALSE}
 # devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
@@
 library(parallel)
+library(gridExtra)
 ```
 
-```{r plotting}
-library(gridExtra)
+```{r optional_plotting_config}
 library(ggthemes)
@@
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+```
 
-palette <- ptol_pal()(6)
+```{r colors}
+#palette <- ptol_pal()(6)
 palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
@@
 
-```{r pomdp_sims}
+```{r overest}
 set.seed(12345)
@@
 ```
-
-```{r  fig.width=5, fig.height=7, fig.cap=""}
-prior <- 
-  ggplot(priors,aes(state, belief, color = policy)) + 
-  geom_line(lwd=1)  + 
-  scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1)) +
-  ggtitle("POMDP priors") + 
-  xlab("Estimated stock biomass") + 
-  theme(legend.position = "none")
-
-harvest <- policy_df %>%
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line(lwd=1)  +
-  scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
-  ggtitle("Harvest view") + 
-  theme(legend.position="none") + 
-  xlab("Estimated stock biomass")
-
-escapement <- policy_df %>%
-  ggplot(aes(state, escapement, col=policy)) + 
-  geom_line(lwd=1)  +
-  scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
-  xlab("Estimated stock biomass") +
-  ggtitle("Escapement view") + 
-  theme(legend.position = "bottom")
-
-gridExtra::grid.arrange(harvest, escapement, prior,
-  layout_matrix = rbind(1,1,2,2, 3))
-
-```
-
-```{r}
-
-```

### 2018-01-26 manuscripts/manuscript.Rmd
WIP

@@
 
-# Model
+# The Decision Problem
 
 We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.
-To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),
-
-$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$
-
-where $x_t$ is the current stock size, $h_t$ the harvest chosen that year, with parameters $r$ giving the individual growth rate, and $K$ the carrying capacity and $\varepsilon$ representing stochastic recruitment in a variable environment.
-We will assume for simplicity $\varepsilon \sim \mathcal{N}(1, \sigma_g)$.
-
-This and similar models are widely used in large scale analyses across diverse stocks (e.g.
-[@Costello2016; @Britten2017]), and form the basis for much of bioeconomic theory [@Clark1990].
-As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arsing only because of some particular subtle assumption hidden in a more complex model.
-
-However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
-Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
-@].
-As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.
-
-# The Decision Problem
 
@@
 
+## Population model
+
+To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),
+
+$$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$$
+
+where $x_t$ is the current stock size, $h_t$ the harvest chosen that year, with parameters $r$ giving the individual growth rate, and $K$ the carrying capacity and $\varepsilon$ representing stochastic recruitment in a variable environment.
+We will assume for simplicity $\varepsilon \sim \mathcal{N}(1, \sigma_g)$.
+This and similar models are widely used in large scale analyses across diverse stocks (e.g.
+[@Costello2016; @Britten2017]), and form the basis for much of bioeconomic theory [@Clark1990].
+As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arsing only because of some particular subtle assumption hidden in a more complex model.
+
+However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
+Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
+@].
+As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.
+
 <!-- MDP, POMDP -->
 
-# Current theory and practice
+# Current Theory and Practice
 
@@
 
-This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis, and does not solve the time-dependent optimization problem above.
-In particular, maximum sustainable yield will not be optimal whenever the stock is away from $B_{MSY}$.
-@Clark1973 shows that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*.
-The time-depednent optimal stategy is impatient: it seeks to achieve the optimal yield as quickly as possible.
-Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, mortalities greater than $F_{MSY}$ can be optimal to bring the stock back to $B_{MSY}$ more quickly.
+This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis which does not solve the time-dependent optimization problem above.
+In particular, this maximum sustainable yield (MSY) strategy will not be optimal whenever the stock is away from $B_{MSY}$.
+@Clark1973 deomonstrated (assuming no uncertainty in measurement or stochasticity in population dynamics) that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*.
+We will return to this approach in a moment.
+
+## Uncertainty in Practice: MSY & TAC
+
+Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].
+Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].
+Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.
+Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Overfishing Limit, (OFL).
+Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment.
+To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.
+To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
+To distinguish this approach from MSY, we will refer to this approach as a TAC policy.
+This more closely represents a heuristic [e.g.
+@Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.
+Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.
+Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.
+
+## Optimal management
+
+So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of of paradoxes.
+Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible.
+Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.
 (This intution must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
@@
 While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.
-
-<!-- Should TAC go here? -->
-
-The literature of optimal harvests under uncertainty is one of paradoxes, which can only be resolved by a more comprehensive treatment of measurement uncertainty.
+Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above.
+Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarise here.
 
@@
 
-The first of these we shall refer to as Reed's Paradox, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.
+The first of these we shall refer to as Reed's Paradox for stochastic growth, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.
 This suprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.
@@
 
-> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
-of the problem increases markedly if this assumption is relaxed.)
+> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period.
+(The mathematical difficulty of the problem increases markedly if this assumption is relaxed.)
 
-By leveraging algorithms (and also computational hardware) developed over two decades later are we able to crack this mathematical difficulty Clark refers to directly without this "simplyfying assumption."  Unforunately for @Clark1986, this "simplifying assumption" they made instead only deepened the paradox, leading them to find policies that become *less* cautious as uncertainty increases:
+Unforunately for @Clark1986, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:
 
@@
 
-Using a different but still flawed assumption nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observed with some concern:
+Relying on a quite different but still flawed assumption nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observed with some concern:
 
@@
 
-While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
+Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than rejct it.
+Others have also attempted to introduce measurement error, but always relying on various approximations that similarly alter the problem.
+The most common such assumption is to assume from the outset that the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
 Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.
 In addition to resolving this paradox, we will demonstrate this direct solution also out-performs current more heuristic rules of thumb such as those currently used in fisheries management to account for uncertainty.
-
-## Uncertainty in Practice: TAC
-
-Fortunately, management practice has so far ignored this paradoxical suggestion that uncertainty should be met with greater rather than reduced fishing quotas.
-Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].
-Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].
-Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.
-Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Overfishing Limit, (OFL).
-Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment.
-To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.
-To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
-To distinguish this approach from MSY, we will refer to this approach as a TAC policy.
-This more closely represents a heuristic [e.g.
-@Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.
-Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.
-Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.
 
@@
   summarise(mean = mean(state), sd = sd(state)) %>% 
-  mutate(sigma_m = as.character(sigma_m)) ->
+  mutate(sigma_m = as.character(sigma_m)) %>%
+  ungroup() ->
 fig1_df
@@
 
-```{r econ, fig.width=7, fig.height=3.5, fig.cap="\\label{econ}"}
+```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$.
+Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{{sims}}.
+The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error.
+The more cautious TAC is not economically optimal but largely unimpacted by increasing error, while POMDP attains consistently high economic yield despite the increasing uncertainty.
+\\label{econ}"}
 sims %>%
@@
 As before, environmental stochasticity is set at $\sigma_g = 0.15$.
-In the absense of measurement uncertainty, constant escapement (CE) is optimal, though POMDP performs nearly as well, while TAC underperforms.
-As measurement uncertainty increases, the economic value of CE declines, remaining better than TAC at modest measurement error $\sigma_m = 0.1$ but worse at higher uncertainty.
+In the absense of measurement uncertainty, constant escapement (CE) is optimal (as per @Reed1979), while the heuristic caution built into the total allowable catch (TAC) strategy (at 80% MSY) results in a sub-optimal economic yield.
+The presense of stochasticity in recruitment ($\sigma_g = 0.15$), also contributes to the reduction in yield under TAC.
+Though CE is quite robust to the environmental stochasticity level, this strategy proves very sensitive to increasing measurement error, showing sharp declines in economic value, falling below the TAC economic value at $\sigma_m = 0.15$.
 Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure \ref{sims}), these have little impact on the economic value due to the discount rate.
@@
 Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.
+In contrast to the declining economic performance of CE and the consistently sub-optimal economic yield for TAC, the POMDP strategy continues to generate economic yields at approximately the optimal level (as attained by CE in the absense of measurement error) despite the increasingly uncertain measurements.
+This demonstrates that the reduced risk of stock collapse and higher average stock biomass attained by the POMDP strategy in the presence of high measurement error, Figure \ref{sims}) is not the result of a trivial reduction in harvesting under all circumstances, but rather, evidence of a more nuanced strategy that manages to account for the uncertainty in measurement while maintaining a reasonable harvest.
 
@@
 
-```{r overest, fig.width=4, fig.height=3.5, fig.cap = "\\label{overest}"}
+```{r overest, fig.width=4, fig.height=3.5, fig.cap = "Average fish biomass when the POMDP strategy assumes a high level of measurement uncertainty, while simulations reflect perfect measurements.
+Despite this overestimation, biomass under POMDP strategy closely tracks the optimal biomass under CE in the absence of measurement error.
+\\label{overest}"}
 overest_sims %>%
@@
   geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
-  coord_cartesian(ylim = c(0, 1)) 
+  coord_cartesian(ylim = c(0, 1))  + 
+  theme(legend.position = "bottom") +
+  scale_color_manual(values = colors) +
+  scale_fill_manual(values = colors) 
 ```
 
-Figure \ref{overest} summarizes the results of the same simulations under the scenario in which the POMDP approach assumes a measurement error of $\sigma_m = 0.15$, when in fact all simulated measurements are made without error ($\sigma_m = 0$).
+The POMDP strategy is also robust to overestimation of the level of measurement uncertainty.
+In the simulation results shown in Figures \ref{sims} and \ref{econ}, we have assumed that the level of measurement uncertainty $\sigma_m$ was known, and saw that ignoring this uncertainty (as the CE policy does) has significant negative impact on ecological and economic outcomes.
+If the level of measurement uncertainty is not known precisely, _overestimating_ the measurement error while using the POMDP strategy provides a precautionary approach that can nevertheless achieve nearly optimal ecological and economic outcomes.
+Figure \ref{overest} summarizes the results of the same simulations as before, but under the scenario in which the POMDP approach assumes a measurement error of $\sigma_m = 0.15$, when in fact all simulated measurements are made without error ($\sigma_m = 0$).
 This represents an extreme case of ovestimating the measurement error.
 Stochastic growth remains the same as before, $\sigma_g = 0.15$, and simulations under TAC and CE policies are shown for comparison.
-When measurement error is absent, Reed's proofs hold that the CE solution is optimal -- interestingly, the POMDP solutions track almost exactly the CE solution despite the misplaced assumption that measurements are quite poor.
-As we have already seen, TAC solution is insufficiently cautious for this level of stochasticicty in recruitment, resulting in overexploitation and long-term decline.
-The ability of the POMDP solution to perform nearly optimally even when significantly overestimating the level of measurement uncertainty contrasts sharply to the significant declines from ignoring measurement uncertatinty seen in the CE solutions in Figure \ref{sim}.
+When measurement error is absent, Reed's proofs hold and the CE strategy is optimal.
+Figure \ref{overest} shows that the POMDP outcomes track almost exactly the CE outcomes despite the misplaced assumption that measurements are quite poor.
+(As we have already seen, the TAC stategy is insufficiently cautious for this level of stochasticicty in recruitment, resulting in overexploitation and long-term decline).
+The ability of the POMDP solution to perform nearly optimally even when significantly overestimating the level of measurement uncertainty contrasts sharply to the significant declines from ignoring measurement uncertatinty seen in the CE solutions in Figure \ref{sims}.
 This demonstrates that while we may not know precisely the level of measurement error, we achieve far better outcomes overestimating measurement uncertainty than underestimating it.
@@
 
-```{r policy, fig.width=5, fig.height=7, fig.cap="\\label{policy}", message=FALSE, warning=FALSE}
+```{r policy, fig.width=5, fig.height=6, fig.cap="Comparison of the harvest and escapement policies under each strategy as a function of the observed stock size.
+Escapement refers to the expected fraction of fish left in the sea, $x-h$, while harvest refers to the target catch; two different conventions for plotting the same action.
+Uniquely, the POMDP policy plot will depend not only on the observed stock size, but also depends on prior information (bottom panel) as determined from any prior observations and actions.
+Depending on this prior it may harvest less or more than the other policies given an identical observation.\\label{policy}", message=FALSE, warning=FALSE}
 policies_w_priors %>% 
   gather(panel, value, -state, -policy) %>%
-  ggplot(aes(state, value, col=policy)) + 
+  rename(method = policy) %>%
+  ggplot(aes(state, value, col=method)) + 
   geom_line(lwd=1)  +
   facet_wrap(~panel, ncol=1, scales = "free_y") +
+  xlab("observed stock size") +
+  ylab("policy") +
   scale_color_manual(values = colors) +
-  coord_cartesian(xlim = c(0,1), ylim = c(0,.8)) + 
-  theme(legend.position = "bottom")
+#  theme(legend.position = "bottom") +
+  coord_cartesian(xlim = c(0,1), ylim = c(0,.8))
+
 ```
@@
 To better understand the differences in performance of these strategies and resolve the paradox of uncertainty, we must take a closer look at how the specific action recommended by each policy compares given the same observation.
-Figure \ref{policy} shows the action taken by each strategy in response to a measurement of the stock size.
+Figure \ref{policy} shows the action taken by each strategy in response to a measurement of the stock size (biomass estimate).
+These plots show, for any possible observations, which strategy will attempt to harvest most and which will attempt to harvest least.
 The first panel shows the policy in terms of expected escapement, $S = Y - H$ for observed stock $Y$ and harvest $H$, consistent with most of the relevant control and bioeconomics literature [e.g.
@@
 These plots clearly illustrate the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.
-Viewing the same policies in terms of escapement, makes it clear that only CE strategy corresponds to keeping escapement constant.
+Escapement CE strategy corresponds to keeping escapement constant.
 This is important, as theoretical work ever since the proofs of @Reed1979 often attempts to simplify the calculations by assuming the optimal policy must be a constant-escapement type strategy [e.g.
@@
 So far we have focused on simulation and an examination of the relative policies under different prior information.
-Figure \ref{histroical} compares the harvest level that would have been recommmended by each of the strategies we have considered here against the historically observed catch recorded for two commericially fished stocks: Argentine Hake and European Plaice.
+Figure \ref{historical} compares the harvest level that would have been recommmended by each of the strategies we have considered here against the historically observed catch recorded for two commericially fished stocks: Argentine Hake and European Plaice.
 Historical estimates of biomass and catch are taken from the R.A.
@@
 Here, the POMDP solution falls in between these extremes, almost mirroring which-ever of the two policies is more conservative: the POMDP solution matches the dips in harvest taken by CE to allow the stock to recover most quickly, but does not track the doubling of harvests recommended by CE in the mid 1980, increasingly only to the more modest harvest recommended by the TAC policy.
-
 Once again, we see that POMDP solution provides a more consistently precautionary policy than either the over-simplified optimal control solution of CE or the more rule-of-thumb approach represented by TAC.
@@
 
-## Discussion
+# Discussion
 
-Using modern algorithms we have been able to crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk a long-standing paradox that increased measurement uncertainty should be met with larger harvests.
-We have illustrated that the POMDP approach introduced here performs consistently better across a range of stochastic growth and measurement error levels, while previous solutions based on either classic optimal control solutions or heuristic rules-of-thumb give significantly worse outcomes in both ecological and economic terms.
+Using modern algorithms we have been able to crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk long-standing paradoxes that increased measurement uncertainty should be met with larger harvests.
+We have demonstrated that both existing optimal control approaches and existing precautionary approaches result in overexploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.
+These results have important implications for both global fisheries and ecological management more generally.
+In marine fisheries, evidence of such over-exploitation has already been well documented [@Worm2006; @Worm2009; @Costello2016].
+Our results suggest that measurement uncertainty may contribute to this pattern, and that reality of measurement errors may frustrate stock recovery under current best practices such as MSY, TAC, and CE.
+More generally, the study and application of strategies such as MSY and CE are in no way limited to the decisions in fisheries quotas, but underpin much of conservation and natural resource management in forests and othe terrestrial ecosystems as well [e.g.
+@Clark1990].
+Our results underscore the realization that measurement uncertainty is not negligible.
+Neither optimization approaches which ignore uncertainty (like CE) or heuristic approaches to account for that uncertainty (like TAC) are sufficient for this task.
+Rather, adaptive approaches such as POMDP, which make use of previous estimates to continually adjust prior beliefs, are required to avoid negative ecological and economic outcomes.
 Nevertheless, it may be tempting to see these results as either trivial or impractical, but a closer inspection reveals they are neither.
 
-First, It may seem obvious that POMDP would out-proform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.
+First, it may seem obvious that POMDP would out-proform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.
 Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accomidate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.
@@
 ^1: Given the sophisticated policies often in place to determine actual TAC levels, one might object to our characterization of this as merely a rule-of-thumb.
-We use this term merely to distinguish it from the optimal control approach -- that despite technical precision, TAC adjustments are set by optimal control
+We use this term merely to distinguish it from the optimal control approach -- that despite technical precision, TAC adjustments are set by a hueristic instead of the result of some dynamic optimization algorithm as in CE and POMDP.
 
-If it is not obvious that POMDP solution would do significantly better, one might still ask if such approaches are practical.
-The importance of simple and effective rules of thumb in conservation management has been well documented [e.g.
-@Possingham].
-Against this we make the following observations.
-First, the strategies we have compared against are not purely of academic interest but underpin both current practice and management aspirations.
-Together, MSY and TAC-based management reflects much of modern best practice as it is understood in fisheries management and elsewhere; US salmon stocks are one of few examples using a constant escapement-based approach instead.
-While real-world application of these approaches frequently involves more complex models, the models we have considered here are the ones under which these strategies were designed and tested, and there is little reason to believe they would perform better under more complex cases such as age-structured models that do not always meet those assumptions [e.g.
+Second, we must ask if the policies derived from POMDP are practical.
+As with constant escapement, the POMDP solutions we considered here result in more year-to-year variation in harvest, and include the possibilty of fishing moritorium to allow stocks to recover -- decisions that many view as economically or socially unacceptable.
+While resolving this debate is beyond the scope of this paper, a few observations are relevant: (A) Stocks already managaed under constant escapement, such as US salmon fisheries, are already subject to the possibility of these moritoriums.
+(B) It is straight-forward to introduce constraints on the minimal acceptable harvest level and/or include the additional costs of closing a fishery in the utility function optimized by POMDP.
+(C) Fisheries today are not managed at either their maximum ecological or economic potential.
+Taken as a whole, most ocean fisheries yield a net negative economic return [@FAO2009], with fishing persisting only with the support of government subsidies [@Arnason2012], meanwhile some 68% of global fisheries are in poor biological condition [@Costello2016].
+While the causes for this are both complex and numerous, we have seen that both the economic inefficiency and ecological health suffer under TAC/MSY and CE based management.
+@Arnason2012 and @Costello2016 have both argued that a rights-based approach to fisheries management (RBFM) would lead to self adoption of precisely such optimal policies.
+
+Third, are the examples considered here too simple?
+The strategies we have compared against underpin both current practice and management aspirations.
+Together, MSY and TAC-based management reflects much of modern best practice as it is understood in fisheries management and elsewhere.
+While real-world application of these approaches frequently involves more complex models, the models we have considered here are the very ones under which these strategies were designed and tested, and there is little reason to believe they would perform better under more complex cases such as age-structured models that do not always meet those assumptions [e.g.
 @Holden2015].
@@
 
-**Are moritoriums pratical?**
-Fisheries today are not managed at either their maximum ecological or economic potential.  
-Taken as a whole, most ocean fisheries yield a net negative economic return
-[@FAO2009], with fishing persisting only with the support of government subsidies [@Arnason2012], 
-meanwhile some 68% of global fisheries are in poor biological condition [@Costello2016].
-While the causes for this are both complex and numerous, we have seen that both the economic
-inefficiency and ecological health suffer under TAC/MSY and CE based management.
+Fourth and last, is the approach taken here too complex to be feasible?
+The importance of simple and effective rules of thumb in conservation management has been well documented [e.g.
+@Possingham].
+Sophisticated and computationally intensive approaches such as POMDP may seem improbable at a time when many areas of natural resource management, even simple, rule-of-thumb based methods struggle to take root.
+And yet, in this same age we rely on far more complex and computationally intensive algorithms to perform far more trivial tasks such as determining what advertisements we see.
+Surely a planet that can afford such complex optimization for how advertising space is allocated on a small screen can one day manage it's natural resources and conservation efforts with such tools?
 
-**Is computationally intensive management feasible?**
-Lastly, one may dismiss the potential of POMDP approaches as being too computationally infeasible for everyday management practices.
-In many areas of natural resource management, even simple, rule-of-thumb based methods struggle to take root.
-In such a world, how can complex artificial intelligence algorithms requiring significant computational resources to execute play any role?
-And yet, in that same world far more complex and demanding algorithms are run every day to provide features such as face recognition to smart phone applications.
-If such resources and expertise can now facilitate every day social decisions with only a tap, surely we can one day harness the same complexity and power to facilitate conservation and resource management.
+<!--
 
-The year-to-year variation in quotas necessary under either a constant escapement policy
-or a policy in which quotas are set by POMDP can seem impractical from a top-down
-management perspective, particularly when compared to the simplicity of fixing a mortality of fishing $F_{MSY}$.
-However, the economist's perspective reveals that this may not be as unrealistic as it may first appear.
+## Future directions
 
+-->
+
+<!--
 Taken as a whole, most ocean fisheries yield a net negative economic return
@@
 the value of ecosystem services from fish not harvested [@Halpern2013].  
+-->
 

### 2018-01-27 manuscripts/manuscript.Rmd
WIP

@@
 As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP).
+The constant escapement (CE) strategy considered here is the result of such an MDP or SDP analysis.
 Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).
-The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process.
+The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process, but cannot be solved using SDP directly.
 The POMDP problem for fisheries question considered here can be summarized as follows:
@@
 Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.
-As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms.
+While the solution method for MDP problems, stochastic dynamic programming (SDP) is well known and readily implemented in ecological problems [e.g @Mangel1988; @Marescot2013], the POMDP problem cannot be efficiently solved using SDP.
 While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
 @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
-POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+POMDP algorithms remain a very active area in artificial intellegence [.
+@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
 Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
@@
 The mechanics of this process are well documented in the extensive literature on POMDPs [e.g.
-@pomdpstuff; @stuff; @more], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.
+@Smallwood1973; @Sondik1978; @Kurniawati2008; @Williams2011], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.
 
@@
 More generally, the study and application of strategies such as MSY and CE are in no way limited to the decisions in fisheries quotas, but underpin much of conservation and natural resource management in forests and othe terrestrial ecosystems as well [e.g.
-@Clark1990].
+@Clark2000].
 Our results underscore the realization that measurement uncertainty is not negligible.
@@
 
-Third, are the examples considered here too simple?
+Third: are the examples considered here too simple?
 The strategies we have compared against underpin both current practice and management aspirations.
@@
 
-Fourth and last, is the approach taken here too complex to be feasible?
+Fourth: is the approach taken here too complex to be feasible?
 The importance of simple and effective rules of thumb in conservation management has been well documented [e.g.
-@Possingham].
+@Chades2011].
 Sophisticated and computationally intensive approaches such as POMDP may seem improbable at a time when many areas of natural resource management, even simple, rule-of-thumb based methods struggle to take root.
@@
 
-<!--
+<!-- Adaptive management -->
 
@@
 
--->
+Several limitations that have been studied in fully observed (MDP) optimal
+decision problems, such as parameter uncertainty [e.g. @Ludwig1982], model 
+uncertainty [@Williams2001; @Boettiger2015] and adaptive management [e.g. @Walters1976]
+remain largely open challenges for partially observed systems. Future work 
+could extend this analysis to more complex models, such as those with age structure,
+as [@Holden2015] does for the fully observed case. Another limiting assumption 
+common to MDPs and POMDPs is that of stationary dynamics: that the population 
+dynamics equation itself is not changing over time. In reality, forces such as
+climate change and other forms of environmental variations violate this
+assumption  [@Britten2017]. Direct approaches such as @Fackler2014's
+adaptation of Mixed Observability MDP [@Ong2010] do not scale to the number
+of states and actions considered here. A value of information (VOI) analysis
+for POMDP,such as @Johnson2015 shows for MDP models, could identify when it
+is worthwhile to actively reduce measurement error.
 

### 2018-01-27 appendix/appendixB.Rmd
yaml and tidy

@@
 ---
+title: "Appendix B: Code for analysis comparing to historical data"
+author: "Carl Boettiger"
+date: "`r Sys.Date()`"
 output: pdf_document
@@
 
-Computes model parameter estimates on each `stockid` in RAM (after normalizing data) using nimble. Then, estimate common stock measurement (Bmsy, MSY, PGY).
+Computes model parameter estimates for selected stocks in RAM using NIMBLE. estimate $B_{MSY}$
 
@@
 
-```{r plotting}
+```{r plotting, include = FALSE}
 ## This chunk is optional, just tweaks plot appearance / fonts

### 2018-01-27 appendix/appendix_harvestfirst.Rmd
yaml and tidy

@@
 ---
-title: "POMDP comparisons across sigma_m, sigma_g"
+title: "Appendix A: Code and supplementary figures for simulation and policy analyis"
 author: "Carl Boettiger"
 date: "`r Sys.Date()`"
-output: 
-  github_document: default
-  pdf_document: default
+output: pdf_document
 ---
@@
 ```{r library, message=FALSE}
-# devtools::install_github("boettiger-lab/sarsop")  ## install package first if necessary.
+## install package first if necessary.
+# devtools::install_github("boettiger-lab/sarsop") 
 library(sarsop)       # the main POMDP package
@@
 
-```{r optional_plotting_config}
+```{r optional_plotting_config, include=FALSE}
 library(ggthemes)
@@
 #palette <- ptol_pal()(6)
-palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
-
+palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45")
 colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 

### 2018-01-29 appendix/appendixB.Rmd
update appendix with clean names, a bit more text

@@
 
-```{r aethetics, include=FALSE}
-#library(printr)
-knitr::opts_chunk$set(message=FALSE, comment="", cache=TRUE)
+```{r knitr, include=FALSE}
+knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE, dev="cairo_pdf")
 ```
@@
 extrafont::loadfonts()
-knitr::opts_chunk$set( dev="cairo_pdf")
 hrbrthemes::import_roboto_condensed() 
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
-#palette <- ptol_pal()(6)
-```
 
-```{r}
-palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
+palette <- c("#D9661F", "#3B7EA1", "#6C3302",
+             "#FDB515", "#00B0DA",  "#CFDD45")
 colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
@@
                       "catch"))
+scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
+scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
 ```

### 2018-01-29 appendix/appendix_harvestfirst.Rmd
update appendix with clean names, a bit more text

@@
 output: pdf_document
+bibliography: ../manuscripts/citations.bib
+
 ---
 
-```{r aethetics, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE)
+# Overview
+
+1. Model definition:
+2. Deterministic Solutions
+3. Discretization and the POMDP solution
+4. Simulation
+
+```{r knitr, include=FALSE}
+knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf")
 ```
@@
 extrafont::loadfonts()
-knitr::opts_chunk$set( dev="cairo_pdf")
 hrbrthemes::import_roboto_condensed() 
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
-```
 
-```{r colors}
+## Set colors
 #palette <- ptol_pal()(6)
 palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45")
-colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
+colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black", palette[6]), 
                     c("TAC", "POMDP", "MSY", "CE", 
@@
                       "biomass",
-                      "catch"))
+                      "catch",
+                      "SDP"))
+## Set as default palette so we can omit scale_color_manual in every plot call
+## https://stackoverflow.com/a/44307181/258662
+scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
+scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
 ```
@@
 
-## Basic deterministic model
+# Basic deterministic model
 
@@
 
-## Unlike classic Graham-Schaefer, this assumes harvest occurs right after assessment, before recruitment
+## Unlike classic Graham-Schaefer, this assumes harvest 
+## occurs right after assessment, before recruitment
 f <- function(x, h){ 
@@
 
-## Calculating MSY
+# Calculating MSY & TAC
 
@@
 #B_MSY <- S_star     # recruit first, as in classic Graham-Schaefer
-B_MSY <- f(S_star,0) # harvest first, we observe the population at B_msy + h
+B_MSY <- f(S_star,0) # harvest first, we observe the population at B_MSY + h
 
@@
 F_MSY <- MSY / B_MSY  
-F_PGY = 0.8 * F_MSY
+F_TAC = 0.8 * F_MSY
 ```
@@
 ```{r}
-
-msy_policy <- function(x) F_MSY * x
-pgy_policy <- function(x) F_PGY * x
+MSY_policy <- function(x) F_MSY * x
+TAC_policy <- function(x) F_TAC * x
 
@@
       action[t] <- policy(state[t])
-      obs[t] <- state[t] - action[t]  # if we observe after harvest but before recruitment
+      obs[t] <- state[t] - action[t]  # we observe->harvest->recruitment
       state[t+1] <- f(state[t], action[t]) 
@@
 det_sims <- 
-  list(msy = msy_policy, 
-       pgy = pgy_policy,
-       det = escapement_policy) %>% 
+  list(MSY = MSY_policy, 
+       TAC = TAC_policy,
+       CE = escapement_policy) %>% 
   map_df(do_det_sim, 
@@
 More conservative rules, such as a harvest set to 80% of MSY result in faster recovery of the stock than under MSY, but slower than under constant escapement.
-Due to the reduced maximum harvest rate, such rules lead to stock returning to a value higher than $B_${MSY}$.
+Due to the reduced maximum harvest rate, such rules lead to stock returning to a value higher than $B_{MSY}$.
 
-```{r fig.width=3.5, fig.height=3.5}
+```{r fig.width=5, fig.height=5}
 det_sims %>%
-  mutate(method = fct_recode(method, 
-                             "CE" = "det",
-                             "TAC" = "pgy",
-                             "MSY" = "msy")) %>%
   ggplot(aes(time, state, col=method)) + 
   geom_line(lwd=1) + 
-  scale_color_manual(values = colors) +
   coord_cartesian(ylim = c(0, 1)) + 
@@
 
-##  Introduce a discrete grid
+#  Introduce a discrete grid
 
@@
 policies <- data.frame(
-#  det = index(pmax(f(states,0) - S_star,0), actions), # obs,recruit,harv, f(x_t) - h_t
-  det = index(escapement_policy(states),     actions), # obs,harv,recruit, f(x_t - h_t)
-  msy = index(msy_policy(states),               actions),
-  pgy = index(pgy_policy(states),               actions))
+#  CE = index(pmax(f(states,0) - S_star,0), actions), # obs,recruit,harv, f(x_t) - h_t
+  CE = index(escapement_policy(states),     actions), # obs,harv,recruit, f(x_t - h_t)
+  MSY = index(MSY_policy(states),               actions),
+  TAC = index(TAC_policy(states),               actions))
 ```
@@
 
-We compute POMDP matrices for a range of `sigma_g` and `sigma_m` values:
+We compute POMDP matrices for a range of `sigma_g` and `sigma_m` values.
+(We consider `sigma_g = 0.02` rather than precisely zero here to avoid a degeneracy in the discretized tranisition matrix.
+Calculations for the precise deterministic solution can be solved without discretization as shown above, and match the results of this nearly-zero nosie level in the discrete simulation, as expected.)
 
@@
 dir.create(log_dir)
-
-## POMDP solution 
-## (slow, >10,000 seconds per scenario for discount ~ 0.95;
-## needs much longer for discount ~ 0.99)
 system.time(
@@
 
+------
+
+# Simulations
+
+Having calculated the POMDP solution in terms of these `alpha` vectors, we can easily compute the POMDP policy for any given prior and update the prior given further observations according to Bayes rule.
+We assume a uniform prior belief at the start of the POMDP simulation.
+These tasks are perfromed by the `sim_pomdp` fucntion.
+Simulating the static strategies (CE, MSY, TAC) is easier, since the policy is uniquely determined by the observed state; we simply need to simulate the stochastic growth and process and measurement with error at each time step an then compute the corrsponding policy.
+The `sim_pomdp` function performs such static simulations when given a pre-specified policy instead of a set of `alpha` vectors.
+
 ## Simulating the static policies under uncertainty
@@
                        reps = reps)$df %>% 
-              mutate(method = "pomdp") # include a column labeling method
+              mutate(method = "POMDP") # include a column labeling method
            },
@@
 
-Combine the resulting data frames
+We then combine the resulting data frames, transition (mutate) the units from grid indices (`state`, `action` $\in 1 ...
+N$) to continuous values (`states`, `actions` $\in [0,2]$), and select necessary columns before writing the data out to a file.
 
@@
 
-sim_col_types <- paste0(gsub("^(\\w).*", "\\1", sapply(unname(sims), class)), collapse = "")
 write_csv(sims, file.path(log_dir, "sims.csv"))
@@
 
-We the results varying over different noise intensities, sigma_g, and sigma_m.  Figure 1 of the main text considers the case of sigma_g = 0.05, sigma_m = 0.1
+We the results varying over different noise intensities, `sigma_g`, and `sigma_m`.
+Note that Figure 1 of the main text shows only the case of `sigma_g = 0.15`, `sigma_m = 0.1` and omits MSY to simplify the presentation.
 
 ```{r}
-sim_col_types <-  "inicnnn"
-sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = sim_col_types)
+sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnnn")
 
@@
 
-##  Figure S3
+##  Figure S3: Economic value
 
-Economic Value
+Economic value is shown under varying noise levels under each strategy.
 
@@
   ungroup() %>%
-  mutate(method = fct_recode(method,
-                             "MSY" = "msy",
-                             "CE" = "det",
-                             "TAC" = "pgy",
-                             "POMDP" = "pomdp")) %>%
 
@@
   theme(legend.position = "none") + 
-  scale_color_manual(values = colors) +
-  scale_fill_manual(values = colors) +
   ylab("expected net present value")
@@
 ```
-
------
 
@@
 However, the POMDP approach is quite robust to overestimation of the measurement error.
-As an extreme example of this, we consider the case where the POMDP solution assumes the largest measurement error level considered here, $\sigma_m = 0.15$, while performing simulations in which measurements occur without error.
-In such a scenario, the
+As an extreme example of this, we consider the case where the POMDP solution assumes the largest measurement error level considered here, $\sigma_m = 0.15$, while performing simulations in which measurements occur without error:
 
@@
             reps = reps)$df %>% 
-              mutate(method = "pomdp") # include a column labeling method
+              mutate(method = "POMDP") # include a column labeling method
            },
@@
 
-Combine the resulting data frames
+Combine the POMDP simulations results from over-estimating measurement error with the previous results from the static policies for comparison:
 
@@
 
-overest %>% filter(method != "pomdp", sigma_m == "0")  %>%
+overest %>% filter(method != "POMDP", sigma_m == "0")  %>%
   bind_rows(
-  overest %>% filter(method == "pomdp", sigma_m == "0.15")) %>%
+  overest %>% filter(method == "POMDP", sigma_m == "0.15")) %>%
   select(-sigma_m, -sigma_g) %>%
-  mutate(method = fct_recode(method,
-                             "MSY" = "msy",
-                             "CE" = "det",
-                             "TAC" = "pgy",
-                             "POMDP" = "pomdp")) %>%
+
   write_csv(file.path(log_dir, "overest_sims.csv"))
@@
 
-## Policy plots
+--------
+
+# Policy plots
 
@@
 
-Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (deterministic) recruitment would land us right at it $B_{MSY}$.
+Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (CEerministic) recruitment would land us right at it $B_{MSY}$.
 This corresponds to constant escapement.
@@
 $K/2 = 0.5$).
-PGY harvests are always smaller than MSY harvests, but unlike the deterministic optimal solution, PGY and MSY solutions never go to zero.
+TAC harvests are always smaller than MSY harvests, but unlike the deterministic optimal solution, TAC and MSY solutions never go to zero.
 
-```{r}
-policy_table %>% 
+```{r fig.width = 5, fig.height=7}
+h_plot <- policy_table %>% 
   ggplot(aes(state, harvest, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+  geom_line(lwd=1)  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))  +
+  theme(legend.position = "none")
+
+e_plot <- policy_table   %>% 
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line(lwd=1) + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8)) +
+  theme(legend.position = "bottom")
+
+gridExtra::grid.arrange(h_plot, e_plot, layout_matrix = rbind(1,2))
 ```
 
-Note that when recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.
+When recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.
 The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles when we plot in terms of escapement instead of harvest (as is common in the optimal control literature in fisheries, e.g.
@@
 
-Note that under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population
-is harvested back down to that size.
+Under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population is harvested back down to that size.
 Note that even stocks observed at states slightly below $B_{MSY} = 0.5$ achieve this target escapement, since we are following the classic Graham Shaeffer formulation here where we observe first, then recruitment happens before harvest, and thus we see the population at smaller size than we harvest it.
@@
 
-```{r}
-policy_table   %>% 
-  ggplot(aes(state, escapement, col=policy)) + 
-  geom_line() + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-
-```
-
-### Policies under uncertainty
+### Policies under uncertainty: S = D
 
@@
 If want to compare these to an MDP or POMDP policy, we must specify the level of uncertainty.
-
 In the absence of measurement uncertainty this is straight forward.
@@
 
-First we grab the transition matrix we have already defined for small `sigma_g`:
+First we grab the transition matrix we have already defined when `sigma_g = 0.1`, and then solve the Stochastic Dynamic Programming solution (SDP) for the (fully observed) Markov Decision Process (MDP) using the `MDPtoolbox` package:
 
 ```{r}
-i <- meta %>% filter(sigma_g == 0.02, sigma_m ==0) %>% pull(scenario) %>% as.integer()
-m <- models[[i]]
-```
-
-With no observation uncertainty, we can solve numerically for the optimal policy with stochastic dynamic programming
-
-```{r}
-mdp <- MDPtoolbox::mdp_policy_iteration(m$transition, m$reward, discount) 
-```
-
-Adding this to the plot we see the result is identical to the deterministic case:
-
-```{r}
-bind_rows(policy_table,
-  data.frame(state = states, 
-             policy = "sdp 0.02",
-             harvest = actions[mdp$policy],
-             stringsAsFactors = FALSE) %>%
-  mutate(escapement = state- harvest)) %>% 
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
-```
-
-Repeating this for larger stochasticity, we get a slightly more conservative result:
-
-```{r}
-i <- meta %>% filter(sigma_g == 0.15, sigma_m ==0) %>% pull(scenario) %>% as.integer()
+i <- meta %>% filter(sigma_g == 0.1, sigma_m ==0) %>%
+  pull(scenario) %>% as.integer()
 m <- models[[i]]
@@
 
+We plot the resulting policy in comparison with the other static policies, which shows that even for such a large noise we get a stochastic constant escapement equal to that of of the deterministic CE calculation:
+
 ```{r}
 bind_rows(policy_table,
-  data.frame(state = states, 
-             policy = "sdp 0.15",
+  data_frame(state = states, 
+             policy = "SDP",
              harvest = actions[mdp$policy],
-             stringsAsFactors = FALSE)) %>% 
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line()  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))
+             escapement = states - actions[mdp$policy])) %>% 
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line(lwd=1)  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8)) + 
+  labs(caption = "SDP solution when sigma_g = 0.1")
+```
+
+Repeating this larger stochastic growth `sigma_g = 0.15`, shows only slight deviation from CE ($S > D$, as @Reed1979 proves for sufficiently large noise to violate the self-sustaining criterion.
+Technically since our noise model in normally distributed none of our populations are self-sustaining for infinite time, but as these simulations show, in practice that deviation is very small.)
+
+```{r}
+i <- meta %>% filter(sigma_g == 0.15, sigma_m ==0) %>%
+  pull(scenario) %>% as.integer()
+m <- models[[i]]
+mdp <- MDPtoolbox::mdp_policy_iteration(m$transition, m$reward, discount)
+
+bind_rows(policy_table,
+  data_frame(state = states, 
+             policy = "SDP",
+             harvest = actions[mdp$policy],
+             escapement = states - actions[mdp$policy])) %>% 
+  ggplot(aes(state, escapement, col=policy)) + 
+  geom_line(lwd=1)  + 
+  coord_cartesian(xlim = c(0,K), ylim = c(0, .8)) + 
+  labs(caption = "SDP solution when sigma_g = 0.15")
 ```
@@
 ```{r}
-i <- meta %>% filter(sigma_g == 0.15, sigma_m ==0.1) %>% pull(scenario) %>% as.integer()
+i <- meta %>% 
+  filter(sigma_g == 0.15, sigma_m == 0.1) %>% 
+  pull(scenario) %>% as.integer()
 m <- models[[i]]
@@
 
-For more realistic set of priors, we will consider priors centered at the target $B_{MSY}$ size (or $S^*$ in the language of Reed), at half $B_{MSY}$, and at 1.5 times $B_{MSY}$, each with a standard deviation of $\sigma_m = 0.1$ (i.e.
+For more realistic set of priors, we will consider priors centered at the target $B_{MSY}$ size (or $S^*$ in the language of Reed), at half $B_{MSY}$ (i.e.
+$K/4$), and at $\tfrac{3}{4} K$, each with a standard deviation of $\sigma_m = 0.1$ (i.e.
 the uncertainty around a single observation of a stock at that size.)
@@
 ```{r}
-i_star <- which.min(abs(states - S_star))
-i_low <- which.min(abs(states - 0.25 * K))
-i_high <- which.min(abs(states - 0.75 * K))
-
-prior_star <- m$observation[,i_star,1]
-prior_low <- m$observation[,i_low,1]
-prior_high <- m$observation[,i_high,1]
+prior_star <- m$observation[, which.min(abs(states - S_star)),1]
+prior_low <- m$observation[,which.min(abs(states - 0.25 * K)),1]
+prior_high <- m$observation[,which.min(abs(states - 0.75 * K)),1]
 
@@
 all_policies <- bind_rows(policy_table, pomdp_policies)
-
 write_csv(all_policies, file.path(log_dir, "all_policies.csv"))
@@
 
-POMDP policies depend on the prior belief.
+Plot these policies in terms of both Harvest and Escapement:
 
-```{r fig.width=5, fig.height=7, fig.cap=""}
+```{r}
+recode_policies <- 
+  all_policies %>%
+  filter(policy %in% c("CE", "MSY", "TAC", "low", "high")) %>%
+  mutate(policy = 
+    fct_recode(policy, 
+               "POMDP: low prior" = "low", 
+               "POMDP: high prior" = "high"))
+```
+
+```{r}
+recode_policies %>%
+  ggplot(aes(state, escapement, col = policy)) + 
+  geom_line(lwd=1) + 
+  coord_cartesian(xlim = c(0,1), ylim=c(0,.8))
+```
+
+```{r}
+recode_policies %>%
+  ggplot(aes(state, harvest, col = policy)) + 
+  geom_line(lwd=1) + 
+  coord_cartesian(xlim = c(0,1), ylim=c(0,.6))
+```
+
+```{r, include=FALSE}
+## Data for manuscript; combine prior distribution data with policy data
 priors <- 
@@
   mutate(prior = prior*10)
+write_csv(priors, file.path(log_dir, "priors.csv"))
+```
 
+```{r}
+p3 <- read_csv(file.path(log_dir, "priors.csv")) %>%
+ggplot(aes(state, prior, col = policy)) + 
+  geom_line(lwd=1) +
+  coord_cartesian(xlim = c(0,1))
+p3
+```
+
+```{r include = FALSE}
 policy_df <- all_policies %>% 
-  filter(policy %in% c("low", "high", "det", "pgy")) %>%
+  filter(policy %in% c("low", "high", "CE", "TAC")) %>%
   mutate(policy = fct_recode(policy,
-                             "CE" = "det",
-                             "TAC" = "pgy",
                              "POMDP: low prior" = "low",
@@
                              )) 
-
 policies_w_priors <- right_join(priors, policy_df, by = c("state", "policy"))
@@
 
-```{r fig.width=5, fig.height=7, fig.cap=""}
+```{r fig.width=5, fig.height=7, include = FALSE}
 policies_w_priors %>% 
@@
   facet_wrap(~panel, ncol=1, scales = "free_y") +
-  scale_color_manual(values = colors) +
   coord_cartesian(xlim = c(0,1), ylim = c(0,.8)) + 

### 2018-01-29 appendix/appendixB.Rmd
appendixB tidy

@@
 
-Computes model parameter estimates for selected stocks in RAM using NIMBLE. estimate $B_{MSY}$
+Computes model parameter estimates for selected stocks in RAM using NIMBLE.
 
 ```{r knitr, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE, cache=TRUE, dev="cairo_pdf")
+knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf")
 ```
@@
 ## This chunk is optional, just tweaks plot appearance / fonts
+library(printr)
 library(ggthemes)
@@
 if(!file.exists("ramlegacy.zip")){
-download.file("https://depts.washington.edu/ramlegac/wordpress/databaseVersions/RLSADB_v3.0_(assessment_data_only)_excel.zip",
+download.file(paste0(
+  "https://depts.washington.edu/ramlegac/wordpress/databaseVersions/",
+  "RLSADB_v3.0_(assessment_data_only)_excel.zip"),
   "ramlegacy.zip")
@@
 ## Model does not estimate sigma_m; data is insufficient to do so.
-## Note that RAM stock estimates are not always direct measurements, can be artifically smooth, underestimating sigma_g
-
 gs_code  <- nimble::nimbleCode({
-
   r ~ dunif(0, 2)
@@
   }
-
 })
@@
   samples <- samples[-burnin,1:(length(inits) - 1)] # drop raised vars, burnin
-
   #gather(samples) %>% ggplot() + geom_density(aes(value)) + facet_wrap(~key, scale='free')
@@
 
-<!--
-Do the estimates make sense?  Consider simulation from models, under this historical harvests, compared to historical trajectories
-
-```{r}
+```{r include = FALSE, eval = FALSE}
+## Do the estimates make sense?  Consider simulation from models, 
+## under this historical harvests, compared to historical trajectories
 examples %>% ungroup() %>%
@@
 ```
-
--->
 
@@
 
-# Decision Policies
+# Calculations of the Decision Policies for Historical Data
 
@@
 
-## All parameter values combinations for which we want solutions
+Consider all parameter values combinations for which we want solutions (both species at each of three possible levels of measurement uncertainty; though we will focus on the `0.1` level for simplicity as overall pattern is the same at `0.15`):
 
@@
 
-## Create the models
+Create the model matrices (transition, observation, and reward matrix):
 
@@
 ```
+
+Here's the slowest part: computing POMDP alpha vectors.
 
@@
   left_join(
-    ## use the estimated sigma_m, r, K pars recorded in the logged meta.csv.  Technically
-    ## these should be the same as the above, but without random seed re-running the nimble
-    ## code but not re-running these can create mis-matches.  
     select(meta, sigma_m, commonname, scenario),   
@@
 
-  tibble(S_star, F_MSY = MSY / B_MSY, F_PGY = 0.8 * F_MSY, 
+  tibble(S_star, F_MSY = MSY / B_MSY, F_TAC = 0.8 * F_MSY, 
          commonname = P$commonname, N = P$N)
@@
 ```
+
 Convert example data into discrete index space.
@@
          catch = index(scaled_catch, actions))  %>% 
-  left_join(policy_pars) %>% left_join(pars) %>%
+  left_join(policy_pars) %>% 
+  left_join(pars) %>%
   ungroup() 
-ex
 ```
 
+Static policy calculations:
+
 ```{r} 
-det_f <- function(S_star, r, K, i) index(pmax(gs(r[[1]],K[[1]])(states,0) - S_star[[1]],0), actions)[i]
-msy_f <- function(F_MSY, i) index(states * F_MSY[[1]], actions)[i]
-pgy_f <- function(F_PGY, i) index(states * F_PGY[[1]], actions)[i]
-
+CE_f <- function(S_star, r, K, i) 
+  index(pmax(gs(r[[1]],K[[1]])(states,0) - S_star[[1]],0), actions)[i]
+MSY_f <- function(F_MSY, i) index(states * F_MSY[[1]], actions)[i]
+TAC_f <- function(F_TAC, i) index(states * F_TAC[[1]], actions)[i]
 rescale <- function(x, N) states[x]*N
@@
   group_by(commonname) %>% 
-  mutate(det =  det_f(S_star, r, K, biomass),
-         msy =  msy_f(F_MSY, biomass),
-         pgy =  pgy_f(F_PGY, biomass)) %>%
-  select(year, biomass, catch, det, msy, pgy, commonname, N) %>%
+  mutate(CE =  CE_f(S_star, r, K, biomass),
+         MSY =  MSY_f(F_MSY, biomass),
+         TAC =  TAC_f(F_TAC, biomass)) %>%
+  select(year, biomass, catch, CE, MSY, TAC, commonname, N) %>%
   gather(model, stock, -year, -commonname, -N) %>%
@@
 
-## Plots
-
-```{r}
-historical %>%
-  ggplot(aes(year, stock, col=model)) +
-  geom_line(lwd=1) +
-  scale_color_manual(values = c("black", "grey", "#D9661F", "#00B0DA", "#CFDD45")) +
-  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
-```
-
-## POMDP historical
+Compute POMDP policy for historical data:
 
@@
 
+Join records:
+
 ```{r}
@@
   mutate(model = recode(sigma_m, 
-                        "0" = "Reed",
+                        "0" = "CE",
                         "0.1" = "POMDP", 
@@
 
+Final plot, as in paper but including MSY:
+
 ```{r}
 ram_ex <- read_csv(file.path(log_dir, "ram_ex.csv"))
-
 ram_ex %>%
-  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy", "msy")) %>%
-   mutate(model = fct_recode(model,
-                             "MSY" = "msy",
-                             "CE" = "Reed",
-                             "TAC" = "pgy"
-                             )) %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "CE", "TAC", "MSY")) %>%
   ggplot(aes(year, stock, col=model)) +

### 2018-01-29 manuscripts/manuscript.Rmd
update sims

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7,  dev="cairo_pdf")
 library(tidyverse)
@@
 library(hrbrthemes)
-
 extrafont::loadfonts()
-knitr::opts_chunk$set( dev="cairo_pdf")
+knitr::opts_chunk$set()
 hrbrthemes::import_roboto_condensed() 
 ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
-```
-
-```{r plot_colors}
-#palette <- ggthemes::ptol_pal()(6)
-palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors
-
+palette <- c("#D9661F", "#3B7EA1",  "#6C3302", 
+             "#FDB515", "#00B0DA",  "#CFDD45")
 colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
@@
                       "catch"))
-
+scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
+scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
 ```
@@
 How have invasive species transformed native communities?
-A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.
-In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.
-In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.
+A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods.
+In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and under-served by available tools.
+In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these existing approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.
 This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.
@@
 Approaches for decision making under uncertainty in ecological systems can be divided into two camps: approaches based in "optimal control" and usually favored by natural resource economists, and approaches based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].
-While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknolwedged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.
-Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigour of optimization to resolve a long-standing paradox and suggest a more robust approach to management.
+While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.
+Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.
 
@@
 
-A few subtlies arise in this seemlingly simple problem statement that we must address.
+A few subtleties arise in this seemingly simple problem statement that we must address.
 First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?
@@
 
-To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),
+To facilitate tractability and interpretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),
 
@@
 
-This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis which does not solve the time-dependent optimization problem above.
+This *constant yield* (constant mortality) solution thus corresponds to an equilibrium analysis which does not solve the time-dependent optimization problem above.
 In particular, this maximum sustainable yield (MSY) strategy will not be optimal whenever the stock is away from $B_{MSY}$.
-@Clark1973 deomonstrated (assuming no uncertainty in measurement or stochasticity in population dynamics) that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*.
+@Clark1973 demonstrated (assuming no uncertainty in measurement or stochasticity in population dynamics) that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*.
 We will return to this approach in a moment.
@@
 Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.
-Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Overfishing Limit, (OFL).
+Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-fishing Limit, (OFL).
 Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment.
@@
 Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.
-(This intution must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
+(This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
 This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.
@@
 Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above.
-Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarise here.
+Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarize here.
 
@@
 The first of these we shall refer to as Reed's Paradox for stochastic growth, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.
-This suprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.
+This surprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.
 This also provides the manager with a remarkably convenient mechanism for determining the optimal harvest policy: rather than rely on computationally intensive Stochastic Dynamic Programming to determine the optimal policy, a simple derivative will suffice, which is all that is needed to solve the maximum of the deterministic problem.
@@
 
-Unforunately for @Clark1986, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:
+Unfortunately for @Clark1986, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:
 
@@
 
-Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than rejct it.
+Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than reject it.
 Others have also attempted to introduce measurement error, but always relying on various approximations that similarly alter the problem.
@@
 
-- Utility function: $U(x_t,a_t)$, the utility recieved at time $t$ for taking action $a_t$, given that the system is in state $x_t$.
+- Utility function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.
 For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.
@@
 @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
-POMDP algorithms remain a very active area in artificial intellegence [.
-@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+POMDP algorithms remain a very active area in artificial intelligence [.
+@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
 Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
@@
   select(-value) %>%
-  filter(method != "msy", sigma_g == "0.15") %>%
-  mutate(method = fct_recode(method, 
-                             "CE" = "det",
-                             "TAC" = "pgy",
-                             "POMDP" = "pomdp")) %>%
+  filter(method != "MSY", sigma_g == "0.15") %>%
   group_by(time, method, sigma_m) %>%
@@
   facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
-  scale_color_manual(values = colors) +
-  scale_fill_manual(values = colors) +
   coord_cartesian(ylim = c(0, 1)) + 
@@
 The first panel of Figure \ref{sims} shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.
-Without mesasurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
+Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
 As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.
@@
 In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.
-Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC stategy rather than the CE strategy that lead to the highest overfishing rate.
+Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC strategy rather than the CE strategy that lead to the highest over-fishing rate.
 In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.
@@
   select(time, value, rep, method, sigma_m, sigma_g) %>%
-  filter(sigma_g == "0.15", method != "msy") %>%
+  filter(sigma_g == "0.15", method != "MSY") %>%
   group_by(rep, method, sigma_m, sigma_g) %>%
@@
   ungroup() %>%
-  mutate(method = fct_recode(method,
-                             "CE" = "det",
-                             "TAC" = "pgy",
-                             "POMDP" = "pomdp")) %>%
 
@@
 
-This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also bourne out in terms of economic performance.
+This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also borne out in terms of economic performance.
 Figure \ref{econ} shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.
 As before, environmental stochasticity is set at $\sigma_g = 0.15$.
-In the absense of measurement uncertainty, constant escapement (CE) is optimal (as per @Reed1979), while the heuristic caution built into the total allowable catch (TAC) strategy (at 80% MSY) results in a sub-optimal economic yield.
-The presense of stochasticity in recruitment ($\sigma_g = 0.15$), also contributes to the reduction in yield under TAC.
+In the absence of measurement uncertainty, constant escapement (CE) is optimal (as per @Reed1979), while the heuristic caution built into the total allowable catch (TAC) strategy (at 80% MSY) results in a sub-optimal economic yield.
+The presence of stochasticity in recruitment ($\sigma_g = 0.15$), also contributes to the reduction in yield under TAC.
 Though CE is quite robust to the environmental stochasticity level, this strategy proves very sensitive to increasing measurement error, showing sharp declines in economic value, falling below the TAC economic value at $\sigma_m = 0.15$.
@@
 Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.
-In contrast to the declining economic performance of CE and the consistently sub-optimal economic yield for TAC, the POMDP strategy continues to generate economic yields at approximately the optimal level (as attained by CE in the absense of measurement error) despite the increasingly uncertain measurements.
+In contrast to the declining economic performance of CE and the consistently sub-optimal economic yield for TAC, the POMDP strategy continues to generate economic yields at approximately the optimal level (as attained by CE in the absence of measurement error) despite the increasingly uncertain measurements.
 This demonstrates that the reduced risk of stock collapse and higher average stock biomass attained by the POMDP strategy in the presence of high measurement error, Figure \ref{sims}) is not the result of a trivial reduction in harvesting under all circumstances, but rather, evidence of a more nuanced strategy that manages to account for the uncertainty in measurement while maintaining a reasonable harvest.
@@
   coord_cartesian(ylim = c(0, 1))  + 
-  theme(legend.position = "bottom") +
-  scale_color_manual(values = colors) +
-  scale_fill_manual(values = colors) 
+  theme(legend.position = "bottom") 
 ```
@@
 Figure \ref{overest} summarizes the results of the same simulations as before, but under the scenario in which the POMDP approach assumes a measurement error of $\sigma_m = 0.15$, when in fact all simulated measurements are made without error ($\sigma_m = 0$).
-This represents an extreme case of ovestimating the measurement error.
+This represents an extreme case of overestimating the measurement error.
 Stochastic growth remains the same as before, $\sigma_g = 0.15$, and simulations under TAC and CE policies are shown for comparison.
@@
 Figure \ref{overest} shows that the POMDP outcomes track almost exactly the CE outcomes despite the misplaced assumption that measurements are quite poor.
-(As we have already seen, the TAC stategy is insufficiently cautious for this level of stochasticicty in recruitment, resulting in overexploitation and long-term decline).
-The ability of the POMDP solution to perform nearly optimally even when significantly overestimating the level of measurement uncertainty contrasts sharply to the significant declines from ignoring measurement uncertatinty seen in the CE solutions in Figure \ref{sims}.
+(As we have already seen, the TAC strategy is insufficiently cautious for this level of stochasticity in recruitment, resulting in over-exploitation and long-term decline).
+The ability of the POMDP solution to perform nearly optimally even when significantly overestimating the level of measurement uncertainty contrasts sharply to the significant declines from ignoring measurement uncertainty seen in the CE solutions in Figure \ref{sims}.
 This demonstrates that while we may not know precisely the level of measurement error, we achieve far better outcomes overestimating measurement uncertainty than underestimating it.
@@
   ylab("policy") +
-  scale_color_manual(values = colors) +
 #  theme(legend.position = "bottom") +
@@
 Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
-While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach.
+While the catch quota under both TAC and CE strategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach.
 This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.
@@
 When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.
-POMDP formalizes this intutition by capturing the information from all previous observations into a *prior belief*.
+POMDP formalizes this intuition by capturing the information from all previous observations into a *prior belief*.
 This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.
@@
 Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where chance local extinction becomes possible.
-It is precisely this assymetry: that too much overharvesting leads to an irreversible state of extinction, while too much underharvesting is always reversable (modulo some lost revenue) that lay behind Clark's original intutition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.
-Constant escapement is particularly suscpetible to overharvesting starting from stock sizes much higher than $B_{MSY}$ since its bang-bang optimal solution attempts to bring stocks back to target level as fast as possible.
+It is precisely this asymmetry: that too much over-harvesting leads to an irreversible state of extinction, while too much under-harvesting is always reversible (modulo some lost revenue) that lay behind Clark's original intuition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.
+Constant escapement is particularly susceptible to over-harvesting starting from stock sizes much higher than $B_{MSY}$ since its bang-bang optimal solution attempts to bring stocks back to target level as fast as possible.
 CE is not so vulnerable to collapse from small stock sizes, since it shuts down all harvests once estimates fall below $B_{MSY}$.
@@
 ram_ex %>%
-  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
-   mutate(model = fct_recode(model,
-                             "CE" = "Reed",
-                             "TAC" = "pgy"
-                             )) %>%
+  filter(model %in% c("biomass", "catch", "POMDP", "CE", "TAC")) %>%
   ggplot(aes(year, stock, col=model)) +
   geom_line(lwd=1) +
-  scale_color_manual(values = colors) +
   facet_wrap(~commonname, scales = "free", ncol=1)  
@@
 So far we have focused on simulation and an examination of the relative policies under different prior information.
-Figure \ref{historical} compares the harvest level that would have been recommmended by each of the strategies we have considered here against the historically observed catch recorded for two commericially fished stocks: Argentine Hake and European Plaice.
+Figure \ref{historical} compares the harvest level that would have been recommended by each of the strategies we have considered here against the historically observed catch recorded for two commercially fished stocks: Argentine Hake and European Plaice.
 Historical estimates of biomass and catch are taken from the R.A.
@@
 In both stocks, historical catch almost always exceeds that recommended by any of our strategies, falling closer to the MSY value estimated (see appendix).
-In Argentine Hake, the persistent declines result in both POMDP and CE strategies quicly closing down the fishery in an attempt to let the stock recover to a more productive target biomass, while TAC persists with merely a reduced harvest.
+In Argentine Hake, the persistent declines result in both POMDP and CE strategies quickly closing down the fishery in an attempt to let the stock recover to a more productive target biomass, while TAC persists with merely a reduced harvest.
 A small recovery of biomass in 1995 is met by an immediate uptick in harvest under CE, while the POMDP response is more conservative.
@@
 Using modern algorithms we have been able to crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk long-standing paradoxes that increased measurement uncertainty should be met with larger harvests.
-We have demonstrated that both existing optimal control approaches and existing precautionary approaches result in overexploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.
+We have demonstrated that both existing optimal control approaches and existing precautionary approaches result in over-exploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.
 These results have important implications for both global fisheries and ecological management more generally.
@@
 Our results suggest that measurement uncertainty may contribute to this pattern, and that reality of measurement errors may frustrate stock recovery under current best practices such as MSY, TAC, and CE.
-More generally, the study and application of strategies such as MSY and CE are in no way limited to the decisions in fisheries quotas, but underpin much of conservation and natural resource management in forests and othe terrestrial ecosystems as well [e.g.
+More generally, the study and application of strategies such as MSY and CE are in no way limited to the decisions in fisheries quotas, but underpin much of conservation and natural resource management in forests and other terrestrial ecosystems as well [e.g.
 @Clark2000].
@@
 
-First, it may seem obvious that POMDP would out-proform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.
-Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accomidate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.
+First, it may seem obvious that POMDP would out-preform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.
+Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accommodate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.
 If stochasticity doesn't matter, it is easy to believe that measurement error would be minor at best.
@@
 ^1: Given the sophisticated policies often in place to determine actual TAC levels, one might object to our characterization of this as merely a rule-of-thumb.
-We use this term merely to distinguish it from the optimal control approach -- that despite technical precision, TAC adjustments are set by a hueristic instead of the result of some dynamic optimization algorithm as in CE and POMDP.
+We use this term merely to distinguish it from the optimal control approach -- that despite technical precision, TAC adjustments are set by a heuristic instead of the result of some dynamic optimization algorithm as in CE and POMDP.
 
 Second, we must ask if the policies derived from POMDP are practical.
-As with constant escapement, the POMDP solutions we considered here result in more year-to-year variation in harvest, and include the possibilty of fishing moritorium to allow stocks to recover -- decisions that many view as economically or socially unacceptable.
-While resolving this debate is beyond the scope of this paper, a few observations are relevant: (A) Stocks already managaed under constant escapement, such as US salmon fisheries, are already subject to the possibility of these moritoriums.
+As with constant escapement, the POMDP solutions we considered here result in more year-to-year variation in harvest, and include the possibility of fishing moratorium to allow stocks to recover -- decisions that many view as economically or socially unacceptable.
+While resolving this debate is beyond the scope of this paper, a few observations are relevant: (A) Stocks already managed under constant escapement, such as US salmon fisheries, are already subject to the possibility of these moratoriums.
 (B) It is straight-forward to introduce constraints on the minimal acceptable harvest level and/or include the additional costs of closing a fishery in the utility function optimized by POMDP.
@@
 The relatively poor ecological and economic performance of these common best-practices relative to the alternative approach of POMDP is thus reason enough for concern.
-This issue is further underscored by the reality of both long-running ecological detioriation [@Worm2006; @Worm2009] and economic costs of global fisheries.
+This issue is further underscored by the reality of both long-running ecological deterioration [@Worm2006; @Worm2009] and economic costs of global fisheries.
 

### 2018-01-29 manuscripts/manuscript.Rmd
draft

@@
 bibliography: citations.bib
+#journal: "The American Naturalist"
+date: "`r Sys.Date()`"
 csl: the-american-naturalist.csl
@@
 layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column
+header-includes: |
+        \journal{The American Naturalist}
 
@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7,  dev="cairo_pdf")
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7,  dev="cairo_pdf", cache = TRUE)
 library(tidyverse)

### 2018-01-30 appendix/appendixA.Rmd
tweaks to appendix A

--- 2018-01-29 appendix/appendix_harvestfirst.Rmd
@@
 date: "`r Sys.Date()`"
-output: pdf_document
+output: 
+  pdf_document:
+    toc: true
+    toc_depth: 2
 bibliography: ../manuscripts/citations.bib
@@
 
-# Overview
-
-1. Model definition:
-2. Deterministic Solutions
-3. Discretization and the POMDP solution
-4. Simulation
-
 ```{r knitr, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf")
+knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf", echo=TRUE)
 ```
 
+The POMDP calculations presented here are performed using the `sarsop` R package, which we have written to provide a general purpose implementation of the SARSOP algorithm of @Kurniawati2008.
+The version used in this analysis can be installed from GitHub using the command: `devtools::install_github("boettiger-lab/sarsop@v0.5.0")`.
+
 ```{r library, message=FALSE}
-## install package first if necessary.
-# devtools::install_github("boettiger-lab/sarsop") 
-library(sarsop)       # the main POMDP package
+devtools::install_github("boettiger-lab/sarsop@v0.5.0")
+library(sarsop)       # Our main POMDP package
 library(tidyverse)    # for munging and plotting
@@
 
-# Basic deterministic model
+# Definitions: Population Model and Utility function
 
@@
 
-Utility (reward) function.  (Note that setting a smaller discount will take longer to converge to smooth POMDP solution, but result in `S_star` closer to the simple `B_MSY`.)
+A simply utility (reward) function: a fixed price per fish with no cost to fishing.
+Alternate models can easily be considered, we focus on this simple utility throughout as this is the choice typically focused on in the optimal control literature, implicitly assumed by the basic MSY model, and does not change the general results.
+Note that setting a smaller discount will take longer to converge to smooth POMDP solution, but result in `S_star` closer to the simple `B_MSY`.
 
@@
 
-# Calculating MSY & TAC
+# Calculating MSY, TAC, and CE
 
-This uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.
+The approach below uses a generic optimization routine to find the stock size at which the maximum growth rate is achieved.
 (Note that this also depends on the discount rate since future profits are worth proportionally less than current profits).
@@
                    c(0, 2*K))$minimum
-S_star
 ```
@@
 
-#  Introduce a discrete grid
+# Solving the POMDP problem
+
+##  Introduce a discrete grid
 
@@
 
-# POMDP Model
+## Define the POMDP matrices
 
 We compute POMDP matrices for a range of `sigma_g` and `sigma_m` values.
-(We consider `sigma_g = 0.02` rather than precisely zero here to avoid a degeneracy in the discretized tranisition matrix.
-Calculations for the precise deterministic solution can be solved without discretization as shown above, and match the results of this nearly-zero nosie level in the discrete simulation, as expected.)
+(We consider `sigma_g = 0.02` rather than precisely zero here to avoid a degeneracy in the discretized transition matrix.
+Calculations for the precise deterministic solution can be solved without discretization as shown above, and match the results of this nearly-zero noise level in the discrete simulation, as expected.)
 
@@
 
-## POMDP solution
+## Compute the POMDP solution
 
@@
 We assume a uniform prior belief at the start of the POMDP simulation.
-These tasks are perfromed by the `sim_pomdp` fucntion.
-Simulating the static strategies (CE, MSY, TAC) is easier, since the policy is uniquely determined by the observed state; we simply need to simulate the stochastic growth and process and measurement with error at each time step an then compute the corrsponding policy.
+These tasks are performed by the `sim_pomdp` function.
+Simulating the static strategies (CE, MSY, TAC) is easier, since the policy is uniquely determined by the observed state; we simply need to simulate the stochastic growth and process and measurement with error at each time step an then compute the corresponding policy.
 The `sim_pomdp` function performs such static simulations when given a pre-specified policy instead of a set of `alpha` vectors.
@@
 
-##  Figure S2
+##  Figure S2: Full simulation plots
 
@@
 true <- 3 # sigma_g = 0.15, sigma_m = 0
-source("../appendix/pomdp_overestimate.R")
+source(system.file("examples/pomdp_overestimate.R", package="sarsop")) # portable example code
 pomdp_overest_sims <- 
@@
 ```{r}
-
 overest <-
@@
   select(-sigma_m, -sigma_g) %>%
-
   write_csv(file.path(log_dir, "overest_sims.csv"))
@@
   coord_cartesian(ylim = c(0, 1))
-
 ```
@@
 # Policy plots
+
+## Static policies
+
+First we gather the static policies (CE, TAC, MSY) into a table and convert into terms of harvest and escapement.
 
@@
   mutate(escapement = state - harvest)
-
 ```
 
-Note that when recruitment occurs before harvest, $f(x) - h$, to get to $X_t = B_{MSY}$ as fast as possible, we actually want to harvest a little bit when X is below $B_{MSY}$, so that rather than over-shooting $B_{MSY}$ (CEerministic) recruitment would land us right at it $B_{MSY}$.
-This corresponds to constant escapement.
-The discrete grid makes these appear slightly stepped.
+TAC harvests are always smaller than MSY harvests, but unlike the deterministic optimal solution, TAC and MSY solutions never go to zero.  The discrete grid makes these appear slightly stepped.
 
-Note that the deterministic solution crosses the MSY solution at an observed value of $B_{MSY}$ (i.e.
-$K/2 = 0.5$).
-TAC harvests are always smaller than MSY harvests, but unlike the deterministic optimal solution, TAC and MSY solutions never go to zero.
+## Policies under uncertainty: S = D
 
-```{r fig.width = 5, fig.height=7}
-h_plot <- policy_table %>% 
-  ggplot(aes(state, harvest, col=policy)) + 
-  geom_line(lwd=1)  + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8))  +
-  theme(legend.position = "none")
-
-e_plot <- policy_table   %>% 
-  ggplot(aes(state, escapement, col=policy)) + 
-  geom_line(lwd=1) + 
-  coord_cartesian(xlim = c(0,K), ylim = c(0, .8)) +
-  theme(legend.position = "bottom")
-
-gridExtra::grid.arrange(h_plot, e_plot, layout_matrix = rbind(1,2))
-```
-
-When recruitment happens before harvest, escapement is $f(x_t) - h_t$, not $x_t - h_t$.
-The effect of a continuous function map $f$ on a discrete grid is also visible as slight wiggles when we plot in terms of escapement instead of harvest (as is common in the optimal control literature in fisheries, e.g.
-@Sethi2005).
-
-Under the optimal solution, escapement is effectively constant at $B_{MSY} = 0.5$: for all states above a certain size the population is harvested back down to that size.
-Note that even stocks observed at states slightly below $B_{MSY} = 0.5$ achieve this target escapement, since we are following the classic Graham Shaeffer formulation here where we observe first, then recruitment happens before harvest, and thus we see the population at smaller size than we harvest it.
-In classical escapement analysis, observations are usually indexed instead to occur immediately before harvests, and this inflection point occurs right at $B_{MSY}$.
-
-### Policies under uncertainty: S = D
-
-In strategies whose policies are shown in the above plots all ignore both stochasticity and measurement error.
-If want to compare these to an MDP or POMDP policy, we must specify the level of uncertainty.
-In the absence of measurement uncertainty this is straight forward.
+Because @Reed1979 has proven that the optimal escapement in the stochastic case is equivalent to the deterministic case, $S = D$, we have simply relied on the deterministic calculation for constant escapement.
+We can instead calculate the optimally policy for such a stochastic but fully observed Markov Decision Process (MDP) directly given the discritized model transition matrix using Stochastic Dynamic Programming (SDP).
 @Reed1979 essentially tells us that for small growth noise (satisfying or approximately satisfying Reed's self-sustaining condition) that the stochastic optimal policy is equal to the deterministic optimal policy.
@@
 
-First we grab the transition matrix we have already defined when `sigma_g = 0.1`, and then solve the Stochastic Dynamic Programming solution (SDP) for the (fully observed) Markov Decision Process (MDP) using the `MDPtoolbox` package:
+First we grab the transition matrix we have already defined when `sigma_g = 0.1`, and then solve the SDP using the `MDPtoolbox` package:
 
@@
 Repeating this larger stochastic growth `sigma_g = 0.15`, shows only slight deviation from CE ($S > D$, as @Reed1979 proves for sufficiently large noise to violate the self-sustaining criterion.
-Technically since our noise model in normally distributed none of our populations are self-sustaining for infinite time, but as these simulations show, in practice that deviation is very small.)
+Technically since our noise model in normally distributed none of our populations are self-sustaining for infinite time, but that deviation is quite small, as seen here.
+Note also the solution is still a constant-escapement type.)
 
@@
 
-### Comparing POMDP Policies
+## Comparing POMDP Policies
 
@@
 The optimal solution for the partially observed system must also reflect all prior observations of the system, not merely the most recent observation, as the system is not Markovian in the observed state variable.
-We summarize this history as a prior "belief"" about the state, which is updated according to Bayes rule after each observation.
-(Note that @Sethi2005 fails to realize this and plots solutions with measurement uncertainty without reference to the prior, which explains their counter-intuitive finding that increased uncertainty should result in increased harvest rates).
+We summarize this history as a prior "belief" about the state, which is updated according to Bayes rule after each observation.
+Note that @Sethi2005 plots solutions with measurement uncertainty without any reference to the prior.
+Bayes Law tells us that without a statement of prior belief over the true state, $P(x)$ we cannot go from the measurement model that defines the probability of observing $y$ given a true state$x$; $P(y | x)$, into the probability of the true state given the measurement, $P(x | y)$.
+Further, this prior $P(x)$ should be updated with respect to each subsequent observation, we cannot simply assume it is a fixed value as @Sethi2005.
+This explains their counter-intuitive finding that increased uncertainty should result in increased harvest rates with increased uncertainty.
 
-Let us look at the POMDP solutions under various priors focusing on the case of moderate uncertainty, $\sigma_g = \sigma_m = 0.1$.
+Let us look at the POMDP solutions under various priors focusing on the case of moderate uncertainty, $\sigma_g = 0.15$ and $\sigma_m = 0.1$.
 (Recall we have already solved the POMDP solution for this model in the simulations above, as defined by the `alpha` vectors, so we can quickly load that solution now.)
@@
 
-For more realistic set of priors, we will consider priors centered at the target $B_{MSY}$ size (or $S^*$ in the language of Reed), at half $B_{MSY}$ (i.e.
+For a representative set of priors, we will consider priors centered at the target $B_{MSY}$ size (or $S^*$ in the language of Reed), at half $B_{MSY}$ (i.e.
 $K/4$), and at $\tfrac{3}{4} K$, each with a standard deviation of $\sigma_m = 0.1$ (i.e.
@@
 ```{r}
-recode_policies %>%
+p1 <- recode_policies %>%
   ggplot(aes(state, escapement, col = policy)) + 
@@
   coord_cartesian(xlim = c(0,1), ylim=c(0,.8))
+
+p1
 ```
@@
 ```{r}
-recode_policies %>%
+p2 <- recode_policies %>%
   ggplot(aes(state, harvest, col = policy)) + 
@@
   coord_cartesian(xlim = c(0,1), ylim=c(0,.6))
+
+p2
 ```
@@
 
+## Evolution of the POMDP prior belief
+
+To get a sense of how the prior belief (and consequently the POMDP policy) is continually adapting over the course of a simulation (unlike the static policies), it is instructive to examine a single simulation rather than the ensemble of replicates considered above.
+
+```{r}
+set.seed(12345)
+Tmax <- 20
+sim <- sim_pomdp(m$transition, m$observation, m$reward, discount, 
+                 unif_prior, x0 = 10, Tmax = Tmax, alpha = alpha)
+```
+
+```{r}
+sim$df %>%
+  select(-value) %>%
+  mutate(state = states[state], action = actions[action], obs = observations[obs]) %>%
+  gather(variable, stock, -time) %>%
+  ggplot(aes(time, stock, color = variable)) + 
+  geom_line(lwd=1)  + scale_color_manual(values = unname(colors))
+```
+
+Here we plot the evolution of the prior belief in state over the course of the above simulation in increasingly darker curves.
+Note that the the beginning belief is uniform, but quickly tightens into a distribution centered around the region of the most recent observation.
+Note that this belief distribution never converges: as the true state constantly changes in due to stochasticity in population recruitment, so too does the belief distribution constantly change in response to new observations of that true state.
+
+```{r}
+sim$state_posterior %>% 
+  data.frame(time = 1:Tmax) %>%
+  filter(time %in% seq(1,Tmax, by = 3)) %>%
+  gather(state, probability, -time, factor_key =TRUE) %>% 
+  mutate(state = states[as.numeric(state)]) %>% 
+  ggplot(aes(state, probability, group = time, alpha = time)) + 
+  geom_line(lwd = 1) + 
+  coord_cartesian(xlim = c(0,1.5))
+
+```
+
 ```{r include = FALSE}
@@
 policies_w_priors <- right_join(priors, policy_df, by = c("state", "policy"))
-
 write_csv(policies_w_priors, file.path(log_dir, "policies_w_priors.csv"))
-
 ```
@@
 ```
+
+# References

### 2018-01-30 manuscripts/manuscript.Rmd
Revisions from milad

@@
 title: "The importance of measurement uncertainty in ecological management"
+
 author:
-  - name: "Milad Memarzadeh"
-    affiliation: a
-  - name: "Carl Boettiger"
-    affiliation: a, 1
-address:
-  - code: a
-    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
+  - name: "Authors removed during peer review"
+
+#  - name: "Milad Memarzadeh"
+#    affiliation: a
+#  - name: "Carl Boettiger"
+#    affiliation: a, 1
+#address:
+#  - code: a
+#    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
 
 abstract: |
+
   We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.
@@
         \journal{The American Naturalist}
+        \usepackage{lineno}
+        \linenumbers
+        \modulolinenumbers[3]
+        \usepackage{setspace}
+        \doublespacing
 
@@
 knitr::opts_chunk$set()
-hrbrthemes::import_roboto_condensed() 
-ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+#hrbrthemes::import_roboto_condensed() 
+#ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+ggplot2::theme_set(hrbrthemes::theme_ipsum())
 palette <- c("#D9661F", "#3B7EA1",  "#6C3302", 
@@
 For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax.
-Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,
+Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size subject to some measurement uncertainty $\xi_t$
 
@@
 
-Given an estimated recruitment model along with this measurement of the stock size, the manager must set a harvest quota $h_t$ for that year in order to maximize their expected long term (net) present value:
+For simplicity we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.
+Given an estimated recruitment model along with this measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term (net) present value:
 
-$$\max_{ \lbrace h_t \rbrace} \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$
+$$\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$
 
@@
 
-A second subtly is in the nature of the decision problem itself: the optimization problem must always consider the set ${h_t}$ all future actions simultaneously, not merely the next action.
-This is a game of chess, where it can be optimal to make costly short term actions (such as closing the fishery) if they pay off in the long run.
-This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a single harvest quota $H_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.
+A second subtly is in the nature of the decision problem itself: note that the manager in not myopic, thinking only about the next year, but rather considers the sequence of all future actions ${h_t}$.
+While the manager will revisit this sequence each year in light of new observations, it is this ability of the decision problem to think ahead that allows it to tolerate short-term costs (e.g.
+reduced harvests in the current season) for larger future payoffs.
+In this, ecological management is a game of chess, always thinking several moves ahead.
+This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a fixed mortality $F_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.
 Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.
@@
 
-$$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$$
+$$x_{t+1} = f(x_t, h_t) = \varepsilon_t x_t r  \left(1 - \frac{x_t}{K}\right) - h_t $$
 
@@
 Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
-@].
+@Dichmont2016].
 As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.
@@
 To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.
-To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
+To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under the MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
 To distinguish this approach from MSY, we will refer to this approach as a TAC policy.
@@
 
-So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of of paradoxes.
+So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of paradoxes.
 Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible.
@@
 
-## Clark's Paradox: $M < D$
+## Clark's Paradox
 
@@
 - Transition function (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.
-In our context, this relationship is given by the Graham-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $\mathcal{N_T}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.
+In our context, this relationship is given by the Gordon-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.
 
-- Observation function: $O(x_t,y_t,a_t)$ the probability of observing state $y$ given a system in state $x$.
+- Observation function: $O(x_t,y_t,a_{t-1})$ the probability of observing state $y_t$ given a system in state $x_t$.
 In principle, the action chosen can influence the precision of the observation.
-In our case, we simply assume normally distributed errors around the true state, $y \sim \mathcal{N_T}(x, \sigma_m)$, truncated at zero to exclude negative population sizes.
+In our case, we simply assume normally distributed errors around the true state, $y_t \sim \mathcal{N}(x_t, \sigma_m)$, truncated at zero to exclude negative population sizes.
 
 - Utility function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.
-For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.
+For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t, a_t)$, indicating that realized harvest cannot be negative.
 This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation.
@@
 @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
-POMDP algorithms remain a very active area in artificial intelligence [.
-@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+POMDP algorithms remain a very active area in artificial intelligence [@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
 Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
@@
 ```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$.
-Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{{sims}}.
+Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{sims}.
 The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error.
@@
 This difference between underestimating and overestimating accounts for the poor performance of the CE solution under large measurement uncertainty, where it over-harvests whenever measurements are too large.
-Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where chance local extinction becomes possible.
+Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where the chance of local extinction becomes possible.
 It is precisely this asymmetry: that too much over-harvesting leads to an irreversible state of extinction, while too much under-harvesting is always reversible (modulo some lost revenue) that lay behind Clark's original intuition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.
@@
 Posterior distributions for parameters for the Graham-Schaefer model are estimated with uncertainty through Markov Chain Monte Carlo using `nimble` [@nimble] on the historical data, illustrating how this process might be done in more complex models as well (details and code in Supplement, Appendix B).
-$B_{MSY} and corresponding TAC and CE policies are calculated based on posterior mean estimates of the model parameters, along with POMDP solution assuming a 10% measurement error rate.
+$B_{MSY}$ and corresponding TAC and CE policies are calculated based on posterior mean estimates of the model parameters, along with POMDP solution assuming a 10% measurement error rate.
 Though measurement error could be estimated directly from the raw data, this would not reflect the true measurement uncertainty arising from the stock assessment process.
@@
 As we have seen, the TAC and CE harvest policies are uniquely determined by the observed stock size relative to $B_{MSY}$, but the POMDP policy must be re-calculated each time step to reflect both the prior observation and prior action.
-Note that since we cannot replay history under these alternative policies, so the prior action taken is always the historical catch for that year and not the catch recommended by the POMDP policy.
 
@@
 It is similarly tempting to dismiss the better performance of POMDP against a more rule-of-thumb[^1] based strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.
-While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to eek out both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.
+While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to deliver both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.
 These results highlight the difficulty of finding effective heuristic rules and underscore the power and utility of more sophisticated strategies such as POMDP which can make the best use of all available information.
@@
 of states and actions considered here. A value of information (VOI) analysis
-for POMDP,such as @Johnson2015 shows for MDP models, could identify when it
+for POMDP [e.g. @Johnson2015; @Memarzadeh2016b], could identify when it
 is worthwhile to actively reduce measurement error.

### 2018-01-30 appendix/appendixA.Rmd
fix typo

@@
 The POMDP calculations presented here are performed using the `sarsop` R package, which we have written to provide a general purpose implementation of the SARSOP algorithm of @Kurniawati2008.
-The version used in this analysis can be installed from GitHub using the command: `devtools::install_github("boettiger-lab/sarsop@v0.5.0")`.
+The version used in this analysis can be installed from GitHub using the command: `devtools::install_github("boettiger-lab/sarsop@0.5.0")`.
 
 ```{r library, message=FALSE}
-devtools::install_github("boettiger-lab/sarsop@v0.5.0")
+#devtools::install_github("boettiger-lab/sarsop@0.5.0")
 library(sarsop)       # Our main POMDP package
@@
 true <- 3 # sigma_g = 0.15, sigma_m = 0
-source(system.file("examples/pomdp_overestimate.R", package="sarsop")) # portable example code
+source(system.file("examples/pomdp_overestimates.R", package="sarsop")) # portable example code
 pomdp_overest_sims <- 

### 2018-01-30 manuscripts/manuscript.Rmd

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7,  dev="cairo_pdf", cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
 library(tidyverse)
@@
 library(ggplot2)
-library(Cairo)
+#library(Cairo)
 library(extrafont)

### 2018-01-30 appendix/appendix_recruit_first.Rmd
fix log dir

--- 2018-01-04 appendix/appendix.Rmd
@@
 options(mc.cores=parallel::detectCores())
+log_dir <- "../data/observe_recruit_harvest" # Store the computed solution files here
+
 ```
@@
         mutate(scenario  = as.character(1:length(sigma_m)))
-log_dir <- "appendix_alphas" # Store the computed solution files here
 meta
@@
 Alternately, readers can skip the evaluation of this code chunk and read the cached solution from the `policyx` file using the `*_from_log` functions that follow:
-
-```{r}
-options(mc.cores=1)
-```
 

### 2018-01-30 manuscripts/AmNat_response_to_rev.Rmd
WIP

@@
 We appreciate the careful and constructive comments of the editor and reviewers, which have allowed us to re-write the manuscript to both better communicate both the method and broader implications of the results shown here.
-We also now address directly the significant differences between the extensive literature on the optimal management of fisheries to
-
-We appreciate the time and effort of reviewer 1 in exploring these results in further simulation, but were unable to understand precisely what the reviewer had implemented.
-From the description provided, it appears the reviewer's results rather underscore our point about the importance of observation uncertainty, an uncertainty which does not appear to be fully present in the reviewer's simulation, as we discuss below.
-In short, if the stock size is unknown (due to measurement uncertainty), how does one enforce the safe-guard that total allowable catch quota or harvest not exceed 90% the stock?
-As our work shows, it precisely the presence of such assumptions
-
-We have expanded our results to include a comparisons to policies that much more closely resemble the implementation in actual fisheries
-
-and believe our revisions should now address each of the concerns raised.  In particular,
 
@@
 5 of the manuscript,  the optimal escapement level is instead higher in the presence of measurement error.
+Reviewer 1's description of their simulations is brief enough that I would not expect the authors to build their entire revision around an effort to respond to this comment.
+If the authors have indeed allowed for 90 percent harvest, however, it is important that they test whether their results are different if they place an upper limit on the fraction of the stock that they allow to be harvested.
+I also agree with Reviewer 1's comment that the methods are not that easy to follow.
+That said, to my understanding, the escapements in fig.
+2 are not constant, and so Reviewer 1's comment that a constant escapement cannot lead to the collapse of a stock seems to me to be irrelevant, but again I may be wrong, and my uncertainty emphasizes the lack of clarity in the writing, especially with respect to the Methods.
 
-We are afraid that Reviewer 1's concern is mistaken
+The simulations performed by Reviewer 1 are not the same as in our example, but are nevertheless very enlightening as they highlight a very important issue that is far more general than fisheries.
+Reviewer 1's simulations have provided another example of the paradox of measurement error that we now highlight much more clearly in the previous work of @Reed1979, @Clark1986, @Sethi2005, and others: Reviewer 1 makes a "simplifying assumption" and paradoxically concludes that in the face of higher measurement error we should increase fishing pressure!
+A key conclusion from our analysis is the demonstration that this common but paradoxical claim is in fact false.
 
-> Reviewer 1's description of their simulations is brief enough that I would not expect the authors to build their entire revision around an effort to respond to this comment.
-If the authors have indeed allowed for 90 percent harvest, however, it is important that they test whether their results are different if they place an upper limit on the fraction of the stock that they allow to be harvested.
+To assume harvest cannot exceed the standing stock is to assume (for any model without an Allee threshold) that humans cannot collapse a stock regardless of how hard they fish.
+Though this assumption that extinction is impossible is often mathematically convenient, we are hardly the first to argue that it is both biologically unrealistic and can lead to ill-advised management practices
+
+> An important tacit assumption in Reed’s analysis, as in the other works referred
+to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.
+
+The reviewer's approach is not essentially equivalent to our approach; as the editor already observes we do not assume that the optimal policy will involve constant escapement, nor does it.
+Nor is the reviewer's simulation the same as in @Clark1986 or @Sethi2005, since neither of those results involve constant escapement, though they
+
+Reviewer 1 proposes that we assume local extinction is impossible regardless of inaccuracies in measurement error.
+This assumption has been previously made in the literature for mathematical convenience, as we discuss it is in fact
 
@@
 Y = 0 is the only (non-negative) solution; you cannot be sure that *any* harvest level is guatenteed to be safe.
-
-> I also agree with Reviewer 1's comment that the methods are not that easy to follow.
-That said, to my understanding, the escapements in fig.
-2 are not constant, and so Reviewer 1's comment that a constant escapement cannot lead to the collapse of a stock seems to me to be irrelevant, but again I may be wrong, and my uncertainty emphasizes the lack of clarity in the writing, especially with respect to the Methods.
 
@@
 
+We thank the Reviewer for this careful parallel investigation, which was very useful in focusing the manuscript on some of the key results.
+
+While based only on this description provided we are not able to precisely replicate Reviewer 1's analysis, we can observe serveral important ways in which it is quite different to the optimization problem we consider.
+
+We have now completely re-written our description and provided both a stand-alone R package and fully executable scripts implementing our methods.
+However, we do not attempt to provide a stand-alone introduction to Partially Observed Markov Decision Processes (POMDPs), which, all though less well known in ecology and evolutionary studies than their (fully observed) MDP cousins (which ecologists commonly refer to by their solution method, Stochastic Dynamic Programming or SDP)
+
 As the editor observes, this short discription does not permit us to properly understand what the reviewer has done to obtain
@@
 or somewhat larger than the 'target' escapement.  Without perfect measurements, it is clearly impossible
-to achieve the target escapement exactly.  Overestimates of the stock size can lead to stock collapse
+to achieve the target escapement exactly. If, by chance, a string of consecutive measurements repeatedly over-estimates
+the current stock size, a constant-escapement strategy will repeatedly over-harvest the stock.
 
-Without perfect measurements, the true escapement will vary under a
+In a model without an Allee effect, such as the classic Gordon-Schaefer model we focus on here, successive over-harvesting
+can reduce a stock to an
 
@@
 
-The reviewer is entirely correct about this -- while a rigorous constant escapement policy shuts down
+The reviewer is entirely correct about this, and we greatly appreciate the reviewer raising this important issue.
+In addition to the previous analysis which focused only on comparison to a constant escapement policy, we now
+also include comparisons to a more realistic "Total Allowable Catch", or TAC strategy.
+
+-- while a rigorous constant escapement policy shuts down
 all harvest below B_MSY (the constant escapement level), most fisheries actually continue to harvest
@@
 suseptible
-
-Moreover, many fisheries do not
 
@@
 
-We had no intention of suggesting that measurement uncertainty was under-represented in stock assessments; indeed it is precisely the *prevalance* with which stock assessment models account for uncertainty that motivates
+We had no intention of suggesting that measurement uncertainty was under-represented in *stock assessments*; indeed it is precisely the *prevalance* with which stock assessment models account for uncertainty that motivates this work.
+We have extensively revised the manuscript to be much clearer on this point.
+Our assertion is that these estimates of the uncertainty associated with a measurement of the stock size to which the reviewer refers are not properly accounted for in the process of setting a quota.
 
@@
 
-This is an excellent point which we now discuss properly.
-Our approach highlights that strategies such as Pretty Good Yield, taking 80% of the maximum sustainable yield, are not sufficient solutions to the challenge of measurement uncertainty, since they are still based on fundamentally deterministic premises.
-
-As our analysis shows, unless it is possible to be certain beyond any shadow of a doubt that one has
+This is an excellent point which we now develop in the discussion:
 
@@
 But at least in those cases one can reasonably assert an average
-
-Stock assessments give us
-
-Measurement errors accumulate.
-
-Pretty Good Yield, taking 80% of the maximum sustainable yield, reflects the same
-
-is defined as a percentage of maximum sustainable yield, which is an essentially deterministic concept.  It is impossible to be sure that
 
@@
 
-Thank you, fixed. (TBD)
+Agreed.
 
 > The authors discuss volatility of different solutions and would be served well to quantify it. E.g., lines 365-369
-
-Thank you, fixed (TBD)

### 2018-01-30 manuscripts/manuscript.Rmd
WIP

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = TRUE)
 library(tidyverse)
@@
 knitr::opts_chunk$set()
-#hrbrthemes::import_roboto_condensed() 
-#ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
-ggplot2::theme_set(hrbrthemes::theme_ipsum())
+hrbrthemes::import_roboto_condensed() 
+ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
+#ggplot2::theme_set(hrbrthemes::theme_ipsum())
 palette <- c("#D9661F", "#3B7EA1",  "#6C3302", 

### 2018-01-30 appendix/appendixA.Rmd
no author (double-blind)

@@
 title: "Appendix A: Code and supplementary figures for simulation and policy analyis"
-author: "Carl Boettiger"
 date: "`r Sys.Date()`"

### 2018-01-30 appendix/appendix_recruit_first.Rmd
no author (double-blind)

@@
 title: "POMDP comparisons across sigma_m, sigma_g"
-author: "Carl Boettiger"
 date: "`r Sys.Date()`"
@@
 
-write_csv(sims, "appendix_files/sims.csv")
+write_csv(sims, file.path(log_dir, "sims.csv"))
 ```
@@
 
-write_csv(policy_table, "appendix_files/policy_table.csv")
+write_csv(policy_table, file.path(log_dir, "policy_table.csv"))
 ```

### 2018-01-30 appendix/appendix_recruit_first.Rmd
namespace

@@
   select(-value) %>% 
-  rename(uniform = policy) %>% 
+  dplyr::rename(uniform = policy) %>% 
   gather(policy, harvest, -state) %>%

### 2018-01-31 appendix/appendix_recruit_first.Rmd
don't render recruit_first v of appendix
not sure why is failing on drone, but should fix drone builds.

--- 2018-01-30 appendix/appendix_recruit_first.Rmd
@@
 
+```{r include = FALSE}
+## Not in use, turn rendering off
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, eval=FALSE)
+```
+
 ```{r message=FALSE}
@@
 ggplot2::theme_set(theme_bw())
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
-
 options(mc.cores=parallel::detectCores())

### 2018-01-31 manuscripts/manuscript.Rmd
WIP

@@
 abstract: |
-
-  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.
-Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to intuition and the precautionary principle.
-Through the use of cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.
+  Ecological management and decision-making typically focus on uncertainty about the future, but surprisingly little is known about how to account for uncertainty of the present: that is, the realities of having only partial or imperfect measurements.
+Our two primary paradigms for handling decisions under uncertainty -- the precautionary principle and optimal control -- have so far given entirely controdictory results.
+This paradox is best illustrated in the example of fisheries management, where many ideas that guide our thinking about ecological decision making were first developed.
+There, we find that simplistic optimal control approaches have repeatedly concluded that a manager should increase catch quotas when faced with greater uncertainty about the standing stock biomass.
+Meanwhile, most current best practices take a more precautionary approach, decreasing catch quotas by a fixed amount to account for uncertainty.
+Using comparisons to both simulated and historical catch data, we find that neither approach is sufficient to avoid stock collapses under moderate observational uncertainty.
+Using cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.
 In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning.
+These results have significant implications for how we handle uncertainty in both fisheries and ecological management more generally.
+Dealing with uncertainty in decisions requires approaches which are both adaptive and explicit about uncertainty, such as POMDPs.
 
@@
 
-However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
+It is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
 Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
@@
 @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
-POMDP algorithms remain a very active area in artificial intelligence [@Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
-Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods.
-For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.
+POMDP algorithms have a substantial literature starting almost contemporaneously to the work of @Reed1979 (see foundational papers by @Smallwood1973 and @Sondik1978) and remain an active area in artificial intelligence [e.g.
+@Kaelbling1998; @Pineau2003].
+By adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+As the details for solving such problems numerically are already well documented we will not rehash them here, but instead provide a preliminary R package [@sarsop-pkg] which provides an implementation of the SARSOP algorithm and the fisheries POMDP problem, which can be used to replicate any of the results presented here and explore other variations.
+Details of the analysis presented here including annotated code to reproduce and further explore all of the following results is provided in the supplementary material.
 
@@
 @Smallwood1973; @Sondik1978; @Kurniawati2008; @Williams2011], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.
+Figure S8 illustrates how the prior belief evolves in response to subsequent observations and actions over the course of an individual simulation.
 
@@
 
-**A resolution to the paradox**
+**A resolution to the paradox.**
+
 Herein lies our resolution to the paradox of uncertainty.
@@
 Myers Legacy Stock Assessment database [@RAM].
-Posterior distributions for parameters for the Graham-Schaefer model are estimated with uncertainty through Markov Chain Monte Carlo using `nimble` [@nimble] on the historical data, illustrating how this process might be done in more complex models as well (details and code in Supplement, Appendix B).
+Posterior distributions for parameters for the Gordon-Schaefer model are estimated with uncertainty through Markov Chain Monte Carlo using `nimble` [@nimble] on the historical data, illustrating how this process might be done in more complex models as well (details and code in Supplement, Appendix B).
 $B_{MSY}$ and corresponding TAC and CE policies are calculated based on posterior mean estimates of the model parameters, along with POMDP solution assuming a 10% measurement error rate.
@@
 These historical examples provide another useful lens to compare how the different strategies respond in the face of fluctuations in real data, rather than model simulations.
-In both stocks, historical catch almost always exceeds that recommended by any of our strategies, falling closer to the MSY value estimated (see appendix).
+In both stocks, historical catch almost always exceeds that recommended by any of our strategies, falling closer to the MSY value estimated (see Appendix B).
 In Argentine Hake, the persistent declines result in both POMDP and CE strategies quickly closing down the fishery in an attempt to let the stock recover to a more productive target biomass, while TAC persists with merely a reduced harvest.
@@
 In marine fisheries, evidence of such over-exploitation has already been well documented [@Worm2006; @Worm2009; @Costello2016].
-Our results suggest that measurement uncertainty may contribute to this pattern, and that reality of measurement errors may frustrate stock recovery under current best practices such as MSY, TAC, and CE.
-More generally, the study and application of strategies such as MSY and CE are in no way limited to the decisions in fisheries quotas, but underpin much of conservation and natural resource management in forests and other terrestrial ecosystems as well [e.g.
+Our results suggest that measurement uncertainty may contribute to this pattern, as such errors lead to over-exploitation under current best practices such as MSY, TAC or CE, which have previously been thought to be sufficient to ensure the recovery of stocks world-wide [@Costello2016].
+
+Our results also have important consequences for ecological management more generally.
+The study and application of management strategies such as MSY and CE underpin much of conservation and natural resource management in forest management and many other terrestrial ecosystems [e.g.
 @Clark2000].
-Our results underscore the realization that measurement uncertainty is not negligible.
-Neither optimization approaches which ignore uncertainty (like CE) or heuristic approaches to account for that uncertainty (like TAC) are sufficient for this task.
-Rather, adaptive approaches such as POMDP, which make use of previous estimates to continually adjust prior beliefs, are required to avoid negative ecological and economic outcomes.
+Descision making under uncertainty is a central challenge for ecology and conservation biology [@Polasky2011], which we have long attempted to address through either simplified optimal control models or more heuristic approaches such as the precautionary principle and resilience thinking [@Fischer2009].
+We have illustrated how both of these approaches can fail to provide an adequate strategy when faced with measurement uncertainty.
+To avoid this pitfall, a strategy must be dynamic to make best use of the available information without assuming that information is perfect.
+The simplisitic optimal control approach of CE fails by placing too much confidence in measurements even when they are significantly at odds with prior expectations.
+Similarly, the TAC approach also fails due to it's reliance on a simple one-size-fits-all strategy, not reducing catch quotas sufficiently when stocks are too low but also failing to increase them when stocks are higher.
+It is precisely this ability to adapt the policy in response to available information, as illustrated in Figure \ref{policy}, that makes the POMDP approach successful.
 Nevertheless, it may be tempting to see these results as either trivial or impractical, but a closer inspection reveals they are neither.
@@
 
-## Future directions
+## Conclusions and Future directions
 

### 2018-01-31 manuscripts/AmNat_response_to_rev.Rmd
WIP

@@
 
-We appreciate the careful and constructive comments of the editor and reviewers, which have allowed us to re-write the manuscript to both better communicate both the method and broader implications of the results shown here.
+We appreciate the careful and constructive comments of the editor and reviewers, which have allowed us to re-write the manuscript to address each of the issues raised here.
+
+Reviewer 1's primary objection essentially re-asserts the "paradox of measurement uncertainty" that previous literature has also asserted: that uncertainty should somehow make less cautious.
+We appreciate the reviewer highlighting the importance of explaning this central issue properly, and our revisions should make it clear how this erroneous and potentially dangerous conclusion arises from simplifying assumptions.
+Reviewer 2's primary objection is the missed opportunity of focusing too narrowly on fisheries.
+We agree entirely with the reviewer that the fundamental issues here are germane to ecological management and not specific to fisheries, and have revised the presentation substantially to address this.
+Both reviewers raised concerns about the clarity of the methods; so in addition to rewritten discussion of methods we now provide fully executable code to replicate the analysis shown here.
+Lastly but most importantly, both reviewers raise the observation that few fisheries are actually managed using a strict constant escapement (CE) policy.
+We have introduced more realistic management strategies based on regulating fishing mortality through maximum sustainable yield (MSY) and total allowable catch (TAC), as additional comparisions.
+TAC illustrates a different, more heuristic approach to uncertainty than seen in the optimal control methods that provides some valuable additional insights into the challenge of accounting for measurement error.
+We think this results in a much more interesting mansucript which should be relevant to the broad audience of the American Naturalist.
+
+Below we provide more detailed responses to each of the concerns raised and how they have been addressed in this revision.
 
@@
 
-we now emphasize that making decisions in face of imperfect information is by no means an issue isolated solely to fisheries, and that our focus on fisheries example is merely for the same reason introductory texts on ecological management have so often turned to fisheries: no other example of ecological managment has a rich or foundational a history in our understanding of ecosystem management and decision-theory.
-We now discuss the broader concept of uncertain decision making in the introduction and highlight examples where these issues most readily appear in so many other ecological applications.
+We now emphasize that making decisions in face of imperfect information is by no means an issue isolated solely to fisheries, which merely serves as a concrete example where relevant data can be easily obtained.
+We also note that fisheries management has long been a crucible in which ecological management theory and practice has been forged before becoming widely applied in other ecological contexts.
 
@@
 This caveat aside, my understanding of Reviewer 1's comment is that the Reviewer ran a set of simulations which were essentially the same as in the manuscript, except that the Reviewer explicitly excluded the possibility of 90 percent harvest of the stock population.
-In the reviewer's simulations, the optimal escapement level is lower in the presence of measurement error, whereas in fig.
+
+The analysis of Reviewer 1 is very interesting, and finds essentially the same paradoxical result as earlier published studies (e.g. @Clark1986; @Sethi2005); as the reviewer summarizes:
+
+> In every case the optimal escapement goal was lower when observation error was included.
+
+This is a paradoxical conclusion that defies the precautionary principle, and as such would have important ramifications in ecological management in many applications beyond fisheries.
+Further, though the reviewer's analysis is not identical to either the approaches of @Clark1986 or @Sethi2005 (neither assume constant escapement and both find this paradoxical result in only some but not all cases), we believe our analysis provides both a refutation of this result and a clear intutition as to where this misleading conclusion has come from.
+
+As the editor notes, we suspect that Reviewer 1's simulations are not indeed identical except in the way mentioned, since they appear to assume the optimal policy can be specified in terms of a single constant escapement.
+Our results and in general escapement need not be constant but will depend on the observed stock size an on previous measurements, as we now emphasize.
+Also as the editor notes, the Reviewer's description of what they have done is extremely brief and thus difficult to be sure we have characterized precisely, but nevertheless, we believe our analysis sheds some light on why the reviewer finds this paradoxical conclusion while our analysis does not.
+
+The escapement recommended by our POMDP solution is not constant, but depends both on prior information and the current estimate of the stock size, as illustrated in Figure 4.
+Assumptions that simplify the calculation of the policy by ignoring these aspects can thus reach different conclusions.
+@Clark1986 assumes that the stock size is known perfectly at the end of the prior period, @Sethi2005 implicitly assumes that the prior belief is always uniform rather than reflecting previous measurements, which puts erroneous probability weight on the chance the stock is in fact very high.
+
+Figure 4 offers a clear explanation for when it is optimal to increase harvest in response to increased uncertainty: the POMDP harvest only exceeds the CE harvest when the measured stock size is erroneously low.
+That is, when the prior belief would have suggested the observation should be much higher, and therefore the measurement is most likely an error under-estimating the true stock size.
+This logic is entirely intuitive and is bourne out in the resulting POMDP policies.
+
+>  In the reviewer's simulations, the optimal escapement level is lower in the presence of measurement error, whereas in fig.
 5 of the manuscript,  the optimal escapement level is instead higher in the presence of measurement error.
@@
 If the authors have indeed allowed for 90 percent harvest, however, it is important that they test whether their results are different if they place an upper limit on the fraction of the stock that they allow to be harvested.
-I also agree with Reviewer 1's comment that the methods are not that easy to follow.
+
+The reviewer and editor have raised an important issue in the question of whether we permit the harvest quota to be set to a value higher than the stock.
+There are two problems with this.
+First, it's not clear how the reviewer implements this assumption in the face of measurement uncertainty: how can one be certain a quota is less than 90% of stock size X without knowing what X is?
+One can obviously enforce this with respect to the measured stock size but not the true stock size.
+More importantly, this assumption amounts to the assertion that it is mathematically impossible to overfish a stock to the point of collapse.
+However low the stock size might get, local extinction is impossible.
+In this world-view, there will always be some tiny fraction of ato-fishes from which the population can recover.
+Assuming that local extinction is impossible will indeed bias any optimal decision model to be more agressive.
+Obviously this also has impliciations for ecological models in general, conservation policy would look different if we assumed extinction was impossible.
+@Clark1986 already recognized that a very similar assumption in @Reed1979 derivation that stochastic models needed no additional caution ($S=D$) was likely due to the fact that accidentally extinction was impossible (see manuscript, line 150).
+We do not adopt this assumption.
+Obviously real world populations cannot be indefinitely reduced by only a fractional percentage.
+In our simulations, though we consider a Ricker model, stock abundances are divided into 100 distinct bins, with the smallest bin corresponding to a stock size of zero.
+Thus it is possible for overharvesting or stochastic environment, or some combination of both, to lead a population to permanenently collapse once it has gotten small enough.
+
+> I also agree with Reviewer 1's comment that the methods are not that easy to follow.
 That said, to my understanding, the escapements in fig.
@@
 
-The simulations performed by Reviewer 1 are not the same as in our example, but are nevertheless very enlightening as they highlight a very important issue that is far more general than fisheries.
-Reviewer 1's simulations have provided another example of the paradox of measurement error that we now highlight much more clearly in the previous work of @Reed1979, @Clark1986, @Sethi2005, and others: Reviewer 1 makes a "simplifying assumption" and paradoxically concludes that in the face of higher measurement error we should increase fishing pressure!
-A key conclusion from our analysis is the demonstration that this common but paradoxical claim is in fact false.
-
-To assume harvest cannot exceed the standing stock is to assume (for any model without an Allee threshold) that humans cannot collapse a stock regardless of how hard they fish.
-Though this assumption that extinction is impossible is often mathematically convenient, we are hardly the first to argue that it is both biologically unrealistic and can lead to ill-advised management practices
-
-> An important tacit assumption in Reed’s analysis, as in the other works referred
-to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.
-
-The reviewer's approach is not essentially equivalent to our approach; as the editor already observes we do not assume that the optimal policy will involve constant escapement, nor does it.
-Nor is the reviewer's simulation the same as in @Clark1986 or @Sethi2005, since neither of those results involve constant escapement, though they
-
-Reviewer 1 proposes that we assume local extinction is impossible regardless of inaccuracies in measurement error.
-This assumption has been previously made in the literature for mathematical convenience, as we discuss it is in fact
-
-When measurements are *uncertain*, it is impossible *not* to allow for a 90% harvest of the stock.
-Indeed, this cuts to  the very heart of why measurement uncertainty in *decisions* is so important!
-If X is a random number from a log-normal distribution of parameters mu and sigma, how can one choose a number Y to be certain that Y < 90% X?
-Y = 0 is the only (non-negative) solution; you cannot be sure that *any* harvest level is guatenteed to be safe.
-
-The editor is absolutely correct, POMDP solutions do not result in a constant escapement, as we stated.  Unfortunately, Reviewer 1 does not provide any description of the methods they used.
-
-We have revised the
+We agree entirely with this, and hope that the rewritten description, together with complete code, helps make this somewhat clearer.
 
@@
 
-We have revised this section significantly to address both grammar and presentation.
-
-The formal definitions and methodology of POMDPs has already been well-documented in the appopriate literature, and there is no reason for us to rehash the all the precise technical details here, which only serves to obscure the most salient concept.
-The point here is that there is an immensely well-developed area of theory and tools to address this problem of decision-making under imperfect information.
-Though much of that research has come from other fields and is not currently familiar to the readers, we believe we now present enough context to understand the problem.
+The presentation of all of this material has been entirely re-written in what we hope you will find is a much more appropriate and accessible style.
+To avoid needless formalism, we do not rehash the technical details of solving Partially Observed Markov Decision Processes but instead point to the extensive literature already existing on the subject.
+Instead, we focus on describing the nature of the problem and how it differs from analyses such as the perfectly observed Markov Decision Process that provides the basis for the Constant Escapement strategy (@Reed1979) and from other Bayesian hierarchical modeling (e.g.
+partially observed Markov processes, pomp, or Hidden Markov Models, HMMs) which are not decisison processes at all, but rather a prerequiste part of choosing & paramaterizing a model, not using said model to make a decision.
 
@@
 
+> The graph below shows a typical result.
+The thin line is with no observation error, the ticker line with the observation error.
+In every case the optimal escapement goal was lower when observation error was included.
+
 We thank the Reviewer for this careful parallel investigation, which was very useful in focusing the manuscript on some of the key results.
-
-While based only on this description provided we are not able to precisely replicate Reviewer 1's analysis, we can observe serveral important ways in which it is quite different to the optimization problem we consider.
-
-We have now completely re-written our description and provided both a stand-alone R package and fully executable scripts implementing our methods.
-However, we do not attempt to provide a stand-alone introduction to Partially Observed Markov Decision Processes (POMDPs), which, all though less well known in ecology and evolutionary studies than their (fully observed) MDP cousins (which ecologists commonly refer to by their solution method, Stochastic Dynamic Programming or SDP)
-
-As the editor observes, this short discription does not permit us to properly understand what the reviewer has done to obtain
-different results, but as the editor already noted, the comments make it clear that it is not analgous to our approach,
-since our approach does not assume nor does it find a constant escapement policy to be optimal.  
-Moreover, the remark about restricting harvest to 90% of population size is very instructive, as it cuts to the heart of the problem of measurement uncertainty. 
-Clearly such a restriction is only possible if the manager has access to perfect observations of the stock size,
-an thus any simulation that enforces this restriction is implicitly ignoring the reality of measurement error.
-
-There are only two ways a manager can be certain that harvest does not exceed 90% of stock: (1) don't harvest, or (2) measure the stock size exactly, (or at least, exactly measure some absolute lower bound of possible stock sizes).
-
-As the literature has observed consistently for some time, (e.g. see an eloquent discussion in Clark & Kirkwood 1987) 
-if we only consider policies in which accidental population extinction by overharvesting is impossible, 
-(e.g. because we assume a priori harvest cannot exceed 90% of the stock), 
-there is every reason to suspect this will result in artifically less cautious policies.  If we begin with 
-the assumption that there's just no possible way overharvesting could destroy a stock, then the risks of
-overharvesting are drastically reduced: net present value need only reflect the loss of revenue while a stock
-rebuilds, and not the much greater loss that would result from extinction.  Because even the stochastic theory
-such as Reed 1979 made effectively the same assumption (e.g. assumes the stock is measured perfectly before
-harvest, so that it would never be rational to attempt a harvest)
+We believe the differences in our respective results can be explained by the differences in our methods as discussed in our reply to the Editor's discussion of this issue.
+In short, we believe the reviewer's seemingly counter-intuitive result that increased uncertainty should be met with an action (lowering escapement) that would increase the risk of stock collapse rather then the intuitive response to be cautious by reducing harvest, is simply an artefact of the reviewer's assumption that stock collapse is impossible, and thus increased harvesting poses no real risk.
 
@@
 
-As we discuss above, this is once again fundamental feature of measurement uncertainty -- and again one
-that was observed long before our paper and is well established by the literature we discuss at some length.
-If a fixed escapement policy is applied to a measurement that is uncertain, the true stock will be somewhat
-larger or somewhat smaller than the observed value, and thus the true escapement can be somewhat smaller
-or somewhat larger than the 'target' escapement.  Without perfect measurements, it is clearly impossible
-to achieve the target escapement exactly. If, by chance, a string of consecutive measurements repeatedly over-estimates
-the current stock size, a constant-escapement strategy will repeatedly over-harvest the stock.
-
-In a model without an Allee effect, such as the classic Gordon-Schaefer model we focus on here, successive over-harvesting
-can reduce a stock to an
-
-The novelty of our result is how to deal with this issue.
+We agree that the results and methodoloy could have been much more clearly presented.
+We hope the rewritten presentation of these results clarifies these issues.
+In particular, note that once you introduce measurement error a fixed escapement policy can only achieve a fixed *expected* escapement, you never know what the actual escapement is because you never know exactly how many fish you started with.
+As measurement errors are equally likely to be too high or too low, a run of "heads", so to speak, would result in consistently overharvesting and continued decline (see discussion around line 363.)
 
@@
 also include comparisons to a more realistic "Total Allowable Catch", or TAC strategy.
-
--- while a rigorous constant escapement policy shuts down
-all harvest below B_MSY (the constant escapement level), most fisheries actually continue to harvest
-even when the stock drops below the target stock size, though potentially at some decreasing rate,
-and do not shut down harvest until the population is significantly below this level, commonly B_MSY / 2, if at all.
-
-As mentioned above, we have now revised all of our analyses to consider these realistic harvest policies,
-and have added a substantial discussion addressing the differences between such constant harvest (or constant effort /
-fishing mortality) policies and the constant-escapement policies we have hitherto focused upon.  
-Because these policies permit harvesting below B_MSY, it is not surprising that we find them even more
-suseptible
 
@@
 
-We thank the reviewer for this careful and constructive review, and agree entirely with the recommendation that the fundamental issues here (managing with imperfect information) as well as some other themes (tipping points, management with constrained resources) are far more general.
-We have now tried to state these implications more clearly.
+We thank the reviewer for this careful and constructive review, and agree entirely with the recommendation that the fundamental issues here are not limited to the problem of fishing quotas butare germane to much of ecological decision making.
+We have now tried both to set up the generic nature of the issue of measurement uncertainty (see Introduction) and discuss its implications beyond fisheries much more explicitly (e.g.
+lines 423 - 437).
+We have however removed the treatment of the tipping point model in order to focus explicitly on the role of uncertainty in the more familiar case of logistic growth, where the dramatic consequences of measurement uncertainty may be less inutitive and cannot be dismissed merely by questioning the frequency of tipping points occurring in real systems.
+We hope to address the problem of tipping points more explicitly and more fully in later work.
 
@@
 
-This is an excellent point which we now develop in the discussion:
-
-How do we determine the maximum sustainable yield in the first place?
-This number comes from a deterministic stock-recruitment model.
-In stochastic models, and even in many deterministic models with age and stage structure, there is no such thing as a maximum sustainable yield, since the optimal yield must fluctuate with the population dynamics.
-But at least in those cases one can reasonably assert an average
+This is an excellent point which we now develop in the discussion, lines 456-470.
 
@@
 
- As the reviewer suggests here and as was recognized by Fischer et al (2009) among others, one advantage of the optimal control approach is that it is generally more transparent and unbiased algorithmic way to reach a decision, which to some extent sidesteps the inherently more subjective existing methods, where this issue is acutely felt.
-For instance, our value function is a cold-blooded profit equation which places no value in leaving fish in the sea beyond their future return as harvest.
-We also ignore more realistic supply-and-demand effects as a conservative assumption, since the decreasing price of higher supply would dampen the incentive for larger harvests (see comparisons in appendix).
+We now include a formal comparison against TAC-based management policy (and also MSY based policy in the supplement), which reflect this aspect of today's actual management polcies which are not updated annually but apply a constant fishing effort (mortality).
+We also discuss several of the practical objections to implementing POMDP in practice and how it might be modified to address more of these concerns, e.g.
+lines 456-490.
 
@@
 
-Excellent observation! We have added the following line to emphasize this more clearly:
+Excellent observation! Yes, due to the influence of the prior belief, POMDP is slightly more catious when a stock is rebuilding from small sizes
 
@@
 
+We agree entirely, and now present the full discussion of the methods and models before going into the results.
+
 > In several places the authors use loose language by referring to marine ecosystem dynamics when in fact they are treating population dynamics.
 
-Agreed.
+Agreed, this has been addressed in the rewrite.
 
 > The authors discuss volatility of different solutions and would be served well to quantify it. E.g., lines 365-369
+
+We believe this is much better illustrated by the visualization of these solutions in Figure 5, and introducing a statistic to quantify this would only needlessly confuse that discussion.

### 2018-01-31 appendix/appendixA.Rmd
Figure numbers, sp

@@
 
-```{r optional_plotting_config, include=FALSE}
+```{r optional_plotting_config, cache = FALSE, include=FALSE}
 library(ggthemes)
@@
 
-## Unlike classic Graham-Schaefer, this assumes harvest 
+## Unlike classic Gordon-Schaefer, this assumes harvest 
 ## occurs right after assessment, before recruitment
@@
 
-```{r fig.width=5, fig.height=5}
+```{r S1, fig.width=5, fig.height=5,fig.cap = "MSY, TAC, and CE policies under deterministic simulation.
+MSY and CE converge to the same standing biomass long term, but CE reaches this level more quickly.
+TAC always harvests less than MSY, resulting in a higher standing biomass but lower long-term yeild.
+Note that POMDP is identical to CE when no uncertainty is present."}
 det_sims %>%
@@
 
-```{r eval=FALSE}
+```{r eval=!file.exists(file.path(log_dir, "meta.csv"))}
 dir.create(log_dir)
@@
 
-```{r pomdp_sims}
+```{r pomdp_sims, eval=!file.exists(file.path(log_dir, "pomdp_sims.csv"))}
 set.seed(12345)
@@
            .id = "scenario")
+
+write_csv(pomdp_sims, file.path(log_dir, "pomdp_sims.csv"))
 ```
@@
 ```{r}
+pomdp_sims <- read_csv(file.path(log_dir, "pomdp_sims.csv")) %>% 
+  mutate(scenario = as.character(scenario))
+
 sims <- bind_rows(static_sims, pomdp_sims) %>%
@@
 
-```{r}
+```{r S2, fig.width=8, fig.cap="Simulations as in the main text, Figure 1, shown across varying growth noise levels and including MSY strategy for comparison."}
 sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnnn")
@@
 
-```{r}
+```{r S3, fig.cap="Economic value as in Fig 2 main text, shown across different levels of environmental noise, including MSY strategy."}
 sims %>%
@@
 
-```{r overest}
+```{r overest, eval=!file.exists(file.path(log_dir, "pomdp_overest_sims.csv"))}
 set.seed(12345)
@@
   )
+write_csv(pomdp_overest_sims, file.path(log_dir, "pomdp_overest_sims.csv"))
 ```
@@
 ```{r}
+pomdp_overest_sims <- read_csv(file.path(log_dir, "pomdp_overest_sims.csv")) %>% 
+  mutate(scenario = as.character(scenario))
+
 overest <-
@@
 
-```{r}
+```{r S4, fig.cap="over-estimating measurement uncertainty using POMDP, as in Fig 3 main text."}
 read_csv(file.path(log_dir, "overest_sims.csv")) %>%
@@
 
-```{r}
+```{r S5, fig.cap="The CE policy is identical to the solution found by SDP under very small noise.
+MSY and TAC can be seen to correspond to lower escapements at small stock sizes, since they continue to harvest stocks that are below $B_{MSY}$."}
 bind_rows(policy_table,
@@
 
-```{r}
+```{r S6, fig.cap="For large growth noise, the escapement under SDP is still constant but very slightly higher than the (deterministic) CE solution, $S>D$."}
 i <- meta %>% filter(sigma_g == 0.15, sigma_m ==0) %>%
@@
   geom_line(lwd=1) + 
-  coord_cartesian(xlim = c(0,1), ylim=c(0,.8))
+  coord_cartesian(xlim = c(0,1), ylim=c(0,.8)) +
+  theme(legend.position = "none")
 
-p1
-```
-
-```{r}
 p2 <- recode_policies %>%
@@
   geom_line(lwd=1) + 
-  coord_cartesian(xlim = c(0,1), ylim=c(0,.6))
+  coord_cartesian(xlim = c(0,1), ylim=c(0,.6)) + 
+  theme(legend.position = "bottom")
 
-p2
 ```
@@
 
-```{r}
+```{r S7, fig.height=10, fig.width=5}
 p3 <- read_csv(file.path(log_dir, "priors.csv")) %>%
@@
   geom_line(lwd=1) +
-  coord_cartesian(xlim = c(0,1))
-p3
+  coord_cartesian(xlim = c(0,1)) + 
+  theme(legend.position = "none")
+
+gridExtra::grid.arrange(p1, p2, p3, ncol=1)
 ```
@@
 
-```{r}
+```{r S8, fig.cap="Observation, True state, and Harvest quota over the course of a single simulation under POMDP management."}
 sim$df %>%
@@
 
-```{r}
+```{r S9, fig.cap="The evolution of the POMDP prior belief distribution over the course of the simulation."}
 sim$state_posterior %>% 

### 2018-01-31 manuscripts/manuscript.Rmd

@@
 These plots show, for any possible observations, which strategy will attempt to harvest most and which will attempt to harvest least.
-The first panel shows the policy in terms of expected escapement, $S = Y - H$ for observed stock $Y$ and harvest $H$, consistent with most of the relevant control and bioeconomics literature [e.g.
-@Reed1979; @Clark1986; @Sethi2005].
-For comparison, the second panel shows the same policies but in terms of harvest $H$.
+For comparison purposes, we plot policies both in terms of expected escapement, $S = Y - H$ as is typical in the optimal control literature, [e.g.
+@Reed1979; @Clark1986; @Sethi2005], and also directly in terms of the harvest quota $H$.
 These plots clearly illustrate the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.
-Escapement CE strategy corresponds to keeping escapement constant.
-This is important, as theoretical work ever since the proofs of @Reed1979 often attempts to simplify the calculations by assuming the optimal policy must be a constant-escapement type strategy [e.g.
-@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
-Note that both CE and TAC policies are independent of the levels of uncertainty $\sigma_m$ and $\sigma_g$.
-TAC implicitly accounts for some uncertainty through its 80% MSY definition -- comparable plots with MSY can be found in the Supplemental Material.
+(Comparable plots with MSY can be found in the Supplemental Material).
 In principle, the CE policy could depend on $\sigma_g$, but as @Reed1979 proved, the constant escapement level with stochasticity is the same as in the deterministic case, S = D, unless the noise level is quite high (comparison plots in Supplemental material.)  In contrast to this, our POMDP plots depend very much on both the choice of $\sigma_m$ and $\sigma_g$.
@@
 
-Different attention to prior events fundamentally drives the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
+Differences in attention to prior beliefs,determined by prior observations, drive the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
 While the catch quota under both TAC and CE strategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach.
@@
   geom_line(lwd=1) +
-  facet_wrap(~commonname, scales = "free", ncol=1)  
+  facet_wrap(~commonname, scales = "free", ncol=1)  +
+  ylab("stock biomass (metric tonnes)")
 ```
@@
 Despite this volatility, the TAC level remains relatively level across all five decades, rising or dipping only slightly with changes in stock size.
-In contrast, the CE harvest tracks this volatility almost in perfect mirror.
+In contrast, the CE harvest tracks this volatility almost exactly.
 Here, the POMDP solution falls in between these extremes, almost mirroring which-ever of the two policies is more conservative: the POMDP solution matches the dips in harvest taken by CE to allow the stock to recover most quickly, but does not track the doubling of harvests recommended by CE in the mid 1980, increasingly only to the more modest harvest recommended by the TAC policy.

### 2018-02-01 manuscripts/AmNat_response_to_rev.Rmd
response to reviewers

@@
   pdf_document:
-    keep_tex: true
     fig_caption: true
+
 bibliography: citations.bib
+header-includes: |
+  \usepackage{times}
+  \usepackage{etoolbox}
+  \usepackage[dvipsnames]{xcolor}
+  \usepackage[framemethod=tikz]{mdframed}
+  \let\oldquote\quote
+  \let\endoldquote\endquote
+  \renewenvironment{quote}
+    { \vspace{3.0\parskip} \oldquote
+    \mdframed[backgroundcolor=gray!10]
+    \it\small }
+    {\endmdframed \endoldquote }
+  \mdfsetup{linewidth=0pt,roundcorner=10pt,skipbelow=5pt,skipabove=10pt}
 ---
 
-ASSOCIATE EDITOR'S RECOMMENDATION
+## ASSOCIATE EDITOR'S RECOMMENDATION
 
-Dear Professor Bronstein; Prof Dwyer,
-
+> Dear Professor Bronstein; Prof Dwyer,
+> 
 > My recommendation is that manuscript 57722 be declined without prejudice.
@@
 Reviewer 1's primary objection essentially re-asserts the "paradox of measurement uncertainty" that previous literature has also asserted: that uncertainty should somehow make less cautious.
-We appreciate the reviewer highlighting the importance of explaning this central issue properly, and our revisions should make it clear how this erroneous and potentially dangerous conclusion arises from simplifying assumptions.
+We appreciate the reviewer highlighting the importance of explaining this central issue properly, and our revisions should make it clear how this erroneous and potentially dangerous conclusion arises from simplifying assumptions.
 Reviewer 2's primary objection is the missed opportunity of focusing too narrowly on fisheries.
@@
 Lastly but most importantly, both reviewers raise the observation that few fisheries are actually managed using a strict constant escapement (CE) policy.
-We have introduced more realistic management strategies based on regulating fishing mortality through maximum sustainable yield (MSY) and total allowable catch (TAC), as additional comparisions.
+We have introduced more realistic management strategies based on regulating fishing mortality through maximum sustainable yield (MSY) and total allowable catch (TAC), as additional comparisons.
 TAC illustrates a different, more heuristic approach to uncertainty than seen in the optimal control methods that provides some valuable additional insights into the challenge of accounting for measurement error.
-We think this results in a much more interesting mansucript which should be relevant to the broad audience of the American Naturalist.
+We think this results in a much more interesting manuscript which should be relevant to the broad audience of the American Naturalist.
 
@@
 This is a paradoxical conclusion that defies the precautionary principle, and as such would have important ramifications in ecological management in many applications beyond fisheries.
-Further, though the reviewer's analysis is not identical to either the approaches of @Clark1986 or @Sethi2005 (neither assume constant escapement and both find this paradoxical result in only some but not all cases), we believe our analysis provides both a refutation of this result and a clear intutition as to where this misleading conclusion has come from.
+Further, though the reviewer's analysis is not identical to either the approaches of @Clark1986 or @Sethi2005 (neither assume constant escapement and both find this paradoxical result in only some but not all cases), we believe our analysis provides both a refutation of this result and a clear intuition as to where this misleading conclusion has come from.
 
@@
 That is, when the prior belief would have suggested the observation should be much higher, and therefore the measurement is most likely an error under-estimating the true stock size.
-This logic is entirely intuitive and is bourne out in the resulting POMDP policies.
+This logic is entirely intuitive and is borne out in the resulting POMDP policies.
 
@@
 One can obviously enforce this with respect to the measured stock size but not the true stock size.
-More importantly, this assumption amounts to the assertion that it is mathematically impossible to overfish a stock to the point of collapse.
+More importantly, this assumption amounts to the assertion that it is mathematically impossible to over-fish a stock to the point of collapse.
 However low the stock size might get, local extinction is impossible.
 In this world-view, there will always be some tiny fraction of ato-fishes from which the population can recover.
-Assuming that local extinction is impossible will indeed bias any optimal decision model to be more agressive.
-Obviously this also has impliciations for ecological models in general, conservation policy would look different if we assumed extinction was impossible.
+Assuming that local extinction is impossible will indeed bias any optimal decision model to be more aggressive.
+Obviously this also has implications for ecological models in general, conservation policy would look different if we assumed extinction was impossible.
 @Clark1986 already recognized that a very similar assumption in @Reed1979 derivation that stochastic models needed no additional caution ($S=D$) was likely due to the fact that accidentally extinction was impossible (see manuscript, line 150).
@@
 In our simulations, though we consider a Ricker model, stock abundances are divided into 100 distinct bins, with the smallest bin corresponding to a stock size of zero.
-Thus it is possible for overharvesting or stochastic environment, or some combination of both, to lead a population to permanenently collapse once it has gotten small enough.
+Thus it is possible for over-harvesting or stochastic environment, or some combination of both, to lead a population to permanently collapse once it has gotten small enough.
 
@@
 Instead, we focus on describing the nature of the problem and how it differs from analyses such as the perfectly observed Markov Decision Process that provides the basis for the Constant Escapement strategy (@Reed1979) and from other Bayesian hierarchical modeling (e.g.
-partially observed Markov processes, pomp, or Hidden Markov Models, HMMs) which are not decisison processes at all, but rather a prerequiste part of choosing & paramaterizing a model, not using said model to make a decision.
+partially observed Markov processes, pomp, or Hidden Markov Models, HMMs) which are not decision processes at all, but rather a prerequisite part of choosing & paramaterizing a model, not using said model to make a decision.
 
@@
 That said, the complexity of the technical issues is quite high, and so it may be simpler to submit the manuscript to a journal that specializes in mathematical biology or in fisheries management.
+> -- Greg Dwyer
 
-Greg Dwyer
+---
 
-xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
+# Reviewer #1:
 
-Reviewer #1:
-
-For complete review, see the attached PDF, which begins as follows:
-
+> For complete review, see the attached PDF, which begins as follows:
+>
 > I attempted to evaluate the papers overall conclusions by a very simple simulation and these results
@@
 year.
-
+> 
 > The graph below shows a typical result.
@@
 We believe the differences in our respective results can be explained by the differences in our methods as discussed in our reply to the Editor's discussion of this issue.
-In short, we believe the reviewer's seemingly counter-intuitive result that increased uncertainty should be met with an action (lowering escapement) that would increase the risk of stock collapse rather then the intuitive response to be cautious by reducing harvest, is simply an artefact of the reviewer's assumption that stock collapse is impossible, and thus increased harvesting poses no real risk.
+In short, we believe the reviewer's seemingly counter-intuitive result that increased uncertainty should be met with an action (lowering escapement) that would increase the risk of stock collapse rather then the intuitive response to be cautious by reducing harvest, is simply an artifact of the reviewer's assumption that stock collapse is impossible, and thus increased harvesting poses no real risk.
 
@@
 
-We agree that the results and methodoloy could have been much more clearly presented.
+We agree that the results and methodology could have been much more clearly presented.
 We hope the rewritten presentation of these results clarifies these issues.
 In particular, note that once you introduce measurement error a fixed escapement policy can only achieve a fixed *expected* escapement, you never know what the actual escapement is because you never know exactly how many fish you started with.
-As measurement errors are equally likely to be too high or too low, a run of "heads", so to speak, would result in consistently overharvesting and continued decline (see discussion around line 363.)
+As measurement errors are equally likely to be too high or too low, a run of "heads", so to speak, would result in consistently over-harvesting and continued decline (see discussion around line 363.)
 
@@
 
-xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
+---
 
-Reviewer #2:
+# Reviewer #2:
 
@@
 
-We thank the reviewer for this careful and constructive review, and agree entirely with the recommendation that the fundamental issues here are not limited to the problem of fishing quotas butare germane to much of ecological decision making.
+We thank the reviewer for this careful and constructive review, and agree entirely with the recommendation that the fundamental issues here are not limited to the problem of fishing quotas but are germane to much of ecological decision making.
 We have now tried both to set up the generic nature of the issue of measurement uncertainty (see Introduction) and discuss its implications beyond fisheries much more explicitly (e.g.
 lines 423 - 437).
-We have however removed the treatment of the tipping point model in order to focus explicitly on the role of uncertainty in the more familiar case of logistic growth, where the dramatic consequences of measurement uncertainty may be less inutitive and cannot be dismissed merely by questioning the frequency of tipping points occurring in real systems.
+We have however removed the treatment of the tipping point model in order to focus explicitly on the role of uncertainty in the more familiar case of logistic growth, where the dramatic consequences of measurement uncertainty may be less intuitive and cannot be dismissed merely by questioning the frequency of tipping points occurring in real systems.
 We hope to address the problem of tipping points more explicitly and more fully in later work.
@@
 
-We now include a formal comparison against TAC-based management policy (and also MSY based policy in the supplement), which reflect this aspect of today's actual management polcies which are not updated annually but apply a constant fishing effort (mortality).
+We now include a formal comparison against TAC-based management policy (and also MSY based policy in the supplement), which reflect this aspect of today's actual management polices which are not updated annually but apply a constant fishing effort (mortality).
 We also discuss several of the practical objections to implementing POMDP in practice and how it might be modified to address more of these concerns, e.g.
@@
 
-Excellent observation! Yes, due to the influence of the prior belief, POMDP is slightly more catious when a stock is rebuilding from small sizes
+Excellent observation!
+Yes, due to the influence of the prior belief, POMDP is slightly more cautious when a stock is rebuilding from small sizes.
+We now speak to this in our discussion of Figure 5.
+We agree with the reviewer's general observation that the POMDP approach allows for more careful tracking of stock size despite measurement error, which in turn leads to both faster rebuilding of stocks and higher economic yields.
 
@@
 We believe this is much better illustrated by the visualization of these solutions in Figure 5, and introducing a statistic to quantify this would only needlessly confuse that discussion.
+
+# References

### 2018-02-01 manuscripts/AmNat_response_to_rev.Rmd

@@
 As the editor notes, we suspect that Reviewer 1's simulations are not indeed identical except in the way mentioned, since they appear to assume the optimal policy can be specified in terms of a single constant escapement.
-Our results and in general escapement need not be constant but will depend on the observed stock size an on previous measurements, as we now emphasize.
+Our results and in general escapement need not be constant but will depend on the observed stock size and on previous measurements, as we now emphasize.
 Also as the editor notes, the Reviewer's description of what they have done is extremely brief and thus difficult to be sure we have characterized precisely, but nevertheless, we believe our analysis sheds some light on why the reviewer finds this paradoxical conclusion while our analysis does not.
@@
 Obviously real world populations cannot be indefinitely reduced by only a fractional percentage.
-In our simulations, though we consider a Ricker model, stock abundances are divided into 100 distinct bins, with the smallest bin corresponding to a stock size of zero.
+In our simulations, though we consider a Gordon-Schaefer model, stock abundances are divided into 100 distinct bins, with the smallest bin corresponding to a stock size of zero.
 Thus it is possible for over-harvesting or stochastic environment, or some combination of both, to lead a population to permanently collapse once it has gotten small enough.
@@
 We believe the differences in our respective results can be explained by the differences in our methods as discussed in our reply to the Editor's discussion of this issue.
-In short, we believe the reviewer's seemingly counter-intuitive result that increased uncertainty should be met with an action (lowering escapement) that would increase the risk of stock collapse rather then the intuitive response to be cautious by reducing harvest, is simply an artifact of the reviewer's assumption that stock collapse is impossible, and thus increased harvesting poses no real risk.
+In short, we believe the reviewer's seemingly counter-intuitive result that increased uncertainty should be met with an action (lowering escapement) that would increase the risk of stock collapse rather than the intuitive response to be cautious by reducing harvest, is simply an artifact of the reviewer's assumption that stock collapse is impossible, and thus increased harvesting poses no real risk.
 

### 2018-02-01 manuscripts/manuscript.Rmd
Sup Fig numbering

@@
 
-$$x_{t+1} = f(x_t, h_t) = \varepsilon_t x_t r  \left(1 - \frac{x_t}{K}\right) - h_t $$
+$$x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) $$
 
 where $x_t$ is the current stock size, $h_t$ the harvest chosen that year, with parameters $r$ giving the individual growth rate, and $K$ the carrying capacity and $\varepsilon$ representing stochastic recruitment in a variable environment.
+We have written this in terms of $x_t - h_t$ to underscore that we assume the convention of observe, harvest, recruit.
 We will assume for simplicity $\varepsilon \sim \mathcal{N}(1, \sigma_g)$.
@@
 @Smallwood1973; @Sondik1978; @Kurniawati2008; @Williams2011], but for our purposes it is sufficient to observe how this evolving prior belief serves to continually adjust the POMDP policy.
-Figure S8 illustrates how the prior belief evolves in response to subsequent observations and actions over the course of an individual simulation.
+Supplemental Figure S9 (Appendix A) illustrates how the prior belief evolves in response to subsequent observations and actions over the course of an individual simulation.
 
@@
 
-## Conclusions and Future directions
+## Future directions
 

### 2018-02-01 appendix/appendixA.Rmd

@@
     toc_depth: 2
+    keep_tex: true
 bibliography: ../manuscripts/citations.bib
@@
 ---
+\renewcommand{\thefigure}{S\arabic{figure}}
 
@@
 
-```{r library, message=FALSE}
+```{r library, message=FALSE,cache=FALSE}
 #devtools::install_github("boettiger-lab/sarsop@0.5.0")
@@
 
-```{r S7, fig.height=10, fig.width=5}
+```{r S7, fig.height=10, fig.width=5, fig.cap = "policy plots, as in manuscript, figure 4. Here we include a comparison to MSY as well."}
 p3 <- read_csv(file.path(log_dir, "priors.csv")) %>%

### 2018-02-01 manuscripts/manuscript.Rmd
manuscript, also separate figure files

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = TRUE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = FALSE)
 library(tidyverse)

### 2018-02-01 manuscripts/manuscript.tex
tex version

\documentclass[3p]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\bibliographystyle{elsarticle-harv}
\biboptions{sort&compress} % For natbib
\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={The importance of measurement uncertainty in ecological management},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header
\journal{The American Naturalist} \usepackage{lineno} \linenumbers
\modulolinenumbers[3]
\usepackage{setspace} \doublespacing

\begin{document}
\begin{frontmatter}

  \title{The importance of measurement uncertainty in ecological management}
    \author[]{Authors removed during peer review}

  \begin{abstract}
  Ecological management and decision-making typically focus on uncertainty
  about the future, but surprisingly little is known about how to account
  for uncertainty of the present: that is, the realities of having only
  partial or imperfect measurements. Our two primary paradigms for
  handling decisions under uncertainty -- the precautionary principle and
  optimal control -- have so far given entirely controdictory results.
  This paradox is best illustrated in the example of fisheries management,
  where many ideas that guide our thinking about ecological decision
  making were first developed. There, we find that simplistic optimal
  control approaches have repeatedly concluded that a manager should
  increase catch quotas when faced with greater uncertainty about the
  standing stock biomass. Meanwhile, most current best practices take a
  more precautionary approach, decreasing catch quotas by a fixed amount
  to account for uncertainty. Using comparisons to both simulated and
  historical catch data, we find that neither approach is sufficient to
  avoid stock collapses under moderate observational uncertainty. Using
  cutting-edge point-based partially observed Markov decision process
  (POMDP) methods, we demonstrate how this paradox arises from flaws in
  the standard management theory, which contributes to over-exploitation
  of fisheries and increased probability of economic and ecological
  collapse. In contrast, we find POMDP-based management avoids such
  over-exploitation while also generating higher economic value through
  adaptive learning. These results have significant implications for how
  we handle uncertainty in both fisheries and ecological management more
  generally. Dealing with uncertainty in decisions requires approaches
  which are both adaptive and explicit about uncertainty, such as POMDPs.
  \end{abstract}
   \begin{keyword} Measurement uncertainty \sep POMDP \sep Decision theory \sep Fisheries \sep Conservation \sep \end{keyword}
 \end{frontmatter}

Imperfect information is ubiquitous in ecological management and
conservation decision making. While the pressing concerns of global
change have put the spotlight on \emph{forecasting}, i.e.~the
uncertainty of the future (e.g. Petchey et al. 2015), management
decisions must also contend with uncertainty of the \emph{present}: How
many fish are in the sea today? What regions harbor the most fragile
biodiversity? How have invasive species transformed native communities?
A recognition of the importance of uncertainty as well as sophisticated
methods and tools to do so has become increasingly widespread in the
context of estimating ecological models from available data, thanks in
large part to the rise of approaches such as Hierarchical Bayesian
Modeling, along with convenient software tools for implementing these
methods. In sharp contrast to that shift, the treatment of uncertainty
in decisions -- how the uncertainty in models should be reflected in
scientific recommendations, remains far less developed, more
qualitative, and under-served by available tools. In this paper we
illustrate how uncertainty in measurement has a dramatic negative impact
on outcomes under a variety of these existing approaches, which can be
avoided through the use of more formal approaches that have since been
developed in other fields that can likewise be implemented in software
tools. This insight both resolves a long-standing paradox in fisheries
management and has important implications for how ecosystem management
deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems
can be divided into two camps: approaches based in ``optimal control''
and usually favored by natural resource economists, and approaches based
in heuristic methods such as scenario planning, resilience thinking, and
precautionary rules of thumb, more commonly found in both ecological
literature and actual practice (Fischer et al. 2009; Polasky et al.
2011). While researchers have for some time recognized the need to unify
the transparent and quantitative algorithmic approach of optimal control
with the greater complexity and uncertainty of real ecosystems that is
acknowledged by heuristic methods, computational barriers to doing so
have hither-to stymied this progress. Here, we illustrate how the
limitations of these approaches has been manifested in the example of
fisheries management, and present a new approach that can combine the
reality of measurement uncertainty and the rigor of optimization to
resolve a long-standing paradox and suggest a more robust approach to
management.

Fisheries conservation and management has long been both a crucible and
proving ground for the theory of ecological management more generally,
including topics such as adaptive management (Walters and Hilborn 1978),
ecosystem-based management (Levin and Lubchenco 2008) and resilience
thinking (Holling 1973; May 1977), while also giving rise to the
sub-discipline of resource economics (Gordon and Press 1954; Schaefer
1954; Beverton and Holt 1957), thanks to its global relevance, long
history, and readily available data. Consequently, understanding the
impact of measurement uncertainty in this rich and well-studied context
will have implications for both resource management and conservation
efforts in other domains, many of which draw on concepts established or
tested in fisheries ecosystems. In particular, the approach here
illustrates the limitations of both over-simple optimal control
approaches and rule-of-thumb precautionary approaches, and offers
instead a more general strategy for dealing with measurement uncertainty
adaptively.

We compare the results of managing simulated fish stocks under the
existing optimal control approach, a heuristic approach that more
closely resembles today's management practices, and a newly proposed
approach that provides a more rigorous treatment of uncertainty.

\section{The Decision Problem}\label{the-decision-problem}

We consider the management problem of setting catch quotas for a marine
fishery in the face of imperfect information about the current stock
size and uncertainty about future recruitment.

We seek to determine the sequence of actions \(h_t\) for
\(t \in [1, ... \infty]\) that maximize the net present value
(discounted sum of all future profits) of the fishery. We will denote
the discount rate \(\gamma\). For simplicity we will once again follow
classic theory and assume a fixed price for fish, (equivalently,
measuring our value in units of discounted fish rather than discounted
dollars), though again this assumption is easy to relax. Each year, the
manager also obtains an estimate \(y_t\) of the true \(x_t\) stock size
subject to some measurement uncertainty \(\xi_t\)

\[y_t = \xi_t x_t\]

For simplicity we will assume measurement uncertainty is normally
distributed with standard deviation \(\sigma_m\),
\(\xi \sim \mathcal{N}(1, \sigma_m)\). Given an estimated recruitment
model along with this measurement of the stock size, the manager must
choose the set of a harvest quotas \(\lbrace h_t \rbrace\) for
\(t \in [0, 1, \infty]\) that maximizes expected long term (net) present
value:

\[\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}\]

A few subtleties arise in this seemingly simple problem statement that
we must address. First, given our discrete-time formulation of the
recruitment process, it is necessary to decide if the measurement
\(y_t\) happens before or after harvest \(h_t\): that is do we: measure,
recruit, harvest, or measure, harvest, recruit? Following convention in
the optimal control literature (e.g. (Reed 1979; Clark 1990; Sethi et
al. 2005)), we will assume the latter; measurement occurs immediately
before harvest, followed by recruitment.

A second subtly is in the nature of the decision problem itself: note
that the manager in not myopic, thinking only about the next year, but
rather considers the sequence of all future actions \({h_t}\). While the
manager will revisit this sequence each year in light of new
observations, it is this ability of the decision problem to think ahead
that allows it to tolerate short-term costs (e.g.~reduced harvests in
the current season) for larger future payoffs. In this, ecological
management is a game of chess, always thinking several moves ahead. This
differs from the problem solved in theory as well as the practice of
maximum sustainable yield, which seeks to identify only a fixed
mortality \(F_{MSY}\) applied for all time, rather than a \emph{policy}
of varying quotas \(h_t\) depending on the stock assessment \(y_t\).
Understanding this distinction is key both to understanding the
differences between current theory and current best practice in
fisheries management.

\subsection{Population model}\label{population-model}

To facilitate tractability and interpretation, we will focus on the
well-studied Gordon-Schaefer model (Gordon and Press 1954; Schaefer
1954) of logistic population growth (stock recruitment),

\[x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) \]

where \(x_t\) is the current stock size, \(h_t\) the harvest chosen that
year, with parameters \(r\) giving the individual growth rate, and \(K\)
the carrying capacity and \(\varepsilon\) representing stochastic
recruitment in a variable environment. We have written this in terms of
\(x_t - h_t\) to underscore that we assume the convention of observe,
harvest, recruit. We will assume for simplicity
\(\varepsilon \sim \mathcal{N}(1, \sigma_g)\). This and similar models
are widely used in large scale analyses across diverse stocks (e.g.
(Costello et al. 2016; Britten et al. 2017)), and form the basis for
much of bioeconomic theory (Clark 1990). As such, it will be easier to
compare our results against intuition and classic theory, and avoid the
possibility of differences arsing only because of some particular subtle
assumption hidden in a more complex model.

It is important to remember that both existing methods we discuss and
the POMDP approach proposed here can or are already applied to more
complex models as well, including those with age or stage structure.
Likewise, while we consider the model and it's parameters as ``given,''
it should be understood to mean that this in practice comes as the
result of some model choice and statistical estimation from historical
data, frequently accounting for uncertainty in both the process and the
measurements (e.g. Dichmont et al. 2016). As we shall see here, having
included uncertainty in the model \emph{estimation} in no way excuses us
from having to also deal once again with that uncertainty in the
decision process.

\section{Current Theory and Practice}\label{current-theory-and-practice}

Gordon and Press (1954) and Schaefer (1954) independently showed in the
same year that the maximum sustainable yield for the model bearing their
names is achieved by reducing the stock to the population to the size at
which it obtains the maximum growth rate which is known as Biomass at
Maximum Sustainable Yield, \(B_{MSY}\). For the Gordon-Schaefer model
(and many others), this is achieved at \(B_{MSY} = K/2\). They observed
that fishing at a \emph{constant yield} (that is, an individual fish
mortality, \(F\), such that harvest \(H = F \cdot B\)) of
\(F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}\)) will eventually lead to a
population that converges to the biomass \(B_{MSY}\) and produces the
maximum sustainable harvest, \(H_{MSY} = r K / 4\) in this model. From
here, theory and practice diverge.

This \emph{constant yield} (constant mortality) solution thus
corresponds to an equilibrium analysis which does not solve the
time-dependent optimization problem above. In particular, this maximum
sustainable yield (MSY) strategy will not be optimal whenever the stock
is away from \(B_{MSY}\). Clark (1973) demonstrated (assuming no
uncertainty in measurement or stochasticity in population dynamics) that
the optimal time-dependent strategy is not one of constant yield, but
rather of \emph{constant escapement}. We will return to this approach in
a moment.

\subsection{Uncertainty in Practice: MSY \&
TAC}\label{uncertainty-in-practice-msy-tac}

Maximum Sustainable Yield (MSY) remains the basis of international law
(including the UN, IWC, IATTC, ICCAT, ICNAF; Mace (2001)) and a familiar
standard of management in terrestrial ecosystems as well as aquatic
(Clark 1990). Critics have for some time observed the limitations of
harvesting at MSY in face of uncertainty in stock sizes and population
dynamics (Larkin 1977; Botsford et al. 1997). Many US fisheries reflect
this uncertainty through a series of adjustments that effectively reduce
the target fishing mortality level to reflect this uncertainty.
Typically, a stock assessment model provides a best estimate of
\(B_{MSY}\) and corresponding mortality \(F_{MSY}\) is used to define
the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower
level is set as the Allowable Biological Catch (ABC), reflecting
uncertainty in the stock assessment. To reflect possible uncertainty
between reported and actual catch, the ABC is reduced somewhat further
to define the Total Allowable Catch, (TAC), which forms the basic unit
of management for many such fisheries. To reflect this process, our
analysis will also consider policies in which the harvest quota is set
at 80\% of the level expected under the MSY policy:
\(H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t\), for a biomass estimated at
\(B(t)\) in year \(t\). To distinguish this approach from MSY, we will
refer to this approach as a TAC policy. This more closely represents a
heuristic (e.g. Hilborn 2010) or resilience-based approach (sensu
Fischer et al. 2009) than an optimization-based policy. Importantly,
this approach shares the fundamentally stationary assumptions of an MSY
policy by defining a constant mortality rather than a dynamic policy.
Thus, despite being more cautious overall and generating lower economic
yield than a constant escapement policy, TAC policies continue to
harvest at non-zero rates even if a stock falls below \(B_{MSY}\), while
the constant escapement policy does not.

\subsection{Optimal management}\label{optimal-management}

So far, attempts to provide a more formal basis for managing uncertainty
than the heuristic adjustment of catch limits described above have
largely foundered in a series of paradoxes. Without uncertainty, the
theoretical policies are quite intuitive: The result derived by Clark
(1973) for an optimal dynamic strategy to replace the equilibrium
solution of MSY can be summarized as: If the stock is most productive at
\(B_{MSY}\), then obtain \(B_{MSY}\) as quickly as possible. Thus, at
any level below \(B_{MSY}\), the economically optimal thing to do is to
completely shut down the fishery, while above this stock, harvests
greater than \(H_{MSY}\) will be needed to bring the stock back to
\(B_{MSY}\). (This intuition must be adjusted slightly in the case of
economic discounting of future profits, but is otherwise quite general,
see Clark (1990)). This strategy is known as constant escapement, since
a constant sock size \(B_{MSY}\) escapes harvest each year. While both
converge to the same long-term biomass and long-term yield, these
differences away from equilibrium between the mortality-based approach
that dominates fisheries management practice and the escapement-based
approach that became the focus of theory will result in different
behaviors under uncertainty. Intuitively, any uncertainty, either in the
population growth model (stochasticity), or measurement error, should
seem to justify harvesting fewer fish, as is built into the TAC limits
discussed above. Yet formal treatments of these issues have so far
required approximations, and these approximations have driven
paradoxical conclusions which we briefly summarize here.

\subsection{\texorpdfstring{Reed's Paradox:
\(S = D\)}{Reed's Paradox: S = D}}\label{reeds-paradox-s-d}

The first of these we shall refer to as Reed's Paradox for stochastic
growth, owing to the mathematical proof provided in Reed (1979) which
demonstrates that, under sufficiently general assumptions, the optimal
escapement \(S\) for a population under stochastic growth, is identical
to the optimal escapement \(D\) of a deterministic population,
\(S = D = B_{MSY}\). This surprising result suggests that in going from
a world where a manager has perfect knowledge of absolutely everything
into a scenario where the manager faces considerable uncertainty about
the future state of the world, no additional precaution is needed. This
also provides the manager with a remarkably convenient mechanism for
determining the optimal harvest policy: rather than rely on
computationally intensive Stochastic Dynamic Programming to determine
the optimal policy, a simple derivative will suffice, which is all that
is needed to solve the maximum of the deterministic problem.

\subsection{Clark's Paradox}\label{clarks-paradox}

Clark and Kirkwood (1986) was among the first attempts to resolve Reed's
Paradox. Clark and Kirkwood (1986) (quite correctly, as we will see),
identified the crux of Reed's Paradox as the absence of measurement
uncertainty:

\begin{quote}
An important tacit assumption in Reed's analysis, as in the other works
referred to above, is that the recruitment level X is known accurately
prior to the harvest decision, {[}\ldots{}{]} In the case of fishery
resources, the stock level X is almost never known very accurately,
owing to the difficulty of observing fish in their natural environment.
\end{quote}

Unfortunately, Clark and Kirkwood (1986) was unable to solve the
resulting problem exactly, but had to adopt an almost equally
troublesome assumption:

\begin{quote}
For reasons of tractability, we shall adopt the simplifying assumption
that the escapement level S, is known exactly at the end of that period.
(The mathematical difficulty of the problem increases markedly if this
assumption is relaxed.)
\end{quote}

Unfortunately for Clark and Kirkwood (1986), this ``simplifying
assumption'' serves to squash most of the measurement error, and their
results instead only deepened the paradox, finding policies that become
even \emph{less} cautious as uncertainty increases:

\begin{quote}
{[}Our{]} results appear to contradict the conventional wisdom of
renewable resource management, under which high uncertainty would call
for increased caution in the setting of quotas.
\end{quote}

Relying on a quite different but still flawed assumption nearly two
decades later, Sethi et al. (2005) largely confirm Clark's Paradox,
which they likewise observed with some concern:

\begin{quote}
It may seem counter-intuitive that a measurement error causes lower
expected escapements below the deterministic fishery closure threshold.
\end{quote}

Despite these notes of caution, both Clark and Kirkwood (1986) and Sethi
et al. (2005) ultimately attempt to rationalize this counter-intuitive
conclusion rather than reject it. Others have also attempted to
introduce measurement error, but always relying on various
approximations that similarly alter the problem. The most common such
assumption is to assume from the outset that the solution must be of
`constant-escapement' type (Ludwig and Walters 1981; Roughgarden and
Smith 1996; Engen et al. 1997; Moxnes 2003). Here we will illustrate the
use of powerful modern algorithms that will permit a more direct
solution to the case of measurement uncertainty. In addition to
resolving this paradox, we will demonstrate this direct solution also
out-performs current more heuristic rules of thumb such as those
currently used in fisheries management to account for uncertainty.

\subsection{POMDPs: An optimal treatment of measurement
uncertainty}\label{pomdps-an-optimal-treatment-of-measurement-uncertainty}

As discussed above, absent the short-cut provided by Reed's theorems --
solving for the optimal management policy in a stochastically
fluctuating population -- involves a Markov Decision Process (MDP),
commonly referred to in ecological literature by its solution method,
Stochastic Dynamic Programming (SDP). The constant escapement (CE)
strategy considered here is the result of such an MDP or SDP analysis.
Once measurement uncertainty is introduced, the problem becomes a
Partially Observed Markov Decision Process (POMDP, not be confused with
a partially observed or hidden Markov model, HMM, in which no decision
process is involved). The POMDP problem is posed identically to that of
the MDP problem, with the addition of an observation process, but cannot
be solved using SDP directly. The POMDP problem for fisheries question
considered here can be summarized as follows:

\begin{itemize}
\item
  Transition function (state equation): \(T(x_t, x_{t+1}, a_t)\): the
  probability that a system is in state \(x_{t+1}\) at time \(t+1\)
  given that it began in state \(x_t\) at time \(t\) and the manager
  took action \(a_t\). In our context, this relationship is given by the
  Gordon-Schaefer stock recruitment function \(f\) with normally
  distributed growth uncertainty
  \(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero
  to exclude negative population sizes.
\item
  Observation function: \(O(x_t,y_t,a_{t-1})\) the probability of
  observing state \(y_t\) given a system in state \(x_t\). In principle,
  the action chosen can influence the precision of the observation. In
  our case, we simply assume normally distributed errors around the true
  state, \(y_t \sim \mathcal{N}(x_t, \sigma_m)\), truncated at zero to
  exclude negative population sizes.
\item
  Utility function: \(U(x_t,a_t)\), the utility received at time \(t\)
  for taking action \(a_t\), given that the system is in state \(x_t\).
  For simplicity of analysis, we will simply set the utility to be equal
  to the harvested stock: \(U(x_t, a_t) = \min(x_t, a_t)\), indicating
  that realized harvest cannot be negative. This choice ensures that in
  the case of no uncertainty (\(\sigma_g = \sigma_m = 0\)), the optimal
  solution matches that expected under a simple MSY calculation. More
  realistic utility functions may include diminishing returns with
  increasing harvest (supply and demand effects), and the cost of
  fishing, both of which act to suppress large harvests. By focusing on
  a simple utility we can be sure that our comparison to MSY is driven
  by the treatment of uncertainty rather than merely differing economic
  assumptions.
\end{itemize}

The optimization problem is to select the action \(a_t\) that will
maximize the net present utility over all time. Future utility may be
discounted by a factor \(\gamma\), so that a value \(V\) in \(t\) year
is valued at \(\gamma^t V\) today. Numerically, each of these functions
are defined over a discrete set of possible states, observations, and
actions, and can thus be represented as a collection of matrices or
tensors. While the solution method for MDP problems, stochastic dynamic
programming (SDP) is well known and readily implemented in ecological
problems (e.g Mangel and Clark 1988; Marescot et al. 2013), the POMDP
problem cannot be efficiently solved using SDP. While not unknown to the
conservation literature (Williams 2011), algorithms for POMDP have
historically scaled quite poorly, and their application has been
restricted to contexts with only a handful of possible states and
actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight 2014; Fackler
and Pacifici 2014). POMDP algorithms have a substantial literature
starting almost contemporaneously to the work of Reed (1979) (see
foundational papers by Smallwood and Sondik (1973) and Sondik (1978))
and remain an active area in artificial intelligence (e.g. Kaelbling et
al. 1998; Pineau et al. 2003). By adapting recent algorithmic
developments (Kurniawati et al. 2008) from that field we are able to
find solutions for the considerably more complex problems such as the
case of fisheries management, which requires on the order of 100 states
and actions to provide sufficient numerical resolution. As the details
for solving such problems numerically are already well documented we
will not rehash them here, but instead provide a preliminary R package
(Boettiger et al. 2018) which provides an implementation of the SARSOP
algorithm and the fisheries POMDP problem, which can be used to
replicate any of the results presented here and explore other
variations. Details of the analysis presented here including annotated
code to reproduce and further explore all of the following results is
provided in the supplementary material.

\section{Results}\label{results}

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/sims-1.pdf}
\caption{Average fish biomass under different management strategies
under increasing levels of measurement uncertainty. Each plot the mean
stock size over time across 100 replicate simulations under each policy:
constant escapement (CE), Total allowable catch (TAC = 80\% MSY), and
the proposed partially observed Markov decision process (POMDP) method.
Measurement error increases as a normal distribution with standard
deviation 0, 0.1, or 0.15, as indicated at the top of the panel.
Environmental stochasticity is fixed a standard deviation of 0.15 in
each panel. Carrying capacity K normalized to 1, r = 0.75. Additional
environmental noise levels and comparison to MSY rather than TAC can be
found in the supplementary material. \label{sims}}
\end{figure}

Figure \ref{sims} shows average fish biomass across 100 replicate
simulations under three different management strategies: constant
escapement (CE), total allowable catch (TAC; equal to 80\% MSY), and
partially observed Markov decision process (POMDP) management. Each
successive panel shows a subsequently higher level of measurement error,
from \(\sigma_m = 0\), \(\sigma_m = 0.1\), and \(\sigma_m = 0.15\), as
indicated. Standard deviation from the mean across replicate simulations
is shown as faint colored bands, indicating significant variation due to
stochasticity between individual replicates. In each panel shown in
Figure \ref{sims} stochastic recruitment (environmental noise) is set to
a moderate \(\sigma_g = 0.15\). In the absence of either environmental
noise or measurement error (Supplemental material, Figure S1) POMDP, CE,
and MSY would converge to stock at the \(B_{MSY}\), while TAC would
maintain the stock at a slightly higher level. The first panel of Figure
\ref{sims} shows that the the introduction of stochastic growth has a
significant negative impact on the TAC strategy (This impact is even
more severe for MSY, which always harvests strictly more than TAC by
definition, as, shown in Supplemental material Figure S2), which results
in an average biomass significantly lower than \(B_{MSY}\). Without
measurement error, the CE and POMDP strategies are nearly identical with
both approximately maintaining the stock at \(B_{MSY}\) despite the
significant environmental stochasticity. As measurement error increases
in the subsequent panels, CE and TAC strategies perform increasingly
poorly, while the POMDP continues to maintain the average stock close to
\(B_{MSY}\). Notably, CE is significantly more impacted by measurement
uncertainty than TAC, with CE averaging even lower biomass than TAC
under moderate measurement uncertainty.

Figure \ref{sims} confirms that the precautionary approach represented
by the TAC does indeed prove more robust to the problem of measurement
error than the optimization solution represented by CE. In general, we
expect that the rigid assumptions required of optimization will lead to
worse outcomes when those assumptions are not met than we see under the
corresponding heuristic approach. Yet it is important to bear in mind
that the opposite pattern was observed in the case of environmental
stochasticity, in which it was the heuristic TAC strategy rather than
the CE strategy that lead to the highest over-fishing rate. In contrast,
the POMDP approach successfully handles both stochasticity and
observation error, maintaining the stock near \(B_{MSY}\) levels that
produce the highest yield.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/econ-1.pdf}
\caption{Expected net present value of the fishery across strategies
under increasing levels of measurement error \(\sigma_m\). Net present
value is th sum of all future harvests discounted over time and averaged
across 100 replicates simulations under each strategy, as shown in
Figure \ref{sims}. The value under constant escapement (CE) is optimal
when measurements are perfect but decreases rapidly with increasing
measurement error. The more cautious TAC is not economically optimal but
largely unimpacted by increasing error, while POMDP attains consistently
high economic yield despite the increasing uncertainty. \label{econ}}
\end{figure}

This pattern in management success in ecological terms (the relatively
recovery and maintenance of fish biomass) is also borne out in terms of
economic performance. Figure \ref{econ} shows the mean net present value
(averaging across replicates and discounting future profits by the
discount factor \(\gamma\)) for these same simulations at increasing
levels of measurement uncertainty. As before, environmental
stochasticity is set at \(\sigma_g = 0.15\). In the absence of
measurement uncertainty, constant escapement (CE) is optimal (as per
Reed (1979)), while the heuristic caution built into the total allowable
catch (TAC) strategy (at 80\% MSY) results in a sub-optimal economic
yield. The presence of stochasticity in recruitment
(\(\sigma_g = 0.15\)), also contributes to the reduction in yield under
TAC. Though CE is quite robust to the environmental stochasticity level,
this strategy proves very sensitive to increasing measurement error,
showing sharp declines in economic value, falling below the TAC economic
value at \(\sigma_m = 0.15\). Though the risk of stock collapse does
increase with increasing measurement error under TAC (as seen by the
mean declines in Figure \ref{sims}), these have little impact on the
economic value due to the discount rate. A smaller discount rate would
penalize unlikely but not improbable collapses more, since a significant
amount of time is required to realize those rare events. Supplemental
Figure S3 summarizes these economic trends across different noise values
\(\sigma_g\) and includes comparison to a simple MSY policy. In contrast
to the declining economic performance of CE and the consistently
sub-optimal economic yield for TAC, the POMDP strategy continues to
generate economic yields at approximately the optimal level (as attained
by CE in the absence of measurement error) despite the increasingly
uncertain measurements. This demonstrates that the reduced risk of stock
collapse and higher average stock biomass attained by the POMDP strategy
in the presence of high measurement error, Figure \ref{sims}) is not the
result of a trivial reduction in harvesting under all circumstances, but
rather, evidence of a more nuanced strategy that manages to account for
the uncertainty in measurement while maintaining a reasonable harvest.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/overest-1.pdf}
\caption{Average fish biomass when the POMDP strategy assumes a high
level of measurement uncertainty, while simulations reflect perfect
measurements. Despite this overestimation, biomass under POMDP strategy
closely tracks the optimal biomass under CE in the absence of
measurement error. \label{overest}}
\end{figure}

The POMDP strategy is also robust to overestimation of the level of
measurement uncertainty. In the simulation results shown in Figures
\ref{sims} and \ref{econ}, we have assumed that the level of measurement
uncertainty \(\sigma_m\) was known, and saw that ignoring this
uncertainty (as the CE policy does) has significant negative impact on
ecological and economic outcomes. If the level of measurement
uncertainty is not known precisely, \emph{overestimating} the
measurement error while using the POMDP strategy provides a
precautionary approach that can nevertheless achieve nearly optimal
ecological and economic outcomes. Figure \ref{overest} summarizes the
results of the same simulations as before, but under the scenario in
which the POMDP approach assumes a measurement error of
\(\sigma_m = 0.15\), when in fact all simulated measurements are made
without error (\(\sigma_m = 0\)). This represents an extreme case of
overestimating the measurement error. Stochastic growth remains the same
as before, \(\sigma_g = 0.15\), and simulations under TAC and CE
policies are shown for comparison. When measurement error is absent,
Reed's proofs hold and the CE strategy is optimal. Figure \ref{overest}
shows that the POMDP outcomes track almost exactly the CE outcomes
despite the misplaced assumption that measurements are quite poor. (As
we have already seen, the TAC strategy is insufficiently cautious for
this level of stochasticity in recruitment, resulting in
over-exploitation and long-term decline). The ability of the POMDP
solution to perform nearly optimally even when significantly
overestimating the level of measurement uncertainty contrasts sharply to
the significant declines from ignoring measurement uncertainty seen in
the CE solutions in Figure \ref{sims}. This demonstrates that while we
may not know precisely the level of measurement error, we achieve far
better outcomes overestimating measurement uncertainty than
underestimating it. This also underscores the observation that POMDP
policy is quite robust to the details of the uncertainty. We can get a
better understanding for the performance of POMDP in these simulations
by looking more closely at how any individual decision under a POMDP
strategy compares to the action chosen by the current alternatives.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/policy-1.pdf}
\caption{Comparison of the harvest and escapement policies under each
strategy as a function of the observed stock size. Escapement refers to
the expected fraction of fish left in the sea, \(x-h\), while harvest
refers to the target catch; two different conventions for plotting the
same action. Uniquely, the POMDP policy plot will depend not only on the
observed stock size, but also depends on prior information (bottom
panel) as determined from any prior observations and actions. Depending
on this prior it may harvest less or more than the other policies given
an identical observation.\label{policy}}
\end{figure}

To better understand the differences in performance of these strategies
and resolve the paradox of uncertainty, we must take a closer look at
how the specific action recommended by each policy compares given the
same observation. Figure \ref{policy} shows the action taken by each
strategy in response to a measurement of the stock size (biomass
estimate). These plots show, for any possible observations, which
strategy will attempt to harvest most and which will attempt to harvest
least. For comparison purposes, we plot policies both in terms of
expected escapement, \(S = Y - H\) as is typical in the optimal control
literature, (e.g. Reed 1979; Clark and Kirkwood 1986; Sethi et al.
2005), and also directly in terms of the harvest quota \(H\). These
plots clearly illustrate the contrast between the constant escapement
(CE) strategy and the precautionary Total Allowable Catch (TAC)
strategy: CE sets harvest strictly to zero for stocks estimated at
biomass below \(B_{MSY}\), while TAC permits a modest harvest of even
very small stocks. (Comparable plots with MSY can be found in the
Supplemental Material). In principle, the CE policy could depend on
\(\sigma_g\), but as Reed (1979) proved, the constant escapement level
with stochasticity is the same as in the deterministic case, S = D,
unless the noise level is quite high (comparison plots in Supplemental
material.) In contrast to this, our POMDP plots depend very much on both
the choice of \(\sigma_m\) and \(\sigma_g\). If \(\sigma_m = 0\), they
reduce exactly to the CE solution. Here we show the corresponding
policies for POMDP solutions focusing on \(\sigma_g = 0.15\) as above,
with a modest measurement uncertainty \(\sigma_m = 0.1\). Alternate
combinations of measurement uncertainty can be found in the appendix but
do not change the general pattern. In addition to an explicit dependence
on the measurement error, our POMDP solutions depend on another piece of
information: any prior observations of the stock size.

Differences in attention to prior beliefs,determined by prior
observations, drive the differences between the POMDP strategy and the
other strategies and can resolve the uncertainty paradox. While the
catch quota under both TAC and CE strategies can be completely
determined given the most recent observation of the stock size by using
the policy curves shown in Figure \ref{policy}, this is not the case for
the POMDP approach. This fundamental difference is key to understanding
the difference in performance and resolving the paradox of uncertainty.
The POMDP policy cannot be specified by the most recent observation
alone. Instead, the POMDP policy depends on all prior observations, not
just the most recent. The reason for this complexity comes from the
Markov property. Observations of the state in the perfectly observed
system satisfy the Markov property: once we have measured the current
biomass exactly, we cannot get any better estimate of the current stock
size by studying older measurements. When measurements are uncertain
this is no longer the case: intuitively, by comparing the most recent
measurement to previous observations we may be able to infer when any
given measurement is unusually high or unusually low. POMDP formalizes
this intuition by capturing the information from all previous
observations into a \emph{prior belief}. This prior belief is updated
after every subsequent action and observation in accordance with Bayes
Law. The mechanics of this process are well documented in the extensive
literature on POMDPs (e.g. Smallwood and Sondik 1973; Sondik 1978;
Kurniawati et al. 2008; Williams 2011), but for our purposes it is
sufficient to observe how this evolving prior belief serves to
continually adjust the POMDP policy. Supplemental Figure S9 (Appendix A)
illustrates how the prior belief evolves in response to subsequent
observations and actions over the course of an individual simulation.

Figure \ref{policy} shows two separate policy curves (in terms of
harvest and escapement) for the same POMDP solution given two different
prior belief distributions (panel 3, priors). While both priors shown
express considerable uncertainty about the precise stock size prior to
the most recent observation, the lower prior is centered at a value
\(\tfrac{1}{4}K\) stock size, while the high prior is centered at a
value of \(\tfrac{3}{4}K\), relative to a (post-harvest, before
observation) target size of \(B_{MSY} = K/2\). The POMDP policy is
determined by the combination of this prior information and the most
recent observation, as indicated by the two different POMDP curves for
harvest/escapement shown corresponding to the different priors. As with
the other policies, the higher the most recent observation (x axis) the
higher the POMDP recommended harvest. Yet unlike the alternative
strategies, the POMDP solution always reflects the prior information.
Consequently, relative to constant escapement (that is, no measurement
error), the POMDP with low prior starts harvesting only at higher stock
sizes and always harvests less. In contrast under the high prior, the
POMDP always harvests at the same or higher level than the constant
escapement solution.

\textbf{A resolution to the paradox.}

Herein lies our resolution to the paradox of uncertainty. Previous work
created this paradox by suggesting that increased harvest rates
(decreased target escapement) would often be the rational response to
increased uncertainty. The exact solutions from the POMDP reveal that
this is only an accident of the assumptions: it is indeed true that
under certain circumstances, harvest levels should increase relative to
the case of no uncertainty, but \emph{only when prior knowledge suggests
the stock size should be much higher than the most recent estimate would
suggest}. In the POMDP solution, all information must be put into its
historical (and constantly updated) context. When a measurement roughly
matches the expectation of this prior context, Figure \ref{policy} shows
that the optimal response from POMDP is roughly comparable or slightly
more cautious to the harvest under no uncertainty, and not more
aggressive as the paradox would suggest. Measurements that exceed
expectations are tempered with some skepticism: while the CE solution is
willing to meet a high stock measurement with a large harvest, the POMDP
solutions increase harvest more cautiously.

This difference between underestimating and overestimating accounts for
the poor performance of the CE solution under large measurement
uncertainty, where it over-harvests whenever measurements are too large.
Even though underestimating is equally likely under the measurement
uncertainty model, sooner or later a run of ``heads'', a sequence of
overestimations relative to the true stock, can drive stocks to very low
levels where the chance of local extinction becomes possible. It is
precisely this asymmetry: that too much over-harvesting leads to an
irreversible state of extinction, while too much under-harvesting is
always reversible (modulo some lost revenue) that lay behind Clark's
original intuition that there was something fishy about Reed's result
that \(S = D\): that uncertainty required no extra caution. Constant
escapement is particularly susceptible to over-harvesting starting from
stock sizes much higher than \(B_{MSY}\) since its bang-bang optimal
solution attempts to bring stocks back to target level as fast as
possible. CE is not so vulnerable to collapse from small stock sizes,
since it shuts down all harvests once estimates fall below \(B_{MSY}\).
Stock collapse under TAC can also be driven by such a string of heads,
but is unlikely at high stock sizes since harvest never exceeds
\(H_{MSY}\). In both cases, some measurement uncertainty interacts with
inherent stochasticity, which provides a continued source of variation
to sizes above \(B_{MSY}\), where CE strategy is most vulnerable, and
below \(B_{MSY}\), where TAC is most vulnerable.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/historical-1.pdf}
\caption{Comparisons of harvest level that would be recommended by
policies considered here relative to historical harvest levels in two
commercial fish stocks. POMDP solution assumes a measurement error of
10\%. \label{historical}}
\end{figure}

\textbf{Historical examples and implementation}.

So far we have focused on simulation and an examination of the relative
policies under different prior information. Figure \ref{historical}
compares the harvest level that would have been recommended by each of
the strategies we have considered here against the historically observed
catch recorded for two commercially fished stocks: Argentine Hake and
European Plaice. Historical estimates of biomass and catch are taken
from the R.A. Myers Legacy Stock Assessment database (Ricard et al.
2011). Posterior distributions for parameters for the Gordon-Schaefer
model are estimated with uncertainty through Markov Chain Monte Carlo
using \texttt{nimble} (de Valpine et al. 2017) on the historical data,
illustrating how this process might be done in more complex models as
well (details and code in Supplement, Appendix B). \(B_{MSY}\) and
corresponding TAC and CE policies are calculated based on posterior mean
estimates of the model parameters, along with POMDP solution assuming a
10\% measurement error rate. Though measurement error could be estimated
directly from the raw data, this would not reflect the true measurement
uncertainty arising from the stock assessment process. Each strategy is
then compared to historical observations to determine a recommended
harvest. As we have seen, the TAC and CE harvest policies are uniquely
determined by the observed stock size relative to \(B_{MSY}\), but the
POMDP policy must be re-calculated each time step to reflect both the
prior observation and prior action.

These historical examples provide another useful lens to compare how the
different strategies respond in the face of fluctuations in real data,
rather than model simulations. In both stocks, historical catch almost
always exceeds that recommended by any of our strategies, falling closer
to the MSY value estimated (see Appendix B). In Argentine Hake, the
persistent declines result in both POMDP and CE strategies quickly
closing down the fishery in an attempt to let the stock recover to a
more productive target biomass, while TAC persists with merely a reduced
harvest. A small recovery of biomass in 1995 is met by an immediate
uptick in harvest under CE, while the POMDP response is more
conservative. The example of European Plaice illustrates more volatility
in the stock size, revealing further differences in the strategies.
Despite this volatility, the TAC level remains relatively level across
all five decades, rising or dipping only slightly with changes in stock
size. In contrast, the CE harvest tracks this volatility almost exactly.
Here, the POMDP solution falls in between these extremes, almost
mirroring which-ever of the two policies is more conservative: the POMDP
solution matches the dips in harvest taken by CE to allow the stock to
recover most quickly, but does not track the doubling of harvests
recommended by CE in the mid 1980, increasingly only to the more modest
harvest recommended by the TAC policy. Once again, we see that POMDP
solution provides a more consistently precautionary policy than either
the over-simplified optimal control solution of CE or the more
rule-of-thumb approach represented by TAC. Though we have seen that the
POMDP solution need not always be more conservative (lower harvest) than
these strategies, it makes use of the all prior observations to tune the
level of caution appropriately.

\section{Discussion}\label{discussion}

Using modern algorithms we have been able to crack the nut of
measurement uncertainty in the optimal management of marine fisheries
and debunk long-standing paradoxes that increased measurement
uncertainty should be met with larger harvests. We have demonstrated
that both existing optimal control approaches and existing precautionary
approaches result in over-exploitation and long term declines of fish
stocks, as well as reduced economic value, while in contrast, the POMDP
approach is able to deliver both ecological and economic recovery
despite uncertainty in measurements. These results have important
implications for both global fisheries and ecological management more
generally. In marine fisheries, evidence of such over-exploitation has
already been well documented (Worm et al. 2006, 2009; Costello et al.
2016). Our results suggest that measurement uncertainty may contribute
to this pattern, as such errors lead to over-exploitation under current
best practices such as MSY, TAC or CE, which have previously been
thought to be sufficient to ensure the recovery of stocks world-wide
(Costello et al. 2016).

Our results also have important consequences for ecological management
more generally. The study and application of management strategies such
as MSY and CE underpin much of conservation and natural resource
management in forest management and many other terrestrial ecosystems
(e.g. Clark and Mangel 2000). Descision making under uncertainty is a
central challenge for ecology and conservation biology (Polasky et al.
2011), which we have long attempted to address through either simplified
optimal control models or more heuristic approaches such as the
precautionary principle and resilience thinking (Fischer et al. 2009).
We have illustrated how both of these approaches can fail to provide an
adequate strategy when faced with measurement uncertainty. To avoid this
pitfall, a strategy must be dynamic to make best use of the available
information without assuming that information is perfect. The
simplisitic optimal control approach of CE fails by placing too much
confidence in measurements even when they are significantly at odds with
prior expectations. Similarly, the TAC approach also fails due to it's
reliance on a simple one-size-fits-all strategy, not reducing catch
quotas sufficiently when stocks are too low but also failing to increase
them when stocks are higher. It is precisely this ability to adapt the
policy in response to available information, as illustrated in Figure
\ref{policy}, that makes the POMDP approach successful. Nevertheless, it
may be tempting to see these results as either trivial or impractical,
but a closer inspection reveals they are neither.

First, it may seem obvious that POMDP would out-preform more primitive
optimal control solutions such as CE, but this was hardly a forgone
conclusion. Indeed, the previous example of Reed (1979) had proven that
the complex stochastic dynamic programming solutions required to
accommodate stochastic dynamics would perform no better than the simple
maximization based purely on the deterministic model. If stochasticity
doesn't matter, it is easy to believe that measurement error would be
minor at best. In fact, we find that once we account for measurement
error, the effects of both stochasticity and measurement have a
significant influence over the optimal policy. It is similarly tempting
to dismiss the better performance of POMDP against a more
rule-of-thumb{[}\^{}1{]} based strategy like the TAC scenario considered
here (fishing at 80\% MSY), but once again this was not a forgone
conclusion. While a better economic return might reasonably be expected
of the optimization approach, it was not obvious that POMDP would also
be at times more cautious than the rule of thumb, able to deliver both
higher economic returns by harvesting more when it was safe to do so,
and higher ecological recovery by harvesting even less than TAC when
necessary. These results highlight the difficulty of finding effective
heuristic rules and underscore the power and utility of more
sophisticated strategies such as POMDP which can make the best use of
all available information.

\^{}1: Given the sophisticated policies often in place to determine
actual TAC levels, one might object to our characterization of this as
merely a rule-of-thumb. We use this term merely to distinguish it from
the optimal control approach -- that despite technical precision, TAC
adjustments are set by a heuristic instead of the result of some dynamic
optimization algorithm as in CE and POMDP.

Second, we must ask if the policies derived from POMDP are practical. As
with constant escapement, the POMDP solutions we considered here result
in more year-to-year variation in harvest, and include the possibility
of fishing moratorium to allow stocks to recover -- decisions that many
view as economically or socially unacceptable. While resolving this
debate is beyond the scope of this paper, a few observations are
relevant: (A) Stocks already managed under constant escapement, such as
US salmon fisheries, are already subject to the possibility of these
moratoriums. (B) It is straight-forward to introduce constraints on the
minimal acceptable harvest level and/or include the additional costs of
closing a fishery in the utility function optimized by POMDP. (C)
Fisheries today are not managed at either their maximum ecological or
economic potential. Taken as a whole, most ocean fisheries yield a net
negative economic return (World Bank and FAO 2009), with fishing
persisting only with the support of government subsidies (Arnason 2012),
meanwhile some 68\% of global fisheries are in poor biological condition
(Costello et al. 2016). While the causes for this are both complex and
numerous, we have seen that both the economic inefficiency and
ecological health suffer under TAC/MSY and CE based management. Arnason
(2012) and Costello et al. (2016) have both argued that a rights-based
approach to fisheries management (RBFM) would lead to self adoption of
precisely such optimal policies.

Third: are the examples considered here too simple? The strategies we
have compared against underpin both current practice and management
aspirations. Together, MSY and TAC-based management reflects much of
modern best practice as it is understood in fisheries management and
elsewhere. While real-world application of these approaches frequently
involves more complex models, the models we have considered here are the
very ones under which these strategies were designed and tested, and
there is little reason to believe they would perform better under more
complex cases such as age-structured models that do not always meet
those assumptions (e.g. Holden and Conrad 2015). The relatively poor
ecological and economic performance of these common best-practices
relative to the alternative approach of POMDP is thus reason enough for
concern. This issue is further underscored by the reality of both
long-running ecological deterioration (Worm et al. 2006, 2009) and
economic costs of global fisheries.

Fourth: is the approach taken here too complex to be feasible? The
importance of simple and effective rules of thumb in conservation
management has been well documented (e.g. Chadès et al. 2011).
Sophisticated and computationally intensive approaches such as POMDP may
seem improbable at a time when many areas of natural resource
management, even simple, rule-of-thumb based methods struggle to take
root. And yet, in this same age we rely on far more complex and
computationally intensive algorithms to perform far more trivial tasks
such as determining what advertisements we see. Surely a planet that can
afford such complex optimization for how advertising space is allocated
on a small screen can one day manage it's natural resources and
conservation efforts with such tools?

\subsection{Future directions}\label{future-directions}

Several limitations that have been studied in fully observed (MDP)
optimal decision problems, such as parameter uncertainty (e.g. Ludwig
and Walters 1982), model uncertainty (Williams 2001; Boettiger et al.
2015) and adaptive management (e.g. Walters and Hilborn 1976) remain
largely open challenges for partially observed systems. Future work
could extend this analysis to more complex models, such as those with
age structure, as (Holden and Conrad 2015) does for the fully observed
case. Another limiting assumption common to MDPs and POMDPs is that of
stationary dynamics: that the population dynamics equation itself is not
changing over time. In reality, forces such as climate change and other
forms of environmental variations violate this assumption (Britten et
al. 2017). Direct approaches such as Fackler and Pacifici (2014)'s
adaptation of Mixed Observability MDP (Ong et al. 2010) do not scale to
the number of states and actions considered here. A value of information
(VOI) analysis for POMDP (e.g. Johnson and Williams 2015; Memarzadeh and
Pozzi 2016), could identify when it is worthwhile to actively reduce
measurement error.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-Arnason2012}{}
Arnason, R. 2012. Property rights in fisheries: How much can individual
transferable quotas accomplish? Review of Environmental Economics and
Policy 6:217--236.

\hypertarget{ref-Beverton1957}{}
Beverton, R., and S. Holt. 1957. On the Dynamics of Exploited Fish
Populations. Chapman; Hall, London.

\hypertarget{ref-Boettiger2015}{}
Boettiger, C., M. Mangel, and S. Munch. 2015. Avoiding tipping points in
fisheries management through Gaussian process dynamic programming.
Proceedings of the Royal Society B: Biological Sciences
282:20141631--20141631.

\hypertarget{ref-sarsop-pkg}{}
Boettiger, C., J. Ooms, and M. Memarzadeh. 2018. Sarsop: Approximate
pomdp planning software.

\hypertarget{ref-Botsford1997}{}
Botsford, L. W., J. C. Castilla, and C. Peterson. 1997. The management
of fisheries and marine ecosystems. Science 277:509--515.

\hypertarget{ref-Britten2017}{}
Britten, G., M. Dowd, L. Kanary, and B. Worm. 2017. Extended fisheries
recovery timelines in a changing environment. Nature Communications.

\hypertarget{ref-Chades2011}{}
Chadès, I., T. G. Martin, S. Nicol, M. A. Burgman, H. P. Possingham, and
Y. M. Buckley. 2011. General rules for managing and surveying networks
of pests, diseases, and endangered species. Proceedings of the National
Academy of Sciences 108:8323--8.

\hypertarget{ref-Chades2008}{}
Chadès, I., E. McDonald-Madden, M. a McCarthy, B. Wintle, M. Linkie, and
H. P. Possingham. 2008. When to stop managing or surveying cryptic
threatened species. Proceedings of the National Academy of Sciences
105:13936--40.

\hypertarget{ref-Clark1973}{}
Clark, C. W. 1973. Profit maximization and the extinction of animal
species. Journal of Political Economy 81:950--961.

\hypertarget{ref-Clark1990}{}
Clark, C. W. 1990. Mathematical Bioeconomics: The Optimal Management of
Renewable Resources, 2nd Edition. Wiley-Interscience.

\hypertarget{ref-Clark1986}{}
Clark, C. W., and G. P. Kirkwood. 1986. On uncertain renewable resource
stocks: Optimal harvest policies and the value of stock surveys. Journal
of Environmental Economics and Management 13:235--244.

\hypertarget{ref-Clark2000}{}
Clark, C. W., and M. Mangel. 2000. Dynamic state variable models in
ecology. Oxford University Press, Oxford.

\hypertarget{ref-Costello2016}{}
Costello, C., D. Ovando, T. Clavelle, C. K. Strauss, R. Hilborn, M. C.
Melnychuk, T. A. Branch, et al. 2016. Global fishery prospects under
contrasting management regimes. Proceedings of the National Academy of
Sciences 113:5125--5129.

\hypertarget{ref-nimble}{}
de Valpine, P., D. Turek, C. J. Paciorek, C. Anderson-Bergman, T.
Duncan, and R. Bodik. 2017. Programming With Models: Writing Statistical
Algorithms for General Model Structures With NIMBLE. Journal of
Computational and Graphical Statistics 26:403--413.

\hypertarget{ref-Dichmont2016}{}
Dichmont, C. M., R. A. Deng, A. E. Punt, J. Brodziak, Y.-J. Chang, J. M.
Cope, J. N. Ianelli, et al. 2016. A review of stock assessment packages
in the united states. Fisheries Research 183:447--460.

\hypertarget{ref-Engen1997}{}
Engen, S., R. Lande, and B. Sæther. 1997. Harvesting strategies for
fluctuating populations based on uncertain population estimates. Journal
of Theoretical Biology 186:201--212.

\hypertarget{ref-Fackler2014b}{}
Fackler, P., and R. Haight. 2014. Monitoring as a partially observable
decision problem. Resource and Energy Economics 37:226--241.

\hypertarget{ref-Fackler2014}{}
Fackler, P., and K. Pacifici. 2014. Addressing structural and
observational uncertainty in resource management. Environmental
Management 133:27--36.

\hypertarget{ref-Fischer2009}{}
Fischer, J., G. D. Peterson, T. a Gardner, L. J. Gordon, I. Fazey, T.
Elmqvist, A. Felton, et al. 2009. Integrating resilience thinking and
optimisation for conservation. Trends in ecology \& evolution
24:549--54.

\hypertarget{ref-Gordon1954}{}
Gordon, H. S., and C. Press. 1954. The Economic Theory of a
Common-Property Resource: The Fishery. Journal of Political Economy
62:124--142.

\hypertarget{ref-Hilborn2010}{}
Hilborn, R. 2010. Pretty Good Yield and exploited fishes. Marine Policy
34:193--196.

\hypertarget{ref-Holden2015}{}
Holden, M., and J. Conrad. 2015. Optimal escapement in stage-structured
fisheries with environmental stochasticity. Mathematical biosciences
269:76--85.

\hypertarget{ref-Holling1973}{}
Holling, C. S. 1973. Resilience and Stability of Ecological Systems.
Annual Review of Ecology and Systematics 4:1--23.

\hypertarget{ref-Johnson2015}{}
Johnson, F. A., and B. K. Williams. 2015. A Decision-Analytic Approach
to Adaptive Resource Management. Pages 61--84 \emph{in} C. R. Allen and
A. S. Garmestani, eds. Adaptive management of social-ecological systems.
Springer Netherlands, Dordrecht.

\hypertarget{ref-Kaelbling1998}{}
Kaelbling, L. P., M. L. Littman, and A. R. Cassandra. 1998. Planning and
Acting in Partially Observable Stochastic Domains. Artificial
Intelligence 101:99--134.

\hypertarget{ref-Kurniawati2008}{}
Kurniawati, H., D. Hsu, and W. S. Lee. 2008. SARSOP : Efficient
Point-Based POMDP Planning by Approximating Optimally Reachable Belief
Spaces. Proceedings of Robotics: Science and Systems IV.

\hypertarget{ref-Larkin1977}{}
Larkin, P. a. 1977. An Epitaph for the Concept of Maximum Sustained
Yield. Transactions of the American Fisheries Society 106:1--11.

\hypertarget{ref-Levin2008}{}
Levin, S. A., and J. Lubchenco. 2008. Resilience, Robustness, and Marine
Ecosystem-based Management. BioScience 58:27--32.

\hypertarget{ref-Ludwig1981}{}
Ludwig, D., and C. Walters. 1981. Measurement errors and uncertainty in
parameter estimates for stock and recruitment. Journal of Canadian
Fisheries and Aquatic Sciences 38:711--720.

\hypertarget{ref-Ludwig1982}{}
Ludwig, D., and C. J. Walters. 1982. Optimal harvesting with imprecise
parameter estimates. Ecological Modelling 14:273--292.

\hypertarget{ref-Mace2001}{}
Mace, P. M. 2001. A new role for MSY in single-species and
ecosystem\(\backslash\)rapproaches to fisheries stock assessment and
management. Fish and Fisheries 2:2--32.

\hypertarget{ref-Mangel1988}{}
Mangel, M., and C. W. Clark. 1988. Dynamic Modeling in Behavioral
Ecology. (J. Krebs \& T. Clutton-Brock, eds.). Princeton University
Press, Princeton.

\hypertarget{ref-Marescot2013}{}
Marescot, L., G. Chapron, I. Chadès, P. L. Fackler, C. Duchamp, E.
Marboutin, and O. Gimenez. 2013. Complex decisions made simple: a primer
on stochastic dynamic programming. Methods in Ecology and Evolution
n/a--n/a.

\hypertarget{ref-May1977}{}
May, R. M. 1977. Thresholds and breakpoints in ecosystems with a
multiplicity of stable states. Nature 269:471--477.

\hypertarget{ref-Memarzadeh2016b}{}
Memarzadeh, M., and M. Pozzi. 2016. Value of information in sequential
decision making: component inspection, permanent monitoring and
system-level scheduling. Reliability Engineering \& System Safety
154:137--151.

\hypertarget{ref-Moxnes2003}{}
Moxnes, E. 2003. Uncertain measurements of renewable resources:
approximations, harvesting policies and value of accuracy. Journal of
Environmental Economics and Management 45:85--108.

\hypertarget{ref-Ong2010}{}
Ong, S., S. Png, D. Hsu, and W. Lee. 2010. Planning under uncertainty
for robotic tasks with mixed observability. The International Journal of
Robotics Research 29:1053--1068.

\hypertarget{ref-Petchy2015}{}
Petchey, O. L., M. Pontarp, T. M. Massie, S. Kéfi, A. Ozgul, M.
Weilenmann, G. M. Palamara, et al. 2015. The ecological forecast
horizon, and examples of its uses and determinants. Ecology Letters
18:597--611.

\hypertarget{ref-Pineau2003}{}
Pineau, J., G. Gordon, and S. Thrun. 2003. Point-based value iteration:
An anytime algorithm for POMDPs. IJCAI International Joint Conference on
Artificial Intelligence 1025--1030.

\hypertarget{ref-Polasky2011}{}
Polasky, S., S. R. Carpenter, C. Folke, and B. Keeler. 2011.
Decision-making under great uncertainty: environmental management in an
era of global change. Trends in ecology \& evolution 1--7.

\hypertarget{ref-Reed1979}{}
Reed, W. J. 1979. Optimal escapement levels in stochastic and
deterministic harvesting models. Journal of Environmental Economics and
Management 6:350--363.

\hypertarget{ref-RAM}{}
Ricard, D., C. Minto, O. Jensen, and J. Baum. 2011. Examining the
knowledge base and status of commercially exploited marine species with
the RAM Legacy Stock Assessment Database. Fish and Fisheries
13:380--398.

\hypertarget{ref-Roughgarden1996}{}
Roughgarden, J. E., and F. Smith. 1996. Why fisheries collapse and what
to do about it. Proceedings of the National Academy of Sciences 93:5078.

\hypertarget{ref-Schaefer1954}{}
Schaefer, M. B. 1954. Some aspects of the dynamics of populations
important to the management of the commercial marine fisheries. Bulletin
of the Inter-American Tropical Tuna Commission 1:27--56.

\hypertarget{ref-Sethi2005}{}
Sethi, G., C. Costello, A. Fisher, M. Hanemann, and L. Karp. 2005.
Fishery management under multiple uncertainty. Journal of Environmental
Economics and Management 50:300--318.

\hypertarget{ref-Smallwood1973}{}
Smallwood, R. D., and E. J. Sondik. 1973. The Optimal Control of
Partially Observable Markov Decision Processes over a Finite Horizon.

\hypertarget{ref-Sondik1978}{}
Sondik, E. J. 1978. The Optimal Control of Partially Observable Markov
Processes Over the Infinite Horizon : Discounted Costs. Operations
Research 26:282--304.

\hypertarget{ref-Walters1976}{}
Walters, C. J., and R. Hilborn. 1976. Adaptive Control of Fishing
Systems. Journal of the Fisheries Research Board of Canada 33:145--159.

\hypertarget{ref-Walters1978}{}
---------. 1978. Ecological Optimization and Adaptive Management. Annual
Review of Ecology and Systematics 9:157--188.

\hypertarget{ref-Williams2001}{}
Williams, B. K. 2001. Uncertainty , learning , and the optimal
management of wildlife. Environmental and Ecological Statistics
8:269--288.

\hypertarget{ref-Williams2011}{}
---------. 2011. Resolving structural uncertainty in natural resources
management using POMDP approaches. Ecological Modelling 222:1092--1102.

\hypertarget{ref-FAO2009}{}
World Bank and FAO. 2009. The Sunken Billions. The Economic
Justification for Fisheries Reform.

\hypertarget{ref-Worm2006}{}
Worm, B., E. B. Barbier, N. Beaumont, J. E. Duffy, C. Folke, B. S.
Halpern, J. B. C. Jackson, et al. 2006. Impacts of biodiversity loss on
ocean ecosystem services. Science (New York, N.Y.) 314:787--90.

\hypertarget{ref-Worm2009}{}
Worm, B., R. Hilborn, J. K. Baum, T. A. Branch, J. S. Collie, C.
Costello, M. J. Fogarty, et al. 2009. Rebuilding global fisheries.
Science (New York, N.Y.) 325:578--85.

\end{document}

### 2018-02-01 manuscripts/manuscript.Rmd
fixing typos

@@
   Ecological management and decision-making typically focus on uncertainty about the future, but surprisingly little is known about how to account for uncertainty of the present: that is, the realities of having only partial or imperfect measurements.
-Our two primary paradigms for handling decisions under uncertainty -- the precautionary principle and optimal control -- have so far given entirely controdictory results.
+Our two primary paradigms for handling decisions under uncertainty -- the precautionary principle and optimal control -- have so far given entirely contradictory results.
 This paradox is best illustrated in the example of fisheries management, where many ideas that guide our thinking about ecological decision making were first developed.

### 2018-02-01 manuscripts/manuscript.tex
fixing typos

@@
   handling decisions under uncertainty -- the precautionary principle and
-  optimal control -- have so far given entirely controdictory results.
+  optimal control -- have so far given entirely contradictory results.
   This paradox is best illustrated in the example of fisheries management,
@@
 management in forest management and many other terrestrial ecosystems
-(e.g. Clark and Mangel 2000). Descision making under uncertainty is a
+(e.g. Clark and Mangel 2000). Decision making under uncertainty is a
 central challenge for ecology and conservation biology (Polasky et al.
@@
 information without assuming that information is perfect. The
-simplisitic optimal control approach of CE fails by placing too much
+simplistic optimal control approach of CE fails by placing too much
 confidence in measurements even when they are significantly at odds with

### 2018-02-01 manuscripts/manuscript.Rmd
220 word abstract only

@@
 Our two primary paradigms for handling decisions under uncertainty -- the precautionary principle and optimal control -- have so far given entirely contradictory results.
-This paradox is best illustrated in the example of fisheries management, where many ideas that guide our thinking about ecological decision making were first developed.
-There, we find that simplistic optimal control approaches have repeatedly concluded that a manager should increase catch quotas when faced with greater uncertainty about the standing stock biomass.
-Meanwhile, most current best practices take a more precautionary approach, decreasing catch quotas by a fixed amount to account for uncertainty.
+This paradox is best illustrated in the example of fisheries management, where many ideas that guide thinking about ecological decision making were first developed.
+We find that simplistic optimal control approaches have repeatedly concluded that a manager should increase catch quotas when faced with greater uncertainty about the fish biomass.
+Current best practices take a more precautionary approach, decreasing catch quotas by a fixed amount to account for uncertainty.
 Using comparisons to both simulated and historical catch data, we find that neither approach is sufficient to avoid stock collapses under moderate observational uncertainty.
-Using cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.
-In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning.
+Using partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.
+In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value.
 These results have significant implications for how we handle uncertainty in both fisheries and ecological management more generally.
-Dealing with uncertainty in decisions requires approaches which are both adaptive and explicit about uncertainty, such as POMDPs.
 

### 2018-02-01 manuscripts/manuscript.tex
220 word abstract only

--- 2018-02-01 manuscripts/manuscript.tex
@@
   handling decisions under uncertainty -- the precautionary principle and
-  optimal control -- have so far given entirely controdictory results.
+  optimal control -- have so far given entirely contradictory results.
   This paradox is best illustrated in the example of fisheries management,
-  where many ideas that guide our thinking about ecological decision
-  making were first developed. There, we find that simplistic optimal
-  control approaches have repeatedly concluded that a manager should
-  increase catch quotas when faced with greater uncertainty about the
-  standing stock biomass. Meanwhile, most current best practices take a
-  more precautionary approach, decreasing catch quotas by a fixed amount
-  to account for uncertainty. Using comparisons to both simulated and
-  historical catch data, we find that neither approach is sufficient to
-  avoid stock collapses under moderate observational uncertainty. Using
-  cutting-edge point-based partially observed Markov decision process
-  (POMDP) methods, we demonstrate how this paradox arises from flaws in
-  the standard management theory, which contributes to over-exploitation
-  of fisheries and increased probability of economic and ecological
-  collapse. In contrast, we find POMDP-based management avoids such
-  over-exploitation while also generating higher economic value through
-  adaptive learning. These results have significant implications for how
-  we handle uncertainty in both fisheries and ecological management more
-  generally. Dealing with uncertainty in decisions requires approaches
-  which are both adaptive and explicit about uncertainty, such as POMDPs.
+  where many ideas that guide thinking about ecological decision making
+  were first developed. We find that simplistic optimal control approaches
+  have repeatedly concluded that a manager should increase catch quotas
+  when faced with greater uncertainty about the fish biomass. Current best
+  practices take a more precautionary approach, decreasing catch quotas by
+  a fixed amount to account for uncertainty. Using comparisons to both
+  simulated and historical catch data, we find that neither approach is
+  sufficient to avoid stock collapses under moderate observational
+  uncertainty. Using partially observed Markov decision process (POMDP)
+  methods, we demonstrate how this paradox arises from flaws in the
+  standard theory, which contributes to over-exploitation of fisheries and
+  increased probability of economic and ecological collapse. In contrast,
+  we find POMDP-based management avoids such over-exploitation while also
+  generating higher economic value. These results have significant
+  implications for how we handle uncertainty in both fisheries and
+  ecological management more generally.
   \end{abstract}

### 2018-07-30 manuscripts/manuscript.Rmd
work in progress

@@
 How have invasive species transformed native communities?
-A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods.
+A recognition of the importance of uncertainty has become widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods.
 In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and under-served by available tools.
-In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these existing approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.
+We illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a range of existing approaches, which can be avoided through the use of more formal approaches that have since been refined in fields such as robotics.
 This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.
 
-Approaches for decision making under uncertainty in ecological systems can be divided into two camps: approaches based in "optimal control" and usually favored by natural resource economists, and approaches based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].
-While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.
-Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.
+Approaches for decision making under uncertainty in ecological systems can be divided into two camps: (1) those based upon "optimal control" solutions, and usually favored by natural resource economists, and (2) those based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].
+While researchers have recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have stymied this progress.
+We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.
+We compare the results of managing simulated fish stocks under (A) the existing optimal control approach, (B) a heuristic approach that is more typical of today's management practices, and (C) a newly proposed approach that provides a more rigorous treatment of uncertainty.
 
-Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.
-Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.
-In particular, the approach here illustrates the limitations of both over-simple optimal control approaches and rule-of-thumb precautionary approaches, and offers instead a more general strategy for dealing with measurement uncertainty adaptively.
-
-We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.
+The challenge of making decisions under uncertainty is obviously not unique to fisheries, and the existing methods we critique here can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010], and protecting biodiversity [@Dee2017].
+Yet fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.
+Despite the generality of the models and issue of measurement uncertainty considered here, any application must be grounded in both the history and context of the particular decision problem.
+In light of this, we begin with a brief background on existing decision theory approaches common in fisheries management.
 
@@
 We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.
-
 We seek to determine the sequence of actions $h_t$ for $t \in [1, ...
@@
 
-$$y_t = \xi_t x_t$$
+$$\begin{equation}
+y_t = \xi_t x_t 
+\end{equation}$$
 
@@
 
-$$\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$
+$$\begin{equation}
+\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
+\end{equation}$$
 
@@
 
-$$x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) $$
+$$\begin{equation}
+x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) 
+\end{equation}$$
 
@@
 [@Costello2016; @Britten2017]), and form the basis for much of bioeconomic theory [@Clark1990].
-As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arsing only because of some particular subtle assumption hidden in a more complex model.
+As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arising only because of some particular subtle assumption hidden in a more complex model.
 
 It is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
-Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
+Likewise, while we consider the model and its parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
 @Dichmont2016].
@@
 
+MSE
+
 ## Optimal management
@@
 (This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).
-This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.
-While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.
+This strategy is known as *constant escapement* (CE), since a constant sock size $B_{MSY}$ escapes harvest each year.
+Because CE and MSY converge to the *same long-term biomass and same long-term average yield*, CE can appear to be merely a more formal justification for MSY.
+However, in practice the two approaches are very different: becuase natural populations fluctuate year-to-year, under CE, a different harvest quota must be set each year to achieve the constant escapement.
+While theoretical results have proven that CE and not MSY is optimal for fluctuating populations [@Reed1979], real-world examples of CE-regulated fisheries are limited primarily to salmon.
+As we shall see, despite being insensitive to natural stochasticity, CE is very sensitive to measurement error -- thus it is perhaps not surprising that CE is not found in other marine fisheries where population sizes may be harder to estimate accurately.
+
 Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above.
-Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarize here.
+Instead, existing theory has presented only the paradoxical results that uncertainty either shoud not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty.
 
@@
 
-Unfortunately, @Clark1986 was unable to solve the resulting problem exactly, but had to adopt an almost equally troublesome assumption:
+@Clark1986 were unable to solve the resulting problem exactly, but had to adopt an almost equally troublesome assumption:
 
@@
 
-Unfortunately for @Clark1986, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:
+Unfortunately, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:
 
@@
 
-As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP).
+Why does the introduction of measurement uncertainty make the decision problem so much harder?
+Consider the thought experiment of managing a fish stock where in each of the previous four years stock assessments have put the population size at, say, 51, 54, 49, 51, (say, measured in percent of carrying capacity, $K$).
+Under a constant escapement target of 50, the harvest quota would be 1, 4, 0, 1, respectively.
+Note that this control rule depends only on the measurement for the year in question, which is assumed to be measured without error.
+If we then measure a stock size of 75, an assumption that this measurement is made without error would justify the sudden increase to a quota of 25.
+Yet if we admit the possibility of measurement error, this sudden bump starts to look very suspicious in light of all the previous observations.
+Perfect measurements let us exploit the Markov property of the model -- the most recent measurement tells us all we need to know.
+In contrast, the introduction of measurement uncertainty breaks this Makovian assumption: a rational decision maker would suspect this 75 to be an over-estimate based on prior observations.
+
+As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP) [@Marescot2013].
 The constant escapement (CE) strategy considered here is the result of such an MDP or SDP analysis.
 Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).
-The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process, but cannot be solved using SDP directly.
-The POMDP problem for fisheries question considered here can be summarized as follows:
+
+Precise definitions of both MDP and POMDP problems can be found in @Williams2011, or in classic work from the engineering and robotics community [e.g. @Smallwood1973 and @Sondik1978].
+
+The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process  The POMDP problem for fisheries question considered here can be summarized as follows:
 
@@
 In our context, this relationship is given by the Gordon-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.
+The action $a_t$ represents the harvest quota set for the fishery.
+Attempting to implement a quota the exceeds the true population size results in collapse of the fishery.
+As @Clark1986 eloquently argues, assuming that extinction is impossible regardless of fishing intensity would unreasonably bias the decision problem.
 
@@
 @Kaelbling1998; @Pineau2003].
-By adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
+By adapting more recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
 As the details for solving such problems numerically are already well documented we will not rehash them here, but instead provide a preliminary R package [@sarsop-pkg] which provides an implementation of the SARSOP algorithm and the fisheries POMDP problem, which can be used to replicate any of the results presented here and explore other variations.
@@
 Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement (CE), Total allowable catch (TAC = 80% MSY), and the proposed partially observed Markov decision process (POMDP) method.
+Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.
 Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.
@@
 
-Figure \ref{sims} shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
-Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.
-Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.
-In each panel shown in Figure \ref{sims} stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
-In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.
-The first panel of Figure \ref{sims} shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.
+We consider the average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
+Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
+In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).
+The introduction of stochastic growth has a significant negative impact on the TAC strategy, whith a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).
 Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
@@
 Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.
-
-Figure \ref{sims} confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.
-In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.
-Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC strategy rather than the CE strategy that lead to the highest over-fishing rate.
-In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.
+These results confirm that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE, but is less robust to environmental error.
+In contrast, the POMDP approach successfully handles both sources of uncertainty, maintaining higher stocks, near $B_{MSY}$ level, and producing the highest yield.
 
 ```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$.
-Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{sims}.
+Net present value is the sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{sims}.
 The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error.
@@
 This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also borne out in terms of economic performance.
-Figure \ref{econ} shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.
+We find the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty (Figure \ref{econ} ).
 As before, environmental stochasticity is set at $\sigma_g = 0.15$.
@@
 
-Using modern algorithms we have been able to crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk long-standing paradoxes that increased measurement uncertainty should be met with larger harvests.
-We have demonstrated that both existing optimal control approaches and existing precautionary approaches result in over-exploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.
+Using modern algorithms we have been able to both crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk a long-standing paradox: that increased measurement uncertainty should be met with larger harvests.
+We have demonstrated that both existing optimal control approaches and existing precautionary approaches can result in over-exploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.
 These results have important implications for both global fisheries and ecological management more generally.
@@
 @Clark2000].
-Descision making under uncertainty is a central challenge for ecology and conservation biology [@Polasky2011], which we have long attempted to address through either simplified optimal control models or more heuristic approaches such as the precautionary principle and resilience thinking [@Fischer2009].
+Decision making under uncertainty is a central challenge for ecology and conservation biology [@Polasky2011], which we have long attempted to address through either simplified optimal control models or more heuristic approaches such as the precautionary principle and resilience thinking [@Fischer2009].
 We have illustrated how both of these approaches can fail to provide an adequate strategy when faced with measurement uncertainty.
 To avoid this pitfall, a strategy must be dynamic to make best use of the available information without assuming that information is perfect.
-The simplisitic optimal control approach of CE fails by placing too much confidence in measurements even when they are significantly at odds with prior expectations.
-Similarly, the TAC approach also fails due to it's reliance on a simple one-size-fits-all strategy, not reducing catch quotas sufficiently when stocks are too low but also failing to increase them when stocks are higher.
+Classical optimal control approaches (such as CE) can fail due to misplaced confidence in measurements even when they are significantly at odds with prior expectations.
+Similarly, the TAC approach also fails due to its reliance on a simple one-size-fits-all strategy, not reducing catch quotas sufficiently when stocks are too low but also failing to increase them when stocks are higher.
 It is precisely this ability to adapt the policy in response to available information, as illustrated in Figure \ref{policy}, that makes the POMDP approach successful.
@@
 Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accommodate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.
-If stochasticity doesn't matter, it is easy to believe that measurement error would be minor at best.
+If stochasticity does not matter, it is easy to believe that measurement error would be minor at best.
 In fact, we find that once we account for measurement error, the effects of both stochasticity and measurement have a significant influence over the optimal policy.
-It is similarly tempting to dismiss the better performance of POMDP against a more rule-of-thumb[^1] based strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.
+It is similarly tempting to dismiss the better performance of POMDP against a more heuristic strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.
 While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to deliver both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.
 These results highlight the difficulty of finding effective heuristic rules and underscore the power and utility of more sophisticated strategies such as POMDP which can make the best use of all available information.
-
-^1: Given the sophisticated policies often in place to determine actual TAC levels, one might object to our characterization of this as merely a rule-of-thumb.
-We use this term merely to distinguish it from the optimal control approach -- that despite technical precision, TAC adjustments are set by a heuristic instead of the result of some dynamic optimization algorithm as in CE and POMDP.
 
@@
 And yet, in this same age we rely on far more complex and computationally intensive algorithms to perform far more trivial tasks such as determining what advertisements we see.
-Surely a planet that can afford such complex optimization for how advertising space is allocated on a small screen can one day manage it's natural resources and conservation efforts with such tools?
+Surely a planet that can afford such complex optimization for how advertising space is allocated on a small screen can one day manage its natural resources and conservation efforts with such tools?
 
@@
 inefficiency and ecological health are hurt by current policies, including
-both MSY and it's more precautionary variants, as these all permit some level of harvest
+both MSY and its more precautionary variants, as these all permit some level of harvest
 below $B_{MSY}$.  Economists have for some time observed these inherent 

### 2018-07-30 scratch/plotting-experiments.Rmd
work in progress

---
title: "plotting experiments"
author: "Carl Boettiger"
date: "7/4/2018"
output: github_document
---

```{r}
library(tidyverse)
library(here)
library(fs)
```

```{r}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
#dev=c("cairo_pdf", "png")
library(Cairo)
library(hrbrthemes)
library(extrafont)
library(ggthemes)
hrbrthemes::import_plex_sans()
hrbrthemes::import_roboto_condensed()  
extrafont::loadfonts()
ggplot2::theme_set(hrbrthemes::theme_ipsum())
scale_colour_discrete <- function(...) scale_colour_manual(..., values=ptol_pal()(3))

```

```{r}
log_dir <- fs::path(here(), "data/observe_harvest_recruit") # solution files here
sims <- read_csv(file.path(log_dir, "sims.csv"), col_types = "inicnnn")
```

Try showing replicates individually?

 Based on original, still best

```{r}
sims %>%
  mutate(method = fct_recode(method, "MDP" = "CE", "Heuristic" = "TAC")) %>%
  filter(method != "MSY", sigma_g == "0.15", time <= 50) %>%
  filter(sigma_m != "0.1") %>%
  group_by(time, method, sigma_m) %>%
  summarise(mean = mean(state), sd = sd(state)) %>% 
  mutate(sigma_m = as.character(sigma_m)) %>%
  ungroup() ->
fig1_df
```

```{r fig.width=6, fig.height=6}
fig1_df %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line(lwd=1) + 
  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m)), ncol=1) + 
  theme(legend.position = "bottom") + 
  ylab("Mean biomass") 
```

```{r}
sims %>%
  select(-value) %>%
  filter(method != "MSY", sigma_g == "0.15", time <= 50) %>%
  mutate(method = fct_recode(method, "MDP" = "CE")) %>%
  filter(sigma_m != 0.1, rep %in% 5:20) %>%
  filter(sigma_m > 0.1) %>%
  ggplot(aes(time, state, col=method, group=interaction(rep,method))) + 
  geom_line(alpha= .5) + 
  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) 
```

```{r}
sims %>%
  select(-value) %>%
  mutate(sigma_m = round(sigma_m,2)) %>%
  mutate(sigma_g = round(sigma_g,2)) %>%
  filter(method != "MSY", sigma_g == 0.15, time <= 50) %>%
  mutate(method = fct_recode(method, "MDP" = "CE")) %>%
  filter(sigma_m == 0.15, method == "POMDP", rep %in% 5:20) %>%
  ggplot(aes(time, state, col=method, group=interaction(rep,method))) + 
  geom_line(col = rgb(0.7, 0.1, 0.4, 0.2))

```

### 2018-07-30 reviews/README.Rmd
reviews

---
title: "README.Rmd"
author: "Carl Boettiger"
date: "12/13/2017"
output: github_document
---

# README

Reviews are copyright of the reviewers and cannot be shared without permission.  To keep everything in one place, review material is stored here in encrypted files.

```{r}
library(gpg)
keyring <- gpg_list_keys(secret = TRUE)
cboettig <- keyring[1, "id"]
```

Encrypt `.gitignore`'d plain text, allowing only specified users to decrypt:

```{r}
## letter 1
letter1 <- "review-reply.Rmd"
gpg_letter1 <- "review-reply.Rmd.gpg"

msg <- gpg_encrypt(letter1, receiver = cboettig, signer = cboettig)
writeLines(msg, gpg_letter1)
```

Unencrypt committed gpg versions:

```{r}
writeLines(gpg_decrypt(gpg_letter1), letter1)
```

Encrypt letter 2

```{r}
letter2 <- "review-reply-2.Rmd"
gpg_letter2 <- "review-reply-2.Rmd.gpg"

msg <- gpg_encrypt(letter2, receiver = cboettig, signer = cboettig)
writeLines(msg, gpg_letter2)
```

Unencrypt committed gpg versions:

```{r}
writeLines(gpg_decrypt(gpg_letter2), letter2)
```

### 2018-07-30 manuscripts/scratch/scratch-current.Rmd
scratch

It is precisely this reasoning which explains the broad patterns of recovery
observed in the landmark paper @Costello2016, which compares future of global
fisheries under a Rights Based Fisheries Mangement (RBFM), fishing at MSY,
and "buisness as usual" (BAU) scenario.  Their analysis assumes that a RBFM
approach should achieve the economically optimal policy of constant-escapement,
which alone explains why RBFM should see both faster recovery and higher 
economic yields than MSY, as we have just described. The substantial detail their
analysis introduces with respect to the dynamics of over 4000 individual stocks
as well as additional economic knock-on effects from managing based on quotas
rather than fishing effort permit them to offer quantitatively precise 
estimates regarding the time required and economic value generated by such
recovery, but are not essential to understand the broad pattern that suggests
that nearly all stocks should recover under either an RBFM or MSY-style 
management, but that the recovery would be fastest and the economic benefits
greatest under the former.  That conclusion was already implicit in how
these models are defined.

-------------

A fundamental concept to the management of any natural population is the notion of Maximum Sustainable Yield, MSY. The notion of MSY is rooted in the determinstic models dating back to foundational work in resource economics and ecological management half a century ago [@Berverton; @Scheffer].  Though the concept has been the focus of significant criticism for some time [e.g. @Botsford_nature_piece] that ctitique has been largely on the "Maximum" part, while the existence of a "Sustainable Yield."

## Reed's Paradox: $S = D$

In 1979, a fundamental paper by William Reed [@Reed1979] proved that, under a very general assumptions, the harvest policy for a stochastically fluctuating population was identical to that of the equivalent, deterministic population.  More precisely, Reed shows that the optimal harvest of the stochastically flutuation population should always aim to reduce the population back to a constant size $S$ (termed by Reed as "escapement," while in most fisheries literature uses the term $B_{MSY}$, the biomass at Maximum Sustainable Yield}).  Moreover, this value, $S$, is no more conservative than had we been able to predict future populations with perfect accuracy, as in the determinstic case of constant escapement, $D$.  A notable exception to this result is when the population is not "self-sustaining," that is,
the stochasticity is severe enough to pose an extinction risk to the unhavested population.  In such cases Reed proves there is still a single constant target population size $S$, but that $S > D$, though often by very small amount, reflecting the small probability of a purely chance extinction.

An important corrolary of this result is that suggested a way to side-step the laborious dynamic programming required to find an optimal policy for stochastic populations: the deterministic solution is found trivially from maximum of the stock-recruitment curve. Thus much subsequent attention has focused on the estimating stock-recruitment, including methods that address considerable uncertainty in the data used for this estimate [@], while largely ignoring the subsequent step of how to determine an optimal policy in light of that uncertainty.

The lack of any intuitively precautionary response to the introduction of uncertainty did not go unnoticed: in 1986 Clark and Kirkwood first pointed to the absence of measurement error in the formulation used by Reed and others as fundamental to understanding this paradox.

> An important tacit assumption in Reed’s analysis, as in the other works referred
to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.

Under such an assumption, uncertainty introduced by stochastic growth is effectively purged by each subsequent observation rather than accumulating, and stock extinction can only occur if the manager explicitly decides to harvest the entire stock.

## Clark's Paradox: M < D

Clark and Kirkwood set out to address this problem by considering the case in which stochastic population growth occurs after population size is measured but before harvest, such that manager must act without perfect knowledge of the current system state.  Unfortunately, this problem proved easier to recognize than to solve, as Clark & Kirkwood recognized at the outset:

> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
of the problem increases markedly if this assumption is
relaxed.)

Rather than resolve Reed's Paradox, this assumption compounds the problem, leading to optimal harvests that are generally even higher (i.e. less cautious) than the analgous deterministic policy.  The same counter-intuitive result can also be found in subsequent work [@Sethi2005] which makes a different but equally limiting simplifying assumption (supplement).  Both @Clark1986 and @Sethi2005 remark on the counter-intuitive nature of their results;

> The above results appear to contradict the conventional wisdom of renewable resource management, under which high uncertainty would call for increased
caution in the setting of quotas.

- @Clark1986

> It may seem counter-intuitive that a measurement error causes lower expected escapements below the deterministic fishery closure threshold.

@Sethi2005

## Precautionary approaches

Perhaps recognizing

## A solution

Here, we explore how uncertainty over the present impacts decisions

We shall focus on examples from fisheries, which have long been both incubator and proving ground for most of our more general theories of ecological management, from the and

Data are readily avialable and a compartively strong connection between scientific process of stock assessment and the implementation of policy is already established.  It is also an area which, as we will see, though many models already account for uncertainty, decision strategy still reflects fundamentally more deterministic ideas.

Conclusions:

- failing to account for uncertainty properly can lead to erroneous conclusions
- Prior beliefs (reflecting prior observations) matter.

### 2018-07-31 appendix/appendixB.Rmd
no author (double-blind)
Figure numbers, sp
fix eqn, sp

@@
 title: "Appendix B: Code for analysis comparing to historical data"
-author: "Carl Boettiger"
 date: "`r Sys.Date()`"
@@
 
-Let's filter out missing data, non-matching units, and obvious reporting errors (catch exceeding total spawning biomass), then we rescale each series into the 0,1 by appropriate choice of units:
+Let's filter out missing data, non-matching units, and obvious reporting errors (catch exceeding total spawning biomass), then we re-scale each series into the 0,1 by appropriate choice of units:
 
@@
 ```{r}
-## Classic Graham-Schaefer. Note that recruitment occurs *before* harvest
+## Classic Gordon-Schaefer. Note that recruitment occurs *before* harvest
 gs <- function(r,K){

### 2018-08-01 manuscripts/manuscript.Rmd
fix eqn, sp
revisions...

@@
 
-$$\begin{equation}
+\begin{equation}
 y_t = \xi_t x_t 
-\end{equation}$$
+\end{equation}
 
@@
 
-$$\begin{equation}
+\begin{equation}
 \max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
-\end{equation}$$
+\end{equation}
 
@@
 
-$$\begin{equation}
-x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) 
-\end{equation}$$
+\begin{equation}
+x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) \label{Gordon-Schaefer}
+\end{equation}
 
@@
 
-MSE
+<!--- FIXME MSE discussion HERE!! -->
 
@@
 However, in practice the two approaches are very different: becuase natural populations fluctuate year-to-year, under CE, a different harvest quota must be set each year to achieve the constant escapement.
-While theoretical results have proven that CE and not MSY is optimal for fluctuating populations [@Reed1979], real-world examples of CE-regulated fisheries are limited primarily to salmon.
-As we shall see, despite being insensitive to natural stochasticity, CE is very sensitive to measurement error -- thus it is perhaps not surprising that CE is not found in other marine fisheries where population sizes may be harder to estimate accurately.
 
-Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above.
-Instead, existing theory has presented only the paradoxical results that uncertainty either shoud not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty.
+In the face of a naturally fluctuating population, previous arguments made by @Gordon1954; @Schaefer1954 or @Clark1973 no longer hold.
+The optimal harvest of a population under a model such as the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem (MDP).
+This is not to be confused with the the problem of *estimating* parameters or posteriors of a Markov process model, e.g.
+through typical Bayesian analysis -- recall this is a decision problem, in which the model has already been estimated and now we are tasked with choosing a sequence of future harvests based on that model.
+Such MDP problems are commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP) [@Marescot2013], widely used in both ecological management and animal behavior [@Mangel1985; @Mangel1988].
+However, thanks to a mathematical proof provided by @Reed1979, fisheries managers have largely been able to bi-pass the need for SDP.
+But this convenient result also opened the door to paradox of uncertainty that would persist for the next four decades.
 
@@
 
-The first of these we shall refer to as Reed's Paradox for stochastic growth, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.
+ @Reed1979 proved that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to @Clark1973's optimal escapement $D$ for a deterministic population, $S = D = B_{MSY}$.
 This surprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.
-This also provides the manager with a remarkably convenient mechanism for determining the optimal harvest policy: rather than rely on computationally intensive Stochastic Dynamic Programming to determine the optimal policy, a simple derivative will suffice, which is all that is needed to solve the maximum of the deterministic problem.
 
@@
 Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than reject it.
-Others have also attempted to introduce measurement error, but always relying on various approximations that similarly alter the problem.
-The most common such assumption is to assume from the outset that the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
-Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.
-In addition to resolving this paradox, we will demonstrate this direct solution also out-performs current more heuristic rules of thumb such as those currently used in fisheries management to account for uncertainty.
+Intuitively, uncertainty in the population growth model (stochasticity) or measurement error, would seem to justify harvesting fewer fish.
+Instead, existing optimal control theory has presented only the paradoxical results that uncertainty either shoud not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty.
+Numerous other attempts have been made to resolve this paradox through alternative assumptions,tThe most common being to assume from the outset to look only for 'constant-escapement' type polcies [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
+Owing to the "mathematical difficulty" Clark first observed, none have attempted a direct solution which we will present here.
+
+In practice, TAC-managed fisheries are more common that CE-managed fisheries, at least in the United States.
+CE-based regulation is limited primarily to salmon, species for which population estimates may be more precise and thus closer to Reed's assumption.
+We are not aware of any fishery that has put Clark's Paradox into practice, harvesting more in response to increased uncertainty.
+As we shall illustrate, this is particularly fortunate; though approaches that properly account for both forms of uncertainty can perform even better.
 
@@
 Perfect measurements let us exploit the Markov property of the model -- the most recent measurement tells us all we need to know.
-In contrast, the introduction of measurement uncertainty breaks this Makovian assumption: a rational decision maker would suspect this 75 to be an over-estimate based on prior observations.
+In contrast, the introduction of measurement uncertainty breaks this Markovian assumption: a rational decision maker would suspect this 75 to be an over-estimate based on prior observations.
+Under imperfect observations, we no longer have a Markov Decision Process (MDP) over the observed state space.
+Rather, we have a Partially Observed Markov Decision Process (POMDP).
+A POMDP policy cannot be expressed as a policy function at all (much less as policy function with a single constant-escapement target, as much previous work has assumed).
+Instead, a POMDP policy can be thought of as depending not only on the most recent observation, but the recent observation combined with a *prior belief* about what state the system is in, which itself is determined by all past observations.
+In this manner, the POMDP can consider how much weight to give the most recent observation based on how it compares to prior beliefs.
 
-As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP) [@Marescot2013].
-The constant escapement (CE) strategy considered here is the result of such an MDP or SDP analysis.
-Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).
+Algorithmic innovations from the robotics community have now made it possible to numerically solve non-trivial POMDP problems.
+The need to consider all past observations becomes particularly demanding given the forward-looking nature of a decision problem.
+Not only must we consider all possible sequences of future actions over all possible sequences of future states, but we must add to that all possible sequences of future measurements.
+The computational complexity involved has historically limited POMDP applications to contexts with only a handful of possible states and actions [e.g.
+@Chades2008; @Chades2011; @Fackler2014; @Fackler2014b], insufficient to capture fisheries models appropriately.
+Meanwhile, methods to solve decision problems with imperfect information have advanced steadily in the field of robotics and artificial intellengence.
+Almost contemporaneous with Reed's work,  papers by @Smallwood1973 and @Sondik1978 laid mathematical foundations for algorithms that could efficiently solve POMDP problems [e.g.
+@Kaelbling1998; @Pineau2003].
+More recently, driven by demands in areas such as autonomous vehicle decision making, newer point-based approximation algorithms such as SARSOP [@Kurniawati2008] have made it possible to solve POMDPs with 100s of states and actions.
+We have adapt this algorithm for application to the fisheries context, and provide an efficient implementation as a preliminary R package [@sarsop-pkg] which can be used to replicate and further explore results presented here.
+Details of the analysis presented here including annotated code to reproduce and further explore all of the following results is provided in the supplementary material.
 
-Precise definitions of both MDP and POMDP problems can be found in @Williams2011, or in classic work from the engineering and robotics community [e.g. @Smallwood1973 and @Sondik1978].
-
+Precise definitions of both the (fully observed) MDPs and POMDPs (not to be confused with Markov Process Models and Hidden Markov Process Models -- methods for which estimate the model but do not involve the feedback of any decision process) in an ecological context can be found in @Williams2011, or in classic work from the engineering and robotics community [e.g.
+@Smallwood1973 and @Sondik1978].
+For convenience, we summarize the POMDP definition and common notation here.
 The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process  The POMDP problem for fisheries question considered here can be summarized as follows:
@@
 Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.
-While the solution method for MDP problems, stochastic dynamic programming (SDP) is well known and readily implemented in ecological problems [e.g @Mangel1988; @Marescot2013], the POMDP problem cannot be efficiently solved using SDP.
-While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g.
-@Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].
-POMDP algorithms have a substantial literature starting almost contemporaneously to the work of @Reed1979 (see foundational papers by @Smallwood1973 and @Sondik1978) and remain an active area in artificial intelligence [e.g.
-@Kaelbling1998; @Pineau2003].
-By adapting more recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution.
-As the details for solving such problems numerically are already well documented we will not rehash them here, but instead provide a preliminary R package [@sarsop-pkg] which provides an implementation of the SARSOP algorithm and the fisheries POMDP problem, which can be used to replicate any of the results presented here and explore other variations.
-Details of the analysis presented here including annotated code to reproduce and further explore all of the following results is provided in the supplementary material.
 
@@
 
-Differences in attention to prior beliefs,determined by prior observations, drive the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
+Differences in attention to prior beliefs, determined by prior observations, drive the differences between the POMDP strategy and the other strategies and can resolve the uncertainty paradox.
 While the catch quota under both TAC and CE strategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure \ref{policy}, this is not the case for the POMDP approach.

### 2018-08-16 manuscripts/manuscript.Rmd
WIP

@@
 How have invasive species transformed native communities?
-A recognition of the importance of uncertainty has become widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods.
-In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and under-served by available tools.
-We illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a range of existing approaches, which can be avoided through the use of more formal approaches that have since been refined in fields such as robotics.
-This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.
+A recognition of the importance of uncertainty has become widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods to build better models.
+As we estimate ever more sophisticated models, it may be easy to forget that a model is not itself a decision, particularly when the estimated model includes considerable uncertainty about future outcomes.
+Knowing the rules of the chess board is not the same thing as knowing how to win.
+A model only gives us the rules of the game.
+The science of decision theory helps us know how to play.
+Yet while our models have become much more sophisticated, decision theory in ecological management frequently ignores the very uncertainty we so carefully include when estimating the model.
+Previous shortcuts in decision theory have led to paradoxical results, such as the suggestion that increased uncertainty in fisheries stock sizes should justify unchanged or even increased harvest rates.
+Borrowing decision theory methodology first developed in fields such as robotics, we are able to resolve this paradox, underscoring the importance of appropriately accounting for uncertainty in decisions just as we already do in models.
+This insight has important implications in both fisheries and how ecosystem management deals with uncertainty more generally.
 
@@
 We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.
-We compare the results of managing simulated fish stocks under (A) the existing optimal control approach, (B) a heuristic approach that is more typical of today's management practices, and (C) a newly proposed approach that provides a more rigorous treatment of uncertainty.
+We compare the results of managing simulated fish stocks under (A) the existing optimal control approach, (B) a heuristic approach that is more typical of today's management practices, and (C) a unified approach that provides a more rigorous treatment of uncertainty.
 
-The challenge of making decisions under uncertainty is obviously not unique to fisheries, and the existing methods we critique here can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010], and protecting biodiversity [@Dee2017].
+The challenge of making decisions under uncertainty is obviously not unique to fisheries, and the existing methods we critique here can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010; @Shea1998], disease outbreaks [@Shea2014; @Li2017], and protecting biodiversity [@Dee2017; @Iacona2017].
 Yet fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.
@@
 For simplicity we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.
-Given an estimated recruitment model along with this measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term (net) present value:
+Given an estimated population model along with this (uncertain) measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term utility of the stock:
 
 \begin{equation}
-\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
+\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot U(x_t, h_t) \right \}
 \end{equation}
 
-A few subtleties arise in this seemingly simple problem statement that we must address.
-First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?
+For convenience of presentation but without loss of generality, we will assume this utility is simply equal to the harvest itself; in economic terms, a fixed price for fish without cost on effort:
+
+\begin{equation}
+ U(x_t, h_t) = \min(h_t, x_t)
+\end{equation}
+
+where the $\min$ function ensures no additional reward for attempting to harvest more than the entire stock.
+Existing optimal control techniques (@Reed1979) prove these results are not dependent on the particular details of this utility.
+Using this form allows a more direct comparison to approaches such as maximum sustainable yield (MSY), that do not include an economic model but merely maximize long-term catch.
+In principle, utility could also reflect ecosystem services or other intrinsic values for fish left in the sea and not merely the market for consumption [@Halpern2013].
+
+The solution to this decision problem is conditional on a population model for the underlying ecological process, describing probability the system moves to state $x_{t+1}$ given that it was in state $x_t$ and the manager selected a harvest $h_t$, which we discuss below.
+Such a model would typically be estimated and parameterized from available data, often (but not necessarily) using a hierarchical Bayesian approach that acknowledges the existence of uncertainty in both measurements and process [e.g.
+see @Dichmont2016 for a discussion of how uncertainties are frequently included in stock assessment models].
+However, accounting for measurement uncertainty when going from data to model does not mean we have also accounted for this uncertainty when going from model to decision.
+For example, a harvest decision policy based upon maximum sustainable yield simply depends on the stock size which maximizes the expected growth rate of the model.
+Measurement uncertainty in model estimation can indeed influence the position of that maximum, but as we shall see, that strategy is only optimal in the absence of any uncertainty.
+
+There is one further subtlety arising in this seemingly simple problem statement that we must address.
+Given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?
 Following convention in the optimal control literature (e.g.
 [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.
+This assumption has occasionally been reversed to introduce some additional uncertainty from stochastic recruitment immediately prior to setting the harvest quota [e.g.
+@Clark1986; @Weitzmann2001], which appears to mimic uncertainty in measurements but does not allow uncertainty to accumulate since the measurement step still occurs without error.
 
-A second subtly is in the nature of the decision problem itself: note that the manager in not myopic, thinking only about the next year, but rather considers the sequence of all future actions ${h_t}$.
-While the manager will revisit this sequence each year in light of new observations, it is this ability of the decision problem to think ahead that allows it to tolerate short-term costs (e.g.
+The optimal strategy must think not only about the next year, but rather considers the sequence of all future actions ${h_t}$, under all possible realizations of the underlying model.
+This strategy must also acknowledge that the manager will receive new information each year, and may need to modify actions accordingly.
+It is this ability of the decision problem to think ahead that allows it to tolerate short-term costs (e.g.
 reduced harvests in the current season) for larger future payoffs.
-In this, ecological management is a game of chess, always thinking several moves ahead.
-This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a fixed mortality $F_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.
-Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.
+When Eq (1) is ignored, such optimal control problems are known as Markov Decision Processes or MDPs.
+These are more commonly referred to in the ecological literature by their solution method of Stochastic Dynamic Programming [SDP; @Marescot2013], and widely used in both ecological management and animal behavior [@Mangel1985; @Mangel1988].
+The addition of measurement uncertainty in Eq (1) turns this problem into a Partially Observed Markov Decision Process (POMDP), for which solution methods have not been widely available for large problems until more recently.
+We discuss these decision methods in more detail below.
 
@@
 
-It is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.
-Likewise, while we consider the model and its parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g.
-@Dichmont2016].
-As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.
+It is important to remember that all three decision methods we compare here will use this same model as given.
+In the case of numerical simulations, we will simply pick reasonable fixed parameter values for the model above.
+To illustrate the application to empirical data, we will first estimate these parameters using a hierarchical Markov Chain Monte Carlo (MCMC), and then use the resulting estimates to drive all three decision models.
+As such, the details of how the model is estimated are immaterial to the differences in performance between these decision methods.
+In all cases, the model is considered fixed.
+Technically, it would be possible to re-estimate the model parameters given each new observation during the decision process.
+Adaptations of the decision processes we consider here include this possibility for learning about the model during the decisions.
+Some include the possibility of taking knowingly bad actions, like suddenly increasing harvest rates, to improve learning over parameters.
+We investigate these in subsequent work [@Memarzadeh2018].
+For better comparison with established theory, here, the model is considered fixed.
+In the case of the simulations, the model is known without error (though still stochastic in nature), such that additional observations cannot improve the parameter estimates in any event.
+Despite that obvious simplification, we shall illustrate how existing decision methods fail to account for measurement uncertainty appropriately.
+Even with these simple models, decisions are complex, and knowing the rules of the game is not the same as knowing how to win.
 
@@
 Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.
-Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-fishing Limit, (OFL).
+Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-Fishing Limit, (OFL).
 Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment.
@@
 
-<!--- FIXME MSE discussion HERE!! -->
+It is important to note that actual management strategies are quite diverse.
+Our focus on comparisons TAC, MSY, and Constant Escapement (CE, also refereed to as MEY, see below) reflects only the subset of current practices.
+Another increasingly common approach to address uncertainty in both process and models is known as Management Strategy Evaluation [MSE, @Smith1994;  @Dorner2009; @Bunnefeld2011; @Punt2016].
+MSE performs forward simulations under a suite of pre-selected candidate strategies to determine which strategy achieves the best expected outcomes.
+Crucially, by limiting the analysis to a suite of candidate strategies over a finite number of replicate simulations, this approach is able to accommodate substantially greater complexity and uncertainty than has been possible with more formal optimal control approaches we discuss below.
+Optimal control solutions rely on dynamic programming to consider all possible strategies (all possible sequences of harvest under all possible realizations of stock dynamics).
+As such, MSE can be viewed as an approximation to the optimal control solutions considered below, a notion that has been formalized in the literature on approximate dynamic programming [@Nicol2011].
 
@@
 Because CE and MSY converge to the *same long-term biomass and same long-term average yield*, CE can appear to be merely a more formal justification for MSY.
-However, in practice the two approaches are very different: becuase natural populations fluctuate year-to-year, under CE, a different harvest quota must be set each year to achieve the constant escapement.
+However, in practice the two approaches are very different: because natural populations fluctuate year-to-year, under CE, a different harvest quota must be set each year to achieve the constant escapement.
+Because CE is derived as the economic optimum (assuming perfect measurements), it is sometimes referred to as Maximum Economic Yield, [MEY; @Burgess2018], or as Rights Based Fisheries Management [RBFM, @Costello2016].
 
 In the face of a naturally fluctuating population, previous arguments made by @Gordon1954; @Schaefer1954 or @Clark1973 no longer hold.
-The optimal harvest of a population under a model such as the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem (MDP).
-This is not to be confused with the the problem of *estimating* parameters or posteriors of a Markov process model, e.g.
-through typical Bayesian analysis -- recall this is a decision problem, in which the model has already been estimated and now we are tasked with choosing a sequence of future harvests based on that model.
-Such MDP problems are commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP) [@Marescot2013], widely used in both ecological management and animal behavior [@Mangel1985; @Mangel1988].
-However, thanks to a mathematical proof provided by @Reed1979, fisheries managers have largely been able to bi-pass the need for SDP.
-But this convenient result also opened the door to paradox of uncertainty that would persist for the next four decades.
+As noted above, the optimal harvest of a population under a model such as the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem (MDP) which must be solved with stochastic dynamic programming.
+Thanks to a mathematical proof provided by @Reed1979, fisheries managers have largely been able to bi-pass such computational effort.
+However, this convenient result also opened the door to paradox of uncertainty that would haunt ecological management for at least the next four decades.
 
@@
 Intuitively, uncertainty in the population growth model (stochasticity) or measurement error, would seem to justify harvesting fewer fish.
-Instead, existing optimal control theory has presented only the paradoxical results that uncertainty either shoud not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty.
-Numerous other attempts have been made to resolve this paradox through alternative assumptions,tThe most common being to assume from the outset to look only for 'constant-escapement' type polcies [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
+Instead, existing optimal control theory has presented only the paradoxical results that uncertainty either should not matter at all, or that we should increase rather than decrease harvests in response to increased uncertainty.
+Numerous other attempts have been made to resolve this paradox through alternative assumptions, the most common being to assume from the outset to look only for 'constant-escapement' type policies [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].
 Owing to the "mathematical difficulty" Clark first observed, none have attempted a direct solution which we will present here.
@@
 We are not aware of any fishery that has put Clark's Paradox into practice, harvesting more in response to increased uncertainty.
-As we shall illustrate, this is particularly fortunate; though approaches that properly account for both forms of uncertainty can perform even better.
+As we shall illustrate, this is fortunate; though approaches that properly account for both forms of uncertainty can perform even better.
 
@@
 In contrast, the introduction of measurement uncertainty breaks this Markovian assumption: a rational decision maker would suspect this 75 to be an over-estimate based on prior observations.
-Under imperfect observations, we no longer have a Markov Decision Process (MDP) over the observed state space.
+Under imperfect observations introduced by Eq (1), we no longer have a Markov Decision Process (MDP) over the observed state space.
 Rather, we have a Partially Observed Markov Decision Process (POMDP).
@@
 @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b], insufficient to capture fisheries models appropriately.
-Meanwhile, methods to solve decision problems with imperfect information have advanced steadily in the field of robotics and artificial intellengence.
+Meanwhile, methods to solve decision problems with imperfect information have advanced steadily in the field of robotics and artificial intelligence.
 Almost contemporaneous with Reed's work,  papers by @Smallwood1973 and @Sondik1978 laid mathematical foundations for algorithms that could efficiently solve POMDP problems [e.g.
@@
 In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).
-The introduction of stochastic growth has a significant negative impact on the TAC strategy, whith a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).
+The introduction of stochastic growth has a significant negative impact on the TAC strategy, with a mean biomass significantly lower than $B_{MSY}$ (first panel, Figure \ref{sims}; this impact is even more severe for MSY, see Figure S2).
 Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.
@@
 
-Using modern algorithms we have been able to both crack the nut of measurement uncertainty in the optimal management of marine fisheries and debunk a long-standing paradox: that increased measurement uncertainty should be met with larger harvests.
-We have demonstrated that both existing optimal control approaches and existing precautionary approaches can result in over-exploitation and long term declines of fish stocks, as well as reduced economic value, while in contrast, the POMDP approach is able to deliver both ecological and economic recovery despite uncertainty in measurements.
-These results have important implications for both global fisheries and ecological management more generally.
-In marine fisheries, evidence of such over-exploitation has already been well documented [@Worm2006; @Worm2009; @Costello2016].
-Our results suggest that measurement uncertainty may contribute to this pattern, as such errors lead to over-exploitation under current best practices such as MSY, TAC or CE, which have previously been thought to be sufficient to ensure the recovery of stocks world-wide [@Costello2016].
+Using modern algorithms we have been able to both crack the nut of measurement uncertainty in the optimal management of ecological populations such as marine fisheries, debunking a long-standing paradox created by previous literature which has argued that increased measurement uncertainty should be met with equal or larger harvests [e.g.
+@Reed1979; @Clark1986; @Sethi2005].
+By using powerful algorithms developed first the robotics literature to solve POMDPs [@Kurniawati2008], we have been able to tackle this optimal control problem directly rather than resorting to the simplifying assumptions used in previous work.
+Our solution confirms the intuition of the precautionary principle in a quantitatively precise form: under most (but not all) conditions, *greater uncertainty results in smaller optimal harvests*.
+Despite frequently proposing smaller harvests than existing approaches, the POMDP solution achieves higher long-term economic value as well as higher average stock sizes over the long term.
 
-Our results also have important consequences for ecological management more generally.
-The study and application of management strategies such as MSY and CE underpin much of conservation and natural resource management in forest management and many other terrestrial ecosystems [e.g.
-@Clark2000].
-Decision making under uncertainty is a central challenge for ecology and conservation biology [@Polasky2011], which we have long attempted to address through either simplified optimal control models or more heuristic approaches such as the precautionary principle and resilience thinking [@Fischer2009].
-We have illustrated how both of these approaches can fail to provide an adequate strategy when faced with measurement uncertainty.
-To avoid this pitfall, a strategy must be dynamic to make best use of the available information without assuming that information is perfect.
-Classical optimal control approaches (such as CE) can fail due to misplaced confidence in measurements even when they are significantly at odds with prior expectations.
-Similarly, the TAC approach also fails due to its reliance on a simple one-size-fits-all strategy, not reducing catch quotas sufficiently when stocks are too low but also failing to increase them when stocks are higher.
-It is precisely this ability to adapt the policy in response to available information, as illustrated in Figure \ref{policy}, that makes the POMDP approach successful.
-Nevertheless, it may be tempting to see these results as either trivial or impractical, but a closer inspection reveals they are neither.
+This lesson has clear relevance in the context of fisheries management.
+Here, practice is already wiser than theory, where the optimal control based on the CE results of @Clark1973 and @Reed1979 is limited to fisheries such as salmon where estimates in stock size may be less significant than in purely marine species.
+To our knowledge, no fishery has intentionally put the more aggressive harvest solutions of @Clark1986 and @Sethi2005 into practice.
+Yet this does not lift the specter of these methods, where CE-based predictions feature prominently in global fisheries analyses as the economically optimal strategy and the fastest route to recovery [@Costello2016; @Burgess2018].
+Our results raise serious questions about the optimistic forecasts predicted under these CE-based policies.
+Meanwhile, MSY remains the dominate standard international law [@Mace2001], and many well-managed US fisheries rely on a version of the TAC based precautionary adjustment to MSY to comply with catch limits under Magnuson-Stevens Act.
+Our results show that these strategies are less sensitive to measurement error, but not immune.
+Comparisons against more heuristic decision methods are more difficult: management strategy evaluation [MSE; @Smith1994; @Punt2016] will depend on precisely what management scenarios are considered.
+Under sufficient uncertainty, both the existing optimal and precautionary approaches can lead to declines in biomass and economic returns which could be avoided by POMDP-based management.
 
-First, it may seem obvious that POMDP would out-preform more primitive optimal control solutions such as CE, but this was hardly a forgone conclusion.
-Indeed, the previous example of @Reed1979 had proven that the complex stochastic dynamic programming solutions required to accommodate stochastic dynamics would perform no better than the simple maximization based purely on the deterministic model.
-If stochasticity does not matter, it is easy to believe that measurement error would be minor at best.
-In fact, we find that once we account for measurement error, the effects of both stochasticity and measurement have a significant influence over the optimal policy.
-It is similarly tempting to dismiss the better performance of POMDP against a more heuristic strategy like the TAC scenario considered here (fishing at 80% MSY), but once again this was not a forgone conclusion.
-While a better economic return might reasonably be expected of the optimization approach, it was not obvious that POMDP would also be at times more cautious than the rule of thumb, able to deliver both higher economic returns by harvesting more when it was safe to do so, and higher ecological recovery by harvesting even less than TAC when necessary.
-These results highlight the difficulty of finding effective heuristic rules and underscore the power and utility of more sophisticated strategies such as POMDP which can make the best use of all available information.
+This conclusion has important broader implications for ecological management more generally.
+Decision making under uncertainty is a central challenge for ecology and conservation biology, which frequently pits optimal control based approaches against more heuristic approaches such as resilience thinking, scenario planning, and the precautionary principle [@Polasky2011].
+With economists frequently favoring the precise, stylized approach of optimal control and ecologists embracing the realities of greater complexity and uncertainty echoed in more heuristic approaches, this division can sometimes appear to pit economic outcomes against ecological objectives.
+That perspective is echoed in the paradox discussed here, in which economic optimization ignoring uncertainty has argued in cold mathematical logic for a rejection of the precautionary principle.
+Our results demonstrate that this divide is artificial.
+With sufficiently powerful algorithms we can go some ways to realizing the call of @Fischer2009 to combine more ecologically realistic assumptions such as measurement error with optimal control approaches in ecological management.
+Both sides are vindicated in our solution: economic optimization under POMDP does not contradict the precautionary principle, while under POMDP, optimal control can also achieve better ecological outcomes than blunt precautionary rule (i.e.
+TAC, which reduced all harvests by 20% below MSY target).
+As @Fischer2009 predicted, unifying optimal control with more resilience-inspired approaches is win-win.
 
-Second, we must ask if the policies derived from POMDP are practical.
-As with constant escapement, the POMDP solutions we considered here result in more year-to-year variation in harvest, and include the possibility of fishing moratorium to allow stocks to recover -- decisions that many view as economically or socially unacceptable.
-While resolving this debate is beyond the scope of this paper, a few observations are relevant: (A) Stocks already managed under constant escapement, such as US salmon fisheries, are already subject to the possibility of these moratoriums.
-(B) It is straight-forward to introduce constraints on the minimal acceptable harvest level and/or include the additional costs of closing a fishery in the utility function optimized by POMDP.
-(C) Fisheries today are not managed at either their maximum ecological or economic potential.
-Taken as a whole, most ocean fisheries yield a net negative economic return [@FAO2009], with fishing persisting only with the support of government subsidies [@Arnason2012], meanwhile some 68% of global fisheries are in poor biological condition [@Costello2016].
-While the causes for this are both complex and numerous, we have seen that both the economic inefficiency and ecological health suffer under TAC/MSY and CE based management.
-@Arnason2012 and @Costello2016 have both argued that a rights-based approach to fisheries management (RBFM) would lead to self adoption of precisely such optimal policies.
+If accounting for measurement uncertainty can resolve previous paradoxes and divides, it also raises additional challenges.
+The approach discussed here is vulnerable both to the charge of being too simple to be realistic while also being too complicated to be feasible.
+These issues are as easily recognized in other ecological management contexts, from harvests of wild game or forestry to pest outbreaks, [e.g.
+@Richards1999; @Blackwood2010; @Shea1998; @Shea2014; @Li2017; @Dee2017; @Iacona2017] as in fisheries, where ecological complexity is invariably high and management capacity invariably limited.
+We address each of these in turn.
 
-Third: are the examples considered here too simple?
-The strategies we have compared against underpin both current practice and management aspirations.
-Together, MSY and TAC-based management reflects much of modern best practice as it is understood in fisheries management and elsewhere.
-While real-world application of these approaches frequently involves more complex models, the models we have considered here are the very ones under which these strategies were designed and tested, and there is little reason to believe they would perform better under more complex cases such as age-structured models that do not always meet those assumptions [e.g.
+Are the dynamics considered here too simple?
+Our underlying model is undoubtedly much simpler than many of the age or stage structured models used in modern fisheries management -- aspects whose importance to ecological dynamics have been established well beyond fisheries as well.
+Yet it is critical to bear in mind that this stylized one-dimensional model is used only to compare the relative performance of the different decision strategies considered here (TAC, CE, and POMDP) in order to illustrate the potential importance of uncertainty in measurement.
+There is little reason to believe CE or TAC would perform better under more complex cases such as age-structured models that only further deviate from the assumptions under which they were derived [e.g.
 @Holden2015].
-The relatively poor ecological and economic performance of these common best-practices relative to the alternative approach of POMDP is thus reason enough for concern.
-This issue is further underscored by the reality of both long-running ecological deterioration [@Worm2006; @Worm2009] and economic costs of global fisheries.
+POMDP approaches are in no way incompatible with more complex models, whether that be age structure or even more ambitious non-parametric representations defined directly from data using Gaussian Processes or other machine learning techniques [e.g.
+@Boettiger2015].
+Simple models have always served as tools for comparison and intuition, and serve the same illustrative role here.
 
-Fourth: is the approach taken here too complex to be feasible?
+Is this approach too complex to be feasible?
 The importance of simple and effective rules of thumb in conservation management has been well documented [e.g.
@@
 Sophisticated and computationally intensive approaches such as POMDP may seem improbable at a time when many areas of natural resource management, even simple, rule-of-thumb based methods struggle to take root.
-And yet, in this same age we rely on far more complex and computationally intensive algorithms to perform far more trivial tasks such as determining what advertisements we see.
-Surely a planet that can afford such complex optimization for how advertising space is allocated on a small screen can one day manage its natural resources and conservation efforts with such tools?
-
-<!-- Adaptive management -->
+There is no doubt that the solution method needed here is computationally intensive even for the simple model considered here, and will only become more so under greater ecological complexity.
+Indeed, it was precisely such computational limitations that led to the shortcuts in work such as @Clark1986, with the ensuing potentially misleading conclusions.
+While in the following decades ecologists largely turned away from computationally intensive optimization required for even simple models with stochastic dynamic programming, (while continuing to construct ever more complicated hierarchical models), engineers have chipped away at the decision problem to reveal algorithms such as SARSOP that make much larger dynamic programming problems tractable.
+Yes, we may need even more sophisticated algorithms to accommodate greater ecological complexity, yet we believe such numerical obstacles are no excuse for management by half measures.
+Rather, our example illustrates how ecologists and managers can learn from and leverage the tools developed in other disciplines to better face the challenges of management in a complex and changing world.
 

### 2018-08-16 manuscripts/manuscript.tex
WIP

@@
 % Pandoc header
-\journal{The American Naturalist} \usepackage{lineno} \linenumbers
+\journal{The American Naturalist}
+\usepackage{lineno}
+\linenumbers
 \modulolinenumbers[3]
-\usepackage{setspace} \doublespacing
+\usepackage{setspace}
+\doublespacing
 
@@
   \end{abstract}
-   \begin{keyword} Measurement uncertainty \sep POMDP \sep Decision theory \sep Fisheries \sep Conservation \sep \end{keyword}
+   \begin{keyword} Measurement uncertainty POMDP Decision theory Fisheries Conservation\end{keyword}
  \end{frontmatter}
@@
 biodiversity? How have invasive species transformed native communities?
-A recognition of the importance of uncertainty as well as sophisticated
-methods and tools to do so has become increasingly widespread in the
-context of estimating ecological models from available data, thanks in
-large part to the rise of approaches such as Hierarchical Bayesian
+A recognition of the importance of uncertainty has become widespread in
+the context of estimating ecological models from available data, thanks
+in large part to the rise of approaches such as Hierarchical Bayesian
 Modeling, along with convenient software tools for implementing these
@@
 scientific recommendations, remains far less developed, more
-qualitative, and under-served by available tools. In this paper we
-illustrate how uncertainty in measurement has a dramatic negative impact
-on outcomes under a variety of these existing approaches, which can be
-avoided through the use of more formal approaches that have since been
-developed in other fields that can likewise be implemented in software
-tools. This insight both resolves a long-standing paradox in fisheries
-management and has important implications for how ecosystem management
-deals with uncertainty more generally.
+qualitative, and under-served by available tools. We illustrate how
+uncertainty in measurement has a dramatic negative impact on outcomes
+under a range of existing approaches, which can be avoided through the
+use of more formal approaches that have since been refined in fields
+such as robotics. This insight both resolves a long-standing paradox in
+fisheries management and has important implications for how ecosystem
+management deals with uncertainty more generally.
 
 Approaches for decision making under uncertainty in ecological systems
-can be divided into two camps: approaches based in ``optimal control''
-and usually favored by natural resource economists, and approaches based
-in heuristic methods such as scenario planning, resilience thinking, and
-precautionary rules of thumb, more commonly found in both ecological
-literature and actual practice (Fischer et al. 2009; Polasky et al.
-2011). While researchers have for some time recognized the need to unify
-the transparent and quantitative algorithmic approach of optimal control
+can be divided into two camps: (1) those based upon ``optimal control''
+solutions, and usually favored by natural resource economists, and (2)
+those based in heuristic methods such as scenario planning, resilience
+thinking, and precautionary rules of thumb, more commonly found in both
+ecological literature and actual practice (Fischer et al. 2009; Polasky
+et al. 2011). While researchers have recognized the need to unify the
+transparent and quantitative algorithmic approach of optimal control
 with the greater complexity and uncertainty of real ecosystems that is
 acknowledged by heuristic methods, computational barriers to doing so
-have hither-to stymied this progress. Here, we illustrate how the
-limitations of these approaches has been manifested in the example of
-fisheries management, and present a new approach that can combine the
-reality of measurement uncertainty and the rigor of optimization to
-resolve a long-standing paradox and suggest a more robust approach to
-management.
-
-Fisheries conservation and management has long been both a crucible and
-proving ground for the theory of ecological management more generally,
-including topics such as adaptive management (Walters and Hilborn 1978),
-ecosystem-based management (Levin and Lubchenco 2008) and resilience
-thinking (Holling 1973; May 1977), while also giving rise to the
-sub-discipline of resource economics (Gordon and Press 1954; Schaefer
-1954; Beverton and Holt 1957), thanks to its global relevance, long
-history, and readily available data. Consequently, understanding the
-impact of measurement uncertainty in this rich and well-studied context
-will have implications for both resource management and conservation
-efforts in other domains, many of which draw on concepts established or
-tested in fisheries ecosystems. In particular, the approach here
-illustrates the limitations of both over-simple optimal control
-approaches and rule-of-thumb precautionary approaches, and offers
-instead a more general strategy for dealing with measurement uncertainty
-adaptively.
-
-We compare the results of managing simulated fish stocks under the
-existing optimal control approach, a heuristic approach that more
-closely resembles today's management practices, and a newly proposed
-approach that provides a more rigorous treatment of uncertainty.
-
-\section{The Decision Problem}\label{the-decision-problem}
+have stymied this progress. We illustrate how the limitations of these
+approaches have manifested in the example of fisheries management, and
+present a new approach that can combine the reality of measurement
+uncertainty and the rigor of optimization to resolve a long-standing
+paradox and suggest a more robust approach to management. We compare the
+results of managing simulated fish stocks under (A) the existing optimal
+control approach, (B) a heuristic approach that is more typical of
+today's management practices, and (C) a newly proposed approach that
+provides a more rigorous treatment of uncertainty.
+
+The challenge of making decisions under uncertainty is obviously not
+unique to fisheries, and the existing methods we critique here can be
+found in a wide range of population and ecosystem management challenges,
+including fire ecology (Richards et al. 1999), invasive species
+({\textbf{???}}; Blackwood et al. 2010), disease outbreaks (Shea et al.
+2014; Li et al. 2017), and protecting biodiversity (Dee et al. 2017;
+Iacona et al. 2017). Yet fisheries conservation and management has long
+been both a crucible and proving ground for the theory of ecological
+management more generally, including topics such as adaptive management
+(Walters and Hilborn 1978), ecosystem-based management (Levin and
+Lubchenco 2008) and resilience thinking (Holling 1973; May 1977), while
+also giving rise to the sub-discipline of resource economics (Gordon and
+Press 1954; Schaefer 1954; Beverton and Holt 1957), thanks to its global
+relevance, long history, and readily available data. Despite the
+generality of the models and issue of measurement uncertainty considered
+here, any application must be grounded in both the history and context
+of the particular decision problem. In light of this, we begin with a
+brief background on existing decision theory approaches common in
+fisheries management.
+
+\hypertarget{the-decision-problem}{%
+\section{The Decision Problem}\label{the-decision-problem}}
 
@@
 fishery in the face of imperfect information about the current stock
-size and uncertainty about future recruitment.
-
-We seek to determine the sequence of actions \(h_t\) for
-\(t \in [1, ... \infty]\) that maximize the net present value
-(discounted sum of all future profits) of the fishery. We will denote
-the discount rate \(\gamma\). For simplicity we will once again follow
-classic theory and assume a fixed price for fish, (equivalently,
-measuring our value in units of discounted fish rather than discounted
-dollars), though again this assumption is easy to relax. Each year, the
-manager also obtains an estimate \(y_t\) of the true \(x_t\) stock size
-subject to some measurement uncertainty \(\xi_t\)
-
-\[y_t = \xi_t x_t\]
+size and uncertainty about future recruitment. We seek to determine the
+sequence of actions \(h_t\) for \(t \in [1, ... \infty]\) that maximize
+the net present value (discounted sum of all future profits) of the
+fishery. We will denote the discount rate \(\gamma\). For simplicity we
+will once again follow classic theory and assume a fixed price for fish,
+(equivalently, measuring our value in units of discounted fish rather
+than discounted dollars), though again this assumption is easy to relax.
+Each year, the manager also obtains an estimate \(y_t\) of the true
+\(x_t\) stock size subject to some measurement uncertainty \(\xi_t\)
+
+\begin{equation}
+y_t = \xi_t x_t 
+\end{equation}
 
@@
 
-\[\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}\]
+\begin{equation}
+\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
+\end{equation}
 
@@
 
-\subsection{Population model}\label{population-model}
+\hypertarget{population-model}{%
+\subsection{Population model}\label{population-model}}
 
@@
 
-\[x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) \]
+\begin{equation}
+x_{t+1} = \varepsilon_t (x_t-h_t) r  \left(1 - \frac{(x_t-h_t)}{K}\right) \label{Gordon-Schaefer}
+\end{equation}
 
@@
 compare our results against intuition and classic theory, and avoid the
-possibility of differences arsing only because of some particular subtle
-assumption hidden in a more complex model.
+possibility of differences arising only because of some particular
+subtle assumption hidden in a more complex model.
 
@@
 complex models as well, including those with age or stage structure.
-Likewise, while we consider the model and it's parameters as ``given,''
+Likewise, while we consider the model and its parameters as ``given,''
 it should be understood to mean that this in practice comes as the
@@
 
-\section{Current Theory and Practice}\label{current-theory-and-practice}
+\hypertarget{current-theory-and-practice}{%
+\section{Current Theory and
+Practice}\label{current-theory-and-practice}}
 
@@
 
+\hypertarget{uncertainty-in-practice-msy-tac}{%
 \subsection{Uncertainty in Practice: MSY \&
-TAC}\label{uncertainty-in-practice-msy-tac}
+TAC}\label{uncertainty-in-practice-msy-tac}}
 
@@
 \(B_{MSY}\) and corresponding mortality \(F_{MSY}\) is used to define
-the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower
+the stock Over-Fishing Limit, (OFL). Based on this, a somewhat lower
 level is set as the Allowable Biological Catch (ABC), reflecting
@@
 
-\subsection{Optimal management}\label{optimal-management}
+It is important to note that actual management strategies are quite
+diverse. Our focus on comparisons TAC, MSY, and Constant Escapement (CE,
+also refereed to as MEY, see below) reflects only the subset of current
+practices. Another increasingly common approach to address uncertainty
+in both process and models is known as Management Strategy Evaluation
+(MSE, Smith 1994; {\textbf{???}}; Bunnefeld et al. 2011; Punt et al.
+2016). MSE performs forward simulations under a suite of pre-selected
+candidate strategies to determine which strategy achieves the best
+expected outcomes. Crucially, by limiting the analysis to a suite of
+candidate strategies over a finite number of replicate simulations, this
+approach is able to accommodate substantially greater complexity and
+uncertainty than has been possible with more formal optimal control
+approaches we discuss below. Optimal control solutions rely on dynamic
+programming to consider all possible strategies (all possible sequences
+of harvest under all possible realizations of stock dynamics). As such,
+MSE can be viewed as an approximation to the optimal control solutions
+considered below, a notion that has been formalized in the literature on
+approximate dynamic programming (Nicol and Chadès 2011).
+
+\hypertarget{optimal-management}{%
+\subsection{Optimal management}\label{optimal-management}}
 
@@
 economic discounting of future profits, but is otherwise quite general,
-see Clark (1990)). This strategy is known as constant escapement, since
-a constant sock size \(B_{MSY}\) escapes harvest each year. While both
-converge to the same long-term biomass and long-term yield, these
-differences away from equilibrium between the mortality-based approach
-that dominates fisheries management practice and the escapement-based
-approach that became the focus of theory will result in different
-behaviors under uncertainty. Intuitively, any uncertainty, either in the
-population growth model (stochasticity), or measurement error, should
-seem to justify harvesting fewer fish, as is built into the TAC limits
-discussed above. Yet formal treatments of these issues have so far
-required approximations, and these approximations have driven
-paradoxical conclusions which we briefly summarize here.
-
+see Clark (1990)). This strategy is known as \emph{constant escapement}
+(CE), since a constant sock size \(B_{MSY}\) escapes harvest each year.
+Because CE and MSY converge to the \emph{same long-term biomass and same
+long-term average yield}, CE can appear to be merely a more formal
+justification for MSY. However, in practice the two approaches are very
+different: because natural populations fluctuate year-to-year, under CE,
+a different harvest quota must be set each year to achieve the constant
+escapement. Because CE is derived as the economic optimum (assuming
+perfect measurements), it is sometimes referred to as Maximum Economic
+Yield, (MEY; Burgess et al. 2018), or as Rights Based Fisheries
+Management (RBFM, Costello et al. 2016).
+
+In the face of a naturally fluctuating population, previous arguments
+made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
+longer hold. The optimal harvest of a population under a model such as
+the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an
+example of a Markov Decision Problem (MDP). This is not to be confused
+with the the problem of \emph{estimating} parameters or posteriors of a
+Markov process model, e.g.~through typical Bayesian analysis -- recall
+this is a decision problem, in which the model has already been
+estimated and now we are tasked with choosing a sequence of future
+harvests based on that model. Such MDP problems are commonly referred to
+in ecological literature by its solution method, Stochastic Dynamic
+Programming (SDP) (Marescot et al. 2013), widely used in both ecological
+management and animal behavior (Mangel 1985; Mangel and Clark 1988).
+However, thanks to a mathematical proof provided by Reed (1979),
+fisheries managers have largely been able to bi-pass the need for SDP.
+But this convenient result also opened the door to paradox of
+uncertainty that would persist for the next four decades.
+
+\hypertarget{reeds-paradox-s-d}{%
 \subsection{\texorpdfstring{Reed's Paradox:
-\(S = D\)}{Reed's Paradox: S = D}}\label{reeds-paradox-s-d}
-
-The first of these we shall refer to as Reed's Paradox for stochastic
-growth, owing to the mathematical proof provided in Reed (1979) which
-demonstrates that, under sufficiently general assumptions, the optimal
-escapement \(S\) for a population under stochastic growth, is identical
-to the optimal escapement \(D\) of a deterministic population,
-\(S = D = B_{MSY}\). This surprising result suggests that in going from
-a world where a manager has perfect knowledge of absolutely everything
-into a scenario where the manager faces considerable uncertainty about
-the future state of the world, no additional precaution is needed. This
-also provides the manager with a remarkably convenient mechanism for
-determining the optimal harvest policy: rather than rely on
-computationally intensive Stochastic Dynamic Programming to determine
-the optimal policy, a simple derivative will suffice, which is all that
-is needed to solve the maximum of the deterministic problem.
-
-\subsection{Clark's Paradox}\label{clarks-paradox}
+\(S = D\)}{Reed's Paradox: S = D}}\label{reeds-paradox-s-d}}
+
+Reed (1979) proved that, under sufficiently general assumptions, the
+optimal escapement \(S\) for a population under stochastic growth, is
+identical to Clark (1973)'s optimal escapement \(D\) for a deterministic
+population, \(S = D = B_{MSY}\). This surprising result suggests that in
+going from a world where a manager has perfect knowledge of absolutely
+everything into a scenario where the manager faces considerable
+uncertainty about the future state of the world, no additional
+precaution is needed.
+
+\hypertarget{clarks-paradox}{%
+\subsection{Clark's Paradox}\label{clarks-paradox}}
 
@@
 
-Unfortunately, Clark and Kirkwood (1986) was unable to solve the
-resulting problem exactly, but had to adopt an almost equally
-troublesome assumption:
+Clark and Kirkwood (1986) were unable to solve the resulting problem
+exactly, but had to adopt an almost equally troublesome assumption:
 
@@
 
-Unfortunately for Clark and Kirkwood (1986), this ``simplifying
-assumption'' serves to squash most of the measurement error, and their
-results instead only deepened the paradox, finding policies that become
-even \emph{less} cautious as uncertainty increases:
+Unfortunately, this ``simplifying assumption'' serves to squash most of
+the measurement error, and their results instead only deepened the
+paradox, finding policies that become even \emph{less} cautious as
+uncertainty increases:
 
@@
 et al. (2005) ultimately attempt to rationalize this counter-intuitive
-conclusion rather than reject it. Others have also attempted to
-introduce measurement error, but always relying on various
-approximations that similarly alter the problem. The most common such
-assumption is to assume from the outset that the solution must be of
-`constant-escapement' type (Ludwig and Walters 1981; Roughgarden and
-Smith 1996; Engen et al. 1997; Moxnes 2003). Here we will illustrate the
-use of powerful modern algorithms that will permit a more direct
-solution to the case of measurement uncertainty. In addition to
-resolving this paradox, we will demonstrate this direct solution also
-out-performs current more heuristic rules of thumb such as those
-currently used in fisheries management to account for uncertainty.
-
+conclusion rather than reject it. Intuitively, uncertainty in the
+population growth model (stochasticity) or measurement error, would seem
+to justify harvesting fewer fish. Instead, existing optimal control
+theory has presented only the paradoxical results that uncertainty
+either should not matter at all, or that we should increase rather than
+decrease harvests in response to increased uncertainty. Numerous other
+attempts have been made to resolve this paradox through alternative
+assumptions, the most common being to assume from the outset to look
+only for `constant-escapement' type policies (Ludwig and Walters 1981;
+Roughgarden and Smith 1996; Engen et al. 1997; Moxnes 2003). Owing to
+the ``mathematical difficulty'' Clark first observed, none have
+attempted a direct solution which we will present here.
+
+In practice, TAC-managed fisheries are more common that CE-managed
+fisheries, at least in the United States. CE-based regulation is limited
+primarily to salmon, species for which population estimates may be more
+precise and thus closer to Reed's assumption. We are not aware of any
+fishery that has put Clark's Paradox into practice, harvesting more in
+response to increased uncertainty. As we shall illustrate, this is
+particularly fortunate; though approaches that properly account for both
+forms of uncertainty can perform even better.
+
+\hypertarget{pomdps-an-optimal-treatment-of-measurement-uncertainty}{%
 \subsection{POMDPs: An optimal treatment of measurement
-uncertainty}\label{pomdps-an-optimal-treatment-of-measurement-uncertainty}
-
-As discussed above, absent the short-cut provided by Reed's theorems --
-solving for the optimal management policy in a stochastically
-fluctuating population -- involves a Markov Decision Process (MDP),
-commonly referred to in ecological literature by its solution method,
-Stochastic Dynamic Programming (SDP). The constant escapement (CE)
-strategy considered here is the result of such an MDP or SDP analysis.
-Once measurement uncertainty is introduced, the problem becomes a
-Partially Observed Markov Decision Process (POMDP, not be confused with
-a partially observed or hidden Markov model, HMM, in which no decision
-process is involved). The POMDP problem is posed identically to that of
-the MDP problem, with the addition of an observation process, but cannot
-be solved using SDP directly. The POMDP problem for fisheries question
-considered here can be summarized as follows:
+uncertainty}\label{pomdps-an-optimal-treatment-of-measurement-uncertainty}}
+
+Why does the introduction of measurement uncertainty make the decision
+problem so much harder? Consider the thought experiment of managing a
+fish stock where in each of the previous four years stock assessments
+have put the population size at, say, 51, 54, 49, 51, (say, measured in
+percent of carrying capacity, \(K\)). Under a constant escapement target
+of 50, the harvest quota would be 1, 4, 0, 1, respectively. Note that
+this control rule depends only on the measurement for the year in
+question, which is assumed to be measured without error. If we then
+measure a stock size of 75, an assumption that this measurement is made
+without error would justify the sudden increase to a quota of 25. Yet if
+we admit the possibility of measurement error, this sudden bump starts
+to look very suspicious in light of all the previous observations.
+Perfect measurements let us exploit the Markov property of the model --
+the most recent measurement tells us all we need to know. In contrast,
+the introduction of measurement uncertainty breaks this Markovian
+assumption: a rational decision maker would suspect this 75 to be an
+over-estimate based on prior observations. Under imperfect observations,
+we no longer have a Markov Decision Process (MDP) over the observed
+state space. Rather, we have a Partially Observed Markov Decision
+Process (POMDP). A POMDP policy cannot be expressed as a policy function
+at all (much less as policy function with a single constant-escapement
+target, as much previous work has assumed). Instead, a POMDP policy can
+be thought of as depending not only on the most recent observation, but
+the recent observation combined with a \emph{prior belief} about what
+state the system is in, which itself is determined by all past
+observations. In this manner, the POMDP can consider how much weight to
+give the most recent observation based on how it compares to prior
+beliefs.
+
+Algorithmic innovations from the robotics community have now made it
+possible to numerically solve non-trivial POMDP problems. The need to
+consider all past observations becomes particularly demanding given the
+forward-looking nature of a decision problem. Not only must we consider
+all possible sequences of future actions over all possible sequences of
+future states, but we must add to that all possible sequences of future
+measurements. The computational complexity involved has historically
+limited POMDP applications to contexts with only a handful of possible
+states and actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight
+2014; Fackler and Pacifici 2014), insufficient to capture fisheries
+models appropriately. Meanwhile, methods to solve decision problems with
+imperfect information have advanced steadily in the field of robotics
+and artificial intelligence. Almost contemporaneous with Reed's work,
+papers by Smallwood and Sondik (1973) and Sondik (1978) laid
+mathematical foundations for algorithms that could efficiently solve
+POMDP problems (e.g. Kaelbling et al. 1998; Pineau et al. 2003). More
+recently, driven by demands in areas such as autonomous vehicle decision
+making, newer point-based approximation algorithms such as SARSOP
+(Kurniawati et al. 2008) have made it possible to solve POMDPs with 100s
+of states and actions. We have adapt this algorithm for application to
+the fisheries context, and provide an efficient implementation as a
+preliminary R package (Boettiger et al. 2018) which can be used to
+replicate and further explore results presented here. Details of the
+analysis presented here including annotated code to reproduce and
+further explore all of the following results is provided in the
+supplementary material.
+
+Precise definitions of both the (fully observed) MDPs and POMDPs (not to
+be confused with Markov Process Models and Hidden Markov Process Models
+-- methods for which estimate the model but do not involve the feedback
+of any decision process) in an ecological context can be found in
+Williams (2011), or in classic work from the engineering and robotics
+community (e.g. Smallwood and Sondik 1973, and @Sondik1978). For
+convenience, we summarize the POMDP definition and common notation here.
+The POMDP problem is posed identically to that of the MDP problem, with
+the addition of an observation process The POMDP problem for fisheries
+question considered here can be summarized as follows:
 
@@
   \(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero
-  to exclude negative population sizes.
+  to exclude negative population sizes. The action \(a_t\) represents
+  the harvest quota set for the fishery. Attempting to implement a quota
+  the exceeds the true population size results in collapse of the
+  fishery. As Clark and Kirkwood (1986) eloquently argues, assuming that
+  extinction is impossible regardless of fishing intensity would
+  unreasonably bias the decision problem.
 \item
@@
 actions, and can thus be represented as a collection of matrices or
-tensors. While the solution method for MDP problems, stochastic dynamic
-programming (SDP) is well known and readily implemented in ecological
-problems (e.g Mangel and Clark 1988; Marescot et al. 2013), the POMDP
-problem cannot be efficiently solved using SDP. While not unknown to the
-conservation literature (Williams 2011), algorithms for POMDP have
-historically scaled quite poorly, and their application has been
-restricted to contexts with only a handful of possible states and
-actions (e.g. Chadès et al. 2008, 2011; Fackler and Haight 2014; Fackler
-and Pacifici 2014). POMDP algorithms have a substantial literature
-starting almost contemporaneously to the work of Reed (1979) (see
-foundational papers by Smallwood and Sondik (1973) and Sondik (1978))
-and remain an active area in artificial intelligence (e.g. Kaelbling et
-al. 1998; Pineau et al. 2003). By adapting recent algorithmic
-developments (Kurniawati et al. 2008) from that field we are able to
-find solutions for the considerably more complex problems such as the
-case of fisheries management, which requires on the order of 100 states
-and actions to provide sufficient numerical resolution. As the details
-for solving such problems numerically are already well documented we
-will not rehash them here, but instead provide a preliminary R package
-(Boettiger et al. 2018) which provides an implementation of the SARSOP
-algorithm and the fisheries POMDP problem, which can be used to
-replicate any of the results presented here and explore other
-variations. Details of the analysis presented here including annotated
-code to reproduce and further explore all of the following results is
-provided in the supplementary material.
-
-\section{Results}\label{results}
+tensors.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
 
@@
 the proposed partially observed Markov decision process (POMDP) method.
-Measurement error increases as a normal distribution with standard
-deviation 0, 0.1, or 0.15, as indicated at the top of the panel.
-Environmental stochasticity is fixed a standard deviation of 0.15 in
-each panel. Carrying capacity K normalized to 1, r = 0.75. Additional
-environmental noise levels and comparison to MSY rather than TAC can be
-found in the supplementary material. \label{sims}}
+Standard deviation from the mean across replicate simulations is shown
+as faint colored bands, indicating significant variation due to
+stochasticity between individual replicates. Measurement error increases
+as a normal distribution with standard deviation 0, 0.1, or 0.15, as
+indicated at the top of the panel. Environmental stochasticity is fixed
+a standard deviation of 0.15 in each panel. Carrying capacity K
+normalized to 1, r = 0.75. Additional environmental noise levels and
+comparison to MSY rather than TAC can be found in the supplementary
+material. \label{sims}}
 \end{figure}
 
-Figure \ref{sims} shows average fish biomass across 100 replicate
-simulations under three different management strategies: constant
-escapement (CE), total allowable catch (TAC; equal to 80\% MSY), and
-partially observed Markov decision process (POMDP) management. Each
-successive panel shows a subsequently higher level of measurement error,
+We consider the average fish biomass across 100 replicate simulations
+under three different management strategies: constant escapement (CE),
+total allowable catch (TAC; equal to 80\% MSY), and partially observed
+Markov decision process (POMDP) management. Each suite of simulations is
+considered under three sequentially higher measurement error regimes,
 from \(\sigma_m = 0\), \(\sigma_m = 0.1\), and \(\sigma_m = 0.15\), as
-indicated. Standard deviation from the mean across replicate simulations
-is shown as faint colored bands, indicating significant variation due to
-stochasticity between individual replicates. In each panel shown in
-Figure \ref{sims} stochastic recruitment (environmental noise) is set to
+indicated, while stochastic recruitment (environmental noise) is set to
 a moderate \(\sigma_g = 0.15\). In the absence of either environmental
-noise or measurement error (Supplemental material, Figure S1) POMDP, CE,
-and MSY would converge to stock at the \(B_{MSY}\), while TAC would
-maintain the stock at a slightly higher level. The first panel of Figure
-\ref{sims} shows that the the introduction of stochastic growth has a
-significant negative impact on the TAC strategy (This impact is even
-more severe for MSY, which always harvests strictly more than TAC by
-definition, as, shown in Supplemental material Figure S2), which results
-in an average biomass significantly lower than \(B_{MSY}\). Without
-measurement error, the CE and POMDP strategies are nearly identical with
-both approximately maintaining the stock at \(B_{MSY}\) despite the
-significant environmental stochasticity. As measurement error increases
-in the subsequent panels, CE and TAC strategies perform increasingly
-poorly, while the POMDP continues to maintain the average stock close to
-\(B_{MSY}\). Notably, CE is significantly more impacted by measurement
-uncertainty than TAC, with CE averaging even lower biomass than TAC
-under moderate measurement uncertainty.
-
-Figure \ref{sims} confirms that the precautionary approach represented
-by the TAC does indeed prove more robust to the problem of measurement
-error than the optimization solution represented by CE. In general, we
-expect that the rigid assumptions required of optimization will lead to
-worse outcomes when those assumptions are not met than we see under the
-corresponding heuristic approach. Yet it is important to bear in mind
-that the opposite pattern was observed in the case of environmental
-stochasticity, in which it was the heuristic TAC strategy rather than
-the CE strategy that lead to the highest over-fishing rate. In contrast,
-the POMDP approach successfully handles both stochasticity and
-observation error, maintaining the stock near \(B_{MSY}\) levels that
-produce the highest yield.
+noise or measurement error POMDP, CE, and MSY would converge to stock at
+the \(B_{MSY}\), while TAC would maintain the stock at a slightly higher
+level (see Supplemental material, Figure S1). The introduction of
+stochastic growth has a significant negative impact on the TAC strategy,
+with a mean biomass significantly lower than \(B_{MSY}\) (first panel,
+Figure \ref{sims}; this impact is even more severe for MSY, see Figure
+S2). Without measurement error, the CE and POMDP strategies are nearly
+identical with both approximately maintaining the stock at \(B_{MSY}\)
+despite the significant environmental stochasticity. As measurement
+error increases in the subsequent panels, CE and TAC strategies perform
+increasingly poorly, while the POMDP continues to maintain the average
+stock close to \(B_{MSY}\). Notably, CE is significantly more impacted
+by measurement uncertainty than TAC, with CE averaging even lower
+biomass than TAC under moderate measurement uncertainty. These results
+confirm that the precautionary approach represented by the TAC does
+indeed prove more robust to the problem of measurement error than the
+optimization solution represented by CE, but is less robust to
+environmental error. In contrast, the POMDP approach successfully
+handles both sources of uncertainty, maintaining higher stocks, near
+\(B_{MSY}\) level, and producing the highest yield.
 
@@
 under increasing levels of measurement error \(\sigma_m\). Net present
-value is th sum of all future harvests discounted over time and averaged
-across 100 replicates simulations under each strategy, as shown in
-Figure \ref{sims}. The value under constant escapement (CE) is optimal
-when measurements are perfect but decreases rapidly with increasing
-measurement error. The more cautious TAC is not economically optimal but
-largely unimpacted by increasing error, while POMDP attains consistently
-high economic yield despite the increasing uncertainty. \label{econ}}
+value is the sum of all future harvests discounted over time and
+averaged across 100 replicates simulations under each strategy, as shown
+in Figure \ref{sims}. The value under constant escapement (CE) is
+optimal when measurements are perfect but decreases rapidly with
+increasing measurement error. The more cautious TAC is not economically
+optimal but largely unimpacted by increasing error, while POMDP attains
+consistently high economic yield despite the increasing uncertainty.
+\label{econ}}
 \end{figure}
@@
 recovery and maintenance of fish biomass) is also borne out in terms of
-economic performance. Figure \ref{econ} shows the mean net present value
-(averaging across replicates and discounting future profits by the
-discount factor \(\gamma\)) for these same simulations at increasing
-levels of measurement uncertainty. As before, environmental
+economic performance. We find the mean net present value (averaging
+across replicates and discounting future profits by the discount factor
+\(\gamma\)) for these same simulations at increasing levels of
+measurement uncertainty (Figure \ref{econ} ). As before, environmental
 stochasticity is set at \(\sigma_g = 0.15\). In the absence of
@@
 
-Differences in attention to prior beliefs,determined by prior
+Differences in attention to prior beliefs, determined by prior
 observations, drive the differences between the POMDP strategy and the
@@
 
-\section{Discussion}\label{discussion}
-
-Using modern algorithms we have been able to crack the nut of
-measurement uncertainty in the optimal management of marine fisheries
-and debunk long-standing paradoxes that increased measurement
-uncertainty should be met with larger harvests. We have demonstrated
-that both existing optimal control approaches and existing precautionary
-approaches result in over-exploitation and long term declines of fish
-stocks, as well as reduced economic value, while in contrast, the POMDP
-approach is able to deliver both ecological and economic recovery
-despite uncertainty in measurements. These results have important
-implications for both global fisheries and ecological management more
-generally. In marine fisheries, evidence of such over-exploitation has
-already been well documented (Worm et al. 2006, 2009; Costello et al.
-2016). Our results suggest that measurement uncertainty may contribute
-to this pattern, as such errors lead to over-exploitation under current
-best practices such as MSY, TAC or CE, which have previously been
-thought to be sufficient to ensure the recovery of stocks world-wide
-(Costello et al. 2016).
-
-Our results also have important consequences for ecological management
-more generally. The study and application of management strategies such
-as MSY and CE underpin much of conservation and natural resource
-management in forest management and many other terrestrial ecosystems
-(e.g. Clark and Mangel 2000). Descision making under uncertainty is a
-central challenge for ecology and conservation biology (Polasky et al.
-2011), which we have long attempted to address through either simplified
-optimal control models or more heuristic approaches such as the
-precautionary principle and resilience thinking (Fischer et al. 2009).
-We have illustrated how both of these approaches can fail to provide an
-adequate strategy when faced with measurement uncertainty. To avoid this
-pitfall, a strategy must be dynamic to make best use of the available
-information without assuming that information is perfect. The
-simplisitic optimal control approach of CE fails by placing too much
-confidence in measurements even when they are significantly at odds with
-prior expectations. Similarly, the TAC approach also fails due to it's
-reliance on a simple one-size-fits-all strategy, not reducing catch
-quotas sufficiently when stocks are too low but also failing to increase
-them when stocks are higher. It is precisely this ability to adapt the
-policy in response to available information, as illustrated in Figure
-\ref{policy}, that makes the POMDP approach successful. Nevertheless, it
-may be tempting to see these results as either trivial or impractical,
-but a closer inspection reveals they are neither.
-
-First, it may seem obvious that POMDP would out-preform more primitive
-optimal control solutions such as CE, but this was hardly a forgone
-conclusion. Indeed, the previous example of Reed (1979) had proven that
-the complex stochastic dynamic programming solutions required to
-accommodate stochastic dynamics would perform no better than the simple
-maximization based purely on the deterministic model. If stochasticity
-doesn't matter, it is easy to believe that measurement error would be
-minor at best. In fact, we find that once we account for measurement
-error, the effects of both stochasticity and measurement have a
-significant influence over the optimal policy. It is similarly tempting
-to dismiss the better performance of POMDP against a more
-rule-of-thumb{[}\^{}1{]} based strategy like the TAC scenario considered
-here (fishing at 80\% MSY), but once again this was not a forgone
-conclusion. While a better economic return might reasonably be expected
-of the optimization approach, it was not obvious that POMDP would also
-be at times more cautious than the rule of thumb, able to deliver both
-higher economic returns by harvesting more when it was safe to do so,
-and higher ecological recovery by harvesting even less than TAC when
-necessary. These results highlight the difficulty of finding effective
-heuristic rules and underscore the power and utility of more
-sophisticated strategies such as POMDP which can make the best use of
-all available information.
-
-\^{}1: Given the sophisticated policies often in place to determine
-actual TAC levels, one might object to our characterization of this as
-merely a rule-of-thumb. We use this term merely to distinguish it from
-the optimal control approach -- that despite technical precision, TAC
-adjustments are set by a heuristic instead of the result of some dynamic
-optimization algorithm as in CE and POMDP.
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+Using modern algorithms we have been able to both crack the nut of
+measurement uncertainty in the optimal management of ecological
+populations such as marine fisheries, debunking a long-standing paradox
+created by previous literature which has argued that increased
+measurement uncertainty should be met with equal or larger harvests
+(e.g. Reed 1979; Clark and Kirkwood 1986; Sethi et al. 2005). By using
+powerful algorithms developed first the robotics literature to solve
+POMDPs (Kurniawati et al. 2008), we have been able to tackle this
+optimal control problem directly rather than resorting to the
+simplifying assumptions used in previous work. Our solution confirms the
+initution of the precautionary principle in a quantitatively precise
+form: under most (but not all) conditions, \emph{greater uncertainty
+results in smaller optimal harvests}. Despite frequently proposing
+smaller harvests than existing approaches, the POMDP solution achieves
+higher long-term economic value as well as higher average stock sizes
+over the long term.
+
+This lesson has clear relevance in the context of fisheries management.
+Here, practice is already wiser than theory, where the optimal control
+based on the CE results of Clark (1973) and Reed (1979) is limited to
+fisheries such as salmon where estimates in stock size may be less
+significant than in purely marine species. To our knowledge, no fishery
+has intentionally put the more aggressive harvest solutions of Clark and
+Kirkwood (1986) and Sethi et al. (2005) into practice. Yet this does not
+lift the spectre of these methods, where CE-based predictions feature
+prominently in global fisheries analyses as the economically optimal
+strategy and the fastest route to recovery (Costello et al. 2016;
+Burgess et al. 2018). Our results raise serious questions about the
+optimistic forecasts predicted under these CE-based policies. Meanwhile,
+MSY remains the dominate standard international law (Mace 2001), and
+many well-managed US fisheries rely on a version of the TAC based
+precautionary adjustment to MSY to comply with catch limits under
+Magnuson-Stevens Act. Our results show that these strategies are less
+sensitive to measurement error, but not immune. Comparisons against more
+heuristic decision methods are more difficult: management strategy
+evaluation (MSE; Smith 1994; Punt et al. 2016) will depend on precisely
+what management scenarios are considered. Under sufficient uncertainty,
+both the existing optimal and precautionary approaches can lead to
+declines in biomass and economic returns which could be avoided by
+POMDP-based management.
+
+This conclusion has important broader implications for ecological
+management more generally. Decision making under uncertainty is a
+central challenge for ecology and conservation biology, which frequently
+pits optimal control based approaches against more heuristic approaches
+such as resilience thinking, scenario planning, and the precautionary
+principle (Polasky et al. 2011). With economists frequently favoring the
+precise, stylized approach of optimal control and ecologists embracing
+the realities of greater complexity and uncertainty echoed in more
+heursitic approaches, this division can sometimes appear to pit economic
+outcomes against ecological objectives. That perspective is echoed in
+the paradox discussed here, in which economic optimization ignoring
+uncertainty has argued in cold mathematical logic for a rejection of the
+precationary principle. Our results demonstrate that this divide is
+artificial. With sufficiently powerful algorithms we can go some ways to
+realizing the call of Fischer et al. (2009) to combine more ecologically
+realistic assumptions such as measurement error with optimal control
+approaches in ecological management. Both sides are vindicated in our
+solution: economic optimization under POMDP does not controdict the
+precautionary principle, while under POMDP, optimal control can also
+achieve better ecological outcomes than blunt precautionary rule
+(i.e.~TAC, which reduced all harvests by 20\% below MSY target). As
+Fischer et al. (2009) predicted, unifying optimal control with more
+resilience-inspired approaches is win-win.
+
+If accounting for measurement uncertainty can resolve previous paradoxes
+and divides, it also raises additional challenges. The approach
+discussed here is vulnerable both to the charge of being too simple to
+be realistic while also being too complicated to be feasible. These
+issues are as easily recognized in other ecological management contexts,
+from harvests of wild game or forestry to pest outbreaks, (e.g. Richards
+et al. 1999; {\textbf{???}}; Blackwood et al. 2010; Shea et al. 2014;
+Dee et al. 2017; Iacona et al. 2017; Li et al. 2017) as in fisheries,
+where ecological complexity is invariably high and management capacity
+invariably limited. We address each of these in turn.
+
+Are the dynamics considered here too simple? Our underlying model is
+undoubtably much simpler than many of the age or stage structured models
+used in modern fisheries management -- aspects whose importance to
+ecological dynamics have been established well beyond fisheries as well.
+Yet it is critical to bear in mind that this stylized one-dimensional
+model is used only to compare the relative performance of the different
+decision strategies considered here (TAC, CE, and POMDP) in order to
+illustrate the potential importance of uncertainty in measurement. There
+is little reason to believe CE or TAC would perform better under more
+complex cases such as age-structured models that only further deviate
+from the assumptions under which they were derived (e.g. Holden and
+Conrad 2015). POMDP approaches are in no way incompatible with more
+complex models, whether that be age structure or even more ambitious
+non-parametric representations defined directly from data using Guassian
+Processes or other machine learning techniques (e.g. Boettiger et al.
+2015). Simple models have always served as tools for comparison and
+intuition, and serve the same illustrative role here.
+
+Is this approach too complex to be feasible? The importance of simple
+and effective rules of thumb in conservation management has been well
+documented (e.g. Chadès et al. 2011). Sophisticated and computationally
+intensive approaches such as POMDP may seem improbable at a time when
+many areas of natural resource management, even simple, rule-of-thumb
+based methods struggle to take root. There is no doubt that the solution
+method needed here is computationally intensive even for the simple
+model considered here, and will only become more so under greater
+ecological complexity. Indeed, it was precisely such computational
+limitations that led to the shortcuts in work such as Clark and Kirkwood
+(1986), with the ensuing potentially misleading conclusions. While in
+the following decades ecologists largely turned away from
+computationally intensive optimizations required for even simple models
+with stochastic dynamic programming, (while continuing to construct ever
+more complicated hierarchical models), engineers have chipped away at
+the decision problem to reveal algorithms such as SARSOP that make much
+larger dynamic programming problems tractable. Yes, even more
+sophisticated algorithms will be needed to address the limitations
 
@@
 
-Third: are the examples considered here too simple? The strategies we
-have compared against underpin both current practice and management
-aspirations. Together, MSY and TAC-based management reflects much of
-modern best practice as it is understood in fisheries management and
-elsewhere. While real-world application of these approaches frequently
-involves more complex models, the models we have considered here are the
-very ones under which these strategies were designed and tested, and
-there is little reason to believe they would perform better under more
-complex cases such as age-structured models that do not always meet
-those assumptions (e.g. Holden and Conrad 2015). The relatively poor
-ecological and economic performance of these common best-practices
-relative to the alternative approach of POMDP is thus reason enough for
-concern. This issue is further underscored by the reality of both
-long-running ecological deterioration (Worm et al. 2006, 2009) and
-economic costs of global fisheries.
-
-Fourth: is the approach taken here too complex to be feasible? The
-importance of simple and effective rules of thumb in conservation
-management has been well documented (e.g. Chadès et al. 2011).
-Sophisticated and computationally intensive approaches such as POMDP may
-seem improbable at a time when many areas of natural resource
-management, even simple, rule-of-thumb based methods struggle to take
-root. And yet, in this same age we rely on far more complex and
-computationally intensive algorithms to perform far more trivial tasks
-such as determining what advertisements we see. Surely a planet that can
-afford such complex optimization for how advertising space is allocated
-on a small screen can one day manage it's natural resources and
-conservation efforts with such tools?
-
-\subsection{Future directions}\label{future-directions}
+\hypertarget{future-directions}{%
+\subsection{Future directions}\label{future-directions}}
 
@@
 
-\section*{References}\label{references}
+\hypertarget{references}{%
+\section*{References}\label{references}}
 \addcontentsline{toc}{section}{References}
@@
 \hypertarget{refs}{}
-\hypertarget{ref-Arnason2012}{}
+\leavevmode\hypertarget{ref-Arnason2012}{}%
 Arnason, R. 2012. Property rights in fisheries: How much can individual
@@
 
-\hypertarget{ref-Beverton1957}{}
+\leavevmode\hypertarget{ref-Beverton1957}{}%
 Beverton, R., and S. Holt. 1957. On the Dynamics of Exploited Fish
@@
 
-\hypertarget{ref-Boettiger2015}{}
+\leavevmode\hypertarget{ref-Blackwood2010}{}%
+Blackwood, J. C., A. Hastings, and C. Costello. 2010. Cost-effective
+management of invasive species using linear-quadratic control.
+Ecological Economics 69:519--527.
+
+\leavevmode\hypertarget{ref-Boettiger2015}{}%
 Boettiger, C., M. Mangel, and S. Munch. 2015. Avoiding tipping points in
@@
 
-\hypertarget{ref-sarsop-pkg}{}
+\leavevmode\hypertarget{ref-sarsop-pkg}{}%
 Boettiger, C., J. Ooms, and M. Memarzadeh. 2018. Sarsop: Approximate
@@
 
-\hypertarget{ref-Botsford1997}{}
+\leavevmode\hypertarget{ref-Botsford1997}{}%
 Botsford, L. W., J. C. Castilla, and C. Peterson. 1997. The management
@@
 
-\hypertarget{ref-Britten2017}{}
-Britten, G., M. Dowd, L. Kanary, and B. Worm. 2017. Extended fisheries
-recovery timelines in a changing environment. Nature Communications.
-
-\hypertarget{ref-Chades2011}{}
+\leavevmode\hypertarget{ref-Britten2017}{}%
+Britten, G. L., M. Dowd, L. Kanary, and B. Worm. 2017. Extended
+fisheries recovery timelines in a changing environment. Nature
+Communications 8:15325.
+
+\leavevmode\hypertarget{ref-Bunnefeld2011}{}%
+Bunnefeld, N., E. Hoshino, and E. Milner-Gulland. 2011. Management
+stratefy evaluation: a powerful tool for conservation? Trends in Ecology
+and Evolution 26:441--447.
+
+\leavevmode\hypertarget{ref-Burgess2018}{}%
+Burgess, M., G. McDermott, B. Owashi, L. Peavey Reeves, T. Clavelle, D.
+Ovando, B. Wallace, et al. 2018. Protecting marine mammals, turtles, and
+birds by rebuilding global fisheries. Science 359:1255--1258.
+
+\leavevmode\hypertarget{ref-Chades2011}{}%
 Chadès, I., T. G. Martin, S. Nicol, M. A. Burgman, H. P. Possingham, and
@@
 
-\hypertarget{ref-Chades2008}{}
+\leavevmode\hypertarget{ref-Chades2008}{}%
 Chadès, I., E. McDonald-Madden, M. a McCarthy, B. Wintle, M. Linkie, and
@@
 
-\hypertarget{ref-Clark1973}{}
+\leavevmode\hypertarget{ref-Clark1973}{}%
 Clark, C. W. 1973. Profit maximization and the extinction of animal
@@
 
-\hypertarget{ref-Clark1990}{}
+\leavevmode\hypertarget{ref-Clark1990}{}%
 Clark, C. W. 1990. Mathematical Bioeconomics: The Optimal Management of
@@
 
-\hypertarget{ref-Clark1986}{}
+\leavevmode\hypertarget{ref-Clark1986}{}%
 Clark, C. W., and G. P. Kirkwood. 1986. On uncertain renewable resource
@@
 
-\hypertarget{ref-Clark2000}{}
-Clark, C. W., and M. Mangel. 2000. Dynamic state variable models in
-ecology. Oxford University Press, Oxford.
-
-\hypertarget{ref-Costello2016}{}
+\leavevmode\hypertarget{ref-Costello2016}{}%
 Costello, C., D. Ovando, T. Clavelle, C. K. Strauss, R. Hilborn, M. C.
@@
 
-\hypertarget{ref-nimble}{}
+\leavevmode\hypertarget{ref-Dee2017}{}%
+Dee, L. E., M. De Lara, C. Costello, and S. D. Gaines. 2017. To what
+extent can ecosystem services motivate protecting biodiversity? Ecology
+Letters 20:935--946.
+
+\leavevmode\hypertarget{ref-nimble}{}%
 de Valpine, P., D. Turek, C. J. Paciorek, C. Anderson-Bergman, T.
@@
 
-\hypertarget{ref-Dichmont2016}{}
+\leavevmode\hypertarget{ref-Dichmont2016}{}%
 Dichmont, C. M., R. A. Deng, A. E. Punt, J. Brodziak, Y.-J. Chang, J. M.
@@
 
-\hypertarget{ref-Engen1997}{}
+\leavevmode\hypertarget{ref-Engen1997}{}%
 Engen, S., R. Lande, and B. Sæther. 1997. Harvesting strategies for
@@
 
-\hypertarget{ref-Fackler2014b}{}
+\leavevmode\hypertarget{ref-Fackler2014b}{}%
 Fackler, P., and R. Haight. 2014. Monitoring as a partially observable
@@
 
-\hypertarget{ref-Fackler2014}{}
+\leavevmode\hypertarget{ref-Fackler2014}{}%
 Fackler, P., and K. Pacifici. 2014. Addressing structural and
@@
 
-\hypertarget{ref-Fischer2009}{}
+\leavevmode\hypertarget{ref-Fischer2009}{}%
 Fischer, J., G. D. Peterson, T. a Gardner, L. J. Gordon, I. Fazey, T.
@@
 
-\hypertarget{ref-Gordon1954}{}
+\leavevmode\hypertarget{ref-Gordon1954}{}%
 Gordon, H. S., and C. Press. 1954. The Economic Theory of a
@@
 
-\hypertarget{ref-Hilborn2010}{}
+\leavevmode\hypertarget{ref-Hilborn2010}{}%
 Hilborn, R. 2010. Pretty Good Yield and exploited fishes. Marine Policy
@@
 
-\hypertarget{ref-Holden2015}{}
+\leavevmode\hypertarget{ref-Holden2015}{}%
 Holden, M., and J. Conrad. 2015. Optimal escapement in stage-structured
@@
 
-\hypertarget{ref-Holling1973}{}
+\leavevmode\hypertarget{ref-Holling1973}{}%
 Holling, C. S. 1973. Resilience and Stability of Ecological Systems.
@@
 
-\hypertarget{ref-Johnson2015}{}
+\leavevmode\hypertarget{ref-Iacona2017}{}%
+Iacona, G. D., H. P. Possingham, and M. Bode. 2017. Waiting can be an
+optimal conservation strategy, even in a crisis discipline. Proceedings
+of the National Academy of Sciences 201702111.
+
+\leavevmode\hypertarget{ref-Johnson2015}{}%
 Johnson, F. A., and B. K. Williams. 2015. A Decision-Analytic Approach
@@
 
-\hypertarget{ref-Kaelbling1998}{}
+\leavevmode\hypertarget{ref-Kaelbling1998}{}%
 Kaelbling, L. P., M. L. Littman, and A. R. Cassandra. 1998. Planning and
@@
 
-\hypertarget{ref-Kurniawati2008}{}
+\leavevmode\hypertarget{ref-Kurniawati2008}{}%
 Kurniawati, H., D. Hsu, and W. S. Lee. 2008. SARSOP : Efficient
@@
 
-\hypertarget{ref-Larkin1977}{}
+\leavevmode\hypertarget{ref-Larkin1977}{}%
 Larkin, P. a. 1977. An Epitaph for the Concept of Maximum Sustained
@@
 
-\hypertarget{ref-Levin2008}{}
+\leavevmode\hypertarget{ref-Levin2008}{}%
 Levin, S. A., and J. Lubchenco. 2008. Resilience, Robustness, and Marine
@@
 
-\hypertarget{ref-Ludwig1981}{}
+\leavevmode\hypertarget{ref-Li2017}{}%
+Li, S.-L., O. N. Bjørnstad, M. J. Ferrari, R. Mummah, M. C. Runge, C. J.
+Fonnesbeck, M. J. Tildesley, et al. 2017. Essential information:
+Uncertainty and optimal control of Ebola outbreaks. Proceedings of the
+National Academy of Sciences 114:5659--5664.
+
+\leavevmode\hypertarget{ref-Ludwig1981}{}%
 Ludwig, D., and C. Walters. 1981. Measurement errors and uncertainty in
@@
 
-\hypertarget{ref-Ludwig1982}{}
+\leavevmode\hypertarget{ref-Ludwig1982}{}%
 Ludwig, D., and C. J. Walters. 1982. Optimal harvesting with imprecise
@@
 
-\hypertarget{ref-Mace2001}{}
-Mace, P. M. 2001. A new role for MSY in single-species and
-ecosystem\(\backslash\)rapproaches to fisheries stock assessment and
-management. Fish and Fisheries 2:2--32.
-
-\hypertarget{ref-Mangel1988}{}
+\leavevmode\hypertarget{ref-Mace2001}{}%
+Mace, P. M. 2001. A new role for MSY in single-species and ecosystem
+approaches to fisheries stock assessment and management. Fish and
+Fisheries 2:2--32.
+
+\leavevmode\hypertarget{ref-Mangel1985}{}%
+Mangel, M. 1985. Decision and control in uncertain resource systems 255.
+
+\leavevmode\hypertarget{ref-Mangel1988}{}%
 Mangel, M., and C. W. Clark. 1988. Dynamic Modeling in Behavioral
@@
 
-\hypertarget{ref-Marescot2013}{}
+\leavevmode\hypertarget{ref-Marescot2013}{}%
 Marescot, L., G. Chapron, I. Chadès, P. L. Fackler, C. Duchamp, E.
 Marboutin, and O. Gimenez. 2013. Complex decisions made simple: a primer
-on stochastic dynamic programming. Methods in Ecology and Evolution
-n/a--n/a.
-
-\hypertarget{ref-May1977}{}
+on stochastic dynamic programming. (R. Freckleton, ed.)Methods in
+Ecology and Evolution 4:872--884.
+
+\leavevmode\hypertarget{ref-May1977}{}%
 May, R. M. 1977. Thresholds and breakpoints in ecosystems with a
@@
 
-\hypertarget{ref-Memarzadeh2016b}{}
+\leavevmode\hypertarget{ref-Memarzadeh2016b}{}%
 Memarzadeh, M., and M. Pozzi. 2016. Value of information in sequential
@@
 
-\hypertarget{ref-Moxnes2003}{}
+\leavevmode\hypertarget{ref-Moxnes2003}{}%
 Moxnes, E. 2003. Uncertain measurements of renewable resources:
@@
 
-\hypertarget{ref-Ong2010}{}
+\leavevmode\hypertarget{ref-Nicol2011}{}%
+Nicol, S., and I. Chadès. 2011. Beyond stochastic dynamic programming: a
+heuristic sampling method for optimizing conservation decisions in very
+large state spaces. Methods in Ecology and Evolution 2:221--228.
+
+\leavevmode\hypertarget{ref-Ong2010}{}%
 Ong, S., S. Png, D. Hsu, and W. Lee. 2010. Planning under uncertainty
@@
 
-\hypertarget{ref-Petchy2015}{}
+\leavevmode\hypertarget{ref-Petchy2015}{}%
 Petchey, O. L., M. Pontarp, T. M. Massie, S. Kéfi, A. Ozgul, M.
@@
 
-\hypertarget{ref-Pineau2003}{}
+\leavevmode\hypertarget{ref-Pineau2003}{}%
 Pineau, J., G. Gordon, and S. Thrun. 2003. Point-based value iteration:
@@
 
-\hypertarget{ref-Polasky2011}{}
+\leavevmode\hypertarget{ref-Polasky2011}{}%
 Polasky, S., S. R. Carpenter, C. Folke, and B. Keeler. 2011.
 Decision-making under great uncertainty: environmental management in an
-era of global change. Trends in ecology \& evolution 1--7.
-
-\hypertarget{ref-Reed1979}{}
+era of global change. Trends in Ecology \& Evolution 26:398--404.
+
+\leavevmode\hypertarget{ref-Punt2016}{}%
+Punt, A. E., D. S. Butterworth, C. L. de Moor, J. A. A. De Oliveira, and
+M. Haddon. 2016. Management strategy evaluation: best practices. Fish
+and Fisheries 17:303--334.
+
+\leavevmode\hypertarget{ref-Reed1979}{}%
 Reed, W. J. 1979. Optimal escapement levels in stochastic and
@@
 
-\hypertarget{ref-RAM}{}
+\leavevmode\hypertarget{ref-RAM}{}%
 Ricard, D., C. Minto, O. Jensen, and J. Baum. 2011. Examining the
@@
 
-\hypertarget{ref-Roughgarden1996}{}
+\leavevmode\hypertarget{ref-Richards1999}{}%
+Richards, S. A., H. P. Possingham, and J. Tizard. 1999. Optimal Fire
+Management for Maintaining Community Diversity. Ecological Applications
+9:880--892.
+
+\leavevmode\hypertarget{ref-Roughgarden1996}{}%
 Roughgarden, J. E., and F. Smith. 1996. Why fisheries collapse and what
-to do about it. Proceedings of the National Academy of Sciences 93:5078.
-
-\hypertarget{ref-Schaefer1954}{}
+to do about it. Proceedings of the National Academy of Sciences
+93:5078--5083.
+
+\leavevmode\hypertarget{ref-Schaefer1954}{}%
 Schaefer, M. B. 1954. Some aspects of the dynamics of populations
@@
 
-\hypertarget{ref-Sethi2005}{}
+\leavevmode\hypertarget{ref-Sethi2005}{}%
 Sethi, G., C. Costello, A. Fisher, M. Hanemann, and L. Karp. 2005.
@@
 
-\hypertarget{ref-Smallwood1973}{}
+\leavevmode\hypertarget{ref-Shea2014}{}%
+Shea, K., M. J. Tildesley, M. C. Runge, C. J. Fonnesbeck, and M. J.
+Ferrari. 2014. Adaptive Management and the Value of Information:
+Learning Via Intervention in Epidemiology. PLoS Biology 12:9--12.
+
+\leavevmode\hypertarget{ref-Smallwood1973}{}%
 Smallwood, R. D., and E. J. Sondik. 1973. The Optimal Control of
-Partially Observable Markov Decision Processes over a Finite Horizon.
-
-\hypertarget{ref-Sondik1978}{}
+Partially Observable Markov Processes Over a Finite Horizon. Operations
+Research 21:1071--1088.
+
+\leavevmode\hypertarget{ref-Smith1994}{}%
+Smith, A. 1994. Management strategy evaluation -- the light on the hill.
+Population dynamics for fisheries management 249--253.
+
+\leavevmode\hypertarget{ref-Sondik1978}{}%
 Sondik, E. J. 1978. The Optimal Control of Partially Observable Markov
@@
 
-\hypertarget{ref-Walters1976}{}
+\leavevmode\hypertarget{ref-Walters1976}{}%
 Walters, C. J., and R. Hilborn. 1976. Adaptive Control of Fishing
@@
 
-\hypertarget{ref-Walters1978}{}
+\leavevmode\hypertarget{ref-Walters1978}{}%
 ---------. 1978. Ecological Optimization and Adaptive Management. Annual
@@
 
-\hypertarget{ref-Williams2001}{}
+\leavevmode\hypertarget{ref-Williams2001}{}%
 Williams, B. K. 2001. Uncertainty , learning , and the optimal
@@
 
-\hypertarget{ref-Williams2011}{}
+\leavevmode\hypertarget{ref-Williams2011}{}%
 ---------. 2011. Resolving structural uncertainty in natural resources
@@
 
-\hypertarget{ref-FAO2009}{}
+\leavevmode\hypertarget{ref-FAO2009}{}%
 World Bank and FAO. 2009. The Sunken Billions. The Economic
@@
 
-\hypertarget{ref-Worm2006}{}
-Worm, B., E. B. Barbier, N. Beaumont, J. E. Duffy, C. Folke, B. S.
-Halpern, J. B. C. Jackson, et al. 2006. Impacts of biodiversity loss on
-ocean ecosystem services. Science (New York, N.Y.) 314:787--90.
-
-\hypertarget{ref-Worm2009}{}
-Worm, B., R. Hilborn, J. K. Baum, T. A. Branch, J. S. Collie, C.
-Costello, M. J. Fogarty, et al. 2009. Rebuilding global fisheries.
-Science (New York, N.Y.) 325:578--85.
-
 \end{document}

### 2018-08-17 manuscripts/manuscript.Rmd
revised introduction

@@
 Imperfect information is ubiquitous in ecological management and conservation decision making.
-While the pressing concerns of global change have put the spotlight on *forecasting*, i.e.
+While the pressing concerns of global change have put the spotlight on forecasting, i.e.
 the uncertainty of the future [e.g.
-@Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today?
+@Petchy2015], management decisions must also contend with uncertainty of the present: How many fish are in the sea today?
 What regions harbor the most fragile biodiversity?
-How have invasive species transformed native communities?
-A recognition of the importance of uncertainty has become widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods to build better models.
-As we estimate ever more sophisticated models, it may be easy to forget that a model is not itself a decision, particularly when the estimated model includes considerable uncertainty about future outcomes.
-Knowing the rules of the chess board is not the same thing as knowing how to win.
-A model only gives us the rules of the game.
-The science of decision theory helps us know how to play.
-Yet while our models have become much more sophisticated, decision theory in ecological management frequently ignores the very uncertainty we so carefully include when estimating the model.
-Previous shortcuts in decision theory have led to paradoxical results, such as the suggestion that increased uncertainty in fisheries stock sizes should justify unchanged or even increased harvest rates.
-Borrowing decision theory methodology first developed in fields such as robotics, we are able to resolve this paradox, underscoring the importance of appropriately accounting for uncertainty in decisions just as we already do in models.
-This insight has important implications in both fisheries and how ecosystem management deals with uncertainty more generally.
+How do we make use of a rapidly expanding but opportunistically collected data, such as in citizen science efforts, which may involve strong sampling biases?
+Much of our focus in dealing with uncertainty has been through the lens of statistical approaches to model fitting.
+The rise of approaches such as Hierarchical Bayesian Modeling have given us sophisticated tools to reflect uncertainty in our data and uncertainty in our knowledge of mechanisms and parameters in our models [].
+But estimating a model is not the same as making a decision, particularly when a model leaves us with considerable uncertainty about current and future states.
+A model can only provide the rules of the chess board.
+It is the role of decision theory to provide us with a strategy.
 
+Decision theory has a long history in the ecological literature [e.g @Schaefer1954; @Walters1978; @Clark1990; @Shea1999; @Polasky2011] which has sometimes been overshadowed by recent emphasis on model fitting.
+While in principle a model will be updated throughout the decision process (as we will discuss), it is convenient to think of the model as "given," estimated, with uncertainty, using the best available methods.
+The goal of a decision theory is not maximize some abstract notion of model fit (least squares, likelihood, etc), but a more practical notion of utility (happiness, economic value, or any combination of stakeholder objectives, see @Halpern2013).
 Approaches for decision making under uncertainty in ecological systems can be divided into two camps: (1) those based upon "optimal control" solutions, and usually favored by natural resource economists, and (2) those based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].
@@
 We illustrate how the limitations of these approaches have manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.
-We compare the results of managing simulated fish stocks under (A) the existing optimal control approach, (B) a heuristic approach that is more typical of today's management practices, and (C) a unified approach that provides a more rigorous treatment of uncertainty.
 
@@
 In light of this, we begin with a brief background on existing decision theory approaches common in fisheries management.
+
+In this paper, we examine three strategies to managing a natural population, such as a fishery, under such uncertainty.
+This problem has a long history in both theory and practice, where seemingly innoccous simplifying assumptions have led to a deep paradox in how managers should respond to uncertainty.
+The precautionary principle [@] would seem to suggest that the greater the uncertainty in our models, the more cautious we should be in management policies.
+Yet a well-known mathematical proof by @Reed1979 established that under quite general conditions, the optimal strategy under uncertainty is no different than under no uncertainty at all.
+Attempts to resolve this have since only deepened this paradox, arguing that adding additional sources of uncertainty, such as imperfect measurements, should lead to optimal harvest that increases with greater uncertainty [e.g.
+@Clark1986; @Sethi2005].
+Meanwhile, much of actual practice has (perhaps fortunately, as we shall see) ignored these results in favor of a more heuristic application of the precautionary principle.
+Here, we show that this paradox can be resolved by uncovering previous shortcuts and tackling the optimal management problem head-on, which we can solve thanks to the computational efficiency of sophisticated algorithms developed more recently in the field of robotics.
+This approach allows us to reject this paradox largely in favor of the precautionary principle.
+Further, our comparisons show that widely recognized existing approaches in both theory and practice are not by themselves sufficient to manage a population under substantial uncertainty over the long term, but our recently borrowed approach from robotics literature can under identical conditions sustain both higher biomass and higher economic yields than those approaches.
+This conclusion has important implications not only for fisheries but eocsystem more broadly, underscoring the danger of not correctly accounting for measurement uncertainty when making decisions, and further to the significant opportunity to learn from fields such as robotics when faced with complex and uncertain decisions.
 

### 2018-08-17 manuscripts/manuscript.Rmd
nearly there

@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = 1)
 library(tidyverse)
@@
 Much of our focus in dealing with uncertainty has been through the lens of statistical approaches to model fitting.
-The rise of approaches such as Hierarchical Bayesian Modeling have given us sophisticated tools to reflect uncertainty in our data and uncertainty in our knowledge of mechanisms and parameters in our models [].
+The rise of approaches such as Hierarchical Bayesian Modeling have given us sophisticated tools to reflect uncertainty in our data and uncertainty in our knowledge of mechanisms and parameters in our models [@Ellison2004].
 But estimating a model is not the same as making a decision, particularly when a model leaves us with considerable uncertainty about current and future states.
@@
 
-Decision theory has a long history in the ecological literature [e.g @Schaefer1954; @Walters1978; @Clark1990; @Shea1999; @Polasky2011] which has sometimes been overshadowed by recent emphasis on model fitting.
+Decision theory has a long history in the ecological literature [e.g @Schaefer1954; @Walters1978; @Clark1990; @Shea1998; @Polasky2011] which has sometimes been overshadowed by recent emphasis on model fitting.
 While in principle a model will be updated throughout the decision process (as we will discuss), it is convenient to think of the model as "given," estimated, with uncertainty, using the best available methods.
@@
 This problem has a long history in both theory and practice, where seemingly innoccous simplifying assumptions have led to a deep paradox in how managers should respond to uncertainty.
-The precautionary principle [@] would seem to suggest that the greater the uncertainty in our models, the more cautious we should be in management policies.
+The precautionary principle [@Kriebel2001] would seem to suggest that the greater the uncertainty in our models, the more cautious we should be in management policies.
 Yet a well-known mathematical proof by @Reed1979 established that under quite general conditions, the optimal strategy under uncertainty is no different than under no uncertainty at all.
@@
 This assumption has occasionally been reversed to introduce some additional uncertainty from stochastic recruitment immediately prior to setting the harvest quota [e.g.
-@Clark1986; @Weitzmann2001], which appears to mimic uncertainty in measurements but does not allow uncertainty to accumulate since the measurement step still occurs without error.
+@Clark1986; @Weitzman2002], which appears to mimic uncertainty in measurements but does not allow uncertainty to accumulate since the measurement step still occurs without error.
 
@@
 
+## Simulations
+
+We consider the average fish biomass across 100 replicate simulations of the same stock dynamics (Eq 4) under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
+Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$ (no measurement errors), $\sigma_m = 0.1$ (low error), and $\sigma_m = 0.15$ (moderate error), as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
+Empirical estimates of measurement error in stock sizes vary widely, with estimates often dependent upon methodological assumptions.
+Early estimates have put possible error as high as 50% [@Clark1986] and more recent analyses have suggested an average of closer to 36%, though with cases that may be much higher [@Ralston2011].
+Consequently, range between 0 to 15% provides a probably conservative estimate of typical unceratinty ranges, which we show is nevertheless large enough to drive substantially negative ecological and economic outcomes under management strategies that do not appropriately account for that error over the long run.
+
 # Results
@@
 
-We consider the average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.
-Each suite of simulations is considered under three sequentially higher measurement error regimes, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated, while stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.
 In the absence of either environmental noise or measurement error POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level (see Supplemental material, Figure S1).

### 2018-08-17 manuscripts/manuscript.tex
nearly there

@@
 conservation decision making. While the pressing concerns of global
-change have put the spotlight on \emph{forecasting}, i.e.~the
-uncertainty of the future (e.g. Petchey et al. 2015), management
-decisions must also contend with uncertainty of the \emph{present}: How
-many fish are in the sea today? What regions harbor the most fragile
-biodiversity? How have invasive species transformed native communities?
-A recognition of the importance of uncertainty has become widespread in
-the context of estimating ecological models from available data, thanks
-in large part to the rise of approaches such as Hierarchical Bayesian
-Modeling, along with convenient software tools for implementing these
-methods. In sharp contrast to that shift, the treatment of uncertainty
-in decisions -- how the uncertainty in models should be reflected in
-scientific recommendations, remains far less developed, more
-qualitative, and under-served by available tools. We illustrate how
-uncertainty in measurement has a dramatic negative impact on outcomes
-under a range of existing approaches, which can be avoided through the
-use of more formal approaches that have since been refined in fields
-such as robotics. This insight both resolves a long-standing paradox in
-fisheries management and has important implications for how ecosystem
-management deals with uncertainty more generally.
-
+change have put the spotlight on forecasting, i.e.~the uncertainty of
+the future (e.g. Petchey et al. 2015), management decisions must also
+contend with uncertainty of the present: How many fish are in the sea
+today? What regions harbor the most fragile biodiversity? How do we make
+use of a rapidly expanding but opportunistically collected data, such as
+in citizen science efforts, which may involve strong sampling biases?
+Much of our focus in dealing with uncertainty has been through the lens
+of statistical approaches to model fitting. The rise of approaches such
+as Hierarchical Bayesian Modeling have given us sophisticated tools to
+reflect uncertainty in our data and uncertainty in our knowledge of
+mechanisms and parameters in our models (Ellison 2004). But estimating a
+model is not the same as making a decision, particularly when a model
+leaves us with considerable uncertainty about current and future states.
+A model can only provide the rules of the chess board. It is the role of
+decision theory to provide us with a strategy.
+
+Decision theory has a long history in the ecological literature (e.g
+Schaefer 1954; Walters and Hilborn 1978; Clark 1990; Shea 1998; Polasky
+et al. 2011) which has sometimes been overshadowed by recent emphasis on
+model fitting. While in principle a model will be updated throughout the
+decision process (as we will discuss), it is convenient to think of the
+model as ``given,'' estimated, with uncertainty, using the best
+available methods. The goal of a decision theory is not maximize some
+abstract notion of model fit (least squares, likelihood, etc), but a
+more practical notion of utility (happiness, economic value, or any
+combination of stakeholder objectives, see Halpern et al. (2013)).
 Approaches for decision making under uncertainty in ecological systems
@@
 uncertainty and the rigor of optimization to resolve a long-standing
-paradox and suggest a more robust approach to management. We compare the
-results of managing simulated fish stocks under (A) the existing optimal
-control approach, (B) a heuristic approach that is more typical of
-today's management practices, and (C) a newly proposed approach that
-provides a more rigorous treatment of uncertainty.
+paradox and suggest a more robust approach to management.
 
@@
 found in a wide range of population and ecosystem management challenges,
-including fire ecology (Richards et al. 1999), invasive species
-({\textbf{???}}; Blackwood et al. 2010), disease outbreaks (Shea et al.
-2014; Li et al. 2017), and protecting biodiversity (Dee et al. 2017;
-Iacona et al. 2017). Yet fisheries conservation and management has long
-been both a crucible and proving ground for the theory of ecological
-management more generally, including topics such as adaptive management
-(Walters and Hilborn 1978), ecosystem-based management (Levin and
-Lubchenco 2008) and resilience thinking (Holling 1973; May 1977), while
-also giving rise to the sub-discipline of resource economics (Gordon and
-Press 1954; Schaefer 1954; Beverton and Holt 1957), thanks to its global
-relevance, long history, and readily available data. Despite the
-generality of the models and issue of measurement uncertainty considered
-here, any application must be grounded in both the history and context
-of the particular decision problem. In light of this, we begin with a
-brief background on existing decision theory approaches common in
-fisheries management.
+including fire ecology (Richards et al. 1999), invasive species (Shea
+1998; Blackwood et al. 2010), disease outbreaks (Shea et al. 2014; Li et
+al. 2017), and protecting biodiversity (Dee et al. 2017; Iacona et al.
+2017). Yet fisheries conservation and management has long been both a
+crucible and proving ground for the theory of ecological management more
+generally, including topics such as adaptive management (Walters and
+Hilborn 1978), ecosystem-based management (Levin and Lubchenco 2008) and
+resilience thinking (Holling 1973; May 1977), while also giving rise to
+the sub-discipline of resource economics (Gordon and Press 1954;
+Schaefer 1954; Beverton and Holt 1957), thanks to its global relevance,
+long history, and readily available data. Despite the generality of the
+models and issue of measurement uncertainty considered here, any
+application must be grounded in both the history and context of the
+particular decision problem. In light of this, we begin with a brief
+background on existing decision theory approaches common in fisheries
+management.
+
+In this paper, we examine three strategies to managing a natural
+population, such as a fishery, under such uncertainty. This problem has
+a long history in both theory and practice, where seemingly innoccous
+simplifying assumptions have led to a deep paradox in how managers
+should respond to uncertainty. The precautionary principle (Kriebel et
+al. 2001) would seem to suggest that the greater the uncertainty in our
+models, the more cautious we should be in management policies. Yet a
+well-known mathematical proof by Reed (1979) established that under
+quite general conditions, the optimal strategy under uncertainty is no
+different than under no uncertainty at all. Attempts to resolve this
+have since only deepened this paradox, arguing that adding additional
+sources of uncertainty, such as imperfect measurements, should lead to
+optimal harvest that increases with greater uncertainty (e.g. Clark and
+Kirkwood 1986; Sethi et al. 2005). Meanwhile, much of actual practice
+has (perhaps fortunately, as we shall see) ignored these results in
+favor of a more heuristic application of the precautionary principle.
+Here, we show that this paradox can be resolved by uncovering previous
+shortcuts and tackling the optimal management problem head-on, which we
+can solve thanks to the computational efficiency of sophisticated
+algorithms developed more recently in the field of robotics. This
+approach allows us to reject this paradox largely in favor of the
+precautionary principle. Further, our comparisons show that widely
+recognized existing approaches in both theory and practice are not by
+themselves sufficient to manage a population under substantial
+uncertainty over the long term, but our recently borrowed approach from
+robotics literature can under identical conditions sustain both higher
+biomass and higher economic yields than those approaches. This
+conclusion has important implications not only for fisheries but
+eocsystem more broadly, underscoring the danger of not correctly
+accounting for measurement uncertainty when making decisions, and
+further to the significant opportunity to learn from fields such as
+robotics when faced with complex and uncertain decisions.
 
@@
 distributed with standard deviation \(\sigma_m\),
-\(\xi \sim \mathcal{N}(1, \sigma_m)\). Given an estimated recruitment
-model along with this measurement of the stock size, the manager must
-choose the set of a harvest quotas \(\lbrace h_t \rbrace\) for
-\(t \in [0, 1, \infty]\) that maximizes expected long term (net) present
-value:
+\(\xi \sim \mathcal{N}(1, \sigma_m)\). Given an estimated population
+model along with this (uncertain) measurement of the stock size, the
+manager must choose the set of a harvest quotas \(\lbrace h_t \rbrace\)
+for \(t \in [0, 1, \infty]\) that maximizes expected long term utility
+of the stock:
 
 \begin{equation}
-\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}
+\max_{ \lbrace h_t \rbrace } \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot U(x_t, h_t) \right \}
 \end{equation}
 
-A few subtleties arise in this seemingly simple problem statement that
-we must address. First, given our discrete-time formulation of the
-recruitment process, it is necessary to decide if the measurement
+For convenience of presentation but without loss of generality, we will
+assume this utility is simply equal to the harvest itself; in economic
+terms, a fixed price for fish without cost on effort:
+
+\begin{equation}
+ U(x_t, h_t) = \min(h_t, x_t)
+\end{equation}
+
+where the \(\min\) function ensures no additional reward for attempting
+to harvest more than the entire stock. Existing optimal control
+techniques (Reed (1979)) prove these results are not dependent on the
+particular details of this utility. Using this form allows a more direct
+comparison to approaches such as maximum sustainable yield (MSY), that
+do not include an economic model but merely maximize long-term catch. In
+principle, utility could also reflect ecosystem services or other
+intrinsic values for fish left in the sea and not merely the market for
+consumption (Halpern et al. 2013).
+
+The solution to this decision problem is conditional on a population
+model for the underlying ecological process, describing probability the
+system moves to state \(x_{t+1}\) given that it was in state \(x_t\) and
+the manager selected a harvest \(h_t\), which we discuss below. Such a
+model would typically be estimated and parameterized from available
+data, often (but not necessarily) using a hierarchical Bayesian approach
+that acknowledges the existence of uncertainty in both measurements and
+process (e.g.~see Dichmont et al. 2016 for a discussion of how
+uncertainties are frequently included in stock assessment models).
+However, accounting for measurement uncertainty when going from data to
+model does not mean we have also accounted for this uncertainty when
+going from model to decision. For example, a harvest decision policy
+based upon maximum sustainable yield simply depends on the stock size
+which maximizes the expected growth rate of the model. Measurement
+uncertainty in model estimation can indeed influence the position of
+that maximum, but as we shall see, that strategy is only optimal in the
+absence of any uncertainty.
+
+There is one further subtlety arising in this seemingly simple problem
+statement that we must address. Given our discrete-time formulation of
+the recruitment process, it is necessary to decide if the measurement
 \(y_t\) happens before or after harvest \(h_t\): that is do we: measure,
@@
 al. 2005)), we will assume the latter; measurement occurs immediately
-before harvest, followed by recruitment.
-
-A second subtly is in the nature of the decision problem itself: note
-that the manager in not myopic, thinking only about the next year, but
-rather considers the sequence of all future actions \({h_t}\). While the
-manager will revisit this sequence each year in light of new
-observations, it is this ability of the decision problem to think ahead
-that allows it to tolerate short-term costs (e.g.~reduced harvests in
-the current season) for larger future payoffs. In this, ecological
-management is a game of chess, always thinking several moves ahead. This
-differs from the problem solved in theory as well as the practice of
-maximum sustainable yield, which seeks to identify only a fixed
-mortality \(F_{MSY}\) applied for all time, rather than a \emph{policy}
-of varying quotas \(h_t\) depending on the stock assessment \(y_t\).
-Understanding this distinction is key both to understanding the
-differences between current theory and current best practice in
-fisheries management.
+before harvest, followed by recruitment. This assumption has
+occasionally been reversed to introduce some additional uncertainty from
+stochastic recruitment immediately prior to setting the harvest quota
+(e.g. Clark and Kirkwood 1986; Weitzman 2002), which appears to mimic
+uncertainty in measurements but does not allow uncertainty to accumulate
+since the measurement step still occurs without error.
+
+The optimal strategy must think not only about the next year, but rather
+considers the sequence of all future actions \({h_t}\), under all
+possible realizations of the underlying model. This strategy must also
+acknowledge that the manager will receive new information each year, and
+may need to modify actions accordingly. It is this ability of the
+decision problem to think ahead that allows it to tolerate short-term
+costs (e.g.~reduced harvests in the current season) for larger future
+payoffs. When Eq (1) is ignored, such optimal control problems are known
+as Markov Decision Processes or MDPs. These are more commonly referred
+to in the ecological literature by their solution method of Stochastic
+Dynamic Programming (SDP; Marescot et al. 2013), and widely used in both
+ecological management and animal behavior (Mangel 1985; Mangel and Clark
+1988). The addition of measurement uncertainty in Eq (1) turns this
+problem into a Partially Observed Markov Decision Process (POMDP), for
+which solution methods have not been widely available for large problems
+until more recently. We discuss these decision methods in more detail
+below.
 
@@
 
-It is important to remember that both existing methods we discuss and
-the POMDP approach proposed here can or are already applied to more
-complex models as well, including those with age or stage structure.
-Likewise, while we consider the model and its parameters as ``given,''
-it should be understood to mean that this in practice comes as the
-result of some model choice and statistical estimation from historical
-data, frequently accounting for uncertainty in both the process and the
-measurements (e.g. Dichmont et al. 2016). As we shall see here, having
-included uncertainty in the model \emph{estimation} in no way excuses us
-from having to also deal once again with that uncertainty in the
-decision process.
+It is important to remember that all three decision methods we compare
+here will use this same model as given. In the case of numerical
+simulations, we will simply pick reasonable fixed parameter values for
+the model above. To illustrate the application to empirical data, we
+will first estimate these parameters using a hierarchical Markov Chain
+Monte Carlo (MCMC), and then use the resulting estimates to drive all
+three decision models. As such, the details of how the model is
+estimated are immaterial to the differences in performance between these
+decision methods. In all cases, the model is considered fixed.
+Technically, it would be possible to re-estimate the model parameters
+given each new observation during the decision process. Adaptations of
+the decision processes we consider here include this possibility for
+learning about the model during the decisions. Some include the
+possibility of taking knowingly bad actions, like suddenly increasing
+harvest rates, to improve learning over parameters. We investigate these
+in subsequent work (Memarzadeh and Boettiger 2018). For better
+comparison with established theory, here, the model is considered fixed.
+In the case of the simulations, the model is known without error (though
+still stochastic in nature), such that additional observations cannot
+improve the parameter estimates in any event. Despite that obvious
+simplification, we shall illustrate how existing decision methods fail
+to account for measurement uncertainty appropriately. Even with these
+simple models, decisions are complex, and knowing the rules of the game
+is not the same as knowing how to win.
 
@@
 in both process and models is known as Management Strategy Evaluation
-(MSE, Smith 1994; {\textbf{???}}; Bunnefeld et al. 2011; Punt et al.
+(MSE, Smith 1994; Dorner et al. 2009; Bunnefeld et al. 2011; Punt et al.
 2016). MSE performs forward simulations under a suite of pre-selected
@@
 made by Gordon and Press (1954); Schaefer (1954) or Clark (1973) no
-longer hold. The optimal harvest of a population under a model such as
-the stochastic Gordon Schaefer, Eq \eqref{Gordon-Schaefer}, is an
-example of a Markov Decision Problem (MDP). This is not to be confused
-with the the problem of \emph{estimating} parameters or posteriors of a
-Markov process model, e.g.~through typical Bayesian analysis -- recall
-this is a decision problem, in which the model has already been
-estimated and now we are tasked with choosing a sequence of future
-harvests based on that model. Such MDP problems are commonly referred to
-in ecological literature by its solution method, Stochastic Dynamic
-Programming (SDP) (Marescot et al. 2013), widely used in both ecological
-management and animal behavior (Mangel 1985; Mangel and Clark 1988).
-However, thanks to a mathematical proof provided by Reed (1979),
-fisheries managers have largely been able to bi-pass the need for SDP.
-But this convenient result also opened the door to paradox of
-uncertainty that would persist for the next four decades.
+longer hold. As noted above, the optimal harvest of a population under a
+model such as the stochastic Gordon Schaefer, Eq
+\eqref{Gordon-Schaefer}, is an example of a Markov Decision Problem
+(MDP) which must be solved with stochastic dynamic programming. Thanks
+to a mathematical proof provided by Reed (1979), fisheries managers have
+largely been able to bi-pass such computational effort. However, this
+convenient result also opened the door to paradox of uncertainty that
+would haunt ecological management for at least the next four decades.
 
@@
 response to increased uncertainty. As we shall illustrate, this is
-particularly fortunate; though approaches that properly account for both
-forms of uncertainty can perform even better.
+fortunate; though approaches that properly account for both forms of
+uncertainty can perform even better.
 
@@
 assumption: a rational decision maker would suspect this 75 to be an
-over-estimate based on prior observations. Under imperfect observations,
-we no longer have a Markov Decision Process (MDP) over the observed
-state space. Rather, we have a Partially Observed Markov Decision
-Process (POMDP). A POMDP policy cannot be expressed as a policy function
-at all (much less as policy function with a single constant-escapement
-target, as much previous work has assumed). Instead, a POMDP policy can
-be thought of as depending not only on the most recent observation, but
-the recent observation combined with a \emph{prior belief} about what
-state the system is in, which itself is determined by all past
-observations. In this manner, the POMDP can consider how much weight to
-give the most recent observation based on how it compares to prior
-beliefs.
+over-estimate based on prior observations. Under imperfect observations
+introduced by Eq (1), we no longer have a Markov Decision Process (MDP)
+over the observed state space. Rather, we have a Partially Observed
+Markov Decision Process (POMDP). A POMDP policy cannot be expressed as a
+policy function at all (much less as policy function with a single
+constant-escapement target, as much previous work has assumed). Instead,
+a POMDP policy can be thought of as depending not only on the most
+recent observation, but the recent observation combined with a
+\emph{prior belief} about what state the system is in, which itself is
+determined by all past observations. In this manner, the POMDP can
+consider how much weight to give the most recent observation based on
+how it compares to prior beliefs.
 
@@
 
+\hypertarget{simulations}{%
+\subsection{Simulations}\label{simulations}}
+
+We consider the average fish biomass across 100 replicate simulations of
+the same stock dynamics (Eq 4) under three different management
+strategies: constant escapement (CE), total allowable catch (TAC; equal
+to 80\% MSY), and partially observed Markov decision process (POMDP)
+management. Each suite of simulations is considered under three
+sequentially higher measurement error regimes, from \(\sigma_m = 0\) (no
+measurement errors), \(\sigma_m = 0.1\) (low error), and
+\(\sigma_m = 0.15\) (moderate error), as indicated, while stochastic
+recruitment (environmental noise) is set to a moderate
+\(\sigma_g = 0.15\). Empirical estimates of measurement error in stock
+sizes vary widely, with estimates often dependent upon methodological
+assumptions. Early estimates have put possible error as high as 50\%
+(Clark and Kirkwood 1986) and more recent analyses have suggested an
+average of closer to 36\%, though with cases that may be much higher
+(Ralston et al. 2011). Consequently, range between 0 to 15\% provides a
+probably conservative estimate of typical unceratinty ranges, which we
+show is nevertheless large enough to drive substantially negative
+ecological and economic outcomes under management strategies that do not
+appropriately account for that error over the long run.
+
 \hypertarget{results}{%
@@
 
-We consider the average fish biomass across 100 replicate simulations
-under three different management strategies: constant escapement (CE),
-total allowable catch (TAC; equal to 80\% MSY), and partially observed
-Markov decision process (POMDP) management. Each suite of simulations is
-considered under three sequentially higher measurement error regimes,
-from \(\sigma_m = 0\), \(\sigma_m = 0.1\), and \(\sigma_m = 0.15\), as
-indicated, while stochastic recruitment (environmental noise) is set to
-a moderate \(\sigma_g = 0.15\). In the absence of either environmental
-noise or measurement error POMDP, CE, and MSY would converge to stock at
-the \(B_{MSY}\), while TAC would maintain the stock at a slightly higher
-level (see Supplemental material, Figure S1). The introduction of
-stochastic growth has a significant negative impact on the TAC strategy,
-with a mean biomass significantly lower than \(B_{MSY}\) (first panel,
-Figure \ref{sims}; this impact is even more severe for MSY, see Figure
-S2). Without measurement error, the CE and POMDP strategies are nearly
-identical with both approximately maintaining the stock at \(B_{MSY}\)
-despite the significant environmental stochasticity. As measurement
-error increases in the subsequent panels, CE and TAC strategies perform
-increasingly poorly, while the POMDP continues to maintain the average
-stock close to \(B_{MSY}\). Notably, CE is significantly more impacted
-by measurement uncertainty than TAC, with CE averaging even lower
-biomass than TAC under moderate measurement uncertainty. These results
-confirm that the precautionary approach represented by the TAC does
-indeed prove more robust to the problem of measurement error than the
-optimization solution represented by CE, but is less robust to
-environmental error. In contrast, the POMDP approach successfully
-handles both sources of uncertainty, maintaining higher stocks, near
-\(B_{MSY}\) level, and producing the highest yield.
+In the absence of either environmental noise or measurement error POMDP,
+CE, and MSY would converge to stock at the \(B_{MSY}\), while TAC would
+maintain the stock at a slightly higher level (see Supplemental
+material, Figure S1). The introduction of stochastic growth has a
+significant negative impact on the TAC strategy, with a mean biomass
+significantly lower than \(B_{MSY}\) (first panel, Figure \ref{sims};
+this impact is even more severe for MSY, see Figure S2). Without
+measurement error, the CE and POMDP strategies are nearly identical with
+both approximately maintaining the stock at \(B_{MSY}\) despite the
+significant environmental stochasticity. As measurement error increases
+in the subsequent panels, CE and TAC strategies perform increasingly
+poorly, while the POMDP continues to maintain the average stock close to
+\(B_{MSY}\). Notably, CE is significantly more impacted by measurement
+uncertainty than TAC, with CE averaging even lower biomass than TAC
+under moderate measurement uncertainty. These results confirm that the
+precautionary approach represented by the TAC does indeed prove more
+robust to the problem of measurement error than the optimization
+solution represented by CE, but is less robust to environmental error.
+In contrast, the POMDP approach successfully handles both sources of
+uncertainty, maintaining higher stocks, near \(B_{MSY}\) level, and
+producing the highest yield.
 
@@
 simplifying assumptions used in previous work. Our solution confirms the
-initution of the precautionary principle in a quantitatively precise
+intuition of the precautionary principle in a quantitatively precise
 form: under most (but not all) conditions, \emph{greater uncertainty
@@
 Kirkwood (1986) and Sethi et al. (2005) into practice. Yet this does not
-lift the spectre of these methods, where CE-based predictions feature
+lift the specter of these methods, where CE-based predictions feature
 prominently in global fisheries analyses as the economically optimal
@@
 the realities of greater complexity and uncertainty echoed in more
-heursitic approaches, this division can sometimes appear to pit economic
+heuristic approaches, this division can sometimes appear to pit economic
 outcomes against ecological objectives. That perspective is echoed in
@@
 uncertainty has argued in cold mathematical logic for a rejection of the
-precationary principle. Our results demonstrate that this divide is
+precautionary principle. Our results demonstrate that this divide is
 artificial. With sufficiently powerful algorithms we can go some ways to
@@
 approaches in ecological management. Both sides are vindicated in our
-solution: economic optimization under POMDP does not controdict the
+solution: economic optimization under POMDP does not contradict the
 precautionary principle, while under POMDP, optimal control can also
@@
 from harvests of wild game or forestry to pest outbreaks, (e.g. Richards
-et al. 1999; {\textbf{???}}; Blackwood et al. 2010; Shea et al. 2014;
-Dee et al. 2017; Iacona et al. 2017; Li et al. 2017) as in fisheries,
-where ecological complexity is invariably high and management capacity
+et al. 1999; Shea 1998; Blackwood et al. 2010; Shea et al. 2014; Dee et
+al. 2017; Iacona et al. 2017; Li et al. 2017) as in fisheries, where
+ecological complexity is invariably high and management capacity
 invariably limited. We address each of these in turn.
@@
 Are the dynamics considered here too simple? Our underlying model is
-undoubtably much simpler than many of the age or stage structured models
+undoubtedly much simpler than many of the age or stage structured models
 used in modern fisheries management -- aspects whose importance to
@@
 complex models, whether that be age structure or even more ambitious
-non-parametric representations defined directly from data using Guassian
+non-parametric representations defined directly from data using Gaussian
 Processes or other machine learning techniques (e.g. Boettiger et al.
@@
 the following decades ecologists largely turned away from
-computationally intensive optimizations required for even simple models
+computationally intensive optimization required for even simple models
 with stochastic dynamic programming, (while continuing to construct ever
@@
 the decision problem to reveal algorithms such as SARSOP that make much
-larger dynamic programming problems tractable. Yes, even more
-sophisticated algorithms will be needed to address the limitations
-
-Second, we must ask if the policies derived from POMDP are practical. As
-with constant escapement, the POMDP solutions we considered here result
-in more year-to-year variation in harvest, and include the possibility
-of fishing moratorium to allow stocks to recover -- decisions that many
-view as economically or socially unacceptable. While resolving this
-debate is beyond the scope of this paper, a few observations are
-relevant: (A) Stocks already managed under constant escapement, such as
-US salmon fisheries, are already subject to the possibility of these
-moratoriums. (B) It is straight-forward to introduce constraints on the
-minimal acceptable harvest level and/or include the additional costs of
-closing a fishery in the utility function optimized by POMDP. (C)
-Fisheries today are not managed at either their maximum ecological or
-economic potential. Taken as a whole, most ocean fisheries yield a net
-negative economic return (World Bank and FAO 2009), with fishing
-persisting only with the support of government subsidies (Arnason 2012),
-meanwhile some 68\% of global fisheries are in poor biological condition
-(Costello et al. 2016). While the causes for this are both complex and
-numerous, we have seen that both the economic inefficiency and
-ecological health suffer under TAC/MSY and CE based management. Arnason
-(2012) and Costello et al. (2016) have both argued that a rights-based
-approach to fisheries management (RBFM) would lead to self adoption of
-precisely such optimal policies.
+larger dynamic programming problems tractable. Yes, we may need even
+more sophisticated algorithms to accommodate greater ecological
+complexity, yet we believe such numerical obstacles are no excuse for
+management by half measures. Rather, our example illustrates how
+ecologists and managers can learn from and leverage the tools developed
+in other disciplines to better face the challenges of management in a
+complex and changing world.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Arnason2012}{}%
-Arnason, R. 2012. Property rights in fisheries: How much can individual
-transferable quotas accomplish? Review of Environmental Economics and
-Policy 6:217--236.
-
 \leavevmode\hypertarget{ref-Beverton1957}{}%
@@
 
+\leavevmode\hypertarget{ref-Dorner2009}{}%
+Dorner, B., R. M. Peterman, and Z. Su. 2009. Evaluation of performance
+of alternative management models of Pacific salmon (Oncorhynchus spp.)
+in the presence of climatic change and outcome uncertainty using Monte
+Carlo simulations. (R. Hilborn, ed.)Canadian Journal of Fisheries and
+Aquatic Sciences 66:2199--2221.
+
+\leavevmode\hypertarget{ref-Ellison2004}{}%
+Ellison, A. M. 2004. Bayesian inference in ecology. Ecology Letters
+7:509--520.
+
 \leavevmode\hypertarget{ref-Engen1997}{}%
@@
 
+\leavevmode\hypertarget{ref-Halpern2013}{}%
+Halpern, B. S., C. J. Klein, C. J. Brown, M. Beger, H. S. Grantham, S.
+Mangubhai, M. Ruckelshaus, et al. 2013. Achieving the triple bottom line
+in the face of inherent trade-offs among social equity, economic return,
+and conservation. Proceedings of the National Academy of Sciences
+110:6229--34.
+
 \leavevmode\hypertarget{ref-Hilborn2010}{}%
@@
 
+\leavevmode\hypertarget{ref-Kriebel2001}{}%
+Kriebel, D., J. Tickner, P. Epstein, J. Lemons, R. Levins, E. L.
+Loechler, M. Quinn, et al. 2001. The precautionary principle in
+environmental science. Environmental health perspectives 109:871--6.
+
 \leavevmode\hypertarget{ref-Kurniawati2008}{}%
@@
 
+\leavevmode\hypertarget{ref-Memarzadeh2018}{}%
+Memarzadeh, M., and C. Boettiger. 2018. Adaptive management of
+ecological systems under partial observability. Biological Conservation
+224:9--15.
+
 \leavevmode\hypertarget{ref-Memarzadeh2016b}{}%
@@
 
+\leavevmode\hypertarget{ref-Ralston2011}{}%
+Ralston, S., A. E. Punt, O. Hamel, J. D. Devore, and J. R. Conser. 2011.
+A meta-analytic approach to quantifying scientific uncertainty in stock
+assessments. Fishery Bulletin 217--231.
+
 \leavevmode\hypertarget{ref-Reed1979}{}%
@@
 
+\leavevmode\hypertarget{ref-Shea1998}{}%
+Shea, K. 1998. Management of populations in conservation, harvesting and
+control. Trends in Ecology and Evolution 13:371--375.
+
 \leavevmode\hypertarget{ref-Shea2014}{}%
@@
 
+\leavevmode\hypertarget{ref-Weitzman2002}{}%
+Weitzman, M. L. 2002. Landing Fees vs Harvest Quotas with Uncertain Fish
+Stocks. Journal of Environmental Economics and Management 43:325--338.
+
 \leavevmode\hypertarget{ref-Williams2001}{}%
@@
 
-\leavevmode\hypertarget{ref-FAO2009}{}%
-World Bank and FAO. 2009. The Sunken Billions. The Economic
-Justification for Fisheries Reform.
-
 \end{document}

### 2018-08-17 manuscripts/manuscript.Rmd
here we go

@@
 ---
-title: "The importance of measurement uncertainty in ecological management"
+title: "Resolving a paradox of measurement uncertainty in ecological management"
 
@@
   Ecological management and decision-making typically focus on uncertainty about the future, but surprisingly little is known about how to account for uncertainty of the present: that is, the realities of having only partial or imperfect measurements.
-Our two primary paradigms for handling decisions under uncertainty -- the precautionary principle and optimal control -- have so far given entirely contradictory results.
+Our primary paradigms for handling decisions under uncertainty -- the precautionary principle and optimal control -- have so far given contradictory results.
 This paradox is best illustrated in the example of fisheries management, where many ideas that guide thinking about ecological decision making were first developed.
@@
 keywords:
-  - Measurement uncertainty
-  - POMDP
-  - Decision theory
-  - Fisheries
-  - Conservation
+  - POMDP,
+  - measurement uncertainty,
+  - decision theory,
+  - fisheries,
+  - conservation
 
@@
 ```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = 1)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,  dev="cairo_pdf", cache = TRUE)
 library(tidyverse)
@@
 
-The challenge of making decisions under uncertainty is obviously not unique to fisheries, and the existing methods we critique here can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010; @Shea1998], disease outbreaks [@Shea2014; @Li2017], and protecting biodiversity [@Dee2017; @Iacona2017].
+The challenge of making decisions under uncertainty is obviously not unique to fisheries.
+Existing methods we optimal control methods which similarly assume perfect measurements can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010; @Shea1998], disease outbreaks [@Shea2014; @Li2017], and protecting biodiversity [@Dee2017; @Iacona2017], and in common textbooks and reviews [@Mangel1985; @Clark1990; @Marescot2013].
 Yet fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.
@@
 In this paper, we examine three strategies to managing a natural population, such as a fishery, under such uncertainty.
-This problem has a long history in both theory and practice, where seemingly innoccous simplifying assumptions have led to a deep paradox in how managers should respond to uncertainty.
+This problem has a long history in both theory and practice, where seemingly innocuous simplifying assumptions have led to a deep paradox in how managers should respond to uncertainty.
 The precautionary principle [@Kriebel2001] would seem to suggest that the greater the uncertainty in our models, the more cautious we should be in management policies.
@@
 Further, our comparisons show that widely recognized existing approaches in both theory and practice are not by themselves sufficient to manage a population under substantial uncertainty over the long term, but our recently borrowed approach from robotics literature can under identical conditions sustain both higher biomass and higher economic yields than those approaches.
-This conclusion has important implications not only for fisheries but eocsystem more broadly, underscoring the danger of not correctly accounting for measurement uncertainty when making decisions, and further to the significant opportunity to learn from fields such as robotics when faced with complex and uncertain decisions.
+This conclusion has important implications not only for fisheries but ecosystem more broadly: underscoring the danger of not correctly accounting for measurement uncertainty when making decisions, and further to the significant opportunity to learn from fields such as robotics when faced with complex and uncertain decisions.
 
@@
 We will denote the discount rate $\gamma$.
-For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax.
+For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars).
 Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size subject to some measurement uncertainty $\xi_t$
@@
 
-For simplicity we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.
+For simplicity, we will assume measurement uncertainty is normally distributed with standard deviation $\sigma_m$,  $\xi \sim \mathcal{N}(1, \sigma_m)$.
 Given an estimated population model along with this (uncertain) measurement of the stock size, the manager must choose the set of a harvest quotas $\lbrace h_t \rbrace$ for $t \in [0, 1, \infty]$ that maximizes expected long term utility of the stock:
@@
 
-For convenience of presentation but without loss of generality, we will assume this utility is simply equal to the harvest itself; in economic terms, a fixed price for fish without cost on effort:
+For convenience of presentation, we will assume this utility is simply equal to the harvest itself.
 
@@
 where the $\min$ function ensures no additional reward for attempting to harvest more than the entire stock.
-Existing optimal control techniques (@Reed1979) prove these results are not dependent on the particular details of this utility.
-Using this form allows a more direct comparison to approaches such as maximum sustainable yield (MSY), that do not include an economic model but merely maximize long-term catch.
+In economic terms, the corresponds to a fixed price per fish caught, with no proportional cost on effort required to do so, but the results do not depend on this assumption, as existing analyses have previously established [@Reed1979].
+Using this form allows a more direct comparison to approaches such as maximum sustainable yield (MSY) which do not include an economic model but merely maximize long-term catch.
 In principle, utility could also reflect ecosystem services or other intrinsic values for fish left in the sea and not merely the market for consumption [@Halpern2013].
@@
 Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment.
-To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.
+To reflect possible uncertainty between reported and actual catch, the ABC may be reduced further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.
 To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under the MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$.
@@
 It is important to note that actual management strategies are quite diverse.
-Our focus on comparisons TAC, MSY, and Constant Escapement (CE, also refereed to as MEY, see below) reflects only the subset of current practices.
+Our focus on comparisons TAC, MSY, and Constant Escapement (CE) reflects only the subset of current practices.
 Another increasingly common approach to address uncertainty in both process and models is known as Management Strategy Evaluation [MSE, @Smith1994;  @Dorner2009; @Bunnefeld2011; @Punt2016].
@@
 As such, MSE can be viewed as an approximation to the optimal control solutions considered below, a notion that has been formalized in the literature on approximate dynamic programming [@Nicol2011].
+Future work may be able to leverage the forward-simulation approach used in MSE to extend the POMDP methods introduced here to yet more complex dynamics.
 
@@
 Early estimates have put possible error as high as 50% [@Clark1986] and more recent analyses have suggested an average of closer to 36%, though with cases that may be much higher [@Ralston2011].
-Consequently, range between 0 to 15% provides a probably conservative estimate of typical unceratinty ranges, which we show is nevertheless large enough to drive substantially negative ecological and economic outcomes under management strategies that do not appropriately account for that error over the long run.
+Consequently, range between 0 to 15% provides a probably conservative estimate of typical uncertainty ranges, which we show is nevertheless large enough to drive substantially negative ecological and economic outcomes under management strategies that do not appropriately account for that error over the long run.
 

### 2018-08-17 manuscripts/manuscript.tex
here we go

@@
             pdfauthor={},
-            pdftitle={The importance of measurement uncertainty in ecological management},
+            pdftitle={Resolving a paradox of measurement uncertainty in ecological management},
             colorlinks=true,
@@
 
-  \title{The importance of measurement uncertainty in ecological management}
+  \title{Resolving a paradox of measurement uncertainty in ecological management}
     \author[]{Authors removed during peer review}
@@
   for uncertainty of the present: that is, the realities of having only
-  partial or imperfect measurements. Our two primary paradigms for
-  handling decisions under uncertainty -- the precautionary principle and
-  optimal control -- have so far given entirely contradictory results.
-  This paradox is best illustrated in the example of fisheries management,
-  where many ideas that guide thinking about ecological decision making
-  were first developed. We find that simplistic optimal control approaches
-  have repeatedly concluded that a manager should increase catch quotas
-  when faced with greater uncertainty about the fish biomass. Current best
+  partial or imperfect measurements. Our primary paradigms for handling
+  decisions under uncertainty -- the precautionary principle and optimal
+  control -- have so far given contradictory results. This paradox is best
+  illustrated in the example of fisheries management, where many ideas
+  that guide thinking about ecological decision making were first
+  developed. We find that simplistic optimal control approaches have
+  repeatedly concluded that a manager should increase catch quotas when
+  faced with greater uncertainty about the fish biomass. Current best
   practices take a more precautionary approach, decreasing catch quotas by
@@
   \end{abstract}
-   \begin{keyword} Measurement uncertainty POMDP Decision theory Fisheries Conservation\end{keyword}
+   \begin{keyword} POMDP, measurement uncertainty, decision theory, fisheries, conservation\end{keyword}
  \end{frontmatter}
@@
 The challenge of making decisions under uncertainty is obviously not
-unique to fisheries, and the existing methods we critique here can be
-found in a wide range of population and ecosystem management challenges,
-including fire ecology (Richards et al. 1999), invasive species (Shea
-1998; Blackwood et al. 2010), disease outbreaks (Shea et al. 2014; Li et
-al. 2017), and protecting biodiversity (Dee et al. 2017; Iacona et al.
-2017). Yet fisheries conservation and management has long been both a
+unique to fisheries. Existing methods we optimal control methods which
+similarly assume perfect measurements can be found in a wide range of
+population and ecosystem management challenges, including fire ecology
+(Richards et al. 1999), invasive species (Shea 1998; Blackwood et al.
+2010), disease outbreaks (Shea et al. 2014; Li et al. 2017), and
+protecting biodiversity (Dee et al. 2017; Iacona et al. 2017), and in
+common textbooks and reviews (Mangel 1985; Clark 1990; Marescot et al.
+2013). Yet fisheries conservation and management has long been both a
 crucible and proving ground for the theory of ecological management more
@@
 population, such as a fishery, under such uncertainty. This problem has
-a long history in both theory and practice, where seemingly innoccous
+a long history in both theory and practice, where seemingly innocuous
 simplifying assumptions have led to a deep paradox in how managers
@@
 conclusion has important implications not only for fisheries but
-eocsystem more broadly, underscoring the danger of not correctly
+ecosystem more broadly: underscoring the danger of not correctly
 accounting for measurement uncertainty when making decisions, and
@@
 (equivalently, measuring our value in units of discounted fish rather
-than discounted dollars), though again this assumption is easy to relax.
-Each year, the manager also obtains an estimate \(y_t\) of the true
-\(x_t\) stock size subject to some measurement uncertainty \(\xi_t\)
+than discounted dollars). Each year, the manager also obtains an
+estimate \(y_t\) of the true \(x_t\) stock size subject to some
+measurement uncertainty \(\xi_t\)
 
@@
 
-For simplicity we will assume measurement uncertainty is normally
+For simplicity, we will assume measurement uncertainty is normally
 distributed with standard deviation \(\sigma_m\),
@@
 
-For convenience of presentation but without loss of generality, we will
-assume this utility is simply equal to the harvest itself; in economic
-terms, a fixed price for fish without cost on effort:
+For convenience of presentation, we will assume this utility is simply
+equal to the harvest itself.
 
@@
 where the \(\min\) function ensures no additional reward for attempting
-to harvest more than the entire stock. Existing optimal control
-techniques (Reed (1979)) prove these results are not dependent on the
-particular details of this utility. Using this form allows a more direct
-comparison to approaches such as maximum sustainable yield (MSY), that
-do not include an economic model but merely maximize long-term catch. In
-principle, utility could also reflect ecosystem services or other
-intrinsic values for fish left in the sea and not merely the market for
-consumption (Halpern et al. 2013).
+to harvest more than the entire stock. In economic terms, the
+corresponds to a fixed price per fish caught, with no proportional cost
+on effort required to do so, but the results do not depend on this
+assumption, as existing analyses have previously established (Reed
+1979). Using this form allows a more direct comparison to approaches
+such as maximum sustainable yield (MSY) which do not include an economic
+model but merely maximize long-term catch. In principle, utility could
+also reflect ecosystem services or other intrinsic values for fish left
+in the sea and not merely the market for consumption (Halpern et al.
+2013).
 
@@
 uncertainty in the stock assessment. To reflect possible uncertainty
-between reported and actual catch, the ABC is reduced somewhat further
-to define the Total Allowable Catch, (TAC), which forms the basic unit
-of management for many such fisheries. To reflect this process, our
+between reported and actual catch, the ABC may be reduced further to
+define the Total Allowable Catch, (TAC), which forms the basic unit of
+management for many such fisheries. To reflect this process, our
 analysis will also consider policies in which the harvest quota is set
@@
 It is important to note that actual management strategies are quite
-diverse. Our focus on comparisons TAC, MSY, and Constant Escapement (CE,
-also refereed to as MEY, see below) reflects only the subset of current
-practices. Another increasingly common approach to address uncertainty
-in both process and models is known as Management Strategy Evaluation
-(MSE, Smith 1994; Dorner et al. 2009; Bunnefeld et al. 2011; Punt et al.
-2016). MSE performs forward simulations under a suite of pre-selected
-candidate strategies to determine which strategy achieves the best
-expected outcomes. Crucially, by limiting the analysis to a suite of
-candidate strategies over a finite number of replicate simulations, this
-approach is able to accommodate substantially greater complexity and
-uncertainty than has been possible with more formal optimal control
-approaches we discuss below. Optimal control solutions rely on dynamic
-programming to consider all possible strategies (all possible sequences
-of harvest under all possible realizations of stock dynamics). As such,
-MSE can be viewed as an approximation to the optimal control solutions
-considered below, a notion that has been formalized in the literature on
-approximate dynamic programming (Nicol and Chadès 2011).
+diverse. Our focus on comparisons TAC, MSY, and Constant Escapement (CE)
+reflects only the subset of current practices. Another increasingly
+common approach to address uncertainty in both process and models is
+known as Management Strategy Evaluation (MSE, Smith 1994; Dorner et al.
+2009; Bunnefeld et al. 2011; Punt et al. 2016). MSE performs forward
+simulations under a suite of pre-selected candidate strategies to
+determine which strategy achieves the best expected outcomes. Crucially,
+by limiting the analysis to a suite of candidate strategies over a
+finite number of replicate simulations, this approach is able to
+accommodate substantially greater complexity and uncertainty than has
+been possible with more formal optimal control approaches we discuss
+below. Optimal control solutions rely on dynamic programming to consider
+all possible strategies (all possible sequences of harvest under all
+possible realizations of stock dynamics). As such, MSE can be viewed as
+an approximation to the optimal control solutions considered below, a
+notion that has been formalized in the literature on approximate dynamic
+programming (Nicol and Chadès 2011). Future work may be able to leverage
+the forward-simulation approach used in MSE to extend the POMDP methods
+introduced here to yet more complex dynamics.
 
@@
 (Ralston et al. 2011). Consequently, range between 0 to 15\% provides a
-probably conservative estimate of typical unceratinty ranges, which we
+probably conservative estimate of typical uncertainty ranges, which we
 show is nevertheless large enough to drive substantially negative

### 2018-08-18 manuscripts/manuscript.Rmd
revisions submitted

@@
 ---
-title: "Resolving a paradox of measurement uncertainty in ecological management"
+title: "Resolving the measurement uncertainty paradox in ecological management"
 
@@
 
+## Acknowledgements
+
+The authors wish to thank T. Essington for his insight into fisheries management, and
+the constructive feedback of the editor and anonymous reviewers.
+CB also acknowledges computational resources from NSF's XSEDE Jetstream (DEB160003) and Chameleon cloud platforms, as well as the support by the USDA Hatch project CA-B-INS-0162-H.
+
 <!--

### 2018-08-18 manuscripts/manuscript.tex
revisions submitted

@@
             pdfauthor={},
-            pdftitle={Resolving a paradox of measurement uncertainty in ecological management},
+            pdftitle={Resolving the measurement uncertainty paradox in ecological management},
             colorlinks=true,
@@
 
-  \title{Resolving a paradox of measurement uncertainty in ecological management}
+  \title{Resolving the measurement uncertainty paradox in ecological management}
     \author[]{Authors removed during peer review}
@@
 
+\hypertarget{acknowledgements}{%
+\subsection{Acknowledgements}\label{acknowledgements}}
+
+The authors wish to thank T. Essington for his insight into fisheries
+management, and the constructive feedback of the editor and anonymous
+reviewers. CB also acknowledges computational resources from NSF's XSEDE
+Jetstream (DEB160003) and Chameleon cloud platforms, as well as the
+support by the USDA Hatch project CA-B-INS-0162-H.
+
 \hypertarget{references}{%

### 2018-11-22 manuscripts/manuscript.Rmd
minor edits on 2nd review

@@
 
-The challenge of making decisions under uncertainty is obviously not unique to fisheries.
-Existing methods we optimal control methods which similarly assume perfect measurements can be found in a wide range of population and ecosystem management challenges, including fire ecology [@Richards1999], invasive species [@Blackwood2010; @Shea1998], disease outbreaks [@Shea2014; @Li2017], and protecting biodiversity [@Dee2017; @Iacona2017], and in common textbooks and reviews [@Mangel1985; @Clark1990; @Marescot2013].
+The challenge of making decisions under uncertainty is not unique to fisheries.
+Optimal control-based decision methods which assume perfect measurements are applied to a wide range of population and ecosystem management issues, including fire ecology [@Richards1999], invasive species [@Blackwood2010; @Shea1998], disease outbreaks [@Shea2014; @Li2017], and protecting biodiversity [@Dee2017; @Iacona2017], and in common textbooks and reviews [@Mangel1985; @Clark1990; @Marescot2013].
 Yet fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.
@@
 Further, our comparisons show that widely recognized existing approaches in both theory and practice are not by themselves sufficient to manage a population under substantial uncertainty over the long term, but our recently borrowed approach from robotics literature can under identical conditions sustain both higher biomass and higher economic yields than those approaches.
-This conclusion has important implications not only for fisheries but ecosystem more broadly: underscoring the danger of not correctly accounting for measurement uncertainty when making decisions, and further to the significant opportunity to learn from fields such as robotics when faced with complex and uncertain decisions.
+This conclusion has important implications not only for fisheries but ecosystem more broadly: First, our results highlight that measurement error must be accounted for in both the decision process itself, and not merely in the estimation of model.
+Second, these results underscore the opportunity to learn from engineering fields such as robotics when faced with complex and uncertain decisions that we have previously dismissed as intractable.
 
@@
 where the $\min$ function ensures no additional reward for attempting to harvest more than the entire stock.
-In economic terms, the corresponds to a fixed price per fish caught, with no proportional cost on effort required to do so, but the results do not depend on this assumption, as existing analyses have previously established [@Reed1979].
+In economic terms, this corresponds to a fixed price per fish caught, with no proportional cost on effort required to do so, but the results do not depend on this assumption, as existing analyses have previously established [@Reed1979].
 Using this form allows a more direct comparison to approaches such as maximum sustainable yield (MSY) which do not include an economic model but merely maximize long-term catch.
@@
 
-The solution to this decision problem is conditional on a population model for the underlying ecological process, describing probability the system moves to state $x_{t+1}$ given that it was in state $x_t$ and the manager selected a harvest $h_t$, which we discuss below.
+The solution to this decision problem is conditional on a population model for the underlying ecological process, describing the probability that the system moves to state $x_{t+1}$ given that it was in state $x_t$ and the manager selected a harvest $h_t$, which we discuss below.
 Such a model would typically be estimated and parameterized from available data, often (but not necessarily) using a hierarchical Bayesian approach that acknowledges the existence of uncertainty in both measurements and process [e.g.
@@
 Given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?
-Following convention in the optimal control literature (e.g.
-[@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.
+Following convention in the optimal control literature [e.g.
+@Reed1979; @Clark1990; @Sethi2005], we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.
 This assumption has occasionally been reversed to introduce some additional uncertainty from stochastic recruitment immediately prior to setting the harvest quota [e.g.
@@
 
-@Gordon1954 and @Schaefer1954 independently showed in the same year that the maximum sustainable yield for the model bearing their names is achieved by reducing the stock to the  population to the size at which it obtains the maximum growth rate which is known as Biomass at Maximum Sustainable Yield, $B_{MSY}$.
+@Gordon1954 and @Schaefer1954 independently showed that the maximum sustainable yield is achieved by reducing the stock to the population to the size at which it obtains the maximum growth rate.
+This stock size is known as Biomass at Maximum Sustainable Yield, $B_{MSY}$.
 For the Gordon-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$.
@@
 Thanks to a mathematical proof provided by @Reed1979, fisheries managers have largely been able to bi-pass such computational effort.
-However, this convenient result also opened the door to paradox of uncertainty that would haunt ecological management for at least the next four decades.
+However, this convenient result also opened the door to paradox of uncertainty that has persisted in ecological management for the past four decades.
 
@@
 More recently, driven by demands in areas such as autonomous vehicle decision making, newer point-based approximation algorithms such as SARSOP [@Kurniawati2008] have made it possible to solve POMDPs with 100s of states and actions.
-We have adapt this algorithm for application to the fisheries context, and provide an efficient implementation as a preliminary R package [@sarsop-pkg] which can be used to replicate and further explore results presented here.
-Details of the analysis presented here including annotated code to reproduce and further explore all of the following results is provided in the supplementary material.
+We have adapted this algorithm for application to the fisheries context, and provide an efficient implementation as a preliminary R package [@sarsop-pkg] which can be used to replicate and further explore results presented here.
+Details of the analysis presented here including annotated code to reproduce and further explore all of the following results are provided in the supplementary material.
 
 Precise definitions of both the (fully observed) MDPs and POMDPs (not to be confused with Markov Process Models and Hidden Markov Process Models -- methods for which estimate the model but do not involve the feedback of any decision process) in an ecological context can be found in @Williams2011, or in classic work from the engineering and robotics community [e.g.
-@Smallwood1973 and @Sondik1978].
+@Smallwood1973; @Sondik1978].
 For convenience, we summarize the POMDP definition and common notation here.
@@
 
-- Transition function (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.
+- Transition process (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.
 In our context, this relationship is given by the Gordon-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.
@@
 fig1_df %>% 
-  ggplot(aes(time, mean, col=method, fill=method)) + 
+  mutate(ymax = mean + sd, ymin = pmax(mean - sd, 0)) %>%
+  ggplot(aes(time, mean, col = method, fill = method)) + 
   geom_line(lwd=1) + 
-  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
+  geom_ribbon(aes(ymax = ymax, ymin = ymin), col = NA, alpha = 0.1) +
   facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
@@
 These results confirm that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE, but is less robust to environmental error.
+Many of the replicate simulations under CE experience complete stock extinction under large measurement error, owing to the arbitrarily large harvests permitted by the "bang-bang" nature of the CE policy.
 In contrast, the POMDP approach successfully handles both sources of uncertainty, maintaining higher stocks, near $B_{MSY}$ level, and producing the highest yield.
@@
 ```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$.
-Net present value is the sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{sims}.
+Net present value is the sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, assuming discount factor $\gamma = 0.95$, as shown in Figure \\ref{sims}.
 The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error.
@@
 This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also borne out in terms of economic performance.
-We find the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty (Figure \ref{econ} ).
+We find the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma = 0.95$) for these same simulations at increasing levels of measurement uncertainty (Figure \ref{econ}).
 As before, environmental stochasticity is set at $\sigma_g = 0.15$.
@@
 In contrast to the declining economic performance of CE and the consistently sub-optimal economic yield for TAC, the POMDP strategy continues to generate economic yields at approximately the optimal level (as attained by CE in the absence of measurement error) despite the increasingly uncertain measurements.
-This demonstrates that the reduced risk of stock collapse and higher average stock biomass attained by the POMDP strategy in the presence of high measurement error, Figure \ref{sims}) is not the result of a trivial reduction in harvesting under all circumstances, but rather, evidence of a more nuanced strategy that manages to account for the uncertainty in measurement while maintaining a reasonable harvest.
+This demonstrates that the reduced risk of stock collapse and higher average stock biomass attained by the POMDP strategy in the presence of high measurement error, Figure \ref{sims} is not the result of a trivial reduction in harvesting under all circumstances, but rather, evidence of a more nuanced strategy that manages to account for the uncertainty in measurement while maintaining a reasonable harvest.
 
@@
 Despite this overestimation, biomass under POMDP strategy closely tracks the optimal biomass under CE in the absence of measurement error.
+Solid lines indicate averages over 100 simulations, colored bands indicate +/- one standard deviation.
 \\label{overest}"}
@@
   geom_line(lwd = 1) + 
-  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
+  geom_ribbon(aes(ymax = mean + sd, ymin = pmax(mean-sd,0)), col = NA, alpha = 0.1) +
   coord_cartesian(ylim = c(0, 1))  + 
@@
 This difference between underestimating and overestimating accounts for the poor performance of the CE solution under large measurement uncertainty, where it over-harvests whenever measurements are too large.
-Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where the chance of local extinction becomes possible.
+Even though underestimating is equally likely under the measurement uncertainty model, sooner or later a run of "heads", a sequence of overestimations relative to the true stock, can drive stocks to very low levels where the chance of stock extinction becomes possible.
 It is precisely this asymmetry: that too much over-harvesting leads to an irreversible state of extinction, while too much under-harvesting is always reversible (modulo some lost revenue) that lay behind Clark's original intuition that there was something fishy about Reed's result that $S = D$: that uncertainty required no extra caution.
@@
 @Reed1979; @Clark1986; @Sethi2005].
-By using powerful algorithms developed first the robotics literature to solve POMDPs [@Kurniawati2008], we have been able to tackle this optimal control problem directly rather than resorting to the simplifying assumptions used in previous work.
+By using powerful algorithms developed first in the robotics literature to solve POMDPs [@Kurniawati2008], we have been able to tackle this optimal control problem directly rather than resorting to the simplifying assumptions used in previous work.
 Our solution confirms the intuition of the precautionary principle in a quantitatively precise form: under most (but not all) conditions, *greater uncertainty results in smaller optimal harvests*.

### 2018-11-22 reviews/README.Rmd
minor edits on 2nd review

@@
 author: "Carl Boettiger"
-date: "12/13/2017"
 output: github_document
@@
 ```
+
+Encrypt letter 3
+
+```{r}
+letter3 <- "review-reply-3.Rmd"
+gpg_letter3 <- "review-reply-3.Rmd.gpg"
+
+msg <- gpg_encrypt(letter3, receiver = cboettig, signer = cboettig)
+writeLines(msg, gpg_letter3)
+```
+
+Unencrypt committed gpg versions:
+
+```{r}
+writeLines(gpg_decrypt(gpg_letter3), letter3)
+```

### 2018-11-22 manuscripts/manuscript.Rmd
double-escape in latex in captions

@@
 ```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$.
-Net present value is the sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, assuming discount factor $\gamma = 0.95$, as shown in Figure \\ref{sims}.
+Net present value is the sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, assuming discount factor $\\gamma = 0.95$, as shown in Figure \\ref{sims}.
 The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error.

### 2018-11-22 manuscripts/manuscript.tex
rebuild manuscript

@@
             pdftitle={Resolving the measurement uncertainty paradox in ecological management},
-            colorlinks=true,
+            colorlinks=false,
             urlcolor=blue,
@@
 
-The challenge of making decisions under uncertainty is obviously not
-unique to fisheries. Existing methods we optimal control methods which
-similarly assume perfect measurements can be found in a wide range of
-population and ecosystem management challenges, including fire ecology
-(Richards et al. 1999), invasive species (Shea 1998; Blackwood et al.
-2010), disease outbreaks (Shea et al. 2014; Li et al. 2017), and
-protecting biodiversity (Dee et al. 2017; Iacona et al. 2017), and in
-common textbooks and reviews (Mangel 1985; Clark 1990; Marescot et al.
-2013). Yet fisheries conservation and management has long been both a
-crucible and proving ground for the theory of ecological management more
-generally, including topics such as adaptive management (Walters and
-Hilborn 1978), ecosystem-based management (Levin and Lubchenco 2008) and
-resilience thinking (Holling 1973; May 1977), while also giving rise to
-the sub-discipline of resource economics (Gordon and Press 1954;
-Schaefer 1954; Beverton and Holt 1957), thanks to its global relevance,
-long history, and readily available data. Despite the generality of the
+The challenge of making decisions under uncertainty is not unique to
+fisheries. Optimal control-based decision methods which assume perfect
+measurements are applied to a wide range of population and ecosystem
+management issues, including fire ecology (Richards et al. 1999),
+invasive species (Shea 1998; Blackwood et al. 2010), disease outbreaks
+(Shea et al. 2014; Li et al. 2017), and protecting biodiversity (Dee et
+al. 2017; Iacona et al. 2017), and in common textbooks and reviews
+(Mangel 1985; Clark 1990; Marescot et al. 2013). Yet fisheries
+conservation and management has long been both a crucible and proving
+ground for the theory of ecological management more generally, including
+topics such as adaptive management (Walters and Hilborn 1978),
+ecosystem-based management (Levin and Lubchenco 2008) and resilience
+thinking (Holling 1973; May 1977), while also giving rise to the
+sub-discipline of resource economics (Gordon and Press 1954; Schaefer
+1954; Beverton and Holt 1957), thanks to its global relevance, long
+history, and readily available data. Despite the generality of the
 models and issue of measurement uncertainty considered here, any
@@
 conclusion has important implications not only for fisheries but
-ecosystem more broadly: underscoring the danger of not correctly
-accounting for measurement uncertainty when making decisions, and
-further to the significant opportunity to learn from fields such as
-robotics when faced with complex and uncertain decisions.
+ecosystem more broadly: First, our results highlight that measurement
+error must be accounted for in both the decision process itself, and not
+merely in the estimation of model. Second, these results underscore the
+opportunity to learn from engineering fields such as robotics when faced
+with complex and uncertain decisions that we have previously dismissed
+as intractable.
 
@@
 where the \(\min\) function ensures no additional reward for attempting
-to harvest more than the entire stock. In economic terms, the
+to harvest more than the entire stock. In economic terms, this
 corresponds to a fixed price per fish caught, with no proportional cost
@@
 The solution to this decision problem is conditional on a population
-model for the underlying ecological process, describing probability the
-system moves to state \(x_{t+1}\) given that it was in state \(x_t\) and
-the manager selected a harvest \(h_t\), which we discuss below. Such a
-model would typically be estimated and parameterized from available
-data, often (but not necessarily) using a hierarchical Bayesian approach
-that acknowledges the existence of uncertainty in both measurements and
-process (e.g.~see Dichmont et al. 2016 for a discussion of how
-uncertainties are frequently included in stock assessment models).
-However, accounting for measurement uncertainty when going from data to
-model does not mean we have also accounted for this uncertainty when
-going from model to decision. For example, a harvest decision policy
-based upon maximum sustainable yield simply depends on the stock size
-which maximizes the expected growth rate of the model. Measurement
+model for the underlying ecological process, describing the probability
+that the system moves to state \(x_{t+1}\) given that it was in state
+\(x_t\) and the manager selected a harvest \(h_t\), which we discuss
+below. Such a model would typically be estimated and parameterized from
+available data, often (but not necessarily) using a hierarchical
+Bayesian approach that acknowledges the existence of uncertainty in both
+measurements and process (e.g.~see Dichmont et al. 2016 for a discussion
+of how uncertainties are frequently included in stock assessment
+models). However, accounting for measurement uncertainty when going from
+data to model does not mean we have also accounted for this uncertainty
+when going from model to decision. For example, a harvest decision
+policy based upon maximum sustainable yield simply depends on the stock
+size which maximizes the expected growth rate of the model. Measurement
 uncertainty in model estimation can indeed influence the position of
@@
 recruit, harvest, or measure, harvest, recruit? Following convention in
-the optimal control literature (e.g. (Reed 1979; Clark 1990; Sethi et
-al. 2005)), we will assume the latter; measurement occurs immediately
-before harvest, followed by recruitment. This assumption has
-occasionally been reversed to introduce some additional uncertainty from
-stochastic recruitment immediately prior to setting the harvest quota
-(e.g. Clark and Kirkwood 1986; Weitzman 2002), which appears to mimic
-uncertainty in measurements but does not allow uncertainty to accumulate
-since the measurement step still occurs without error.
+the optimal control literature (e.g. Reed 1979; Clark 1990; Sethi et al.
+2005), we will assume the latter; measurement occurs immediately before
+harvest, followed by recruitment. This assumption has occasionally been
+reversed to introduce some additional uncertainty from stochastic
+recruitment immediately prior to setting the harvest quota (e.g. Clark
+and Kirkwood 1986; Weitzman 2002), which appears to mimic uncertainty in
+measurements but does not allow uncertainty to accumulate since the
+measurement step still occurs without error.
 
@@
 
-Gordon and Press (1954) and Schaefer (1954) independently showed in the
-same year that the maximum sustainable yield for the model bearing their
-names is achieved by reducing the stock to the population to the size at
-which it obtains the maximum growth rate which is known as Biomass at
-Maximum Sustainable Yield, \(B_{MSY}\). For the Gordon-Schaefer model
-(and many others), this is achieved at \(B_{MSY} = K/2\). They observed
-that fishing at a \emph{constant yield} (that is, an individual fish
-mortality, \(F\), such that harvest \(H = F \cdot B\)) of
+Gordon and Press (1954) and Schaefer (1954) independently showed that
+the maximum sustainable yield is achieved by reducing the stock to the
+population to the size at which it obtains the maximum growth rate. This
+stock size is known as Biomass at Maximum Sustainable Yield,
+\(B_{MSY}\). For the Gordon-Schaefer model (and many others), this is
+achieved at \(B_{MSY} = K/2\). They observed that fishing at a
+\emph{constant yield} (that is, an individual fish mortality, \(F\),
+such that harvest \(H = F \cdot B\)) of
 \(F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}\)) will eventually lead to a
@@
 convenient result also opened the door to paradox of uncertainty that
-would haunt ecological management for at least the next four decades.
+has persisted in ecological management for the past four decades.
 
@@
 (Kurniawati et al. 2008) have made it possible to solve POMDPs with 100s
-of states and actions. We have adapt this algorithm for application to
+of states and actions. We have adapted this algorithm for application to
 the fisheries context, and provide an efficient implementation as a
@@
 analysis presented here including annotated code to reproduce and
-further explore all of the following results is provided in the
+further explore all of the following results are provided in the
 supplementary material.
@@
 Williams (2011), or in classic work from the engineering and robotics
-community (e.g. Smallwood and Sondik 1973, and @Sondik1978). For
+community (e.g. Smallwood and Sondik 1973; Sondik 1978). For
 convenience, we summarize the POMDP definition and common notation here.
@@
 \item
-  Transition function (state equation): \(T(x_t, x_{t+1}, a_t)\): the
+  Transition process (state equation): \(T(x_t, x_{t+1}, a_t)\): the
   probability that a system is in state \(x_{t+1}\) at time \(t+1\)
@@
 solution represented by CE, but is less robust to environmental error.
-In contrast, the POMDP approach successfully handles both sources of
+Many of the replicate simulations under CE experience complete stock
+extinction under large measurement error, owing to the arbitrarily large
+harvests permitted by the ``bang-bang'' nature of the CE policy. In
+contrast, the POMDP approach successfully handles both sources of
 uncertainty, maintaining higher stocks, near \(B_{MSY}\) level, and
@@
 value is the sum of all future harvests discounted over time and
-averaged across 100 replicates simulations under each strategy, as shown
-in Figure \ref{sims}. The value under constant escapement (CE) is
-optimal when measurements are perfect but decreases rapidly with
-increasing measurement error. The more cautious TAC is not economically
-optimal but largely unimpacted by increasing error, while POMDP attains
-consistently high economic yield despite the increasing uncertainty.
-\label{econ}}
+averaged across 100 replicates simulations under each strategy, assuming
+discount factor \(\gamma = 0.95\), as shown in Figure \ref{sims}. The
+value under constant escapement (CE) is optimal when measurements are
+perfect but decreases rapidly with increasing measurement error. The
+more cautious TAC is not economically optimal but largely unimpacted by
+increasing error, while POMDP attains consistently high economic yield
+despite the increasing uncertainty. \label{econ}}
 \end{figure}
@@
 across replicates and discounting future profits by the discount factor
-\(\gamma\)) for these same simulations at increasing levels of
-measurement uncertainty (Figure \ref{econ} ). As before, environmental
+\(\gamma = 0.95\)) for these same simulations at increasing levels of
+measurement uncertainty (Figure \ref{econ}). As before, environmental
 stochasticity is set at \(\sigma_g = 0.15\). In the absence of
@@
 collapse and higher average stock biomass attained by the POMDP strategy
-in the presence of high measurement error, Figure \ref{sims}) is not the
+in the presence of high measurement error, Figure \ref{sims} is not the
 result of a trivial reduction in harvesting under all circumstances, but
@@
 closely tracks the optimal biomass under CE in the absence of
-measurement error. \label{overest}}
+measurement error. Solid lines indicate averages over 100 simulations,
+colored bands indicate +/- one standard deviation. \label{overest}}
 \end{figure}
@@
 overestimations relative to the true stock, can drive stocks to very low
-levels where the chance of local extinction becomes possible. It is
+levels where the chance of stock extinction becomes possible. It is
 precisely this asymmetry: that too much over-harvesting leads to an
@@
 (e.g. Reed 1979; Clark and Kirkwood 1986; Sethi et al. 2005). By using
-powerful algorithms developed first the robotics literature to solve
+powerful algorithms developed first in the robotics literature to solve
 POMDPs (Kurniawati et al. 2008), we have been able to tackle this

### 2018-11-22 manuscripts/manuscript.tex
fix a few refs
rebuild pdf

@@
 \leavevmode\hypertarget{ref-sarsop-pkg}{}%
-Boettiger, C., J. Ooms, and M. Memarzadeh. 2018. Sarsop: Approximate
-pomdp planning software.
+Boettiger, C., J. Ooms, and M. Memarzadeh. 2018. sarsop: Approximate
+POMDP Planning Software in R, v0.5.0,
+\textless{}https://github.com/boettiger-lab/sarsop\textgreater{}.
 
@@
 in the presence of climatic change and outcome uncertainty using Monte
-Carlo simulations. (R. Hilborn, ed.)Canadian Journal of Fisheries and
-Aquatic Sciences 66:2199--2221.
+Carlo simulations. Canadian Journal of Fisheries and Aquatic Sciences
+66:2199--2221.
 
@@
 optimal conservation strategy, even in a crisis discipline. Proceedings
-of the National Academy of Sciences 201702111.
+of the National Academy of Sciences 114:201702111.
 
@@
 Marboutin, and O. Gimenez. 2013. Complex decisions made simple: a primer
-on stochastic dynamic programming. (R. Freckleton, ed.)Methods in
-Ecology and Evolution 4:872--884.
+on stochastic dynamic programming. Methods in Ecology and Evolution
+4:872--884.
 

### 2018-11-22 manuscripts/manuscript.Rmd
fix typo

@@
 For the Gordon-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$.
-They observed that fishing at a *constant yield* (that is, an individual fish mortality, $F$, such that harvest $H = F \cdot B$) of $F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}$)  will eventually lead to a population that converges to the biomass $B_{MSY}$ and produces the maximum sustainable harvest, $H_{MSY} = r K / 4$ in this model.
+They observed that fishing at a *constant yield* (that is, an individual fish mortality, $F$, such that harvest is biomass times mortality, $H = F \cdot B$, and thus $F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}$)  will eventually lead to a population that converges to the biomass $B_{MSY}$ and produces the maximum sustainable harvest, $H_{MSY} = r K / 4$ in this model.
 From here, theory and practice diverge.

### 2018-11-22 manuscripts/manuscript.tex
fix typo

@@
 \emph{constant yield} (that is, an individual fish mortality, \(F\),
-such that harvest \(H = F \cdot B\)) of
-\(F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}\)) will eventually lead to a
+such that harvest is biomass times mortality, \(H = F \cdot B\), and
+thus \(F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}\)) will eventually lead to a
 population that converges to the biomass \(B_{MSY}\) and produces the

### 2018-11-23 appendix/appendixA.Rmd
rebuild appendices

@@
 ```
+
+\pagebreak
+
+# Setup
 
@@
 library(hrbrthemes)
-
+library(tictoc)
 extrafont::loadfonts()
@@
 ```{r}
+tic()
 options(mc.cores=parallel::detectCores())
@@
 
+\pagebreak
+
+Total runtime:
+
+```{r}
+toc()
+```
+
+Hardware:
+
+```{r}
+system2("grep", c("MemTotal", "/proc/meminfo"), stdout = TRUE)
+system2('grep', '-m 1 "model name" /proc/cpuinfo', stdout = TRUE)
+```
+
+Software:
+
+```{r}
+sessionInfo()
+```
+
+\pagebreak
+
 # References

### 2018-11-23 appendix/appendixB.Rmd
rebuild appendices

@@
 library(gridExtra)
+library(tictoc)
+tic()
 ```
@@
 ```
+
+\pagebreak
+
+Total runtime:
+
+```{r}
+toc()
+```
+
+Hardware:
+
+```{r}
+system2("grep", c("MemTotal", "/proc/meminfo"), stdout = TRUE)
+system2('grep', '"model name" /proc/cpuinfo', stdout = TRUE)
+```
+
+Software:
+
+```{r}
+sessionInfo()
+```

### 2018-12-03 appendix/appendixA.Rmd
rebuild from scratch

@@
 ```{r knitr, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf", echo=TRUE)
+knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=FALSE, dev="cairo_pdf", echo=TRUE)
 ```
@@
 options(mc.cores=parallel::detectCores())
-log_dir <- "../data/observe_harvest_recruit" # Store the computed solution files here
+log_dir <- "../data/appendixA" # Store the computed solution files here
 
@@
 
-write_csv(det_sims, "../data/observe_harvest_recruit/det_sims.csv")
+write_csv(det_sims, "../data/appendixA/det_sims.csv")
 ```
@@
 
-```{r eval=!file.exists(file.path(log_dir, "meta.csv"))}
-dir.create(log_dir)
+```{r}
+dir.create(log_dir, FALSE)
 system.time(
@@
              discount = discount,
-             precision = 0.00000002,
+             precision = 1e-8,
              timeout = 15000,
@@
 
-```
-
-We can read the stored solution from the log:
-
-```{r}
-meta <- meta_from_log(data.frame(model="gs", discount=discount), log_dir) %>% 
-  mutate(scenario = as.character(scenario)) %>%
-  left_join(meta) %>% 
-  arrange(scenario)
-alphas <- alphas_from_log(meta, log_dir)
 ```
@@
 
-```{r pomdp_sims, eval=!file.exists(file.path(log_dir, "pomdp_sims.csv"))}
+```{r pomdp_sims}
 set.seed(12345)
@@
 
-```{r overest, eval=!file.exists(file.path(log_dir, "pomdp_overest_sims.csv"))}
+```{r overest}
 set.seed(12345)
@@
 
+# System Information
+
 Total runtime:
@@
 
-Hardware:
+### Hardware:
 
@@
 system2("grep", c("MemTotal", "/proc/meminfo"), stdout = TRUE)
-system2('grep', '-m 1 "model name" /proc/cpuinfo', stdout = TRUE)
+system2('grep', '"model name" /proc/cpuinfo', stdout = TRUE)
 ```
 
-Software:
+### Software:
 
 ```{r}
-sessionInfo()
+devtools::session_info()
 ```

### 2018-12-03 appendix/appendixB.Rmd
rebuild from scratch

@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+
 # Section 1: Model estimates from data
@@
 ```{r knitr, include=FALSE}
-knitr::opts_chunk$set(comment="", message=FALSE, warning = FALSE, cache=TRUE, dev="cairo_pdf")
+knitr::opts_chunk$set(comment = "", message = FALSE, 
+                      warning = FALSE, cache = FALSE, 
+                      dev = "cairo_pdf")
 ```
@@
 
-```{r plotting, include = FALSE}
+```{r plotting, include = FALSE, cache = FALSE}
 ## This chunk is optional, just tweaks plot appearance / fonts
@@
                       "catch"))
+## Overwriting is convenient, but breaks any plot not using these names!
 scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
@@
   N <- dim(fish)[1]
-  scaled_data <- data.frame(t = 1:N, y = fish$scaled_biomass, a = fish$scaled_catch)
+  scaled_data <- data.frame(t = 1:N, 
+                            y = fish$scaled_biomass, 
+                            a = fish$scaled_catch)
   data = data.frame(x = scaled_data$y)
@@
 
-```{r include = FALSE, eval = FALSE}
-## Do the estimates make sense?  Consider simulation from models, 
-## under this historical harvests, compared to historical trajectories
-examples %>% ungroup() %>%
-  select(year, scaled_biomass, scaled_catch, commonname) %>%
-  gather(stock, biomass, -year, -commonname) %>%
-  ggplot(aes(year, biomass, col=stock)) + 
-  geom_line() + facet_wrap(~commonname, scales = "free")
-```
-
 ```{r}
@@
 options(mc.cores = parallel::detectCores()) # Reserve ~ 10 GB per core
-log_dir <- "../data/ram_ex"
+log_dir <- "../data/appendixB"
 ```
@@
 
-```{r eval=FALSE}
-dir.create(log_dir)
+```{r message = FALSE, output = "hide"}
+dir.create(log_dir, FALSE)
 options(mc.cores = 1)
@@
              discount = discount,
-             precision = 2e-6,
-             timeout = 25000,
+             precision = 2e-8,
+             timeout = 26000,
              log_dir = log_dir,
@@
 
-```
-
-```{r message=FALSE}
-meta <- meta_from_log(data.frame(model="gs"), log_dir)  %>% 
-  mutate(scenario = as.character(scenario)) %>%
-  left_join(
-    select(meta, sigma_m, commonname, scenario),   
-    by = c("scenario", "sigma_m", "commonname")) %>% 
-  arrange(scenario)
-alphas <- alphas_from_log(meta, log_dir)
-#models <- models_from_log(meta)
 ```
@@
 
-write_csv(sims,file.path(log_dir, "ram_ex.csv"))
+write_csv(sims,file.path(log_dir, "appendixB.csv"))
 ```
@@
 ```{r}
-ram_ex <- read_csv(file.path(log_dir, "ram_ex.csv"))
-ram_ex %>%
+appendixB <- read_csv(file.path(log_dir, "appendixB.csv"))
+appendixB %>%
   filter(model %in% c("biomass", "catch", "POMDP", "CE", "TAC", "MSY")) %>%
@@
 
-Software:
+### Software:
 
 ```{r}
-sessionInfo()
+devtools::session_info()
 ```

### 2018-12-07 manuscripts/manuscript.Rmd
rebuild mansucript locally

@@
 ```{r load_data, message=FALSE, warning=FALSE}
-log_dir <- "../data/observe_harvest_recruit" # solution files here
+log_dir <- "../data/appendixA" # solution files here
 
@@
 ## Fig 5
-ram_ex <- read_csv(file.path(log_dir, "../ram_ex/ram_ex.csv"))
+ram_ex <- read_csv(file.path("../data/appendixB/appendixB.csv"))
 

### 2018-12-07 install.R

@@ -1,1 +1,1 @@
-tinytex::tlmgr_install(c("lineno", "ec", "psnfss"))
+tinytex::tlmgr_install(c("lineno", "ec", "psnfss", "endfloat"))

### 2018-12-07 appendix/appendixA.Rmd
try to cache slow block for travis
fix typo
cache another possibly slow block

@@
 
-```{r pomdp_sims}
+```{r pomdp_sims, eval=!file.exists(file.path(log_dir, "pomdp_sims.csv"))}
+
 set.seed(12345)
@@
 
-```{r overest}
+```{r overest, eval = !file.exists(file.path(log_dir, "pomdp_overest_sims.csv"))}
 set.seed(12345)

### 2018-12-28 appendix/appendixB.Rmd
rebuild appendix B

@@
 library(tictoc)
+library(furrr)
 tic()
@@
 names(ram) <- sheets
-unlink("ramlegacy.zip")
-unlink("RLSADB v3.0 (assessment data only).xlsx")
+
 ramlegacy <- 
@@
 ```{r}
-options(mc.cores = parallel::detectCores()) # Reserve ~ 10 GB per core
+options(mc.cores = 6) # Reserve ~ 10 GB per core
 log_dir <- "../data/appendixB"
@@
 ```{r}
+plan(multiprocess)
+
 models <- 
-  parallel::mclapply(1:dim(meta)[1], 
+  furrr::future_map(1:dim(meta)[1], 
                      function(i){
@@
 dir.create(log_dir, FALSE)
-options(mc.cores = 1)
-
-## POMDP solution (slow, >20,000 seconds per loop memory intensive)
+plan(multiprocess)
+## POMDP solution 
 system.time(
   alphas <- 
-    parallel::mclapply(1:length(models), 
+    furrr::future_map(1:length(models), 
     function(i){
@@
              discount = discount,
-             precision = 2e-8,
-             timeout = 26000,
+             precision = 2e-6,
+             timeout = 25000,
              log_dir = log_dir,
@@
   pars %>% 
-  rowwise() %>% 
-  do(statics(.))
+  transpose() %>% 
+  map_df(statics)
 ```
@@
 
-```{r}
+```{r include=FALSE}
+## allow resizing output font in latex
+def.chunk.hook  <- knitr::knit_hooks$get("chunk")
+knitr::knit_hooks$set(chunk = function(x, options) {
+  x <- def.chunk.hook(x, options)
+  ifelse(options$size != "normalsize", paste0("\\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
+})
+```
+
+```{r  size="tiny"}
 system2("grep", c("MemTotal", "/proc/meminfo"), stdout = TRUE)
@@
 
-```{r}
+```{r size="tiny"}
 devtools::session_info()

### 2018-12-28 manuscripts/manuscript.Rmd
rebuild manuscript

@@
 title: "Resolving the measurement uncertainty paradox in ecological management"
-
 author:
-  - name: "Authors removed during peer review"
-
-#  - name: "Milad Memarzadeh"
-#    affiliation: a
-#  - name: "Carl Boettiger"
-#    affiliation: a, 1
-#address:
-#  - code: a
-#    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
-
+  - name: "Milad Memarzadeh"
+    affiliation: a
+  - name: "Carl Boettiger"
+    affiliation: a, 1
+address:
+  - code: a
+    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
 abstract: |
@@
 These results have significant implications for how we handle uncertainty in both fisheries and ecological management more generally.
-
 keywords:
@@
   - conservation
-
-lineno: true
-
 bibliography: citations.bib
-#journal: "The American Naturalist"
 date: "`r Sys.Date()`"
@@
 layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column
-header-includes: |
+preamble: |
   \journal{The American Naturalist}
-        \usepackage{lineno}
-        \linenumbers
-        \modulolinenumbers[3]
-        \usepackage{setspace}
-        \doublespacing
+  \usepackage[nomarkers]{endfloat}
 
@@
 
-- Transition process (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.
+**Transition** process (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.
 In our context, this relationship is given by the Gordon-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.
@@
 
-- Observation function: $O(x_t,y_t,a_{t-1})$ the probability of observing state $y_t$ given a system in state $x_t$.
+**Observation** function: $O(x_t,y_t,a_{t-1})$ the probability of observing state $y_t$ given a system in state $x_t$.
 In principle, the action chosen can influence the precision of the observation.
@@
 
-- Utility function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.
+**Utility** function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.
 For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t, a_t)$, indicating that realized harvest cannot be negative.
@@
 While in the following decades ecologists largely turned away from computationally intensive optimization required for even simple models with stochastic dynamic programming, (while continuing to construct ever more complicated hierarchical models), engineers have chipped away at the decision problem to reveal algorithms such as SARSOP that make much larger dynamic programming problems tractable.
-Yes, we may need even more sophisticated algorithms to accommodate greater ecological complexity, yet we believe such numerical obstacles are no excuse for management by half measures.
+We may need even more sophisticated algorithms to accommodate greater ecological complexity, yet we believe such numerical obstacles are no excuse for management by half measures.
 Rather, our example illustrates how ecologists and managers can learn from and leverage the tools developed in other disciplines to better face the challenges of management in a complex and changing world.

### 2018-12-28 manuscripts/manuscript.tex
rebuild manuscript

@@
 % Pandoc header
+
 \journal{The American Naturalist}
-\usepackage{lineno}
-\linenumbers
-\modulolinenumbers[3]
-\usepackage{setspace}
-\doublespacing
+\usepackage[nomarkers]{endfloat}
 
@@
   \title{Resolving the measurement uncertainty paradox in ecological management}
-    \author[]{Authors removed during peer review}
+    \author[a]{Milad Memarzadeh}
+
+    \author[a, 1]{Carl Boettiger}
+
+      \address[a]{Dept of Environmental Science, Policy, and Management, University of
+California Berkeley, Berkeley CA 94720-3114, USA}
 
@@
 
-\begin{itemize}
-\item
-  Transition process (state equation): \(T(x_t, x_{t+1}, a_t)\): the
-  probability that a system is in state \(x_{t+1}\) at time \(t+1\)
-  given that it began in state \(x_t\) at time \(t\) and the manager
-  took action \(a_t\). In our context, this relationship is given by the
+\textbf{Transition} process (state equation): \(T(x_t, x_{t+1}, a_t)\):
+the probability that a system is in state \(x_{t+1}\) at time \(t+1\)
+given that it began in state \(x_t\) at time \(t\) and the manager took
+action \(a_t\). In our context, this relationship is given by the
 Gordon-Schaefer stock recruitment function \(f\) with normally
 distributed growth uncertainty
-  \(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero
-  to exclude negative population sizes. The action \(a_t\) represents
-  the harvest quota set for the fishery. Attempting to implement a quota
-  the exceeds the true population size results in collapse of the
-  fishery. As Clark and Kirkwood (1986) eloquently argues, assuming that
-  extinction is impossible regardless of fishing intensity would
-  unreasonably bias the decision problem.
-\item
-  Observation function: \(O(x_t,y_t,a_{t-1})\) the probability of
+\(x_{t+1} \sim \mathcal{N}(f(x_t,a_t), \sigma_g)\), truncated at zero to
+exclude negative population sizes. The action \(a_t\) represents the
+harvest quota set for the fishery. Attempting to implement a quota the
+exceeds the true population size results in collapse of the fishery. As
+Clark and Kirkwood (1986) eloquently argues, assuming that extinction is
+impossible regardless of fishing intensity would unreasonably bias the
+decision problem.
+
+\textbf{Observation} function: \(O(x_t,y_t,a_{t-1})\) the probability of
 observing state \(y_t\) given a system in state \(x_t\). In principle,
-  the action chosen can influence the precision of the observation. In
-  our case, we simply assume normally distributed errors around the true
+the action chosen can influence the precision of the observation. In our
+case, we simply assume normally distributed errors around the true
 state, \(y_t \sim \mathcal{N}(x_t, \sigma_m)\), truncated at zero to
 exclude negative population sizes.
-\item
-  Utility function: \(U(x_t,a_t)\), the utility received at time \(t\)
-  for taking action \(a_t\), given that the system is in state \(x_t\).
-  For simplicity of analysis, we will simply set the utility to be equal
-  to the harvested stock: \(U(x_t, a_t) = \min(x_t, a_t)\), indicating
-  that realized harvest cannot be negative. This choice ensures that in
-  the case of no uncertainty (\(\sigma_g = \sigma_m = 0\)), the optimal
-  solution matches that expected under a simple MSY calculation. More
-  realistic utility functions may include diminishing returns with
-  increasing harvest (supply and demand effects), and the cost of
-  fishing, both of which act to suppress large harvests. By focusing on
-  a simple utility we can be sure that our comparison to MSY is driven
-  by the treatment of uncertainty rather than merely differing economic
+
+\textbf{Utility} function: \(U(x_t,a_t)\), the utility received at time
+\(t\) for taking action \(a_t\), given that the system is in state
+\(x_t\). For simplicity of analysis, we will simply set the utility to
+be equal to the harvested stock: \(U(x_t, a_t) = \min(x_t, a_t)\),
+indicating that realized harvest cannot be negative. This choice ensures
+that in the case of no uncertainty (\(\sigma_g = \sigma_m = 0\)), the
+optimal solution matches that expected under a simple MSY calculation.
+More realistic utility functions may include diminishing returns with
+increasing harvest (supply and demand effects), and the cost of fishing,
+both of which act to suppress large harvests. By focusing on a simple
+utility we can be sure that our comparison to MSY is driven by the
+treatment of uncertainty rather than merely differing economic
 assumptions.
-\end{itemize}
 
@@
 the decision problem to reveal algorithms such as SARSOP that make much
-larger dynamic programming problems tractable. Yes, we may need even
-more sophisticated algorithms to accommodate greater ecological
-complexity, yet we believe such numerical obstacles are no excuse for
-management by half measures. Rather, our example illustrates how
-ecologists and managers can learn from and leverage the tools developed
-in other disciplines to better face the challenges of management in a
-complex and changing world.
+larger dynamic programming problems tractable. We may need even more
+sophisticated algorithms to accommodate greater ecological complexity,
+yet we believe such numerical obstacles are no excuse for management by
+half measures. Rather, our example illustrates how ecologists and
+managers can learn from and leverage the tools developed in other
+disciplines to better face the challenges of management in a complex and
+changing world.
 

### 2018-12-28 manuscripts/manuscript.tex
add license and citation to code repo

@@
 ecological and economic outcomes under management strategies that do not
-appropriately account for that error over the long run.
+appropriately account for that error over the long run. We provide
+complete code and data files to reproduce the analysis shown here in the
+Appendices and as research compendium (Marwick et al. 2018) at
+\url{https://github.com/boettiger-lab/pomdp-intro}, archived at
+(Boettiger and Memarzadeh 2018).
 
@@
 
+\leavevmode\hypertarget{ref-pomdp-intro}{}%
+Boettiger, C., and M. Memarzadeh. 2018. Supplement to: "Resolving the
+measurement uncertainty paradox in ecological management". Zenodo.
+
 \leavevmode\hypertarget{ref-sarsop-pkg}{}%
@@
 
+\leavevmode\hypertarget{ref-Marwick2018}{}%
+Marwick, B., C. Boettiger, and L. Mullen. 2018. Packaging Data
+Analytical Work Reproducibly Using R (and Friends). The American
+Statistician 72:80--88.
+
 \leavevmode\hypertarget{ref-May1977}{}%

### 2018-12-28 manuscripts/manuscript.tex
with DOIs

@@
 Ecological Economics 69:519--527.
+DOI:\href{https://doi.org/10.1016/j.ecolecon.2009.08.029}{10.1016/j.ecolecon.2009.08.029}.
 
@@
 282:20141631--20141631.
+DOI:\href{https://doi.org/10.1098/rspb.2014.1631}{10.1098/rspb.2014.1631}.
 
@@
 Boettiger, C., and M. Memarzadeh. 2018. Supplement to: "Resolving the
-measurement uncertainty paradox in ecological management". Zenodo.
+measurement uncertainty paradox in ecological management". Zenodo,
+DOI:\href{https://doi.org/10.5281/zenodo.2528507}{10.5281/zenodo.2528507}.
 
@@
 \textless{}https://github.com/boettiger-lab/sarsop\textgreater{}.
+DOI:\href{https://doi.org/10.5281/zenodo.1162856}{10.5281/zenodo.1162856}.
 
@@
 Communications 8:15325.
+DOI:\href{https://doi.org/10.1038/ncomms15325}{10.1038/ncomms15325}.
 
@@
 Academy of Sciences 108:8323--8.
+DOI:\href{https://doi.org/10.1073/pnas.1016846108}{10.1073/pnas.1016846108}.
 
@@
 105:13936--40.
+DOI:\href{https://doi.org/10.1073/pnas.0805265105}{10.1073/pnas.0805265105}.
 
@@
 species. Journal of Political Economy 81:950--961.
+DOI:\href{https://doi.org/10.1086/260090}{10.1086/260090}.
 
@@
 of Environmental Economics and Management 13:235--244.
+DOI:\href{https://doi.org/10.1016/0095-0696(86)90024-0}{10.1016/0095-0696(86)90024-0}.
 
@@
 Sciences 113:5125--5129.
+DOI:\href{https://doi.org/10.1073/pnas.1520420113}{10.1073/pnas.1520420113}.
 
@@
 Letters 20:935--946.
+DOI:\href{https://doi.org/10.1111/ele.12790}{10.1111/ele.12790}.
 
@@
 Computational and Graphical Statistics 26:403--413.
+DOI:\href{https://doi.org/10.1080/10618600.2016.1172487}{10.1080/10618600.2016.1172487}.
 
@@
 in the united states. Fisheries Research 183:447--460.
+DOI:\href{https://doi.org/10.1016/j.fishres.2016.07.001}{10.1016/j.fishres.2016.07.001}.
 
@@
 66:2199--2221.
+DOI:\href{https://doi.org/10.1139/F09-144}{10.1139/F09-144}.
 
@@
 7:509--520.
+DOI:\href{https://doi.org/10.1111/j.1461-0248.2004.00603.x}{10.1111/j.1461-0248.2004.00603.x}.
 
@@
 24:549--54.
+DOI:\href{https://doi.org/10.1016/j.tree.2009.03.020}{10.1016/j.tree.2009.03.020}.
 
@@
 Common-Property Resource: The Fishery. Journal of Political Economy
-62:124--142.
+62:124--142. DOI:\href{https://doi.org/10.1086/257497}{10.1086/257497}.
 
@@
 110:6229--34.
+DOI:\href{https://doi.org/10.1073/pnas.1217689110}{10.1073/pnas.1217689110}.
 
@@
 Annual Review of Ecology and Systematics 4:1--23.
+DOI:\href{https://doi.org/10.1146/annurev.es.04.110173.000245}{10.1146/annurev.es.04.110173.000245}.
 
@@
 of the National Academy of Sciences 114:201702111.
+DOI:\href{https://doi.org/10.1073/pnas.1702111114}{10.1073/pnas.1702111114}.
 
@@
 A. S. Garmestani, eds. Adaptive management of social-ecological systems.
-Springer Netherlands, Dordrecht.
+Springer Netherlands, Dordrecht,
+DOI:\href{https://doi.org/10.1007/978-94-017-9682-8_5}{10.1007/978-94-017-9682-8\_5}.
 
@@
 Intelligence 101:99--134.
+DOI:\href{https://doi.org/10.1016/S0004-3702(98)00023-X}{10.1016/S0004-3702(98)00023-X}.
 
@@
 environmental science. Environmental health perspectives 109:871--6.
+DOI:\href{https://doi.org/10.2307/3454986}{10.2307/3454986}.
 
@@
 Yield. Transactions of the American Fisheries Society 106:1--11.
+DOI:\href{https://doi.org/10.1577/1548-8659(1977)106\%3C1:AEFTCO\%3E2.0.CO;2}{10.1577/1548-8659(1977)106\textless{}1:AEFTCO\textgreater{}2.0.CO;2}.
 
@@
 Ecosystem-based Management. BioScience 58:27--32.
+DOI:\href{https://doi.org/10.1641/B580107}{10.1641/B580107}.
 
@@
 National Academy of Sciences 114:5659--5664.
+DOI:\href{https://doi.org/10.1073/pnas.1617482114}{10.1073/pnas.1617482114}.
 
@@
 parameter estimates. Ecological Modelling 14:273--292.
+DOI:\href{https://doi.org/10.1016/0304-3800(82)90023-0}{10.1016/0304-3800(82)90023-0}.
 
@@
 Fisheries 2:2--32.
+DOI:\href{https://doi.org/10.1046/j.1467-2979.2001.00033.x}{10.1046/j.1467-2979.2001.00033.x}.
 
@@
 4:872--884.
+DOI:\href{https://doi.org/10.1111/2041-210X.12082}{10.1111/2041-210X.12082}.
 
@@
 Statistician 72:80--88.
+DOI:\href{https://doi.org/10.1080/00031305.2017.1375986}{10.1080/00031305.2017.1375986}.
 
@@
 multiplicity of stable states. Nature 269:471--477.
+DOI:\href{https://doi.org/10.1038/269471a0}{10.1038/269471a0}.
 
@@
 224:9--15.
+DOI:\href{https://doi.org/10.1016/j.biocon.2018.05.009}{10.1016/j.biocon.2018.05.009}.
 
@@
 large state spaces. Methods in Ecology and Evolution 2:221--228.
+DOI:\href{https://doi.org/10.1111/j.2041-210X.2010.00069.x}{10.1111/j.2041-210X.2010.00069.x}.
 
@@
 18:597--611.
+DOI:\href{https://doi.org/10.1111/ele.12443}{10.1111/ele.12443}.
 
@@
 era of global change. Trends in Ecology \& Evolution 26:398--404.
+DOI:\href{https://doi.org/10.1016/j.tree.2011.04.007}{10.1016/j.tree.2011.04.007}.
 
@@
 and Fisheries 17:303--334.
+DOI:\href{https://doi.org/10.1111/faf.12104}{10.1111/faf.12104}.
 
@@
 Management 6:350--363.
+DOI:\href{https://doi.org/10.1016/0095-0696(79)90014-7}{10.1016/0095-0696(79)90014-7}.
 
@@
 9:880--892.
+DOI:\href{https://doi.org/10.1890/1051-0761(1999)009\%5B0880:OFMFMC\%5D2.0.CO;2}{10.1890/1051-0761(1999)009{[}0880:OFMFMC{]}2.0.CO;2}.
 
@@
 of the Inter-American Tropical Tuna Commission 1:27--56.
+DOI:\href{https://doi.org/10.1007/BF02464432}{10.1007/BF02464432}.
 
@@
 Economics and Management 50:300--318.
+DOI:\href{https://doi.org/10.1016/j.jeem.2004.11.005}{10.1016/j.jeem.2004.11.005}.
 
@@
 control. Trends in Ecology and Evolution 13:371--375.
+DOI:\href{https://doi.org/10.1016/S0169-5347(98)01381-0}{10.1016/S0169-5347(98)01381-0}.
 
@@
 Learning Via Intervention in Epidemiology. PLoS Biology 12:9--12.
+DOI:\href{https://doi.org/10.1371/journal.pbio.1001970}{10.1371/journal.pbio.1001970}.
 
@@
 Systems. Journal of the Fisheries Research Board of Canada 33:145--159.
+DOI:\href{https://doi.org/10.1139/f76-017}{10.1139/f76-017}.
 
@@
 Review of Ecology and Systematics 9:157--188.
+DOI:\href{https://doi.org/10.1146/annurev.es.09.110178.001105}{10.1146/annurev.es.09.110178.001105}.
 
@@
 Stocks. Journal of Environmental Economics and Management 43:325--338.
+DOI:\href{https://doi.org/10.1006/jeem.2000.1181}{10.1006/jeem.2000.1181}.
 
@@
 8:269--288.
+DOI:\href{https://doi.org/10.1023/A:1011395725123}{10.1023/A:1011395725123}.
 
@@
 management using POMDP approaches. Ecological Modelling 222:1092--1102.
+DOI:\href{https://doi.org/10.1016/j.ecolmodel.2010.12.015}{10.1016/j.ecolmodel.2010.12.015}.
 

### 2019-04-11 appendix/appendixB.Rmd
ramlegacy data no longer available
from ramlegacy.org.  Use cached version instead

@@
 ```{r}
-if(!file.exists("ramlegacy.zip")){
-download.file(paste0(
-  "https://depts.washington.edu/ramlegac/wordpress/databaseVersions/",
-  "RLSADB_v3.0_(assessment_data_only)_excel.zip"),
-  "ramlegacy.zip")
-}
-```
-
-```{r}
-path <- unzip("ramlegacy.zip")
+path <- "../data/appendixB/ramlegacy_v3.0.xlsx"
 sheets <- readxl::excel_sheets(path)

### 2019-06-15 appendix/appendixB.Rmd

@@
 ```{r}
-path <- "../data/appendixB/ramlegacy_v3.0.xlsx"
-sheets <- readxl::excel_sheets(path)
-ram <- lapply(sheets, readxl::read_excel, path = path)
-names(ram) <- sheets
+#if(!file.exists("ramlegacy.zip")){
+#download.file(paste0(
+#  "https://depts.washington.edu/ramlegac/wordpress/databaseVersions/",
+#  "RLSADB_v3.0_(assessment_data_only)_excel.zip"),
+#  "ramlegacy.zip")
+#}
+## Use more robust source
+#path <- unzip("ramlegacy.zip")
+#sheets <- readxl::excel_sheets(path)
+#ram <- lapply(sheets, readxl::read_excel, path = path)
+#names(ram) <- sheets
+library(ramlegacy)
+download_ramlegacy("3.0", overwrite=TRUE, quiet = TRUE)
+ram <- load_ramlegacy("3.0")
+```
 
+```{r}
 ramlegacy <- 
